### 2017-06-23 paper-webserver.tex
START: webserver paper

\documentclass[a4,center,fleqn]{NAR}

% Enter dates of publication
\copyrightyear{2008}
\pubdate{31 July 2009}
\pubyear{2009}
\jvolume{37}
\jissue{12}

\bibliography{paper-webserver.bib}

%\articlesubtype{This is the article type (optional)}

\begin{document}

\title{Article title}

\author{%
Corresponding Author\,$^{1,*}$,
First Co-Author\,$^{2}$
and Second Co-Author\,$^2$%
\footnote{To whom correspondence should be addressed.
Tel: +44 000 0000000; Fax: +44 000 0000000; Email: xxx@yyyy.ac.zz}}

\address{%
$^{1}$Affiliation of Corresponding Author
and
$^{2}$Affiliation of Both Co-Authors}
% Affiliation must include:
% Department name, institution name, full road and district address,
% state, Zip or postal code, country

\history{%
Received January 1, 2009;
Revised February 1, 2009;
Accepted March 1, 2009}

\maketitle

\begin{abstract}
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text.
\end{abstract}

\section{Introduction}

The surveilance of bacterial pathogens is routine function of public-health agencies. In \textit{Salmonella enterica} \cite{bell2016recent}, \textit{Escherichia coli} \cite{fratamico2016advances}, and many others pathogens \cite{ronholm2016navigating}, culture-dependent subtyping techniques are being displaced by culture-independent methods based on whole-genome sequencing (WGS). Advances in WGS technologies have allowed the monitoring and sharing of outbreak data is real-time \cite{ronholm2016navigating}.

Subtyping software often requires command-line or programming experience. To make these \textit{in-silico} techniques user-friendly for front-line clinicians and laboratories, a variety of web services have been developed in Denmark \cite{joensen2014real}, and in collaborations \cite{hasman2015detection}. While useful, argubly \cite{fricke2014bacterial}, the lack of bioinformatic standards and common infrastructure technologies have made it difficult to fully leverage the big-data results from WGS.

Here we present Spfy, a platform, for bundling subtyping modules and machine-learning analyses, built using modern containerization, task queues, and exception handling.

Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text.
Text \cite{1}.

Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
\begin{align*}
&\mathrm{Ascorbate} + \mathrm{EDTA} \cdot \mathrm{Fe}^{3+} \to
\hbox{Oxidized ascorbate}
\\
&\mathrm{EDTA} \cdot \mathrm{Fe}^{2+} + \mathrm{H}_2
\mathrm{O}_2 \to
\mathrm{EDTA} \cdot \mathrm{Fe}^{3+} + \cdot
\mathrm{OH} + \mathrm{OH}^-
\end{align*}
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.

Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
% **************************************************************
% Keep this command to avoid text of first page running into the
% first page footnotes
\enlargethispage{-65.1pt}
% **************************************************************

Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text.
Text \cite{2,3}.

Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text.

Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text.
Text \cite{4}.

\section{MATERIALS AND METHODS}

\subsection{Materials subsection one}

Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.

\subsubsection{Materials subsubsection one.}

Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text:
\begin{align}
\mathrm{LD}^r = \frac{\mathrm{LD}}{A_\mathrm{iso}}
= 1.5 S \left( 3 \cos^2 \alpha_i - 1 \right)
\end{align}
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.

\subsection{Materials subsection two}

Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text
(see Figure \ref{NAR-fig1}).

Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text.
\begin{equation*}
\mathrm{LD} \left( t \right) =
\sum\limits_i
a_i \exp \left( \frac{-t}{\tau_i} \right)
\end{equation*}
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text.

\begin{figure}[t]
\begin{center}
\includegraphics{NAR-fig1.eps}
\end{center}
\caption{Caption for figure within column.}
\label{NAR-fig1}
\end{figure}

\section{RESULTS}

\subsection{Results subsection one}

Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text.

\begin{table}[b]
\tableparts{%
\caption{This is a table caption}
\label{table:01}%
}{%
\begin{tabular*}{\columnwidth}{@{}lllll@{}}
\toprule
Col. head 1 & Col. head 2 & Col. head 3 & Col. head 4 & Col. head 5
\\
& (\%) & (s$^{-1}$) & (\%) & (s$^{-1}$)
\\
\colrule
Row 1 & Row 1 & Row 1 & -- & --
\\
Row 2 & Row 2 & Row 2 & Row 2 & Row 2
\\
\botrule
\end{tabular*}%
}
{This is a table footnote}
\end{table}

\subsection{Results subsection two}

Text.  Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text (see Table \ref{table:01}).

Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text (see Figure \ref{NAR-fig2}a).

Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.

\begin{figure*}[t]
\begin{center}
\includegraphics{NAR-fig2.eps}
\end{center}
\caption{Caption for wide figure over two columns.
\textbf{(a)} Left figure.
\textbf{(b)} Right figure (see (a)).
}
\label{NAR-fig2}
\end{figure*}

\subsection{Results subsection three}

Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text.

\section{DISCUSSION}

\subsection{Discussion subsection one}

Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text.

\subsection{Discussion subsection two}

Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text.

Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.

\subsection{Discussion subsection three}

Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text.

Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text.

Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text.

\section{CONCLUSION}

Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text.

\section{ACKNOWLEDGEMENTS}

Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
Text. Text. Text. Text.

\subsubsection{Conflict of interest statement.} None declared.
\newpage

\begin{thebibliography}{4}

% Format for article
\bibitem{1}
Author,A.B. and Author,C. (1992)
Article title.
\textit{Abbreviated Journal Name}, \textbf{5}, 300--330.

% Format for book
\bibitem{2}
Author,D., Author,E.F. and Author,G. (1995)
\textit{Book Title}.
Publisher Name, Publisher Address.

% Format for chapter in book
\bibitem{3}
Author,H. and Author,I. (2005)
Chapter title.
In
Editor,A. and Editor,B. (eds),
\textit{Book Title},
Publisher Name, Publisher Address,
pp.\ 60--80.

% Another article
\bibitem{4}
Author,Y. and Author,Z. (2002)
Article title.
\textit{Abbreviated Journal Name}, \textbf{53}, 500--520.

\end{thebibliography}

\end{document}

### 2017-06-27 paper-webserver.tex
ADD: updates to the first two paras

@@
 
-The surveilance of bacterial pathogens is routine function of public-health agencies.
-In \textit{Salmonella enterica} \cite{bell2016recent}, \textit{Escherichia coli} \cite{fratamico2016advances}, and many others pathogens \cite{ronholm2016navigating}, culture-dependent subtyping techniques are being displaced by culture-independent methods based on whole-genome sequencing (WGS).
-Advances in WGS technologies have allowed the monitoring and sharing of outbreak data is real-time \cite{ronholm2016navigating}.
+The surveilance of bacterial pathogens is a routine function of public-health agencies.
+In \textit{Salmonella enterica} \cite{bell2016recent}, \textit{Escherichia coli} \cite{fratamico2016advances}, and others pathogens \cite{ronholm2016navigating}, culture-dependent subtyping techniques are becoming displaced by culture-independent methods based on whole-genome sequencing (WGS).
+Secondary analyses, the study of relations and historical relevance of genomes, are now possible on large datasets through WGS \cite{ronholm2016navigating}.
+Advances in WGS technologies have also enabled the monitoring and sharing of outbreak information in real-time \cite{ronholm2016navigating}.
 
-Subtyping software often requires command-line or programming experience.
-To make these \textit{in-silico} techniques user-friendly for front-line clinicians and laboratories, a variety of web services have been developed in Denmark \cite{joensen2014real}, and in collaborations \cite{hasman2015detection}.
+Subtyping and other analyses software often requires command-line or programming experience.
+To make these \textit{in-silico} techniques user-friendly for front-line clinicians and laboratories, a variety of web services have been developed in Denmark \cite{joensen2014real}, Canada \cite{whiteside2016superphy} and in collaborations \cite{hasman2015detection}.
 While useful, argubly \cite{fricke2014bacterial}, the lack of bioinformatic standards and common infrastructure technologies have made it difficult to fully leverage the big-data results from WGS.
 
-Here we present Spfy, a platform, for bundling subtyping modules and machine-learning analyses, built using modern containerization, task queues, and exception handling.
+Here we present Spfy, a platform for bundling subtyping modules and machine-learning predictions, built using modern containerization of services, task queues for parrelelizing analyses, and exception handling throughout.
 
@@
 Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text.
+Text. Text. Text{}. Text.
 Text \cite{1}.

### 2017-06-27 paper-webserver.tex
ADD: third para

@@
 
-Here we present Spfy, a platform for bundling subtyping modules and machine-learning predictions, built using modern containerization of services, task queues for parrelelizing analyses, and exception handling throughout.
+Here we present Spfy, a platform for bundling subtyping modules and machine-learning predictions, and built using modern containerization of services, task queues for parrelelizing analyses, and exception handling throughout.
+Spfy offers serotype prediction, identification of antimicrobial resistance genes through the Comprehensive Antibiotic Resistance Database (CARD) \cite{mcarthur2013comprehensive}, and statisical or machine-learning based comparisons of groups based on user-selected restrictions.
 

### 2017-06-27 paper-webserver.tex
ADD: updates to third para

--- 2017-06-27 paper-webserver.tex
@@
 In \textit{Salmonella enterica} \cite{bell2016recent}, \textit{Escherichia coli} \cite{fratamico2016advances}, and others pathogens \cite{ronholm2016navigating}, culture-dependent subtyping techniques are becoming displaced by culture-independent methods based on whole-genome sequencing (WGS).
-Secondary analyses, the study of relations and historical relevance of genomes, are now possible on large datasets through WGS \cite{ronholm2016navigating}.
+Secondary analyses, the study of relations and historical relevance of genomes, are now possible on large datasets due to WGS \cite{ronholm2016navigating}.
 Advances in WGS technologies have also enabled the monitoring and sharing of outbreak information in real-time \cite{ronholm2016navigating}.
 
-Subtyping and other analyses software often requires command-line or programming experience.
-To make these \textit{in-silico} techniques user-friendly for front-line clinicians and laboratories, a variety of web services have been developed in Denmark \cite{joensen2014real}, Canada \cite{whiteside2016superphy} and in collaborations \cite{hasman2015detection}.
+Subtyping and other analyses software often require command-line or programming experience.
+To make such \textit{in-silico} techniques user-friendly for front-line clinicians and laboratories, a variety of web services have been developed in Denmark \cite{joensen2014real}, Canada \cite{whiteside2016superphy}, and in collaborations \cite{hasman2015detection}.
 While useful, argubly \cite{fricke2014bacterial}, the lack of bioinformatic standards and common infrastructure technologies have made it difficult to fully leverage the big-data results from WGS.
 
-Here we present Spfy, a platform for bundling subtyping modules and machine-learning predictions, built using modern containerization of services, task queues for parrelelizing analyses, and exception handling throughout.
+Here we present Spfy, a platform bundling subtyping modules and machine-learning predictions for \textit{E.
+coli}, built using modern containerization of services, task queues for parrelelizing analyses, and exception handling throughout.
+Spfy offers serotype predictions, identification of antimicrobial resistance genes through the Comprehensive Antibiotic Resistance Database (CARD) \cite{mcarthur2013comprehensive}, and statisical or machine-learning based comparisons of groups based on user-selected restrictions.
 

### 2017-06-27 paper-webserver.tex
CHANGE: citation

@@
 Subtyping and other analyses software often require command-line or programming experience.
-To make such \textit{in-silico} techniques user-friendly for front-line clinicians and laboratories, a variety of web services have been developed in Denmark \cite{joensen2014real}, Canada \cite{whiteside2016superphy}, and in collaborations \cite{hasman2015detection}.
+To make such \textit{in-silico} techniques user-friendly for front-line clinicians and laboratories, a variety of web services have been developed in Denmark \cite{joensen2014real}, Taiwan \cite{liu2016construction}, and in collaborations \cite{hasman2015detection}.
 While useful, argubly \cite{fricke2014bacterial}, the lack of bioinformatic standards and common infrastructure technologies have made it difficult to fully leverage the big-data results from WGS.

### 2017-06-27 paper-webserver.tex
ADD: changes to para 3 of intro
FIX: spelling

@@
 
-The surveilance of bacterial pathogens is a routine function of public-health agencies.
+The surveillance of bacterial pathogens is a routine function of public-health agencies.
 In \textit{Salmonella enterica} \cite{bell2016recent}, \textit{Escherichia coli} \cite{fratamico2016advances}, and others pathogens \cite{ronholm2016navigating}, culture-dependent subtyping techniques are becoming displaced by culture-independent methods based on whole-genome sequencing (WGS).
 Secondary analyses, the study of relations and historical relevance of genomes, are now possible on large datasets due to WGS \cite{ronholm2016navigating}.
-Advances in WGS technologies have also enabled the monitoring and sharing of outbreak information in real-time \cite{ronholm2016navigating}.
+Advances in WGS technologies have also enabled the monitoring and sharing of outbreak genomes in real-time \cite{ronholm2016navigating}.
 
@@
 To make such \textit{in-silico} techniques user-friendly for front-line clinicians and laboratories, a variety of web services have been developed in Denmark \cite{joensen2014real}, Taiwan \cite{liu2016construction}, and in collaborations \cite{hasman2015detection}.
-While useful, argubly \cite{fricke2014bacterial}, the lack of bioinformatic standards and common infrastructure technologies have made it difficult to fully leverage the big-data results from WGS.
+While useful, arguably \cite{fricke2014bacterial}, the lack of bioinformatic standards and common infrastructure technologies have made it difficult to fully leverage the big-data results from WGS.
 
 Here we present Spfy, a platform bundling subtyping modules and machine-learning predictions for \textit{E.
-coli}, built using modern containerization of services, task queues for parrelelizing analyses, and exception handling throughout.
-Spfy offers serotype predictions, identification of antimicrobial resistance genes through the Comprehensive Antibiotic Resistance Database (CARD) \cite{mcarthur2013comprehensive}, and statisical or machine-learning based comparisons of groups based on user-selected restrictions.
+coli}, built using modern containerization of services, task queues for parallelizing analyses, and exception handling throughout.
+Spfy offers serotype predictions, identification of antimicrobial resistance genes through the Comprehensive Antibiotic Resistance Database (CARD) \cite{mcarthur2013comprehensive}, and statistical or machine-learning based comparisons of groups based on user-selected restrictions.
+All services are publicly available at \url{https://lfz.corefacility.ca/superphy/spfy/}.
 

### 2017-06-28 paper-webserver.tex
ADD: fourth para

@@
 Here we present Spfy, a platform bundling subtyping modules and machine-learning predictions for \textit{E.
-coli}, built using modern containerization of services, task queues for parallelizing analyses, and exception handling throughout.
-Spfy offers serotype predictions, identification of antimicrobial resistance genes through the Comprehensive Antibiotic Resistance Database (CARD) \cite{mcarthur2013comprehensive}, and statistical or machine-learning based comparisons of groups based on user-selected restrictions.
+coli}, built using containerization of services, task queues for parallelizing analyses, and exception handling throughout.
+Spfy offers serotype predictions, identification of antimicrobial resistance (AMR) genes through the Comprehensive Antibiotic Resistance Database (CARD) \cite{mcarthur2013comprehensive}, and statistical or machine-learning based comparisons of groups based on user-selected restrictions.
 All services are publicly available at \url{https://lfz.corefacility.ca/superphy/spfy/}.
+
+Previously, we developed Superphy \cite{whiteside2016superphy} which determines Shiga toxin subtypes, AMR genes, and statistical significance tests.
+Spfy is a complete rewrite of Superphy integrating our Panseq \cite{laing2010pan} for pan-genome analyses of accessory genomic regions, providing long-term storage in a graph database, and containerization of services using Docker, a modern approach found in many web servers and proposed to mitigate the complexities of bioinformatic pipelines \cite{di2015impact}.
 

### 2017-06-28 paper-webserver.tex
CHANGE: collapse para 3 and 4 together

--- 2017-06-27 paper-webserver.tex
@@
 
-Here we present Spfy, a platform bundling subtyping modules and machine-learning predictions for \textit{E.
-coli}, built using modern containerization of services, task queues for parallelizing analyses, and exception handling throughout.
-Spfy offers serotype predictions, identification of antimicrobial resistance genes through the Comprehensive Antibiotic Resistance Database (CARD) \cite{mcarthur2013comprehensive}, and statistical or machine-learning based comparisons of groups based on user-selected restrictions.
+Previously, we developed Superphy \cite{whiteside2016superphy} a online platform, targeting \textit{E.
+coli}, which determines Shiga toxin subtypes, AMR genes through the Comprehensive Antibiotic Resistance Database (CARD) \cite{mcarthur2013comprehensive}, and statistical significance tests between subgroups in associated metadata.
+Here we present Spfy, a complete rewrite of Superphy integrating our Panseq package \cite{laing2010pan} for pan-genomic analyses of accessory regions, long-term storage of results in a graph database, and containerization of services using Docker, a modern approach found in many web servers and proposed to mitigate the complexities of bioinformatic pipelines \cite{di2015impact}.
+Furthermore, Spfy takes a modular approach to tool integration and distributes analyses over a task queue, allowing modules to run in parallel.
 All services are publicly available at \url{https://lfz.corefacility.ca/superphy/spfy/}.

### 2017-06-28 paper-webserver.tex
ADD: new 4th para - not sure about it

@@
 All services are publicly available at \url{https://lfz.corefacility.ca/superphy/spfy/}.
+
+Spfy enables processing of thousands of genome sequences using multiple modules.
+By using task queue workers, enabled by the Python-based Redis Queue library, performance can scale to available infrastructure.
+As a bioinformatics tool, we also prioritized reliability by integrating the open-source Sentry platform for real-time exception tracking.
+The use of Docker-Compose for orchestrating containers ensures that the entire platform, webservers, databases, and task workers, can be easily replicated by other researchers.
+We present Spfy as a template for brining advance in web application development and deployment to online bioinformatics platforms.
 

### 2017-06-28 paper-webserver.tex
FIX: grammar?

--- 2017-06-28 paper-webserver.tex
@@
 
-Previously, we developed Superphy \cite{whiteside2016superphy} a online platform, targeting \textit{E.
-coli}, which determines Shiga toxin subtypes, AMR genes through the Comprehensive Antibiotic Resistance Database (CARD) \cite{mcarthur2013comprehensive}, and statistical significance tests between subgroups in associated metadata.
+Previously we developed Superphy \cite{whiteside2016superphy}, an online platform, targeting \textit{E.
+coli}, which determines Shiga toxin subtypes, AMR genes using the Comprehensive Antibiotic Resistance Database (CARD) \cite{mcarthur2013comprehensive}, and provides statistical significance tests between subgroups in associated metadata.
 Here we present Spfy, a complete rewrite of Superphy integrating our Panseq package \cite{laing2010pan} for pan-genomic analyses of accessory regions, long-term storage of results in a graph database, and containerization of services using Docker, a modern approach found in many web servers and proposed to mitigate the complexities of bioinformatic pipelines \cite{di2015impact}.
-Furthermore, Spfy takes a modular approach to tool integration and distributes analyses over a task queue, allowing modules to run in parallel.
+Furthermore, Spfy takes a modular approach to tool integration and distributes analyses over task queues, thereby allowing modules to run in parallel.
 All services are publicly available at \url{https://lfz.corefacility.ca/superphy/spfy/}.
+
+Spfy enables processing of thousands of genome sequences using multiple modules.
+By using task queue workers, enabled by the Python-based Redis Queue library, performance can quickly scale to available infrastructure.
+As a bioinformatics tool, we ensured reliability of the platform by integrating the open-source Sentry toolkit for real-time exception tracking.
+The use of Docker-Compose for orchestrating containers ensures that the entire platform, webservers, databases, and task workers, can be easily replicated by other researchers.
+We present Spfy as a template for brining advances in web application development and deployment to online bioinformatics platforms.
 

### 2017-06-28 paper-webserver.tex
ADD: grammer + 5th para

@@
 In \textit{Salmonella enterica} \cite{bell2016recent}, \textit{Escherichia coli} \cite{fratamico2016advances}, and others pathogens \cite{ronholm2016navigating}, culture-dependent subtyping techniques are becoming displaced by culture-independent methods based on whole-genome sequencing (WGS).
-Secondary analyses, the study of relations and historical relevance of genomes, are now possible on large datasets due to WGS \cite{ronholm2016navigating}.
+Secondary analyses, the study of relations and historical relevance of genomes, are now possible on large datasets provided by WGS \cite{ronholm2016navigating}.
 Advances in WGS technologies have also enabled the monitoring and sharing of outbreak genomes in real-time \cite{ronholm2016navigating}.
@@
 
-Previously we developed Superphy \cite{whiteside2016superphy}, an online platform, targeting \textit{E.
-coli}, which determines Shiga toxin subtypes, AMR genes using the Comprehensive Antibiotic Resistance Database (CARD) \cite{mcarthur2013comprehensive}, and provides statistical significance tests between subgroups in associated metadata.
+Previously we developed Superphy \cite{whiteside2016superphy}, an online platform targeting \textit{E.
+coli}, which determines Shiga toxin subtypes, Antimicrobial Resistance (AMR) genes using the Comprehensive Antibiotic Resistance Database (CARD) \cite{mcarthur2013comprehensive}, and provides statistical significance tests between subgroups in associated metadata.
 Here we present Spfy, a complete rewrite of Superphy integrating our Panseq package \cite{laing2010pan} for pan-genomic analyses of accessory regions, long-term storage of results in a graph database, and containerization of services using Docker, a modern approach found in many web servers and proposed to mitigate the complexities of bioinformatic pipelines \cite{di2015impact}.
@@
 Spfy enables processing of thousands of genome sequences using multiple modules.
-By using task queue workers, enabled by the Python-based Redis Queue library, performance can quickly scale to available infrastructure.
-As a bioinformatics tool, we ensured reliability of the platform by integrating the open-source Sentry toolkit for real-time exception tracking.
-The use of Docker-Compose for orchestrating containers ensures that the entire platform, webservers, databases, and task workers, can be easily replicated by other researchers.
+By using task queue workers, enabled by the Python-based Redis Queue library \url{https://github.com/nvie/rq}, performance can quickly scale to available infrastructure.
+As a bioinformatics tool, we ensured reliability of the platform by integrating the open-source Sentry toolkit \url{https://github.com/getsentry/sentry} for real-time exception tracking.
+Containers are orchestrated through Docker-Compose allowing the entire platform: webservers, databases, and task workers, to be easily replicated by other researchers.
 We present Spfy as a template for brining advances in web application development and deployment to online bioinformatics platforms.
+
+We designed Spfy's front-end following the Material Design standards \url{https://material.io/}, released by Google.
+The user interface is implemented with the React JavaScript library \url{https://facebook.github.io/react/}, as a single-page application to allow efficient data-flow without reloading the website.
+Everything is compartmentalized within separate Docker containers where the front-end is networked, through Docker-Compose, to the back-end webserver.
 

### 2017-07-06 paper-webserver.tex
CHANGE: one sentance per line
CHNAGE: position of 5th para moved to 4th
UPDATE: sentance
Added an outline for the intro. Separated Materials / Methods

@@
 \section{Introduction}
+Broad --> narrow
+-Bacterial infections are a cause of human disease -> associated costs
+-Surveillance is used to monitor known pathogens and detect emerging threats, to lower burden of human illness
+-Until recently, surveillance involved traditional reference laboratory techniques (and still does) --> examples of both old and new (WGS), what they measure, how they are used
+-WGS methods have great potential, but some drawbacks (list with references)
+-Some of the drawbacks can be solved by providing pre-computed analyses for routine tasks
+-We have previously created SuperPhy, which helps solve some of the issues
+-Some issues still remain
+-In this study, we have addressed these issues by ... which will provide these benefits
 
@@
 
+\section{MATERIALS AND METHODS}
+We designed Spfy's front-end following the Material Design standards \url{https://material.io/}, released by Google.
+The user interface is implemented with the React JavaScript library \url{https://facebook.github.io/react/}, as a single-page application to allow efficient data-flow without reloading the website.
+Everything is compartmentalized within separate Docker containers where the front-end is networked, through Docker-Compose, to the back-end webserver.
+
 Spfy enables processing of thousands of genome sequences using multiple modules.
@@
 Containers are orchestrated through Docker-Compose allowing the entire platform: webservers, databases, and task workers, to be easily replicated by other researchers.
-We present Spfy as a template for brining advances in web application development and deployment to online bioinformatics platforms.
-
-We designed Spfy's front-end following the Material Design standards \url{https://material.io/}, released by Google.
-The user interface is implemented with the React JavaScript library \url{https://facebook.github.io/react/}, as a single-page application to allow efficient data-flow without reloading the website.
-Everything is compartmentalized within separate Docker containers where the front-end is networked, through Docker-Compose, to the back-end webserver.
+We present Spfy as an integrated predictive genomics toolkit for \textit{E. coli} which leverages advances in web application development and deployment.
 

### 2017-07-13 paper-webserver.tex
CHANGE: sentance

--- 2017-06-28 paper-webserver.tex
@@
 
+We designed Spfy's front-end following the Material Design standards \url{https://material.io/}, released by Google.
+The user interface is implemented with the React JavaScript library \url{https://facebook.github.io/react/}, by Facebook, as a single-page application to allow efficient data-flow without reloading the website.
+Everything is compartmentalized within separate Docker containers where the front-end is networked, through Docker-Compose, to the back-end webserver.
+
 Spfy enables processing of thousands of genome sequences using multiple modules.
@@
 Containers are orchestrated through Docker-Compose allowing the entire platform: webservers, databases, and task workers, to be easily replicated by other researchers.
-We present Spfy as a template for brining advances in web application development and deployment to online bioinformatics platforms.
-
-We designed Spfy's front-end following the Material Design standards \url{https://material.io/}, released by Google.
-The user interface is implemented with the React JavaScript library \url{https://facebook.github.io/react/}, as a single-page application to allow efficient data-flow without reloading the website.
-Everything is compartmentalized within separate Docker containers where the front-end is networked, through Docker-Compose, to the back-end webserver.
+We present Spfy as an integrated predictive genomics toolkit for pathogens / \textit{E. coli} which leverages advances in web application development and deployment.
 

### 2017-07-13 paper-webserver.tex

--- 2017-07-06 paper-webserver.tex
@@
 We designed Spfy's front-end following the Material Design standards \url{https://material.io/}, released by Google.
-The user interface is implemented with the React JavaScript library \url{https://facebook.github.io/react/}, as a single-page application to allow efficient data-flow without reloading the website.
+The user interface is implemented with the React JavaScript library \url{https://facebook.github.io/react/}, by Facebook, as a single-page application to allow efficient data-flow without reloading the website.
 Everything is compartmentalized within separate Docker containers where the front-end is networked, through Docker-Compose, to the back-end webserver.
@@
 Containers are orchestrated through Docker-Compose allowing the entire platform: webservers, databases, and task workers, to be easily replicated by other researchers.
-We present Spfy as an integrated predictive genomics toolkit for \textit{E. coli} which leverages advances in web application development and deployment.
+We present Spfy as an integrated predictive genomics toolkit for pathogens / \textit{E. coli} which leverages advances in web application development and deployment.
 

### 2017-07-20 paper-webserver.tex
ADD: pts for new general 1st para

@@
 -In this study, we have addressed these issues by ... which will provide these benefits
+
+Amongst produce in Canada, \textit{Salmonella enterica} and \textit{Escherichia coli} are the two most common bacetrial infections \cite{kozak2013foodborne}.
+Translate into actual costs: hospitalization adn death \cite{thomas2015estimates}.
+Of which, monetary costs have been known for decades \cite{todd1989costs}.
 

### 2017-07-21 paper-webserver.tex
ADD: bunch of new paras wrt Chads suggestions, still need to expand on drawbacks

@@
 
+% costs of bacterial infections
+With the prevalence of ready-to-eat produce and expansion of global food-supply chains, consumers are a greater risk for bacterial infections in an environment that is more complicated to track than ever before \cite{kozak2013foodborne}.
 Amongst produce in Canada, \textit{Salmonella enterica} and \textit{Escherichia coli} are the two most common bacetrial infections \cite{kozak2013foodborne}.
-Translate into actual costs: hospitalization adn death \cite{thomas2015estimates}.
-Of which, monetary costs have been known for decades \cite{todd1989costs}.
+Foodborne illnesses translate into actual costs: in Canada alone, an estimated 4000 hospitalizations and 105 deaths are thought to be the result of various pathogens \cite{thomas2015estimates}.
+Futhermore, economic costs have been known for decades \cite{todd1989costs}.
 
-The surveillance of bacterial pathogens is a routine function of public-health agencies.
-In \textit{Salmonella enterica} \cite{bell2016recent}, \textit{Escherichia coli} \cite{fratamico2016advances}, and others pathogens \cite{ronholm2016navigating}, culture-dependent subtyping techniques are becoming displaced by culture-independent methods based on whole-genome sequencing (WGS).
+% surveillance
+The surveillance of bacterial pathogens is a routine function of public-health agencies in an effort to lower the burden caused by human illness.
+In \textit{Salmonella enterica} \cite{bell2016recent}, \textit{Escherichia coli} \cite{fratamico2016advances}, and others pathogens \cite{ronholm2016navigating}, culture-dependent subtyping techniques are lately coupled with culture-independent methods based on whole-genome sequencing (WGS) to monitor known pathogens and detect emerging threats.
+
+% traditional reference lab techniques
+Subtyping encompasses various methods to designate particular strains of bacterial isolates.
+In order to monitor outbreaks, platforms must be able to rapidly subtype collected strains and communicate such information to other agencies \cite{ronholm2016navigating}. 
+Traditional reference laboratory techniques handled subtyping through methods such as serotyping, detecting the presence of specific cell surface antigens as tested by antibody reaction, pulsed-field gel electrophoresis (PFGE), to identify known virulence factors (VFs) or antimicrobial resistance genes (AMR genes), or Sanger sequencing, to determine multi-locus sequence types (MLSTs) \cite{ronholm2016navigating}.
+Using PFGE, outbreaks can be tracked by comparing band patterns of bacterial isolates, and is routinely performed in public health agencies \cite{kozak2013foodborne}. 
+The original computational methods to leverage WGS focused on creating \textit{in-silico} approaches to return subtyping results matching that of wet-lab techniques.
+Software such as VirulenceFinder \cite{joensen2014real} compares a pre-computed set of reference sequences encompassing known virulence factor gene sequences against a WGS result.
+Similar approaches have been applied to predict AMR genes \cite{mcarthur2013comprehensive}, O-type, and H-type \cite{whiteside2016superphy}.
+
+% potential of WGS
+WGS has greater potential due to the inclusion of the entire genomes; in effect, this lends WGS-based methods better resolution and context in evaluating the spread of outbreaks \cite{ronholm2016navigating}.
+For example, instead of examining selected single-nucleotide polymorphisms (SNPs) in a subset of housekeeping genes, as done for laboratory-based MLST, newer methods such as core genome MLST (cgMLST) attempt to cover every common gene.
 Secondary analyses, the study of relations and historical relevance of genomes, are now possible on large datasets provided by WGS \cite{ronholm2016navigating}.
@@
 
-Subtyping and other analyses software often require command-line or programming experience.
+% Drawbacks
+However, these WGS-base methods still examine the underlying genome, not necessarily the expressed genes.
+Correspondingly, a greater number of target sequences are identified which complicates risk assessments \cite{ronholm2016navigating}.
+The effectiveness of WGS-based methods is also largely reliant on the accuracy and breadth of reference datasets.
+
+% Solution via pre-computed analyses for routine tasks
+Efforts such as the Comprehensive Antibiotic Resistance Database (CARD) \cite{mcarthur2013comprehensive} attempt to alleviate data concerns by curating a set of known AMR genes.
+In addition, agencies such as the Food and Drug Administration (FDA) (CITE GENOME TRACKR) and the National Center for Biotechnology Information (NCBI) (CITE ???) provide services to openly share genome sequences and associated metadata.
+Recent efforts have been focused on providing modern tools for quality control of reference genomes (CITE SOME PAPER ON QC OF BACETIRA GENOMES), particularly amongst human-genome projects (CITE NGSCHECKMATE, GOOGLE GENOMICS).
+Panseq \cite{laing2010pan}, Roary \cite{page2015roary}, and chewBBACA \url{https://github.com/mickaelsilva/chewBBACA} are available to compute core and accessory regions and identify SNPs across a large collection of reference genomes.
+Taken together, open-access initiatives and shared tooling aim to alleviate the drawbacks of WGS-based methods by pre-computing sufficient reference data.
+
+% why Superphy
+Subtyping and computing reference data often require command-line or programming experience.
 To make such \textit{in-silico} techniques user-friendly for front-line clinicians and laboratories, a variety of web services have been developed in Denmark \cite{joensen2014real}, Taiwan \cite{liu2016construction}, and in collaborations \cite{hasman2015detection}.
@@
 
+% Superphy
 Previously we developed Superphy \cite{whiteside2016superphy}, an online platform targeting \textit{E.
-coli}, which determines Shiga toxin subtypes, Antimicrobial Resistance (AMR) genes using the Comprehensive Antibiotic Resistance Database (CARD) \cite{mcarthur2013comprehensive}, and provides statistical significance tests between subgroups in associated metadata.
-Here we present Spfy, a complete rewrite of Superphy integrating our Panseq package \cite{laing2010pan} for pan-genomic analyses of accessory regions, long-term storage of results in a graph database, and containerization of services using Docker, a modern approach found in many web servers and proposed to mitigate the complexities of bioinformatic pipelines \cite{di2015impact}.
+coli}, which determines Shiga toxin subtypes, Antimicrobial Resistance (AMR) genes using CARD, and provides statistical significance tests between subgroups in associated metadata and pan-genome regions.
+However, we struggled to provide real-time computing of results and to accomdate an ever-growing reference database.
+Furthermore, due to the nature of the various dependencies involved, the platform was difficult for other researchers to independently replicate.
+
+% this study
+Here we present Spfy, a complete rewrite of Superphy which expands on supported features to also provide long-term storage of results in a graph database, and containerization of services using Docker, a modern approach found in many web servers and proposed to mitigate the complexities of bioinformatic pipelines \cite{di2015impact}.
 Furthermore, Spfy takes a modular approach to tool integration and distributes analyses over task queues, thereby allowing modules to run in parallel.
+These changes enable us to provide real-time genome analysis and database storage of all results for upwards of 5000 genomes.
 All services are publicly available at \url{https://lfz.corefacility.ca/superphy/spfy/}.

### 2017-07-24 paper-webserver.tex
ADD: more details

--- 2017-07-13 paper-webserver.tex
@@
 
-The surveillance of bacterial pathogens is a routine function of public-health agencies.
-In \textit{Salmonella enterica} \cite{bell2016recent}, \textit{Escherichia coli} \cite{fratamico2016advances}, and others pathogens \cite{ronholm2016navigating}, culture-dependent subtyping techniques are becoming displaced by culture-independent methods based on whole-genome sequencing (WGS).
+% costs of bacterial infections
+As food-sourcing is globalized, consumers are at a greater risk for bacterial infections in an environment that is more complicated to track than ever before \cite{kozak2013foodborne}.
+\textit{Salmonella enterica} and \textit{Escherichia coli} are amongst the two most common bacetrial infections \cite{kozak2013foodborne} originating from produce.
+Foodborne illnesses translate into actual costs: for example, in Canada alone an estimated 4000 hospitalizations and 105 deaths per year are thought to be the result of various pathogens \cite{thomas2015estimates}.
+Futhermore, the associated economic costs have been known for decades \cite{todd1989costs}.
+
+% surveillance
+The surveillance of bacterial pathogens is a routine function of public-health agencies in an effort to lower the burden caused by human illness.
+In \textit{Salmonella enterica} \cite{bell2016recent}, \textit{Escherichia coli} \cite{fratamico2016advances}, and others pathogens \cite{ronholm2016navigating}, culture-dependent subtyping techniques are lately coupled with culture-independent methods based on whole-genome sequencing (WGS), to monitor known pathogens and detect emerging threats.
+
+% traditional reference lab techniques
+Subtyping encompasses various methods to designate particular strains of bacterial isolates.
+In order to monitor outbreaks, platforms must rapidly subtype collected strains and communicate such information to other agencies \cite{ronholm2016navigating}. 
+Traditional reference laboratory techniques handled subtyping through methods such as serotyping, detecting the presence of specific cell surface antigens as tested by antibody reaction, pulsed-field gel electrophoresis (PFGE), to identify known virulence factors (VFs) or antimicrobial resistance genes (AMR genes), or Sanger sequencing, to determine multi-locus sequence types (MLSTs) \cite{ronholm2016navigating}.
+Using PFGE, outbreaks can be tracked by comparing band patterns of bacterial isolates and is routinely performed in public health agencies \cite{kozak2013foodborne}. 
+The original computational methods to leverage WGS focused on creating \textit{in-silico} approaches to return subtyping results matching those of wet-lab techniques.
+Software such as VirulenceFinder \cite{joensen2014real} compares a pre-collected set of reference sequences encompassing known virulence factor gene sequences against a WGS result.
+Similar approaches have been applied to predict AMR genes \cite{mcarthur2013comprehensive}, O-type, and H-type \cite{whiteside2016superphy}.
+
+% potential of WGS
+WGS has the potential to expand on traditional methods, due to the inclusion of the entire genome in WGS-based analyses; in effect, this lends WGS-based methods better resolution and context in evaluating the spread of outbreaks \cite{ronholm2016navigating}.
+As an example, instead of examining selected single-nucleotide polymorphisms (SNPs) in a subset of housekeeping genes, as is the case for laboratory-based MLST, newer methods such as core genome MLST (cgMLST) attempt to cover every common gene.
 Secondary analyses, the study of relations and historical relevance of genomes, are now possible on large datasets provided by WGS \cite{ronholm2016navigating}.
@@
 
-Subtyping and other analyses software often require command-line or programming experience.
+% Drawbacks - genetics
+However, these WGS-base methods still examine the underlying genetics, not the expressed phenotype.
+The presence of a particular gene sequence does not necessitate expression.
+Correspondingly, a greater number of target sequences are identified which complicates risk assessments \cite{ronholm2016navigating} and laboratory based phenotypic tests are still relatively inexpensive \cite(CITE ROUTINE USE - KOSER).
+The effectiveness of WGS-based methods is also reliant on the accuracy and breadth of reference datasets.
+
+% drawbacks - big data
+Traditionaly, bioinformatics software was developed \textit{ad hoc}, without employing software engineering principles \cite(TRENDS - BREVEREN).
+Such approaches were often script-based, with custom data formats, and only suitable for small collections of data \cite(TRENDS - BREVEREN).
+As WGS-technology expanded the availability of data, systems engineering principles became important as code failure could quickly disrupt entire bioinformatics pipelines \cite(BIOLOGICAL DATA SCIENCES - SCHATZ).
+The field also lags behind other big data industries: bioinformatics software typically struggles to perform computations in real-time, stores data in inflexible databases, doesn't support live visualizations \cite(TRENDS - BREVEREN), and lacks a common application programming interface (API) linking resources \cite(REVIEW - SWAMINATHAN).
+
+% Solution via pre-computed analyses for routine tasks
+Efforts such as the Comprehensive Antibiotic Resistance Database (CARD) \cite{mcarthur2013comprehensive} attempt to alleviate data concerns by curating a set of known AMR genes.
+In addition, agencies such as the Food and Drug Administration (FDA), which hosts GenomeTrakr \cite{allard2016practical}, the National Center for Biotechnology Information (NCBI), which hosts GenBank \cite{doi:10.1093/nar/gks1195}, and the Center for Disease Control and Prevention (CDC), which hosts PulseNet \cite(PULSENET), provide services to share genome sequences and associated metadata either openly or amongst member institutions.
+Recent efforts have been focused on providing modern tools for quality control of reference genomes \cite{parks2015checkm}, particularly amongst human-genome projects \cite{lee2017ngscheckmate}.
+Panseq \cite{laing2010pan}, Roary \cite{page2015roary}, and chewBBACA \url{https://github.com/mickaelsilva/chewBBACA} are available to compute core and accessory regions and identify SNPs across a large collection of reference genomes.
+Taken together, open-access initiatives and shared tooling aim to alleviate the drawbacks of WGS-based methods by pre-computing sufficient reference data.
+
+% solution via modern big data approaches
+For genotypic, WGS-based approaches to be truly compliment laboratory-based phenotypic tests, analyses software must be faster or demonstrate a reasonable link between their genotype and phenotype \cite(CITE ROUTINE USE - KOSER).
+Antimicrobial resistance testing, which is a major component of public health monitoring efforts, typically takes 18-24 hours in a reference laboratory \cite(CITE ROUTINE USE - KOSER).
+Results should also be stored in a tiered structure, with rapid access available for retrieval and visualization while maintaining long-term data storage \cite(BIOLOGICAL DATA SCIENCES - SCHATZ).
+Such an approach would enable developers to address the current limitations \cite(TRENDS - BREVEREN) of bioinformatics developers.
+
+% why Superphy
+Subtyping and reference data generation often require command-line or programming experience.
 To make such \textit{in-silico} techniques user-friendly for front-line clinicians and laboratories, a variety of web services have been developed in Denmark \cite{joensen2014real}, Taiwan \cite{liu2016construction}, and in collaborations \cite{hasman2015detection}.
@@
 
+% Superphy
 Previously we developed Superphy \cite{whiteside2016superphy}, an online platform targeting \textit{E.
-coli}, which determines Shiga toxin subtypes, Antimicrobial Resistance (AMR) genes using the Comprehensive Antibiotic Resistance Database (CARD) \cite{mcarthur2013comprehensive}, and provides statistical significance tests between subgroups in associated metadata.
-Here we present Spfy, a complete rewrite of Superphy integrating our Panseq package \cite{laing2010pan} for pan-genomic analyses of accessory regions, long-term storage of results in a graph database, and containerization of services using Docker, a modern approach found in many web servers and proposed to mitigate the complexities of bioinformatic pipelines \cite{di2015impact}.
-Furthermore, Spfy takes a modular approach to tool integration and distributes analyses over task queues, thereby allowing modules to run in parallel.
+coli}, which determines Shiga toxin subtypes, AMR genes using CARD, and provides statistical significance tests between subgroups in associated metadata and pan-genome regions.
+However, we struggled to provide real-time computing of results and to accomdate an ever-growing reference database.
+Furthermore, due to the nature of the various dependencies involved, the platform was difficult for other researchers to independently replicate.
+
+% this study
+Here we present Spfy, a complete rewrite of Superphy which expands on supported features to also provide long-term storage of results in a graph database, one of the emerging \cite(TRENDS - BREVEREN) database types being used for biological data.
+We also containerize of services using Docker, a modern approach found in many web servers and proposed to mitigate the complexities of bioinformatic pipelines \cite{di2015impact}.
+Docker integration ensures that nested dependency resolution, which typically must be manually installed \cite{doi:10.1093/bioinformatics/btu153,laing2010pan,inouye2014srst2}, are handled automatically.
+Furthermore, Spfy takes a modular approach to tool integration and distrib,utes analyses over task queues, thereby allowing modules to run in parallel and addresses code failures.
+These changes enable us to provide real-time genome analysis and database storage of all results for upwards of 5000 genomes.
 All services are publicly available at \url{https://lfz.corefacility.ca/superphy/spfy/}.

### 2017-07-24 paper-webserver.tex
FIX: citations

@@
 As food-sourcing is globalized, consumers are at a greater risk for bacterial infections in an environment that is more complicated to track than ever before \cite{kozak2013foodborne}.
-\textit{Salmonella enterica} and \textit{Escherichia coli} are amongst the two most common bacetrial infections \cite{kozak2013foodborne} originating from produce.
-Foodborne illnesses translate into actual costs: for example, in Canada alone an estimated 4000 hospitalizations and 105 deaths per year are thought to be the result of various pathogens \cite{thomas2015estimates}.
+\textit{Salmonella enterica} and \textit{Escherichia coli} are amongst the two most common bacetrial infections originating from produce and is a central concern amongst public health efforts \cite{kozak2013foodborne}.
+Foodborne illnesses translate into actual costs: in Canada alone an estimated 4000 hospitalizations and 105 deaths per year are thought to be the result of various pathogens \cite{thomas2015estimates}.
 Futhermore, the associated economic costs have been known for decades \cite{todd1989costs}.
@@
 In order to monitor outbreaks, platforms must rapidly subtype collected strains and communicate such information to other agencies \cite{ronholm2016navigating}. 
-Traditional reference laboratory techniques handled subtyping through methods such as serotyping, detecting the presence of specific cell surface antigens as tested by antibody reaction, pulsed-field gel electrophoresis (PFGE), to identify known virulence factors (VFs) or antimicrobial resistance genes (AMR genes), or Sanger sequencing, to determine multi-locus sequence types (MLSTs) \cite{ronholm2016navigating}.
+Traditional reference laboratory techniques handled subtyping through methods such as serotyping, detecting the presence of specific cell surface antigens as tested by antibody reaction, pulsed-field gel electrophoresis (PFGE), to identify band patterns, known virulence factors (VFs), or antimicrobial resistance genes (AMR genes), or Sanger sequencing, to determine multi-locus sequence types (MLSTs) \cite{ronholm2016navigating}.
 Using PFGE, outbreaks can be tracked by comparing band patterns of bacterial isolates and is routinely performed in public health agencies \cite{kozak2013foodborne}. 
 The original computational methods to leverage WGS focused on creating \textit{in-silico} approaches to return subtyping results matching those of wet-lab techniques.
-Software such as VirulenceFinder \cite{joensen2014real} compares a pre-collected set of reference sequences encompassing known virulence factor gene sequences against a WGS result.
+Software such as VirulenceFinder \cite{joensen2014real} compares a pre-collected set of reference sequences encompassing known VF gene sequences against a WGS result.
 Similar approaches have been applied to predict AMR genes \cite{mcarthur2013comprehensive}, O-type, and H-type \cite{whiteside2016superphy}.
@@
 % potential of WGS
-WGS has the potential to expand on traditional methods, due to the inclusion of the entire genome in WGS-based analyses; in effect, this lends WGS-based methods better resolution and context in evaluating the spread of outbreaks \cite{ronholm2016navigating}.
-As an example, instead of examining selected single-nucleotide polymorphisms (SNPs) in a subset of housekeeping genes, as is the case for laboratory-based MLST, newer methods such as core genome MLST (cgMLST) attempt to cover every common gene.
-Secondary analyses, the study of relations and historical relevance of genomes, are now possible on large datasets provided by WGS \cite{ronholm2016navigating}.
+Due to the inclusion of the entire genome in WGS-based analyses, WGS has the potential to expand on traditional methods; in effect, this lends WGS-based methods better resolution and context in evaluating the spread of outbreaks \cite{ronholm2016navigating}.
+In the case for laboratory-based MLST, instead of examining selected single-nucleotide polymorphisms (SNPs) in a subset of housekeeping genes, newer WGS-based methods such as core genome MLST (cgMLST) attempt to cover every common gene.
+Secondary analyses, the study of relations and historical relevance of genomes, are now possible as a result of the large datasets provided by WGS \cite{ronholm2016navigating}.
 Advances in WGS technologies have also enabled the monitoring and sharing of outbreak genomes in real-time \cite{ronholm2016navigating}.
@@
 % Drawbacks - genetics
-However, these WGS-base methods still examine the underlying genetics, not the expressed phenotype.
-The presence of a particular gene sequence does not necessitate expression.
-Correspondingly, a greater number of target sequences are identified which complicates risk assessments \cite{ronholm2016navigating} and laboratory based phenotypic tests are still relatively inexpensive \cite(CITE ROUTINE USE - KOSER).
+However, these WGS-base methods still examine the underlying genetics, not the expressed phenotype; the presence of a particular gene sequence does not necessitate expression.
+Correspondingly, a greater number of target sequences are identified using WGS approaches which complicates risk assessments \cite{ronholm2016navigating}, and laboratory based phenotypic tests remain relatively inexpensive \cite{koser2012routine}.
 The effectiveness of WGS-based methods is also reliant on the accuracy and breadth of reference datasets.
@@
 % drawbacks - big data
-Traditionaly, bioinformatics software was developed \textit{ad hoc}, without employing software engineering principles \cite(TRENDS - BREVEREN).
-Such approaches were often script-based, with custom data formats, and only suitable for small collections of data \cite(TRENDS - BREVEREN).
-As WGS-technology expanded the availability of data, systems engineering principles became important as code failure could quickly disrupt entire bioinformatics pipelines \cite(BIOLOGICAL DATA SCIENCES - SCHATZ).
-The field also lags behind other big data industries: bioinformatics software typically struggles to perform computations in real-time, stores data in inflexible databases, doesn't support live visualizations \cite(TRENDS - BREVEREN), and lacks a common application programming interface (API) linking resources \cite(REVIEW - SWAMINATHAN).
+Traditionaly, bioinformatics software was developed \textit{ad hoc}, without employing software engineering principles \cite{de2015trends}.
+Such designs were often script-based, with custom data formats, and only suitable for small collections of data \cite{de2015trends}.
+As WGS-technology expanded the availability of data, systems engineering principles became important as code failure could quickly disrupt entire bioinformatics pipelines \cite{schatz2015biological}.
+The field also lags behind other big data industries: bioinformatics software typically struggles to perform computations in real-time, stores data in inflexible databases, doesn't support live visualizations \cite{de2015trends}, and lacks a common application programming interface (API) linking resources \cite{swaminathan2016review}.
 
@@
 % solution via modern big data approaches
-For genotypic, WGS-based approaches to be truly compliment laboratory-based phenotypic tests, analyses software must be faster or demonstrate a reasonable link between their genotype and phenotype \cite(CITE ROUTINE USE - KOSER).
-Antimicrobial resistance testing, which is a major component of public health monitoring efforts, typically takes 18-24 hours in a reference laboratory \cite(CITE ROUTINE USE - KOSER).
-Results should also be stored in a tiered structure, with rapid access available for retrieval and visualization while maintaining long-term data storage \cite(BIOLOGICAL DATA SCIENCES - SCHATZ).
-Such an approach would enable developers to address the current limitations \cite(TRENDS - BREVEREN) of bioinformatics developers.
+For genotypic, WGS-based approaches to truly compliment laboratory-based phenotypic tests, analyses software must be faster or demonstrate a reasonable link between their genotype and phenotype \cite{koser2012routine}.
+Antimicrobial resistance testing, which is a major component of public health monitoring efforts, typically takes 18-24 hours in a reference laboratory \cite{koser2012routine} and thus WGS solutions should perform faster.
+Results should also be stored in a tiered structure, with rapid access available for retrieval and visualization while maintaining long-term data storage \cite{schatz2015biological}.
+Such an approach would enable developers to address the current limitations \cite{de2015trends} of bioinformatics software.
 
@@
 % this study
-Here we present Spfy, a complete rewrite of Superphy which expands on supported features to also provide long-term storage of results in a graph database, one of the emerging \cite(TRENDS - BREVEREN) database types being used for biological data.
-We also containerize of services using Docker, a modern approach found in many web servers and proposed to mitigate the complexities of bioinformatic pipelines \cite{di2015impact}.
+Here we present Spfy, a complete rewrite of Superphy which expands on supported features to also provide long-term storage of results in a graph database - one of the emerging \cite{de2015trends} database types being used for biological data.
+We also containerize services using Docker, a modern approach found in many web servers and proposed to mitigate the complexities of bioinformatic pipelines \cite{di2015impact}.
 Docker integration ensures that nested dependency resolution, which typically must be manually installed \cite{doi:10.1093/bioinformatics/btu153,laing2010pan,inouye2014srst2}, are handled automatically.
-Furthermore, Spfy takes a modular approach to tool integration and distrib,utes analyses over task queues, thereby allowing modules to run in parallel and addresses code failures.
+Furthermore, Spfy takes a modular approach to tool integration and distributes analyses over task queues, thereby allowing modules to run in parallel and address code failures.
 These changes enable us to provide real-time genome analysis and database storage of all results for upwards of 5000 genomes.

### 2017-07-24 paper-webserver.tex
CHANGE: some edits

@@
 % costs of bacterial infections
-As food-sourcing is globalized, consumers are at a greater risk for bacterial infections in an environment that is more complicated to track than ever before \cite{kozak2013foodborne}.
+As food-sourcing is globalized, the consumer risk for bacterial infections is more complicated to track than ever before \cite{kozak2013foodborne}.
 \textit{Salmonella enterica} and \textit{Escherichia coli} are amongst the two most common bacetrial infections originating from produce and is a central concern amongst public health efforts \cite{kozak2013foodborne}.
-Foodborne illnesses translate into actual costs: in Canada alone an estimated 4000 hospitalizations and 105 deaths per year are thought to be the result of various pathogens \cite{thomas2015estimates}.
+Foodborne illnesses translate into actual costs: in Canada alone, an estimated 4000 hospitalizations and 105 deaths per year are thought to be the result of various pathogens \cite{thomas2015estimates}.
 Futhermore, the associated economic costs have been known for decades \cite{todd1989costs}.
@@
 Traditional reference laboratory techniques handled subtyping through methods such as serotyping, detecting the presence of specific cell surface antigens as tested by antibody reaction, pulsed-field gel electrophoresis (PFGE), to identify band patterns, known virulence factors (VFs), or antimicrobial resistance genes (AMR genes), or Sanger sequencing, to determine multi-locus sequence types (MLSTs) \cite{ronholm2016navigating}.
-Using PFGE, outbreaks can be tracked by comparing band patterns of bacterial isolates and is routinely performed in public health agencies \cite{kozak2013foodborne}. 
+Using PFGE, outbreaks can be tracked by comparing band patterns of bacterial isolates, as routinely performed in public health agencies \cite{kozak2013foodborne}. 
 The original computational methods to leverage WGS focused on creating \textit{in-silico} approaches to return subtyping results matching those of wet-lab techniques.
@@
 % potential of WGS
-Due to the inclusion of the entire genome in WGS-based analyses, WGS has the potential to expand on traditional methods; in effect, this lends WGS-based methods better resolution and context in evaluating the spread of outbreaks \cite{ronholm2016navigating}.
+WGS has the potential to expand on traditional methods: due to the inclusion of the entire genome, WGS-based methods offer better resolution and context in evaluating the spread of outbreaks \cite{ronholm2016navigating}.
 In the case for laboratory-based MLST, instead of examining selected single-nucleotide polymorphisms (SNPs) in a subset of housekeeping genes, newer WGS-based methods such as core genome MLST (cgMLST) attempt to cover every common gene.
@@
 % Drawbacks - genetics
-However, these WGS-base methods still examine the underlying genetics, not the expressed phenotype; the presence of a particular gene sequence does not necessitate expression.
+However, these WGS-based methods still examine the underlying genetics, not the expressed phenotype; the presence of a particular gene sequence does not necessitate expression.
 Correspondingly, a greater number of target sequences are identified using WGS approaches which complicates risk assessments \cite{ronholm2016navigating}, and laboratory based phenotypic tests remain relatively inexpensive \cite{koser2012routine}.
@@
 Efforts such as the Comprehensive Antibiotic Resistance Database (CARD) \cite{mcarthur2013comprehensive} attempt to alleviate data concerns by curating a set of known AMR genes.
-In addition, agencies such as the Food and Drug Administration (FDA), which hosts GenomeTrakr \cite{allard2016practical}, the National Center for Biotechnology Information (NCBI), which hosts GenBank \cite{doi:10.1093/nar/gks1195}, and the Center for Disease Control and Prevention (CDC), which hosts PulseNet \cite(PULSENET), provide services to share genome sequences and associated metadata either openly or amongst member institutions.
+In addition, agencies such as the Food and Drug Administration (FDA), which hosts GenomeTrakr \cite{allard2016practical}, the National Center for Biotechnology Information (NCBI), which hosts GenBank \cite{doi:10.1093/nar/gks1195}, and the Center for Disease Control and Prevention (CDC), which hosts PulseNet \cite{swaminathan2001pulsenet}, provide services to share genome sequences and associated metadata either openly or amongst member institutions.
 Recent efforts have been focused on providing modern tools for quality control of reference genomes \cite{parks2015checkm}, particularly amongst human-genome projects \cite{lee2017ngscheckmate}.
@@
 % solution via modern big data approaches
-For genotypic, WGS-based approaches to truly compliment laboratory-based phenotypic tests, analyses software must be faster or demonstrate a reasonable link between their genotype and phenotype \cite{koser2012routine}.
+For genotypic, WGS-based approaches to truly compliment laboratory-based phenotypic tests, analyses software must be faster or there must be a reasonable link between target genes and the expressed phenotype \cite{koser2012routine}.
 Antimicrobial resistance testing, which is a major component of public health monitoring efforts, typically takes 18-24 hours in a reference laboratory \cite{koser2012routine} and thus WGS solutions should perform faster.
-Results should also be stored in a tiered structure, with rapid access available for retrieval and visualization while maintaining long-term data storage \cite{schatz2015biological}.
-Such an approach would enable developers to address the current limitations \cite{de2015trends} of bioinformatics software.
+Storing results in a tiered structure, with rapid access available for retrieval and visualization while maintaining long-term data storage \cite{schatz2015biological}, would enable developers to address the current limitations \cite{de2015trends} of bioinformatics software.
 
@@
 Furthermore, Spfy takes a modular approach to tool integration and distributes analyses over task queues, thereby allowing modules to run in parallel and address code failures.
-These changes enable us to provide real-time genome analysis and database storage of all results for upwards of 5000 genomes.
+These changes enable us to provide real-time genome analysis and database storage of results for upwards of 5000 genomes.
 All services are publicly available at \url{https://lfz.corefacility.ca/superphy/spfy/}.

### 2017-07-27 paper-webserver.tex
DEL: unecessary words

@@
 % costs of bacterial infections
-As food-sourcing is globalized, the consumer risk for bacterial infections is more complicated to track than ever before \cite{kozak2013foodborne}.
-\textit{Salmonella enterica} and \textit{Escherichia coli} are amongst the two most common bacetrial infections originating from produce and is a central concern amongst public health efforts \cite{kozak2013foodborne}.
+Consumers' risks for bacterial infections is more difficult to track due to the global sourcing of food \cite{kozak2013foodborne}.
+Public health efforts monitor bacterial infections, of which \textit{Salmonella enterica} and \textit{Escherichia coli} are amongst the two most common \cite{kozak2013foodborne}.
 Foodborne illnesses translate into actual costs: in Canada alone, an estimated 4000 hospitalizations and 105 deaths per year are thought to be the result of various pathogens \cite{thomas2015estimates}.
 Futhermore, the associated economic costs have been known for decades \cite{todd1989costs}.
-
 % surveillance
-The surveillance of bacterial pathogens is a routine function of public-health agencies in an effort to lower the burden caused by human illness.
-In \textit{Salmonella enterica} \cite{bell2016recent}, \textit{Escherichia coli} \cite{fratamico2016advances}, and others pathogens \cite{ronholm2016navigating}, culture-dependent subtyping techniques are lately coupled with culture-independent methods based on whole-genome sequencing (WGS), to monitor known pathogens and detect emerging threats.
+Public-health agencies routinely look for emerging threats, attempting to lower the burden caused by human illness.
+In \textit{Salmonella enterica} \cite{bell2016recent}, \textit{Escherichia coli} \cite{fratamico2016advances}, and others pathogens \cite{ronholm2016navigating}, culture-dependent subtyping techniques are now coupled with culture-independent methods based on whole-genome sequencing (WGS).
 
 % traditional reference lab techniques
-Subtyping encompasses various methods to designate particular strains of bacterial isolates.
-In order to monitor outbreaks, platforms must rapidly subtype collected strains and communicate such information to other agencies \cite{ronholm2016navigating}. 
+Subtyping encompasses methods designating particular strains of bacterial isolates.
+To monitor outbreaks, platforms must rapidly subtype collected strains and be shareable with other agencies \cite{ronholm2016navigating}. 
 Traditional reference laboratory techniques handled subtyping through methods such as serotyping, detecting the presence of specific cell surface antigens as tested by antibody reaction, pulsed-field gel electrophoresis (PFGE), to identify band patterns, known virulence factors (VFs), or antimicrobial resistance genes (AMR genes), or Sanger sequencing, to determine multi-locus sequence types (MLSTs) \cite{ronholm2016navigating}.
 Using PFGE, outbreaks can be tracked by comparing band patterns of bacterial isolates, as routinely performed in public health agencies \cite{kozak2013foodborne}. 
-The original computational methods to leverage WGS focused on creating \textit{in-silico} approaches to return subtyping results matching those of wet-lab techniques.
-Software such as VirulenceFinder \cite{joensen2014real} compares a pre-collected set of reference sequences encompassing known VF gene sequences against a WGS result.
-Similar approaches have been applied to predict AMR genes \cite{mcarthur2013comprehensive}, O-type, and H-type \cite{whiteside2016superphy}.
+The original computational methods leveraging WGS focused on returning subtyping results matching those of wet-lab techniques.
+Software such as VirulenceFinder \cite{joensen2014real} compares a pre-collected set of reference VF gene sequences against a WGS result.
+Similar approaches are applied to predict AMR genes \cite{mcarthur2013comprehensive}, O-type, and H-type \cite{whiteside2016superphy}.
 
 % potential of WGS
-WGS has the potential to expand on traditional methods: due to the inclusion of the entire genome, WGS-based methods offer better resolution and context in evaluating the spread of outbreaks \cite{ronholm2016navigating}.
-In the case for laboratory-based MLST, instead of examining selected single-nucleotide polymorphisms (SNPs) in a subset of housekeeping genes, newer WGS-based methods such as core genome MLST (cgMLST) attempt to cover every common gene.
-Secondary analyses, the study of relations and historical relevance of genomes, are now possible as a result of the large datasets provided by WGS \cite{ronholm2016navigating}.
-Advances in WGS technologies have also enabled the monitoring and sharing of outbreak genomes in real-time \cite{ronholm2016navigating}.
+WGS expands on traditional methods: due to the inclusion of the entire genome, WGS-based methods offer better resolution and context for evaluating the spread of outbreaks \cite{ronholm2016navigating}.
+In the case of laboratory-based MLST, instead of examining selected single-nucleotide polymorphisms (SNPs) in a subset of housekeeping genes, newer WGS-based methods, such as core genome MLST (cgMLST), attempt to cover every common gene.
+Secondary analyses, the study of relations and historical relevance of genomes, are possible as a result of the large datasets provided by WGS \cite{ronholm2016navigating}.
+Advances in WGS technologies have also enabled the real-time monitoring and sharing of outbreak genomes \cite{ronholm2016navigating}.
 
@@
 However, these WGS-based methods still examine the underlying genetics, not the expressed phenotype; the presence of a particular gene sequence does not necessitate expression.
-Correspondingly, a greater number of target sequences are identified using WGS approaches which complicates risk assessments \cite{ronholm2016navigating}, and laboratory based phenotypic tests remain relatively inexpensive \cite{koser2012routine}.
-The effectiveness of WGS-based methods is also reliant on the accuracy and breadth of reference datasets.
+Correspondingly, a greater number of target sequences are identified using WGS approaches which complicates risk assessments \cite{ronholm2016navigating}, and laboratory based phenotypic tests remain inexpensive \cite{koser2012routine}.
+The effectiveness of WGS-based methods also rely on the accuracy and breadth of reference datasets.
 
 % drawbacks - big data
-Traditionaly, bioinformatics software was developed \textit{ad hoc}, without employing software engineering principles \cite{de2015trends}.
-Such designs were often script-based, with custom data formats, and only suitable for small collections of data \cite{de2015trends}.
-As WGS-technology expanded the availability of data, systems engineering principles became important as code failure could quickly disrupt entire bioinformatics pipelines \cite{schatz2015biological}.
-The field also lags behind other big data industries: bioinformatics software typically struggles to perform computations in real-time, stores data in inflexible databases, doesn't support live visualizations \cite{de2015trends}, and lacks a common application programming interface (API) linking resources \cite{swaminathan2016review}.
+Traditionaly, bioinformatics software was developed \textit{ad hoc} without employing software engineering principles \cite{de2015trends}.
+Tools were often script-based, with custom data formats, and only suitable for small collections of data \cite{de2015trends}.
+As WGS-technology expanded the availability of data, systems engineering principles became important since code failure could quickly disrupt entire bioinformatics pipelines \cite{schatz2015biological}.
+The field also lags behind other big data industries: bioinformatics software still struggles to perform computations in real-time, stores data in inflexible databases, doesn't support live visualizations \cite{de2015trends}, and lacks a common application programming interface (API) linking resources \cite{swaminathan2016review}.
 
@@
 In addition, agencies such as the Food and Drug Administration (FDA), which hosts GenomeTrakr \cite{allard2016practical}, the National Center for Biotechnology Information (NCBI), which hosts GenBank \cite{doi:10.1093/nar/gks1195}, and the Center for Disease Control and Prevention (CDC), which hosts PulseNet \cite{swaminathan2001pulsenet}, provide services to share genome sequences and associated metadata either openly or amongst member institutions.
-Recent efforts have been focused on providing modern tools for quality control of reference genomes \cite{parks2015checkm}, particularly amongst human-genome projects \cite{lee2017ngscheckmate}.
-Panseq \cite{laing2010pan}, Roary \cite{page2015roary}, and chewBBACA \url{https://github.com/mickaelsilva/chewBBACA} are available to compute core and accessory regions and identify SNPs across a large collection of reference genomes.
-Taken together, open-access initiatives and shared tooling aim to alleviate the drawbacks of WGS-based methods by pre-computing sufficient reference data.
+Recent efforts have focused on providing modern tools for quality control of reference genomes \cite{parks2015checkm}, particularly amongst human-genome projects \cite{lee2017ngscheckmate}.
+Panseq \cite{laing2010pan}, Roary \cite{page2015roary}, and chewBBACA \url{https://github.com/mickaelsilva/chewBBACA} are freely available to compute core and accessory regions and identify SNPs across a large collection of reference genomes.
+Taken together, open-access initiatives and shared tooling are lessening the drawbacks of WGS-based methods, by pre-computing sufficient reference data.
 
@@
 To make such \textit{in-silico} techniques user-friendly for front-line clinicians and laboratories, a variety of web services have been developed in Denmark \cite{joensen2014real}, Taiwan \cite{liu2016construction}, and in collaborations \cite{hasman2015detection}.
-While useful, arguably \cite{fricke2014bacterial}, the lack of bioinformatic standards and common infrastructure technologies have made it difficult to fully leverage the big-data results from WGS.
+While useful, arguably \cite{fricke2014bacterial}, the lack of bioinformatic standards and common infrastructure technologies make it difficult to fully leverage the big-data results from WGS.
 
@@
 Previously we developed Superphy \cite{whiteside2016superphy}, an online platform targeting \textit{E.
-coli}, which determines Shiga toxin subtypes, AMR genes using CARD, and provides statistical significance tests between subgroups in associated metadata and pan-genome regions.
+coli} which determines Shiga toxin subtypes, AMR genes using CARD, and provides statistical significance tests between subgroups in associated metadata and pan-genome regions.
 However, we struggled to provide real-time computing of results and to accomdate an ever-growing reference database.
@@
 % this study
-Here we present Spfy, a complete rewrite of Superphy which expands on supported features to also provide long-term storage of results in a graph database - one of the emerging \cite{de2015trends} database types being used for biological data.
+Here we present Spfy, a complete rewrite of Superphy expanding on supported features to also provide long-term storage of results in a graph database - one of the emerging \cite{de2015trends} database types being used for biological data.
 We also containerize services using Docker, a modern approach found in many web servers and proposed to mitigate the complexities of bioinformatic pipelines \cite{di2015impact}.

### 2017-07-28 paper-webserver.tex
CHANGE: some restructuring

@@
 % costs of bacterial infections
-Consumers' risks for bacterial infections is more difficult to track due to the global sourcing of food \cite{kozak2013foodborne}.
-Public health efforts monitor bacterial infections, of which \textit{Salmonella enterica} and \textit{Escherichia coli} are amongst the two most common \cite{kozak2013foodborne}.
-Foodborne illnesses translate into actual costs: in Canada alone, an estimated 4000 hospitalizations and 105 deaths per year are thought to be the result of various pathogens \cite{thomas2015estimates}.
+Foodborne ilnesses are increasingly difficult to track; food is globally sourced and consumers are purchasing more prepared ingrediants \cite{kozak2013foodborne}.
+The resulting consumer infections translate into actual costs: in Canada alone, an estimated 4000 hospitalizations and 105 deaths per year are thought to be the result of various pathogens \cite{thomas2015estimates}.
 Futhermore, the associated economic costs have been known for decades \cite{todd1989costs}.
 % surveillance
-Public-health agencies routinely look for emerging threats, attempting to lower the burden caused by human illness.
-In \textit{Salmonella enterica} \cite{bell2016recent}, \textit{Escherichia coli} \cite{fratamico2016advances}, and others pathogens \cite{ronholm2016navigating}, culture-dependent subtyping techniques are now coupled with culture-independent methods based on whole-genome sequencing (WGS).
+Public-health agencies routinely look for emerging threats, of which \textit{Salmonella enterica} and \textit{Escherichia coli} are amongst the two most common \cite{kozak2013foodborne}.
+In \textit{Salmonella enterica} \cite{bell2016recent}, \textit{Escherichia coli} \cite{fratamico2016advances}, and others pathogens \cite{ronholm2016navigating}, a variety of bioinformatic approaches are now based on whole-genome sequencing (WGS).
 
 % traditional reference lab techniques
-Subtyping encompasses methods designating particular strains of bacterial isolates.
-To monitor outbreaks, platforms must rapidly subtype collected strains and be shareable with other agencies \cite{ronholm2016navigating}. 
+To begin analyzing genomes, researchers designate particular strains of bacterial isolates by identifiers known as subtypes.
 Traditional reference laboratory techniques handled subtyping through methods such as serotyping, detecting the presence of specific cell surface antigens as tested by antibody reaction, pulsed-field gel electrophoresis (PFGE), to identify band patterns, known virulence factors (VFs), or antimicrobial resistance genes (AMR genes), or Sanger sequencing, to determine multi-locus sequence types (MLSTs) \cite{ronholm2016navigating}.
-Using PFGE, outbreaks can be tracked by comparing band patterns of bacterial isolates, as routinely performed in public health agencies \cite{kozak2013foodborne}. 
-The original computational methods leveraging WGS focused on returning subtyping results matching those of wet-lab techniques.
+In public health agencies, PFGE is used to track outbreaks by comparing band patterns of bacterial isolates \cite{kozak2013foodborne}. 
+The original computational methods leveraging WGS returned subtyping results matching those of wet-lab techniques.
 Software such as VirulenceFinder \cite{joensen2014real} compares a pre-collected set of reference VF gene sequences against a WGS result.
@@
 % this study
-Here we present Spfy, a complete rewrite of Superphy expanding on supported features to also provide long-term storage of results in a graph database - one of the emerging \cite{de2015trends} database types being used for biological data.
+Here we present Spfy, a complete rewrite of Superphy to put a graph databases at the core.
+Graph databases focus on describing the relationships between different data, and is one of the emerging \cite{de2015trends} database types being used for biological data.
 We also containerize services using Docker, a modern approach found in many web servers and proposed to mitigate the complexities of bioinformatic pipelines \cite{di2015impact}.
-Docker integration ensures that nested dependency resolution, which typically must be manually installed \cite{doi:10.1093/bioinformatics/btu153,laing2010pan,inouye2014srst2}, are handled automatically.
+Docker integration ensures that software dependencies, which typically must be manually installed \cite{doi:10.1093/bioinformatics/btu153,laing2010pan,inouye2014srst2}, are handled automatically.
 Furthermore, Spfy takes a modular approach to tool integration and distributes analyses over task queues, thereby allowing modules to run in parallel and address code failures.

### 2017-07-28 paper-webserver.tex
CHANGE: rewrite of the traditional wet-labs tech para

@@
 % surveillance
-Public-health agencies routinely look for emerging threats, of which \textit{Salmonella enterica} and \textit{Escherichia coli} are amongst the two most common \cite{kozak2013foodborne}.
+Public-health agencies routinely monitor emerging threats, of which \textit{Salmonella enterica} and \textit{Escherichia coli} are amongst the two most common \cite{kozak2013foodborne}.
 In \textit{Salmonella enterica} \cite{bell2016recent}, \textit{Escherichia coli} \cite{fratamico2016advances}, and others pathogens \cite{ronholm2016navigating}, a variety of bioinformatic approaches are now based on whole-genome sequencing (WGS).
@@
 % traditional reference lab techniques
-To begin analyzing genomes, researchers designate particular strains of bacterial isolates by identifiers known as subtypes.
-Traditional reference laboratory techniques handled subtyping through methods such as serotyping, detecting the presence of specific cell surface antigens as tested by antibody reaction, pulsed-field gel electrophoresis (PFGE), to identify band patterns, known virulence factors (VFs), or antimicrobial resistance genes (AMR genes), or Sanger sequencing, to determine multi-locus sequence types (MLSTs) \cite{ronholm2016navigating}.
-In public health agencies, PFGE is used to track outbreaks by comparing band patterns of bacterial isolates \cite{kozak2013foodborne}. 
-The original computational methods leveraging WGS returned subtyping results matching those of wet-lab techniques.
+To begin analyzing genomes, researchers designate particular strains of bacterial isolates using identifiers known as subtypes.
+Traditional reference laboratory techniques handled subtyping through methods such as serotyping, pulsed-field gel electrophoresis (PFGE), or Sanger sequencing, amongst others \cite{ronholm2016navigating}.
+PFGE is unique amongst the set; restriction enzymes are used to produce DNA fragments visible as specific banding patterns.
+Public health agencies use PFGE to track outbreaks by comparing band patterns of bacterial isolates \cite{kozak2013foodborne}.
+However, as WGS rarely produces complete circular genomes and, in PFGE, cutting by a restriction enzyme is dependent on protein structure, there is not a bioinformatic-analog to PFGE.
+Serotyping uses antibodies to detect the presence of specific cell surface antigens.
+Sanger sequencing is historically used to examine differences in housekeeping genes which are required to maintain cell function, an approach known as multi-locus sequence typing (MLST).
+Unlike PFGE, both serotyping and MLST have WGS-based computational methods which mimick results of the wet-lab techniques.
 Software such as VirulenceFinder \cite{joensen2014real} compares a pre-collected set of reference VF gene sequences against a WGS result.

### 2017-07-28 paper-webserver.tex
CHANGE: slight rewrite of potentials para

@@
 In the case of laboratory-based MLST, instead of examining selected single-nucleotide polymorphisms (SNPs) in a subset of housekeeping genes, newer WGS-based methods, such as core genome MLST (cgMLST), attempt to cover every common gene.
-Secondary analyses, the study of relations and historical relevance of genomes, are possible as a result of the large datasets provided by WGS \cite{ronholm2016navigating}.
-Advances in WGS technologies have also enabled the real-time monitoring and sharing of outbreak genomes \cite{ronholm2016navigating}.
+We can also detect the presence of every known virulence factor (VF) or antimicrobial resistance (AMR) gene in a WGS result.
+Furthermore, using WGS we can study the relations between genomes, changes in the VF or AMR genes present over time, and mutations in gene \cite{ronholm2016navigating}.
+Entire datasets and results can also be shared in real-time \cite{ronholm2016navigating}.
 
 % Drawbacks - genetics
-However, these WGS-based methods still examine the underlying genetics, not the expressed phenotype; the presence of a particular gene sequence does not necessitate expression.
+However, these WGS-based methods only examine the underlying genetics, not the expressed phenotype; the presence of a particular gene sequence does not necessitate expression.
 Correspondingly, a greater number of target sequences are identified using WGS approaches which complicates risk assessments \cite{ronholm2016navigating}, and laboratory based phenotypic tests remain inexpensive \cite{koser2012routine}.

### 2017-07-28 paper-webserver.tex
ADD: rewrite of the bigdata solution to instead focus on pangenomes

@@
 However, these WGS-based methods only examine the underlying genetics, not the expressed phenotype; the presence of a particular gene sequence does not necessitate expression.
-Correspondingly, a greater number of target sequences are identified using WGS approaches which complicates risk assessments \cite{ronholm2016navigating}, and laboratory based phenotypic tests remain inexpensive \cite{koser2012routine}.
+Correspondingly, more target sequences are identified using WGS; if we take a tradition approach to making threat assessments by looking for known VFs or AMR genes, the large number of results make it difficult to guage risk \cite{ronholm2016navigating}.
 The effectiveness of WGS-based methods also rely on the accuracy and breadth of reference datasets.
+Agencies must also weigh the cost of developing people and knowledge in bioinformatic approaches, and laboratory based phenotypic tests remain inexpensive \cite{koser2012routine}.
 
@@
 
-% Solution via pre-computed analyses for routine tasks
+% Solution via software engineering
 Efforts such as the Comprehensive Antibiotic Resistance Database (CARD) \cite{mcarthur2013comprehensive} attempt to alleviate data concerns by curating a set of known AMR genes.
@@
 Recent efforts have focused on providing modern tools for quality control of reference genomes \cite{parks2015checkm}, particularly amongst human-genome projects \cite{lee2017ngscheckmate}.
-Panseq \cite{laing2010pan}, Roary \cite{page2015roary}, and chewBBACA \url{https://github.com/mickaelsilva/chewBBACA} are freely available to compute core and accessory regions and identify SNPs across a large collection of reference genomes.
-Taken together, open-access initiatives and shared tooling are lessening the drawbacks of WGS-based methods, by pre-computing sufficient reference data.
 
-% solution via modern big data approaches
-For genotypic, WGS-based approaches to truly compliment laboratory-based phenotypic tests, analyses software must be faster or there must be a reasonable link between target genes and the expressed phenotype \cite{koser2012routine}.
-Antimicrobial resistance testing, which is a major component of public health monitoring efforts, typically takes 18-24 hours in a reference laboratory \cite{koser2012routine} and thus WGS solutions should perform faster.
+% rewrite of bigdata approach (BIGGER bigdata)
+While efforts such as CARD as important for replicating traditional results and for historical context, WGS can also be used to develop new methods using the larger datasets available.
+To study entire genomes, we need methods of creating reference data encompassing all genes, both core and accessory, in a given phlogeny.
+Panseq \cite{laing2010pan}, Roary \cite{page2015roary}, chewBBACA \url{https://github.com/mickaelsilva/chewBBACA}, and Graphtyper \cite{Eggertsson148403} are freely available to compute these pan-genomic regions.
+These software approaches must be scalable to significantly larger datasets than those found in wet-lab approaches.
+Also, for uptake by front-line clinicians and laboratories, WGS solutions should perform faster; for example, AMR genes should be determined faster than the typical 18-24 hours required in a reference laboratory \cite{koser2012routine}.
 Storing results in a tiered structure, with rapid access available for retrieval and visualization while maintaining long-term data storage \cite{schatz2015biological}, would enable developers to address the current limitations \cite{de2015trends} of bioinformatics software.

### 2017-07-28 paper-webserver.tex
ADD: stuff from Matts abstract

@@
 % why Superphy
-Subtyping and reference data generation often require command-line or programming experience.
-To make such \textit{in-silico} techniques user-friendly for front-line clinicians and laboratories, a variety of web services have been developed in Denmark \cite{joensen2014real}, Taiwan \cite{liu2016construction}, and in collaborations \cite{hasman2015detection}.
+To make bioinformatic approaches user-friendly for front-line clinicians and laboratories, a variety of web services have been developed in Denmark \cite{joensen2014real}, Taiwan \cite{liu2016construction}, and in collaborations \cite{hasman2015detection}.
 While useful, arguably \cite{fricke2014bacterial}, the lack of bioinformatic standards and common infrastructure technologies make it difficult to fully leverage the big-data results from WGS.
+Existing software also lack ways to integrate results from multiple genomes, and examine the connections between them.
 
@@
 Graph databases focus on describing the relationships between different data, and is one of the emerging \cite{de2015trends} database types being used for biological data.
+Spfy takes a modular approach to tool integration and distributes analyses over task queues, thereby allowing modules to run in parallel and address code failures.
+Serotyping, VF, AMR predictions are computed within minutes, and results are efficiently stored within the graph.
+Furthermore, large-scale analyses, such as epidemiological associations between specific genotypes, biomarkers, host, source, and other metadata, and statistically significance testing of genome markers for user-defined groups, are provided in real-time.
 We also containerize services using Docker, a modern approach found in many web servers and proposed to mitigate the complexities of bioinformatic pipelines \cite{di2015impact}.
 Docker integration ensures that software dependencies, which typically must be manually installed \cite{doi:10.1093/bioinformatics/btu153,laing2010pan,inouye2014srst2}, are handled automatically.
-Furthermore, Spfy takes a modular approach to tool integration and distributes analyses over task queues, thereby allowing modules to run in parallel and address code failures.
+Furthermore, 
 These changes enable us to provide real-time genome analysis and database storage of results for upwards of 5000 genomes.

### 2017-07-31 paper-webserver.tex
CHANGE: first para is weird, slight restructuring

@@
 % costs of bacterial infections
-Foodborne ilnesses are increasingly difficult to track; food is globally sourced and consumers are purchasing more prepared ingrediants \cite{kozak2013foodborne}.
-The resulting consumer infections translate into actual costs: in Canada alone, an estimated 4000 hospitalizations and 105 deaths per year are thought to be the result of various pathogens \cite{thomas2015estimates}.
+Foodborne ilnesses are increasingly difficult to track; ingrediants are sourced globally and contamination can occur at any point in the supply-chain.
+The resulting consumer infections translate into actual costs: in Canada, an estimated 4000 hospitalizations and 105 deaths per year are thought to be the result of various pathogens \cite{thomas2015estimates}.
+In the United States, this figure is 55961 hospitalizations and 1351 deaths per year \cite{scallan2011foodborne}.
 Futhermore, the associated economic costs have been known for decades \cite{todd1989costs}.

### 2017-07-31 paper-webserver.tex
ADD: more revisions

@@
 % costs of bacterial infections
-Foodborne ilnesses are increasingly difficult to track; ingrediants are sourced globally and contamination can occur at any point in the supply-chain.
+Foodborne ilnesses are increasingly difficult to track; ingrediants are globally sourced and contamination can occur at any point in the supply-chain.
 The resulting consumer infections translate into actual costs: in Canada, an estimated 4000 hospitalizations and 105 deaths per year are thought to be the result of various pathogens \cite{thomas2015estimates}.
@@
 Public-health agencies routinely monitor emerging threats, of which \textit{Salmonella enterica} and \textit{Escherichia coli} are amongst the two most common \cite{kozak2013foodborne}.
-In \textit{Salmonella enterica} \cite{bell2016recent}, \textit{Escherichia coli} \cite{fratamico2016advances}, and others pathogens \cite{ronholm2016navigating}, a variety of bioinformatic approaches are now based on whole-genome sequencing (WGS).
+In \textit{Salmonella enterica} \cite{bell2016recent}, \textit{Escherichia coli} \cite{fratamico2016advances}, and others pathogens \cite{ronholm2016navigating}, a variety of bioinformatic approaches to surveillance are now based on whole-genome sequencing (WGS).
 
@@
 PFGE is unique amongst the set; restriction enzymes are used to produce DNA fragments visible as specific banding patterns.
-Public health agencies use PFGE to track outbreaks by comparing band patterns of bacterial isolates \cite{kozak2013foodborne}.
-However, as WGS rarely produces complete circular genomes and, in PFGE, cutting by a restriction enzyme is dependent on protein structure, there is not a bioinformatic-analog to PFGE.
-Serotyping uses antibodies to detect the presence of specific cell surface antigens.
-Sanger sequencing is historically used to examine differences in housekeeping genes which are required to maintain cell function, an approach known as multi-locus sequence typing (MLST).
+Public health agencies use PFGE to track outbreaks by comparing band patterns of isolates \cite{kozak2013foodborne}.
+However, as WGS rarely produces complete circular genomes and, in PFGE, cutting by a restriction enzyme is dependent on protein structure, there is no bioinformatic-analog to PFGE.
 Unlike PFGE, both serotyping and MLST have WGS-based computational methods which mimick results of the wet-lab techniques.
-Software such as VirulenceFinder \cite{joensen2014real} compares a pre-collected set of reference VF gene sequences against a WGS result.
-Similar approaches are applied to predict AMR genes \cite{mcarthur2013comprehensive}, O-type, and H-type \cite{whiteside2016superphy}.
+Serotyping uses antibodies to detect the presence of specific cell surface antigens and multi-locus sequence typing (MLST), using Sanger sequencing, examines differences in housekeeping genes which are required to maintain cell function.
+By comparing a pre-collected set of reference VF gene sequences against a WGS result, we can predict wet-lab serotyping results \cite{whiteside2016superphy}, Virulence Factor (VF) results \cite{joensen2014real}, and antimicrobial resistance (AMR) results \cite{mcarthur2013comprehensive}.
 
 % potential of WGS
-WGS expands on traditional methods: due to the inclusion of the entire genome, WGS-based methods offer better resolution and context for evaluating the spread of outbreaks \cite{ronholm2016navigating}.
+WGS also expands on traditional methods: due to the inclusion of the entire genome, WGS-based methods offer better resolution and context for evaluating the spread of outbreaks \cite{ronholm2016navigating}.
 In the case of laboratory-based MLST, instead of examining selected single-nucleotide polymorphisms (SNPs) in a subset of housekeeping genes, newer WGS-based methods, such as core genome MLST (cgMLST), attempt to cover every common gene.
-We can also detect the presence of every known virulence factor (VF) or antimicrobial resistance (AMR) gene in a WGS result.
-Furthermore, using WGS we can study the relations between genomes, changes in the VF or AMR genes present over time, and mutations in gene \cite{ronholm2016navigating}.
+Instead of detecting only a subset of VFs and AMR genes, we can detect every known sequence in a WGS result.
+Furthermore, using WGS we can study the relations between genomes, changes in the VF or AMR genes present over time, and mutations in genes \cite{ronholm2016navigating}.
 Entire datasets and results can also be shared in real-time \cite{ronholm2016navigating}.
@@
 Correspondingly, more target sequences are identified using WGS; if we take a tradition approach to making threat assessments by looking for known VFs or AMR genes, the large number of results make it difficult to guage risk \cite{ronholm2016navigating}.
-The effectiveness of WGS-based methods also rely on the accuracy and breadth of reference datasets.
-Agencies must also weigh the cost of developing people and knowledge in bioinformatic approaches, and laboratory based phenotypic tests remain inexpensive \cite{koser2012routine}.
+The effectiveness of WGS-based methods also rely on the accuracy and breadth of reference datasets and agencies must weigh the cost of developing expertise in bioinformatic approaches, while laboratory based phenotypic tests remain inexpensive \cite{koser2012routine}.
 
@@
 Tools were often script-based, with custom data formats, and only suitable for small collections of data \cite{de2015trends}.
-As WGS-technology expanded the availability of data, systems engineering principles became important since code failure could quickly disrupt entire bioinformatics pipelines \cite{schatz2015biological}.
-The field also lags behind other big data industries: bioinformatics software still struggles to perform computations in real-time, stores data in inflexible databases, doesn't support live visualizations \cite{de2015trends}, and lacks a common application programming interface (API) linking resources \cite{swaminathan2016review}.
+Since WGS-technology expanded the availability of data, systems engineering principles are important since code failure must be handled throughout the entire bioinformatics pipeline \cite{schatz2015biological}.
+The field also lags behind other big-data industries: bioinformatics software still struggles to perform computations in real-time, stores data in inflexible databases, doesn't support live visualizations \cite{de2015trends}, and lacks a common application programming interface (API) linking resources \cite{swaminathan2016review}.
 
@@
 In addition, agencies such as the Food and Drug Administration (FDA), which hosts GenomeTrakr \cite{allard2016practical}, the National Center for Biotechnology Information (NCBI), which hosts GenBank \cite{doi:10.1093/nar/gks1195}, and the Center for Disease Control and Prevention (CDC), which hosts PulseNet \cite{swaminathan2001pulsenet}, provide services to share genome sequences and associated metadata either openly or amongst member institutions.
-Recent efforts have focused on providing modern tools for quality control of reference genomes \cite{parks2015checkm}, particularly amongst human-genome projects \cite{lee2017ngscheckmate}.
+Recent efforts have focused on improving the quality of shared reference genomes \cite{parks2015checkm}, particularly amongst human-genome projects \cite{lee2017ngscheckmate}.
 
@@
 While efforts such as CARD as important for replicating traditional results and for historical context, WGS can also be used to develop new methods using the larger datasets available.
-To study entire genomes, we need methods of creating reference data encompassing all genes, both core and accessory, in a given phlogeny.
+To study entire genomes, we need methods of creating reference data encompassing all genes in a given phlogeny.
 Panseq \cite{laing2010pan}, Roary \cite{page2015roary}, chewBBACA \url{https://github.com/mickaelsilva/chewBBACA}, and Graphtyper \cite{Eggertsson148403} are freely available to compute these pan-genomic regions.
-These software approaches must be scalable to significantly larger datasets than those found in wet-lab approaches.
-Also, for uptake by front-line clinicians and laboratories, WGS solutions should perform faster; for example, AMR genes should be determined faster than the typical 18-24 hours required in a reference laboratory \cite{koser2012routine}.
-Storing results in a tiered structure, with rapid access available for retrieval and visualization while maintaining long-term data storage \cite{schatz2015biological}, would enable developers to address the current limitations \cite{de2015trends} of bioinformatics software.
+These software approaches must be scalable to larger datasets than those found in wet-lab approaches.
+Also, for uptake by front-line clinicians and laboratories, WGS solutions should perform faster; for example, AMR gene prediction should take less than the typical 18-24 hours required in a reference laboratory \cite{koser2012routine}.
+Storing results in a tiered structure, with rapid access available for retrieval and visualization while maintaining long-term data storage \cite{schatz2015biological}, would enable developers to address the current interactivity limitations \cite{de2015trends} of bioinformatics software.
 
@@
 % this study
-Here we present Spfy, a complete rewrite of Superphy to put a graph databases at the core.
+Here we present Spfy, a complete rewrite of Superphy.
+Spfy is backed by a graph database which our analyses modules rely on.
 Graph databases focus on describing the relationships between different data, and is one of the emerging \cite{de2015trends} database types being used for biological data.
@@
 Docker integration ensures that software dependencies, which typically must be manually installed \cite{doi:10.1093/bioinformatics/btu153,laing2010pan,inouye2014srst2}, are handled automatically.
-Furthermore, 
 These changes enable us to provide real-time genome analysis and database storage of results for upwards of 5000 genomes.

### 2017-07-31 paper-webserver.tex
ADD: minor edits

@@
 The resulting consumer infections translate into actual costs: in Canada, an estimated 4000 hospitalizations and 105 deaths per year are thought to be the result of various pathogens \cite{thomas2015estimates}.
-In the United States, this figure is 55961 hospitalizations and 1351 deaths per year \cite{scallan2011foodborne}.
+In the United States, 55961 hospitalizations and 1351 deaths per year \cite{scallan2011foodborne} are due to foodborne illnesses.
 Futhermore, the associated economic costs have been known for decades \cite{todd1989costs}.
@@
 WGS also expands on traditional methods: due to the inclusion of the entire genome, WGS-based methods offer better resolution and context for evaluating the spread of outbreaks \cite{ronholm2016navigating}.
-In the case of laboratory-based MLST, instead of examining selected single-nucleotide polymorphisms (SNPs) in a subset of housekeeping genes, newer WGS-based methods, such as core genome MLST (cgMLST), attempt to cover every common gene.
-Instead of detecting only a subset of VFs and AMR genes, we can detect every known sequence in a WGS result.
-Furthermore, using WGS we can study the relations between genomes, changes in the VF or AMR genes present over time, and mutations in genes \cite{ronholm2016navigating}.
+In laboratory-based MLST, instead of examining selected single-nucleotide polymorphisms (SNPs) in a subset of housekeeping genes, newer WGS-based methods, such as core genome MLST (cgMLST), attempt to cover every common gene.
+WGS approaches also expands on VF and AMR gene detection to cover every known sequence.
+Furthermore, using WGS we can study the relations between genomes, changes in which VF or AMR genes are present over time, and mutations in genes \cite{ronholm2016navigating}.
 Entire datasets and results can also be shared in real-time \cite{ronholm2016navigating}.
@@
 Tools were often script-based, with custom data formats, and only suitable for small collections of data \cite{de2015trends}.
-Since WGS-technology expanded the availability of data, systems engineering principles are important since code failure must be handled throughout the entire bioinformatics pipeline \cite{schatz2015biological}.
-The field also lags behind other big-data industries: bioinformatics software still struggles to perform computations in real-time, stores data in inflexible databases, doesn't support live visualizations \cite{de2015trends}, and lacks a common application programming interface (API) linking resources \cite{swaminathan2016review}.
+Since WGS-technology expanded the availability of data, systems engineering principles are important to handle code failures across the entire bioinformatics pipeline \cite{schatz2015biological}.
+The field also lags behind other big-data industries: bioinformatics software still struggles to perform computations in real-time, stores data in inflexible databases \cite{de2015trends}, and lacks a common application programming interface (API) linking resources \cite{swaminathan2016review}.
 
@@
 In addition, agencies such as the Food and Drug Administration (FDA), which hosts GenomeTrakr \cite{allard2016practical}, the National Center for Biotechnology Information (NCBI), which hosts GenBank \cite{doi:10.1093/nar/gks1195}, and the Center for Disease Control and Prevention (CDC), which hosts PulseNet \cite{swaminathan2001pulsenet}, provide services to share genome sequences and associated metadata either openly or amongst member institutions.
-Recent efforts have focused on improving the quality of shared reference genomes \cite{parks2015checkm}, particularly amongst human-genome projects \cite{lee2017ngscheckmate}.
+Recent efforts have focused on checking the quality of shared reference genomes \cite{parks2015checkm}, particularly amongst human-genome projects \cite{lee2017ngscheckmate}.
 
@@
 While efforts such as CARD as important for replicating traditional results and for historical context, WGS can also be used to develop new methods using the larger datasets available.
-To study entire genomes, we need methods of creating reference data encompassing all genes in a given phlogeny.
+To study entire genomes, we need methods of creating reference sets encompassing all genes in a given phlogeny.
 Panseq \cite{laing2010pan}, Roary \cite{page2015roary}, chewBBACA \url{https://github.com/mickaelsilva/chewBBACA}, and Graphtyper \cite{Eggertsson148403} are freely available to compute these pan-genomic regions.
-These software approaches must be scalable to larger datasets than those found in wet-lab approaches.
+Such software approaches must scale to larger datasets than those found in wet-lab approaches.
 Also, for uptake by front-line clinicians and laboratories, WGS solutions should perform faster; for example, AMR gene prediction should take less than the typical 18-24 hours required in a reference laboratory \cite{koser2012routine}.
-Storing results in a tiered structure, with rapid access available for retrieval and visualization while maintaining long-term data storage \cite{schatz2015biological}, would enable developers to address the current interactivity limitations \cite{de2015trends} of bioinformatics software.
+Storing results in a tiered structure, with rapid access available for retrieval while maintaining long-term data storage \cite{schatz2015biological}, would enable developers to address the current interactivity limitations \cite{de2015trends} of bioinformatics software.
 
@@
 However, we struggled to provide real-time computing of results and to accomdate an ever-growing reference database.
-Furthermore, due to the nature of the various dependencies involved, the platform was difficult for other researchers to independently replicate.
+Furthermore, due to the various dependencies involved, the platform was difficult for other researchers to independently replicate.
 
@@
 Here we present Spfy, a complete rewrite of Superphy.
-Spfy is backed by a graph database which our analyses modules rely on.
+Spfy is backed by a graph database which our analysis modules rely on.
 Graph databases focus on describing the relationships between different data, and is one of the emerging \cite{de2015trends} database types being used for biological data.

### 2017-07-31 paper-webserver.tex
ADD: wording fixes

@@
 Public health agencies use PFGE to track outbreaks by comparing band patterns of isolates \cite{kozak2013foodborne}.
-However, as WGS rarely produces complete circular genomes and, in PFGE, cutting by a restriction enzyme is dependent on protein structure, there is no bioinformatic-analog to PFGE.
+However, as WGS rarely produces complete circular genomes and, in PFGE, cutting by a restriction enzyme is dependent on protein structure, bioinformatic-analogs to PFGE are difficult to develop.
 Unlike PFGE, both serotyping and MLST have WGS-based computational methods which mimick results of the wet-lab techniques.
 Serotyping uses antibodies to detect the presence of specific cell surface antigens and multi-locus sequence typing (MLST), using Sanger sequencing, examines differences in housekeeping genes which are required to maintain cell function.
-By comparing a pre-collected set of reference VF gene sequences against a WGS result, we can predict wet-lab serotyping results \cite{whiteside2016superphy}, Virulence Factor (VF) results \cite{joensen2014real}, and antimicrobial resistance (AMR) results \cite{mcarthur2013comprehensive}.
+By comparing a pre-collected set of reference gene sequences against a WGS result, we can predict wet-lab serotyping results \cite{whiteside2016superphy}, Virulence Factor (VF) results \cite{joensen2014real}, and antimicrobial resistance (AMR) results \cite{mcarthur2013comprehensive}.
+A similar approach can be taken for MLST \cite{larsen2012multilocus}.
 
@@
 % why Superphy
-To make bioinformatic approaches user-friendly for front-line clinicians and laboratories, a variety of web services have been developed in Denmark \cite{joensen2014real}, Taiwan \cite{liu2016construction}, and in collaborations \cite{hasman2015detection}.
+To make bioinformatic approaches user-friendly for front-line clinicians and laboratories, a variety of web services have been developed in Denmark \cite{joensen2014real,larsen2012multilocus}, Taiwan \cite{liu2016construction}, and in collaborations \cite{hasman2015detection}.
 While useful, arguably \cite{fricke2014bacterial}, the lack of bioinformatic standards and common infrastructure technologies make it difficult to fully leverage the big-data results from WGS.

### 2017-07-31 paper-webserver.tex
FIX: verbing

@@
 % costs of bacterial infections
-Foodborne ilnesses are increasingly difficult to track; ingrediants are globally sourced and contamination can occur at any point in the supply-chain.
-The resulting consumer infections translate into actual costs: in Canada, an estimated 4000 hospitalizations and 105 deaths per year are thought to be the result of various pathogens \cite{thomas2015estimates}.
-In the United States, 55961 hospitalizations and 1351 deaths per year \cite{scallan2011foodborne} are due to foodborne illnesses.
-Futhermore, the associated economic costs have been known for decades \cite{todd1989costs}.
+Tracking foodborne ilnesses is increasingly difficult; ingrediants are globally sourced and contamination can occur at any point in the supply-chain.
+The resulting consumer infections translate into actual costs: foodborne illnesses cause an estimated 4000 hospitalizations and 105 deaths per year in Canada \cite{thomas2015estimates}, and 55961 hospitalizations and 1351 deaths per year in the United States \cite{scallan2011foodborne}.
+Futhermore, many studies \cite{todd1989costs,scharff2010health} document the associated financial and productivity costs.
 % surveillance
@@
 Public health agencies use PFGE to track outbreaks by comparing band patterns of isolates \cite{kozak2013foodborne}.
-However, as WGS rarely produces complete circular genomes and, in PFGE, cutting by a restriction enzyme is dependent on protein structure, bioinformatic-analogs to PFGE are difficult to develop.
+However, as WGS rarely produces complete circular genomes and, in PFGE, cutting by a restriction enzyme is dependent on protein structure, developing bioinformatic-analogs to PFGE is difficult.
 Unlike PFGE, both serotyping and MLST have WGS-based computational methods which mimick results of the wet-lab techniques.
@@
 By comparing a pre-collected set of reference gene sequences against a WGS result, we can predict wet-lab serotyping results \cite{whiteside2016superphy}, Virulence Factor (VF) results \cite{joensen2014real}, and antimicrobial resistance (AMR) results \cite{mcarthur2013comprehensive}.
-A similar approach can be taken for MLST \cite{larsen2012multilocus}.
+MLST software uses a similar approach \cite{larsen2012multilocus}.
 
 % potential of WGS
-WGS also expands on traditional methods: due to the inclusion of the entire genome, WGS-based methods offer better resolution and context for evaluating the spread of outbreaks \cite{ronholm2016navigating}.
+WGS also expands on traditional methods: due to the inclusion of the entire genome, WGS-based methods evaluate the spread of outbreaks with better resolution and context \cite{ronholm2016navigating}.
 In laboratory-based MLST, instead of examining selected single-nucleotide polymorphisms (SNPs) in a subset of housekeeping genes, newer WGS-based methods, such as core genome MLST (cgMLST), attempt to cover every common gene.
@@
 However, these WGS-based methods only examine the underlying genetics, not the expressed phenotype; the presence of a particular gene sequence does not necessitate expression.
-Correspondingly, more target sequences are identified using WGS; if we take a tradition approach to making threat assessments by looking for known VFs or AMR genes, the large number of results make it difficult to guage risk \cite{ronholm2016navigating}.
+Correspondingly, more target sequences are identified using WGS; if we take a tradition approach to making threat assessments by looking for known VFs or AMR genes, guaging risk is difficult due to the large number of results \cite{ronholm2016navigating}.
 The effectiveness of WGS-based methods also rely on the accuracy and breadth of reference datasets and agencies must weigh the cost of developing expertise in bioinformatic approaches, while laboratory based phenotypic tests remain inexpensive \cite{koser2012routine}.
@@
 Tools were often script-based, with custom data formats, and only suitable for small collections of data \cite{de2015trends}.
-Since WGS-technology expanded the availability of data, systems engineering principles are important to handle code failures across the entire bioinformatics pipeline \cite{schatz2015biological}.
-The field also lags behind other big-data industries: bioinformatics software still struggles to perform computations in real-time, stores data in inflexible databases \cite{de2015trends}, and lacks a common application programming interface (API) linking resources \cite{swaminathan2016review}.
+Since WGS-technology expanded the availability of data, bioinformatic pipelines are larger and more complicated.
+We must apply systems engineering principles to handle code failures throughout \cite{schatz2015biological}.
+The bioinformatics field also lags behind other big-data industries: software still struggles to perform computations in real-time, stores data in inflexible databases \cite{de2015trends}, and lacks a common application programming interface (API) linking resources \cite{swaminathan2016review}.
 
@@
 To make bioinformatic approaches user-friendly for front-line clinicians and laboratories, a variety of web services have been developed in Denmark \cite{joensen2014real,larsen2012multilocus}, Taiwan \cite{liu2016construction}, and in collaborations \cite{hasman2015detection}.
-While useful, arguably \cite{fricke2014bacterial}, the lack of bioinformatic standards and common infrastructure technologies make it difficult to fully leverage the big-data results from WGS.
-Existing software also lack ways to integrate results from multiple genomes, and examine the connections between them.
+While useful, arguably \cite{fricke2014bacterial}, leveraging the big-data results from WGS is difficult due to the lack of bioinformatic standards and common infrastructure technologies.
+Integrating results from multiple genomes, and examining shared connections, will be important for new software.
 
@@
 However, we struggled to provide real-time computing of results and to accomdate an ever-growing reference database.
-Furthermore, due to the various dependencies involved, the platform was difficult for other researchers to independently replicate.
+Furthermore, due to the various dependencies involved, independently replicating the platform was difficult for other researchers.
 

### 2017-07-31 paper-webserver.tex
FIX: spelling

@@
 % costs of bacterial infections
-Tracking foodborne ilnesses is increasingly difficult; ingrediants are globally sourced and contamination can occur at any point in the supply-chain.
+Tracking foodborne illnesses is increasingly difficult; ingredients are globally sourced and contamination can occur at any point in the supply-chain.
 The resulting consumer infections translate into actual costs: foodborne illnesses cause an estimated 4000 hospitalizations and 105 deaths per year in Canada \cite{thomas2015estimates}, and 55961 hospitalizations and 1351 deaths per year in the United States \cite{scallan2011foodborne}.
-Futhermore, many studies \cite{todd1989costs,scharff2010health} document the associated financial and productivity costs.
+Furthermore, many studies \cite{todd1989costs,scharff2010health} document the associated financial and productivity costs.
 % surveillance
@@
 However, as WGS rarely produces complete circular genomes and, in PFGE, cutting by a restriction enzyme is dependent on protein structure, developing bioinformatic-analogs to PFGE is difficult.
-Unlike PFGE, both serotyping and MLST have WGS-based computational methods which mimick results of the wet-lab techniques.
+Unlike PFGE, both serotyping and MLST have WGS-based computational methods which mimic results of the wet-lab techniques.
 Serotyping uses antibodies to detect the presence of specific cell surface antigens and multi-locus sequence typing (MLST), using Sanger sequencing, examines differences in housekeeping genes which are required to maintain cell function.
@@
 However, these WGS-based methods only examine the underlying genetics, not the expressed phenotype; the presence of a particular gene sequence does not necessitate expression.
-Correspondingly, more target sequences are identified using WGS; if we take a tradition approach to making threat assessments by looking for known VFs or AMR genes, guaging risk is difficult due to the large number of results \cite{ronholm2016navigating}.
-The effectiveness of WGS-based methods also rely on the accuracy and breadth of reference datasets and agencies must weigh the cost of developing expertise in bioinformatic approaches, while laboratory based phenotypic tests remain inexpensive \cite{koser2012routine}.
+Correspondingly, more target sequences are identified using WGS; if we take a tradition approach to making threat assessments by looking for known VFs or AMR genes, gauging risk is difficult due to the large number of results \cite{ronholm2016navigating}.
+The effectiveness of WGS-based methods also rely on the accuracy and breadth of reference datasets and agencies must weigh the cost of developing expertise in bioinformatic approaches, while laboratory based phenotype tests remain inexpensive \cite{koser2012routine}.
 
 % drawbacks - big data
-Traditionaly, bioinformatics software was developed \textit{ad hoc} without employing software engineering principles \cite{de2015trends}.
+Traditionally, bioinformatics software was developed \textit{ad hoc} without employing software engineering principles \cite{de2015trends}.
 Tools were often script-based, with custom data formats, and only suitable for small collections of data \cite{de2015trends}.
@@
 While efforts such as CARD as important for replicating traditional results and for historical context, WGS can also be used to develop new methods using the larger datasets available.
-To study entire genomes, we need methods of creating reference sets encompassing all genes in a given phlogeny.
+To study entire genomes, we need methods of creating reference sets encompassing all genes in a given phylogeny.
 Panseq \cite{laing2010pan}, Roary \cite{page2015roary}, chewBBACA \url{https://github.com/mickaelsilva/chewBBACA}, and Graphtyper \cite{Eggertsson148403} are freely available to compute these pan-genomic regions.
@@
 coli} which determines Shiga toxin subtypes, AMR genes using CARD, and provides statistical significance tests between subgroups in associated metadata and pan-genome regions.
-However, we struggled to provide real-time computing of results and to accomdate an ever-growing reference database.
+However, we struggled to provide real-time computing of results and to accommodate an ever-growing reference database.
 Furthermore, due to the various dependencies involved, independently replicating the platform was difficult for other researchers.

### 2017-07-31 paper-webserver.tex
ADD: more active

@@
 Tracking foodborne illnesses is increasingly difficult; ingredients are globally sourced and contamination can occur at any point in the supply-chain.
-The resulting consumer infections translate into actual costs: foodborne illnesses cause an estimated 4000 hospitalizations and 105 deaths per year in Canada \cite{thomas2015estimates}, and 55961 hospitalizations and 1351 deaths per year in the United States \cite{scallan2011foodborne}.
+Consumers infected by bacteria incur actual costs: foodborne illnesses cause an estimated 4000 hospitalizations and 105 deaths per year in Canada \cite{thomas2015estimates}, and 55961 hospitalizations and 1351 deaths per year in the United States \cite{scallan2011foodborne}.
 Furthermore, many studies \cite{todd1989costs,scharff2010health} document the associated financial and productivity costs.
@@
 Public-health agencies routinely monitor emerging threats, of which \textit{Salmonella enterica} and \textit{Escherichia coli} are amongst the two most common \cite{kozak2013foodborne}.
-In \textit{Salmonella enterica} \cite{bell2016recent}, \textit{Escherichia coli} \cite{fratamico2016advances}, and others pathogens \cite{ronholm2016navigating}, a variety of bioinformatic approaches to surveillance are now based on whole-genome sequencing (WGS).
+In \textit{Salmonella enterica} \cite{bell2016recent}, \textit{Escherichia coli} \cite{fratamico2016advances}, and others pathogens \cite{ronholm2016navigating}, surveillance now employs a variety of bioinformatic approaches based on whole-genome sequencing (WGS).
 
@@
 However, as WGS rarely produces complete circular genomes and, in PFGE, cutting by a restriction enzyme is dependent on protein structure, developing bioinformatic-analogs to PFGE is difficult.
-Unlike PFGE, both serotyping and MLST have WGS-based computational methods which mimic results of the wet-lab techniques.
-Serotyping uses antibodies to detect the presence of specific cell surface antigens and multi-locus sequence typing (MLST), using Sanger sequencing, examines differences in housekeeping genes which are required to maintain cell function.
+Unlike PFGE, both serotyping and MLST have WGS-based computational methods mimicking wet-lab techniques.
+Serotyping uses antibodies to detect the presence of specific cell surface antigens and multi-locus sequence typing (MLST), using Sanger sequencing, examines differences in housekeeping genes required to maintain cell function.
 By comparing a pre-collected set of reference gene sequences against a WGS result, we can predict wet-lab serotyping results \cite{whiteside2016superphy}, Virulence Factor (VF) results \cite{joensen2014real}, and antimicrobial resistance (AMR) results \cite{mcarthur2013comprehensive}.
-MLST software uses a similar approach \cite{larsen2012multilocus}.
+MLST software uses a similar reference-based approach \cite{larsen2012multilocus}.
 
@@
 WGS also expands on traditional methods: due to the inclusion of the entire genome, WGS-based methods evaluate the spread of outbreaks with better resolution and context \cite{ronholm2016navigating}.
-In laboratory-based MLST, instead of examining selected single-nucleotide polymorphisms (SNPs) in a subset of housekeeping genes, newer WGS-based methods, such as core genome MLST (cgMLST), attempt to cover every common gene.
-WGS approaches also expands on VF and AMR gene detection to cover every known sequence.
+Instead of examining selected single-nucleotide polymorphisms (SNPs) in a subset of housekeeping genes as in laboratory-base MLST, newer WGS-based methods, such as core genome MLST (cgMLST), attempt to cover every common gene.
+WGS approaches also expand coverage of VF and AMR gene detection to every known sequence.
 Furthermore, using WGS we can study the relations between genomes, changes in which VF or AMR genes are present over time, and mutations in genes \cite{ronholm2016navigating}.
-Entire datasets and results can also be shared in real-time \cite{ronholm2016navigating}.
+Entire datasets and results can also be shared online \cite{ronholm2016navigating}.
 
 % Drawbacks - genetics
-However, these WGS-based methods only examine the underlying genetics, not the expressed phenotype; the presence of a particular gene sequence does not necessitate expression.
-Correspondingly, more target sequences are identified using WGS; if we take a tradition approach to making threat assessments by looking for known VFs or AMR genes, gauging risk is difficult due to the large number of results \cite{ronholm2016navigating}.
-The effectiveness of WGS-based methods also rely on the accuracy and breadth of reference datasets and agencies must weigh the cost of developing expertise in bioinformatic approaches, while laboratory based phenotype tests remain inexpensive \cite{koser2012routine}.
+However, these WGS-based methods only examine the underlying genetics, not the expressed phenotype.
+Correspondingly, more target sequences are identified using WGS; if we make threat assessments by looking for known VFs or AMR genes, we can not gauge the associated risks properly due to the large number of results \cite{ronholm2016navigating}.
+The effectiveness of WGS-based methods also rely on the accuracy and breadth of reference datasets and agencies must weigh the cost of developing expertise in bioinformatic approaches.
+Meanwhile, laboratory based phenotype tests remain inexpensive \cite{koser2012routine}.
 
@@
 Tools were often script-based, with custom data formats, and only suitable for small collections of data \cite{de2015trends}.
-Since WGS-technology expanded the availability of data, bioinformatic pipelines are larger and more complicated.
+Since WGS-technology expanded the availability of data, bioinformatic pipelines are now larger and more complicated.
 We must apply systems engineering principles to handle code failures throughout \cite{schatz2015biological}.
@@
 % Solution via software engineering
-Efforts such as the Comprehensive Antibiotic Resistance Database (CARD) \cite{mcarthur2013comprehensive} attempt to alleviate data concerns by curating a set of known AMR genes.
-In addition, agencies such as the Food and Drug Administration (FDA), which hosts GenomeTrakr \cite{allard2016practical}, the National Center for Biotechnology Information (NCBI), which hosts GenBank \cite{doi:10.1093/nar/gks1195}, and the Center for Disease Control and Prevention (CDC), which hosts PulseNet \cite{swaminathan2001pulsenet}, provide services to share genome sequences and associated metadata either openly or amongst member institutions.
-Recent efforts have focused on checking the quality of shared reference genomes \cite{parks2015checkm}, particularly amongst human-genome projects \cite{lee2017ngscheckmate}.
+Efforts such as the Comprehensive Antibiotic Resistance Database (CARD) \cite{mcarthur2013comprehensive} address data concerns by curating a set of known AMR genes.
+In addition, agencies such as the Food and Drug Administration (FDA), which hosts GenomeTrakr \cite{allard2016practical}, the National Center for Biotechnology Information (NCBI), which hosts GenBank \cite{doi:10.1093/nar/gks1195}, and the Center for Disease Control and Prevention (CDC), which hosts PulseNet \cite{swaminathan2001pulsenet}, enables sharing of genome sequences and associated metadata either openly or amongst member institutions.
+Recent efforts focused on checking the quality of shared reference genomes \cite{parks2015checkm}, particularly amongst human-genome projects \cite{lee2017ngscheckmate}.
 
 % rewrite of bigdata approach (BIGGER bigdata)
-While efforts such as CARD as important for replicating traditional results and for historical context, WGS can also be used to develop new methods using the larger datasets available.
+While efforts such as CARD are important for replicating traditional results and for historical context, WGS also lets us develop new methods targeting larger datasets.
 To study entire genomes, we need methods of creating reference sets encompassing all genes in a given phylogeny.
-Panseq \cite{laing2010pan}, Roary \cite{page2015roary}, chewBBACA \url{https://github.com/mickaelsilva/chewBBACA}, and Graphtyper \cite{Eggertsson148403} are freely available to compute these pan-genomic regions.
+Panseq \cite{laing2010pan}, Roary \cite{page2015roary}, chewBBACA \url{https://github.com/mickaelsilva/chewBBACA}, and Graphtyper \cite{Eggertsson148403} can compute these pan-genomic regions, and are freely available.
 Such software approaches must scale to larger datasets than those found in wet-lab approaches.
 Also, for uptake by front-line clinicians and laboratories, WGS solutions should perform faster; for example, AMR gene prediction should take less than the typical 18-24 hours required in a reference laboratory \cite{koser2012routine}.
-Storing results in a tiered structure, with rapid access available for retrieval while maintaining long-term data storage \cite{schatz2015biological}, would enable developers to address the current interactivity limitations \cite{de2015trends} of bioinformatics software.
+Storing results in a tiered structure, with rapid access available for retrieval while maintaining long-term data storage \cite{schatz2015biological}, would help address the current interactivity limitations \cite{de2015trends} of bioinformatics software.
 
@@
 Spfy is backed by a graph database which our analysis modules rely on.
-Graph databases focus on describing the relationships between different data, and is one of the emerging \cite{de2015trends} database types being used for biological data.
+Graph databases specifically describe the relationships between different data, and is one of the emerging \cite{de2015trends} database types used for biological data.
 Spfy takes a modular approach to tool integration and distributes analyses over task queues, thereby allowing modules to run in parallel and address code failures.
 Serotyping, VF, AMR predictions are computed within minutes, and results are efficiently stored within the graph.
-Furthermore, large-scale analyses, such as epidemiological associations between specific genotypes, biomarkers, host, source, and other metadata, and statistically significance testing of genome markers for user-defined groups, are provided in real-time.
+Furthermore, large-scale analyses, such as epidemiological associations between specific genotypes, biomarkers, host, source, and other metadata, and statistically significance testing of genome markers for user-defined groups, run in real-time.
 We also containerize services using Docker, a modern approach found in many web servers and proposed to mitigate the complexities of bioinformatic pipelines \cite{di2015impact}.
 Docker integration ensures that software dependencies, which typically must be manually installed \cite{doi:10.1093/bioinformatics/btu153,laing2010pan,inouye2014srst2}, are handled automatically.
-These changes enable us to provide real-time genome analysis and database storage of results for upwards of 5000 genomes.
+These changes enable real-time genome analysis and database storage of results for upwards of 5000 genomes.
 All services are publicly available at \url{https://lfz.corefacility.ca/superphy/spfy/}.

### 2017-07-31 paper-webserver.tex
STOP: edits before sending to chad

@@
 Consumers infected by bacteria incur actual costs: foodborne illnesses cause an estimated 4000 hospitalizations and 105 deaths per year in Canada \cite{thomas2015estimates}, and 55961 hospitalizations and 1351 deaths per year in the United States \cite{scallan2011foodborne}.
-Furthermore, many studies \cite{todd1989costs,scharff2010health} document the associated financial and productivity costs.
+Furthermore, many studies \cite{todd1989costs,scharff2010health} document the associated economic costs.
 % surveillance
 Public-health agencies routinely monitor emerging threats, of which \textit{Salmonella enterica} and \textit{Escherichia coli} are amongst the two most common \cite{kozak2013foodborne}.
-In \textit{Salmonella enterica} \cite{bell2016recent}, \textit{Escherichia coli} \cite{fratamico2016advances}, and others pathogens \cite{ronholm2016navigating}, surveillance now employs a variety of bioinformatic approaches based on whole-genome sequencing (WGS).
+In \textit{Salmonella enterica} \cite{bell2016recent}, \textit{Escherichia coli} \cite{fratamico2016advances}, and others pathogens \cite{ronholm2016navigating}, surveillance now involves a variety of bioinformatic approaches based on whole-genome sequencing (WGS).
 
@@
 However, as WGS rarely produces complete circular genomes and, in PFGE, cutting by a restriction enzyme is dependent on protein structure, developing bioinformatic-analogs to PFGE is difficult.
-Unlike PFGE, both serotyping and MLST have WGS-based computational methods mimicking wet-lab techniques.
-Serotyping uses antibodies to detect the presence of specific cell surface antigens and multi-locus sequence typing (MLST), using Sanger sequencing, examines differences in housekeeping genes required to maintain cell function.
-By comparing a pre-collected set of reference gene sequences against a WGS result, we can predict wet-lab serotyping results \cite{whiteside2016superphy}, Virulence Factor (VF) results \cite{joensen2014real}, and antimicrobial resistance (AMR) results \cite{mcarthur2013comprehensive}.
+Unlike PFGE, both serotyping and MLST have WGS-based computational methods mimicking wet-lab results.
+Serotyping, using antibodies, detects the presence of specific cell surface antigens and multi-locus sequence typing (MLST), using Sanger sequencing, examines differences in housekeeping genes - the genes required to maintain cell function.
+By comparing a pre-collected set of reference gene sequences against a WGS result, bioinformatics software predicts wet-lab serotyping results \cite{whiteside2016superphy}, Virulence Factor (VF) results \cite{joensen2014real}, and antimicrobial resistance (AMR) results \cite{mcarthur2013comprehensive}.
 MLST software uses a similar reference-based approach \cite{larsen2012multilocus}.
@@
 WGS also expands on traditional methods: due to the inclusion of the entire genome, WGS-based methods evaluate the spread of outbreaks with better resolution and context \cite{ronholm2016navigating}.
-Instead of examining selected single-nucleotide polymorphisms (SNPs) in a subset of housekeeping genes as in laboratory-base MLST, newer WGS-based methods, such as core genome MLST (cgMLST), attempt to cover every common gene.
+Instead of examining selected single-nucleotide polymorphisms (SNPs) in a subset of housekeeping genes as in laboratory-based MLST, newer WGS-based methods, such as core genome MLST (cgMLST), attempt to cover every common gene.
 WGS approaches also expand coverage of VF and AMR gene detection to every known sequence.
 Furthermore, using WGS we can study the relations between genomes, changes in which VF or AMR genes are present over time, and mutations in genes \cite{ronholm2016navigating}.
-Entire datasets and results can also be shared online \cite{ronholm2016navigating}.
+Researchers can also share entire datasets and results online \cite{ronholm2016navigating}.
 
@@
 However, these WGS-based methods only examine the underlying genetics, not the expressed phenotype.
-Correspondingly, more target sequences are identified using WGS; if we make threat assessments by looking for known VFs or AMR genes, we can not gauge the associated risks properly due to the large number of results \cite{ronholm2016navigating}.
+Correspondingly, more target sequences are identified using WGS; if health officials make threat assessments by looking for known VFs or AMR genes, they can only estimate the associated risks due to the large number of results \cite{ronholm2016navigating}.
 The effectiveness of WGS-based methods also rely on the accuracy and breadth of reference datasets and agencies must weigh the cost of developing expertise in bioinformatic approaches.
@@
 Tools were often script-based, with custom data formats, and only suitable for small collections of data \cite{de2015trends}.
-Since WGS-technology expanded the availability of data, bioinformatic pipelines are now larger and more complicated.
-We must apply systems engineering principles to handle code failures throughout \cite{schatz2015biological}.
-The bioinformatics field also lags behind other big-data industries: software still struggles to perform computations in real-time, stores data in inflexible databases \cite{de2015trends}, and lacks a common application programming interface (API) linking resources \cite{swaminathan2016review}.
+Since WGS-technology expanded the availability of data, bioinformatic pipelines are now larger and involve linked dependencies.
+Bioinformatics developers must apply systems engineering principles to handle code failures throughout \cite{schatz2015biological}.
+The bioinformatics field also lags behind other big-data industries \cite{de2015trends}: software still struggles to perform computations in real-time, stores data in inflexible databases, and lacks a common application programming interface (API) linking resources \cite{swaminathan2016review}.
 
@@
 % rewrite of bigdata approach (BIGGER bigdata)
-While efforts such as CARD are important for replicating traditional results and for historical context, WGS also lets us develop new methods targeting larger datasets.
-To study entire genomes, we need methods of creating reference sets encompassing all genes in a given phylogeny.
+While efforts such as CARD are important for replicating traditional results and for historical context, researchers can now develop new methods targeting larger datasets from WGS.
+To study entire genomes, the field needs methods of creating reference sets encompassing all genes in a given phylogeny.
 Panseq \cite{laing2010pan}, Roary \cite{page2015roary}, chewBBACA \url{https://github.com/mickaelsilva/chewBBACA}, and Graphtyper \cite{Eggertsson148403} can compute these pan-genomic regions, and are freely available.
@@
 % why Superphy
-To make bioinformatic approaches user-friendly for front-line clinicians and laboratories, a variety of web services have been developed in Denmark \cite{joensen2014real,larsen2012multilocus}, Taiwan \cite{liu2016construction}, and in collaborations \cite{hasman2015detection}.
+To make bioinformatic approaches user-friendly, a variety of web services have been developed in Denmark \cite{joensen2014real,larsen2012multilocus}, Taiwan \cite{liu2016construction}, and in collaborations \cite{hasman2015detection}.
 While useful, arguably \cite{fricke2014bacterial}, leveraging the big-data results from WGS is difficult due to the lack of bioinformatic standards and common infrastructure technologies.
@@
 However, we struggled to provide real-time computing of results and to accommodate an ever-growing reference database.
-Furthermore, due to the various dependencies involved, independently replicating the platform was difficult for other researchers.
+Furthermore, due to the various dependencies involved, other researchers could not easily replicate the platform.
 
@@
 Spfy is backed by a graph database which our analysis modules rely on.
-Graph databases specifically describe the relationships between different data, and is one of the emerging \cite{de2015trends} database types used for biological data.
+Graph databases focus on describing the relationships between different data, and is one of the emerging \cite{de2015trends} database types used for biological data.
 Spfy takes a modular approach to tool integration and distributes analyses over task queues, thereby allowing modules to run in parallel and address code failures.
@@
 Furthermore, large-scale analyses, such as epidemiological associations between specific genotypes, biomarkers, host, source, and other metadata, and statistically significance testing of genome markers for user-defined groups, run in real-time.
-We also containerize services using Docker, a modern approach found in many web servers and proposed to mitigate the complexities of bioinformatic pipelines \cite{di2015impact}.
+We also containerize services using Docker, a modern approach found in many web servers and proposed \cite{di2015impact} to mitigate the complexities of bioinformatic pipelines.
 Docker integration ensures that software dependencies, which typically must be manually installed \cite{doi:10.1093/bioinformatics/btu153,laing2010pan,inouye2014srst2}, are handled automatically.

### 2017-08-01 paper-webserver.tex
intro comments
beginning edits

@@
 % costs of bacterial infections
-Tracking foodborne illnesses is increasingly difficult; ingredients are globally sourced and contamination can occur at any point in the supply-chain.
-Consumers infected by bacteria incur actual costs: foodborne illnesses cause an estimated 4000 hospitalizations and 105 deaths per year in Canada \cite{thomas2015estimates}, and 55961 hospitalizations and 1351 deaths per year in the United States \cite{scallan2011foodborne}.
-Furthermore, many studies \cite{todd1989costs,scharff2010health} document the associated economic costs.
+% We need to state the "big problem" up-front that we are working on, which is not really tracking foodborne illness.
+It is fast analyses of bacterial whole-genomes, providing reference-laboratory tests from in-silico data, predicting phenotype from genotype, and providing a linked database of known genomes that don't need to be re-computed.
+This can be used in the tracking of foodborne illness, but is more an application of the problem we are working on.
+The isolation and characterization of bacterial pathogens are critical for Public Health laboratories to rapidly respond to outbreaks, and to effectively monitor known and emerging pathogens through surveillance programs.
+Within Canada it is estimated that over 4,000 hospitalizations and 100 deaths occur due to bacterial infections \cite{thomas2015estimates}; in the United States, over 55,000 hospitalizations and 1,300 deaths occur per year \cite{scallan2011foodborne}.
+Within North America the total economic burden of these illnesses is estimated to be ACTUAL NUMBER \cite{todd1989costs,scharff2010health}.
+The bacterial species that are most frequently associated with severe disease, and cause the majority of this health burden in (Canada / NA/ Europe / Worldwide ) are ABC.
+%You could combine the Canada / US numbers into a North America number, and include world-wide data as well.
+Always avoid vacuous descriptors like "many", "various" -- use an actual quantifier.
+Eg.
+Most people prefer vanilla to chocolate.
+Fifty-one percent of people prefer vanilla to chocolate.
+Ninety-nine percent of people prefer vanilla to chocolate.
+
+% routinely monitor emerging threats, of which \textit{Salmonella enterica} and \textit{Escherichia coli} are amongst the two most common \cite{kozak2013foodborne}.
+In \textit{Salmonella enterica} \cite{bell2016recent}, \textit{Escherichia coli} \cite{fratamico2016advances}, and (others pathogens --> either state the other pathogens or leave it out entirely) \cite{ronholm2016navigating}
+
 % surveillance
-Public-health agencies routinely monitor emerging threats, of which \textit{Salmonella enterica} and \textit{Escherichia coli} are amongst the two most common \cite{kozak2013foodborne}.
-In \textit{Salmonella enterica} \cite{bell2016recent}, \textit{Escherichia coli} \cite{fratamico2016advances}, and others pathogens \cite{ronholm2016navigating}, surveillance now involves a variety of bioinformatic approaches based on whole-genome sequencing (WGS).
+Until recently, Public-health agencies relied on laboratories tests such as XYZ to characterize bacterial isolates in outbreak and surveillance settings. The previous gold-standard in determining
 
-% traditional reference lab techniques
-To begin analyzing genomes, researchers designate particular strains of bacterial isolates using identifiers known as subtypes.
 Traditional reference laboratory techniques handled subtyping through methods such as serotyping, pulsed-field gel electrophoresis (PFGE), or Sanger sequencing, amongst others \cite{ronholm2016navigating}.

### 2017-08-02 paper-webserver.tex

@@
 % surveillance
-Until recently, Public-health agencies relied on laboratories tests such as XYZ to characterize bacterial isolates in outbreak and surveillance settings. The previous gold-standard in determining
-
-Traditional reference laboratory techniques handled subtyping through methods such as serotyping, pulsed-field gel electrophoresis (PFGE), or Sanger sequencing, amongst others \cite{ronholm2016navigating}.
-PFGE is unique amongst the set; restriction enzymes are used to produce DNA fragments visible as specific banding patterns.
-Public health agencies use PFGE to track outbreaks by comparing band patterns of isolates \cite{kozak2013foodborne}.
-However, as WGS rarely produces complete circular genomes and, in PFGE, cutting by a restriction enzyme is dependent on protein structure, developing bioinformatic-analogs to PFGE is difficult.
-Unlike PFGE, both serotyping and MLST have WGS-based computational methods mimicking wet-lab results.
-Serotyping, using antibodies, detects the presence of specific cell surface antigens and multi-locus sequence typing (MLST), using Sanger sequencing, examines differences in housekeeping genes - the genes required to maintain cell function.
-By comparing a pre-collected set of reference gene sequences against a WGS result, bioinformatics software predicts wet-lab serotyping results \cite{whiteside2016superphy}, Virulence Factor (VF) results \cite{joensen2014real}, and antimicrobial resistance (AMR) results \cite{mcarthur2013comprehensive}.
-MLST software uses a similar reference-based approach \cite{larsen2012multilocus}.
+Until recently, Public-health agencies relied on laboratory tests such as XYZ to characterize bacterial isolates in outbreak and surveillance settings.
+The previous gold-standard in determining strain relatedness was pulsed-field gel electrophoresis (PFGE) {ronholm2016navigating}, which uses rare-cutting restriction enzymes to produce a unique banding pattern for each strain.
+Downsides of PFGE ...
+.
+To further characterize strains ...
+serotyping, which detects the presence of specific cell surface antigens.
+Antimicrobial resistance testing, virulence factor testing ...
+However, current efforts are foucsed on predictive genomics, where the relevant phenotypic information can be determined through examination of the whole-genome sequence.
 
 % potential of WGS
-WGS also expands on traditional methods: due to the inclusion of the entire genome, WGS-based methods evaluate the spread of outbreaks with better resolution and context \cite{ronholm2016navigating}.
-Instead of examining selected single-nucleotide polymorphisms (SNPs) in a subset of housekeeping genes as in laboratory-based MLST, newer WGS-based methods, such as core genome MLST (cgMLST), attempt to cover every common gene.
-WGS approaches also expand coverage of VF and AMR gene detection to every known sequence.
-Furthermore, using WGS we can study the relations between genomes, changes in which VF or AMR genes are present over time, and mutations in genes \cite{ronholm2016navigating}.
-Researchers can also share entire datasets and results online \cite{ronholm2016navigating}.
-
-% Drawbacks - genetics
-However, these WGS-based methods only examine the underlying genetics, not the expressed phenotype.
-Correspondingly, more target sequences are identified using WGS; if health officials make threat assessments by looking for known VFs or AMR genes, they can only estimate the associated risks due to the large number of results \cite{ronholm2016navigating}.
-The effectiveness of WGS-based methods also rely on the accuracy and breadth of reference datasets and agencies must weigh the cost of developing expertise in bioinformatic approaches.
-Meanwhile, laboratory based phenotype tests remain inexpensive \cite{koser2012routine}.
+Whole-genome sequencing (WGS) uses the entire genetic content of an organism, and as such can be used to evaluate the spread of outbreaks with better resolution and context than traditional methods \cite{ronholm2016navigating}.
+However, while these WGS-based methods can fully examine the underlying genetics, the expressed phenotype is more difficult to predict.
+For example, predictions involving reference databases are only as good as the databases themselves.
+Additionally, the cost of developing, approving, and transforming existing workflows from wet-lab to sequence prediction approaches in time consuming and expensive \cite{koser2012routine}.
 
 % drawbacks - big data
-Traditionally, bioinformatics software was developed \textit{ad hoc} without employing software engineering principles \cite{de2015trends}.
-Tools were often script-based, with custom data formats, and only suitable for small collections of data \cite{de2015trends}.
-Since WGS-technology expanded the availability of data, bioinformatic pipelines are now larger and involve linked dependencies.
-Bioinformatics developers must apply systems engineering principles to handle code failures throughout \cite{schatz2015biological}.
-The bioinformatics field also lags behind other big-data industries \cite{de2015trends}: software still struggles to perform computations in real-time, stores data in inflexible databases, and lacks a common application programming interface (API) linking resources \cite{swaminathan2016review}.
+Many previously developed bioinformatics software programs have been developed \textit{ad hoc}, without the use of software engineering principles \cite{de2015trends}.
+Sych tools were often script-based, with custom data formats, and only suitable for small collections of data \cite{de2015trends}.
+While this was acceptable for smaller analyses, bioinformatic pipelines utilizing WGS data are larger and involve linked dependencies, which require the application of systems engineering principles \cite{schatz2015biological}.
+Additionally, many subsets of biology now require the analyses of big-data, where the ability to perform computations in real-time, store data in flexible databases, and utilize a common application programming interface (API) linking resources are required \cite{swaminathan2016review}.
 
 % Solution via software engineering
-Efforts such as the Comprehensive Antibiotic Resistance Database (CARD) \cite{mcarthur2013comprehensive} address data concerns by curating a set of known AMR genes.
-In addition, agencies such as the Food and Drug Administration (FDA), which hosts GenomeTrakr \cite{allard2016practical}, the National Center for Biotechnology Information (NCBI), which hosts GenBank \cite{doi:10.1093/nar/gks1195}, and the Center for Disease Control and Prevention (CDC), which hosts PulseNet \cite{swaminathan2001pulsenet}, enables sharing of genome sequences and associated metadata either openly or amongst member institutions.
-Recent efforts focused on checking the quality of shared reference genomes \cite{parks2015checkm}, particularly amongst human-genome projects \cite{lee2017ngscheckmate}.
+Current programs that address these issues include the Comprehensive Antibiotic Resistance Database (CARD) \cite{mcarthur2013comprehensive}, which addresses data concerns by curating a set of known AMR genes.
+In addition, GenomeTrackr from the Food and Drug Administration (FDA) \cite{allard2016practical}; GenBank from the National Center for Biotechnology Information (NCBI) \cite{doi:10.1093/nar/gks1195}; and PulseNet from the Center for Disease Control and Prevention (CDC) \cite{swaminathan2001pulsenet}, all enable sharing of genome sequences and associated metadata, either openly or amongst member institutions.
+
+%By comparing a pre-collected set of reference gene sequences against a WGS result, bioinformatics software predicts wet-lab serotyping results \cite{whiteside2016superphy}, Virulence Factor (VF) results \cite{joensen2014real}, and antimicrobial resistance (AMR) results \cite{mcarthur2013comprehensive}.
+%MLST software uses a similar reference-based approach \cite{larsen2012multilocus}.
 
 % rewrite of bigdata approach (BIGGER bigdata)
-While efforts such as CARD are important for replicating traditional results and for historical context, researchers can now develop new methods targeting larger datasets from WGS.
-To study entire genomes, the field needs methods of creating reference sets encompassing all genes in a given phylogeny.
-Panseq \cite{laing2010pan}, Roary \cite{page2015roary}, chewBBACA \url{https://github.com/mickaelsilva/chewBBACA}, and Graphtyper \cite{Eggertsson148403} can compute these pan-genomic regions, and are freely available.
-Such software approaches must scale to larger datasets than those found in wet-lab approaches.
-Also, for uptake by front-line clinicians and laboratories, WGS solutions should perform faster; for example, AMR gene prediction should take less than the typical 18-24 hours required in a reference laboratory \cite{koser2012routine}.
-Storing results in a tiered structure, with rapid access available for retrieval while maintaining long-term data storage \cite{schatz2015biological}, would help address the current interactivity limitations \cite{de2015trends} of bioinformatics software.
+Bacterial species, pan-genome ...Panseq \cite{laing2010pan}, Roary \cite{page2015roary}, chewBBACA \url{https://github.com/mickaelsilva/chewBBACA}, and Graphtyper \cite{Eggertsson148403} can compute these pan-genomic regions, and are freely available.
+These software programs are capable of processing X number of genomes per Y days at a cost of Z, which would take a A, B and C to do in a wet-lab.
+
+While these programs are useful for computing novel analyses, storing results in a tiered structure, with rapid access available for retrieval while maintaining long-term data storage \cite{schatz2015biological}, is needed to prevent the constant re-computation of the same data, and to provide pre-computed analyses for common predictive genomic analyses needed by the public health, and research community \cite{de2015trends}.
 
 % why Superphy
-To make bioinformatic approaches user-friendly, a variety of web services have been developed in Denmark \cite{joensen2014real,larsen2012multilocus}, Taiwan \cite{liu2016construction}, and in collaborations \cite{hasman2015detection}.
-While useful, arguably \cite{fricke2014bacterial}, leveraging the big-data results from WGS is difficult due to the lack of bioinformatic standards and common infrastructure technologies.
-Integrating results from multiple genomes, and examining shared connections, will be important for new software.
+Currently, a number of organizations are creating user-friendly web services for common analyses of bacterial genomes.
+These include: NAME \cite{joensen2014real,larsen2012multilocus}, NAME \cite{liu2016construction}, and NAME \cite{hasman2015detection}, which do X, Y and Z respectively.
+
+While useful, leveraging big-data results from WGS is difficult due to the lack of bioinformatic standards and common infrastructure technologies.
+Integrating results from multiple genomes, and examining shared connections, is an area that currently requires more research \cite{fricke2014bacterial},
 
 % Superphy
-Previously we developed Superphy \cite{whiteside2016superphy}, an online platform targeting \textit{E.
-coli} which determines Shiga toxin subtypes, AMR genes using CARD, and provides statistical significance tests between subgroups in associated metadata and pan-genome regions.
-However, we struggled to provide real-time computing of results and to accommodate an ever-growing reference database.
-Furthermore, due to the various dependencies involved, other researchers could not easily replicate the platform.
+We have previously developed Superphy \cite{whiteside2016superphy}, an online platform targeting \textit{E.
+coli}, which provided pre-computed, predictive genomic analyses for determining Shiga toxin subtype, AMR gene profiles, the presence of known virulence factors, and provides predictive biomarkers, both the presence / absence of genomic regions and single nucleotide variants, for bacterial sub-groups based on metadata such as serotype, host source, or location.
+While this tool has been useful for the thousands of pre-computed genomes in its database, the current pace of genome sequencing requires real-time predictive genomic analyses of tens of thousands of genomes, and the long term storage and referencing of these results, something that the original SuperPhy platform was incapable of.
 
-% this study
-Here we present Spfy, a complete rewrite of Superphy.
-Spfy is backed by a graph database which our analysis modules rely on.
-Graph databases focus on describing the relationships between different data, and is one of the emerging \cite{de2015trends} database types used for biological data.
-Spfy takes a modular approach to tool integration and distributes analyses over task queues, thereby allowing modules to run in parallel and address code failures.
-Serotyping, VF, AMR predictions are computed within minutes, and results are efficiently stored within the graph.
-Furthermore, large-scale analyses, such as epidemiological associations between specific genotypes, biomarkers, host, source, and other metadata, and statistically significance testing of genome markers for user-defined groups, run in real-time.
-We also containerize services using Docker, a modern approach found in many web servers and proposed \cite{di2015impact} to mitigate the complexities of bioinformatic pipelines.
-Docker integration ensures that software dependencies, which typically must be manually installed \cite{doi:10.1093/bioinformatics/btu153,laing2010pan,inouye2014srst2}, are handled automatically.
-These changes enable real-time genome analysis and database storage of results for upwards of 5000 genomes.
+In this study, we present an update to the SuperPhy platform, called Spfy.
+It is backed by a graph database and takes a modular approach to tool integration and distributes analyses over task queues, thereby allowing modules to run in parallel in real-time, and address code failures.
+This allows large-scale analyses, such as epidemiological associations between specific genotypes, biomarkers, host, source, and other metadata, and statistically significance testing of genome markers for user-defined groups.
+Lastly, these services are containerized using Docker to mitigate the complexities of the underlying bioinformatic pipelines.
 All services are publicly available at \url{https://lfz.corefacility.ca/superphy/spfy/}.
@@
 We present Spfy as an integrated predictive genomics toolkit for pathogens / \textit{E. coli} which leverages advances in web application development and deployment.
+
+which our analysis modules rely on.
+Graph databases focus on describing the relationships between different data, and is one of the emerging \cite{de2015trends} database types used for biological data.
+Spfy
+
+Serotyping, VF, AMR predictions are computed within minutes, and results are efficiently stored within the graph.
+
+Docker integration ensures that software dependencies, which typically must be manually installed \cite{doi:10.1093/bioinformatics/btu153,laing2010pan,inouye2014srst2}, are handled automatically.
+These changes enable real-time genome analysis and database storage of results for upwards of 5000 genomes.
 

### 2017-08-02 paper-webserver.tex
CHANGE: one sentance per line for git
STOP: about to shift focus to big data as opposed to public health

@@
 The previous gold-standard in determining strain relatedness was pulsed-field gel electrophoresis (PFGE) {ronholm2016navigating}, which uses rare-cutting restriction enzymes to produce a unique banding pattern for each strain.
-Downsides of PFGE ...
-.
-To further characterize strains ...
-serotyping, which detects the presence of specific cell surface antigens.
+However, in \textit{Enterococcus faecium}, PFGE has been shown to misclassify 9 of 132 isolates, when compared to whole-genome sequencing (WGS) based discrimination \cite{pinholt2015multiple}.
+In \textit{Klebsiella pneumoniae} \cite{marsh2015genomic}, \textit{Yersinia enterocolitica} \cite{gilpin2014limitations}, and \textit{Staphylococcus aureus} \cite{doi:10.1093/ofid/ofu096}, WGS was used to discriminate isolates after initial clustering by PFGE resulted in indistinguishable samples.
+Examination of PFGE bands are also subjective, difficult to share \cite{lytsy2017time}, and collative platforms such as PulseNet reported \cite{gilpin2014limitations} that even after collecting 72\% of \textit{Campylobacter jejuni} in a given year in Minnesota (673 cases), 87\% of isolates could not be linked by PFGE pattern.
+To further characterize strains ... serotyping, which detects the presence of specific cell surface antigens.
 Antimicrobial resistance testing, virulence factor testing ...
-However, current efforts are foucsed on predictive genomics, where the relevant phenotypic information can be determined through examination of the whole-genome sequence.
+However, current efforts are focused on predictive genomics, where the relevant phenotypic information can be determined through examination of the whole-genome sequence.
 
 % potential of WGS
-Whole-genome sequencing (WGS) uses the entire genetic content of an organism, and as such can be used to evaluate the spread of outbreaks with better resolution and context than traditional methods \cite{ronholm2016navigating}.
+WGS uses the entire genetic content of an organism, and as such can be used to evaluate the spread of outbreaks with better resolution and context than traditional methods \cite{ronholm2016navigating}.
 However, while these WGS-based methods can fully examine the underlying genetics, the expressed phenotype is more difficult to predict.
@@
 % drawbacks - big data
-Many previously developed bioinformatics software programs have been developed \textit{ad hoc}, without the use of software engineering principles \cite{de2015trends}.
-Sych tools were often script-based, with custom data formats, and only suitable for small collections of data \cite{de2015trends}.
+Many previous bioinformatics software programs have been developed \textit{ad hoc}, without the use of software engineering principles \cite{de2015trends}.
+Such tools were often script-based, with custom data formats, and only suitable for small collections of data \cite{de2015trends}.
 While this was acceptable for smaller analyses, bioinformatic pipelines utilizing WGS data are larger and involve linked dependencies, which require the application of systems engineering principles \cite{schatz2015biological}.

### 2017-08-02 paper-webserver.tex
ADD: initial synth of outline from chad

@@
 \section{Introduction}
-Broad --> narrow
--Bacterial infections are a cause of human disease -> associated costs
--Surveillance is used to monitor known pathogens and detect emerging threats, to lower burden of human illness
--Until recently, surveillance involved traditional reference laboratory techniques (and still does) --> examples of both old and new (WGS), what they measure, how they are used
--WGS methods have great potential, but some drawbacks (list with references)
--Some of the drawbacks can be solved by providing pre-computed analyses for routine tasks
--We have previously created SuperPhy, which helps solve some of the issues
--Some issues still remain
--In this study, we have addressed these issues by ... which will provide these benefits
+% new outline
+% para
+% 1. brief: WGS is standard
+% 2. Big Problem: but tools are for individual analysis
+% 3. 100,000 genomes, (lookup number for Enterobase, GenBank), how do we run analysis on it
+% 4. previous methods (Galaxy/IRIDA: real-time, but no storage of results, other website examples - Denmark?), what problems they addressed
+% 5. Problems that remain / lack of result storage means: recomputation, lost data, can't reference old analyses, not suited for "big data"
+% 6. Solutions in general: store and increment, huga "big data" analyses, parallelization /queues
+% para
+% 1. Our specific problem, previous work
+% 2. Why solving it is important (for Public health / research)
+% 3. how we solved it
+% para
+% 4. benefits: rapid analyses in real-time -> huge comparisons, replace reference labs -> time & money saved, future work -> expand analyses, more genomes, more species
+% 5. analyses modules -> conda -> IRIDA/Galaxy
+% 6. short snippit on website link & github link
 

### 2017-08-03 paper-webserver.tex
ADD: outline for materials and methods
ADD: new first draft of first para

@@
 \begin{abstract}
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text.
+
 \end{abstract}
@@
 % 6. short snippit on website link & github link
+
+% 1. brief: WGS is standard
+Whole genome sequencing (WGS) uses the entire genetic content of an organism, and is the new standard for bacterial surveillance \cite{ronholm2016navigating,lytsy2017time}, research into disease \cite{wang2014whole,yuen2015whole}, and clinical diagnoses \cite{willig2015whole,dewey2014clinical}.
+% 2. Big Problem: but tools are for individual analysis
+Targeted software, such as the Resistance Gene Identifier (RGI) \cite{mcarthur2013comprehensive} for antimicrobial resistance (AMR) gene prediction, Prokka for bacterial genome annotation \cite{doi:10.1093/bioinformatics/btu153}, and integrated platforms, such as the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial} and the Integrated Rapid Infectious Disease Analysis (IRIDA) project \url{http://www.irida.ca/}, all leverage WGS data.
+For bioinformatics software to reach true "big-data" goals, analyses modules must run in real-time and results must be integrated.
+% 3. 100,000 genomes, (lookup number for Enterobase, GenBank), how do we run analysis on it
+WGS generates data at a break-neck speed; for \textit{Escherichia coli} alone, the public genome databases EnteroBase \url{https://enterobase.warwick.ac.uk/} and GenBank \cite{doi:10.1093/nar/gks1195} have 60,206 and 2,779,008 sequences uploaded, respectively.
+Typical bioinformatics software, such as RGI and Prokka, take single files as input, and integrated platforms, such as BAP and IRIDA, build workflows linking different analyses modules.
+BAP and IRIDA begin to solve "big-data" challenges by offering a hosted solution which computes results in real-time, and parallelizes analyses across computing resources.
+For WGS to further benefit surveillance, research, and clinical fields, analyses results should also be integrated and searchable; storage of results would avoid recomputation, persist data, and allow references to old analyses.
+\par
+
+% 1. Our specific problem, previous work
+We've attempted to build big-data platforms which address both real-time analyses and integrated storage of results.
+We have previously developed Superphy \cite{whiteside2016superphy}, an online predictive genomics platform targeting \textit{E.
+coli}.
 
@@
 % potential of WGS
-WGS uses the entire genetic content of an organism, and as such can be used to evaluate the spread of outbreaks with better resolution and context than traditional methods \cite{ronholm2016navigating}.
+, and as such can be used to evaluate the spread of outbreaks with better resolution and context than traditional methods \cite{ronholm2016navigating}.
 However, while these WGS-based methods can fully examine the underlying genetics, the expressed phenotype is more difficult to predict.
@@
 % Superphy
-We have previously developed Superphy \cite{whiteside2016superphy}, an online platform targeting \textit{E.
-coli}, which provided pre-computed, predictive genomic analyses for determining Shiga toxin subtype, AMR gene profiles, the presence of known virulence factors, and provides predictive biomarkers, both the presence / absence of genomic regions and single nucleotide variants, for bacterial sub-groups based on metadata such as serotype, host source, or location.
+, which provided pre-computed, predictive genomic analyses for determining Shiga toxin subtype, AMR gene profiles, the presence of known virulence factors, and provides predictive biomarkers, both the presence / absence of genomic regions and single nucleotide variants, for bacterial sub-groups based on metadata such as serotype, host source, or location.
 While this tool has been useful for the thousands of pre-computed genomes in its database, the current pace of genome sequencing requires real-time predictive genomic analyses of tens of thousands of genomes, and the long term storage and referencing of these results, something that the original SuperPhy platform was incapable of.
@@
 
+% **************************************************************
+% Keep this command to avoid text of first page running into the
+% first page footnotes
+\enlargethispage{-65.1pt}
+% **************************************************************
+
 \section{MATERIALS AND METHODS}
+% outline
+% para
+% problem: predictive analylitcs in minutes - sol'n: RQ
+%	related: packaging of modules in conda
+% para (probably 2 paras)
+% problem: distribute tasks & automate installation - sol'n: docker
+% para(probably 2 for graphs)
+% problem: storage up 10,000+ genomes - sol'n: graph database
+% para
+% problem: responsive ui that's easy to navigate - sol'n: react & material design
+% shortpara
+% problem: CI/testing - sol'n Github, TravisCI, Pytest
+
 We designed Spfy's front-end following the Material Design standards \url{https://material.io/}, released by Google.
@@
 These changes enable real-time genome analysis and database storage of results for upwards of 5000 genomes.
-
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text{}. Text.
-Text \cite{1}.
-
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-\begin{align*}
-&\mathrm{Ascorbate} + \mathrm{EDTA} \cdot \mathrm{Fe}^{3+} \to
-\hbox{Oxidized ascorbate}
-\\
-&\mathrm{EDTA} \cdot \mathrm{Fe}^{2+} + \mathrm{H}_2
-\mathrm{O}_2 \to
-\mathrm{EDTA} \cdot \mathrm{Fe}^{3+} + \cdot
-\mathrm{OH} + \mathrm{OH}^-
-\end{align*}
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-% **************************************************************
-% Keep this command to avoid text of first page running into the
-% first page footnotes
-\enlargethispage{-65.1pt}
-% **************************************************************
-
-Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text.
-Text \cite{2,3}.
-
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text.
-
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text.
-Text \cite{4}.
-
-\section{MATERIALS AND METHODS}
 

### 2017-08-03 paper-webserver.tex
ADD: start of second para

@@
 % 1. Our specific problem, previous work
-We've attempted to build big-data platforms which address both real-time analyses and integrated storage of results.
-We have previously developed Superphy \cite{whiteside2016superphy}, an online predictive genomics platform targeting \textit{E.
-coli}.
+We have previously developed Superphy \cite{whiteside2016superphy}, an online predictive genomics platform targeting \textit{E. coli}.
+Superphy integrates pre-computed results with domain-specific metadata to provide real-time analyses of epidemiology relations over thousands of genomes.
+To scale Superphy for tens of thousands of genomes and for use by front-line clinicians, software pipelines should run in real-time and existing population-wide analyses should automatically integrate new results.
 

### 2017-08-03 paper-webserver.tex
ADD: something looking like a 2nd para

@@
 Superphy integrates pre-computed results with domain-specific metadata to provide real-time analyses of epidemiology relations over thousands of genomes.
-To scale Superphy for tens of thousands of genomes and for use by front-line clinicians, software pipelines should run in real-time and existing population-wide analyses should automatically integrate new results.
+While this tool has been useful for the thousands of pre-computed genomes in its database, the current pace of genome sequencing requires real-time predictive genomic analyses of tens of thousands of genomes, and the long term storage and referencing of these results, something that the original SuperPhy platform was incapable of.
+% 2. Why solving it is important (for Public health / research)
+WGS offers better resolution than traditional strain comparison methods, such as  pulsed-field gel electrophoresis (PFGE) \cite{ronholm2016navigating}, and is set to replace the functions of reference laboratories.
+Though the cost of developing, approving, and transforming existing workflows from wet-lab to sequence prediction approaches in time consuming and expensive \cite{koser2012routine}, performing real-time analyses and linkage to thousands of historical results is only possible through WGS.
+% 3. how we solved it
+In this study, we present an update to the SuperPhy platform, called Spfy.
+Our update rewrites result storage with backing by a graph database and takes a modular approach to tool integration and distributes analyses over task queues, thereby allowing users to submit genomes and modules to run in parallel in real-time, and address code failures.
+All results are stored as a series of linked nodes which enables the platform to build associations between results as they are generated. \par
+
+% 4. benefits: rapid analyses in real-time -> huge comparisons, replace reference labs -> time & money saved, future work -> expand analyses, more genomes, more species
+This allows large-scale analyses, such as epidemiological associations between specific genotypes, biomarkers, host, source, and other metadata, and statistical significance testing of genome markers for user-defined groups.
 
@@
 For example, predictions involving reference databases are only as good as the databases themselves.
-Additionally, the cost of developing, approving, and transforming existing workflows from wet-lab to sequence prediction approaches in time consuming and expensive \cite{koser2012routine}.
 
@@
 , which provided pre-computed, predictive genomic analyses for determining Shiga toxin subtype, AMR gene profiles, the presence of known virulence factors, and provides predictive biomarkers, both the presence / absence of genomic regions and single nucleotide variants, for bacterial sub-groups based on metadata such as serotype, host source, or location.
-While this tool has been useful for the thousands of pre-computed genomes in its database, the current pace of genome sequencing requires real-time predictive genomic analyses of tens of thousands of genomes, and the long term storage and referencing of these results, something that the original SuperPhy platform was incapable of.
 
-In this study, we present an update to the SuperPhy platform, called Spfy.
-It is backed by a graph database and takes a modular approach to tool integration and distributes analyses over task queues, thereby allowing modules to run in parallel in real-time, and address code failures.
-This allows large-scale analyses, such as epidemiological associations between specific genotypes, biomarkers, host, source, and other metadata, and statistically significance testing of genome markers for user-defined groups.
 Lastly, these services are containerized using Docker to mitigate the complexities of the underlying bioinformatic pipelines.

### 2017-08-03 paper-webserver.tex
ADD: some words that look like a 3rd para

@@
 % 4. benefits: rapid analyses in real-time -> huge comparisons, replace reference labs -> time & money saved, future work -> expand analyses, more genomes, more species
-This allows large-scale analyses, such as epidemiological associations between specific genotypes, biomarkers, host, source, and other metadata, and statistical significance testing of genome markers for user-defined groups.
+Integrating task distribution with graph storage allows large-scale analyses, such as epidemiological associations between specific genotypes, biomarkers, host, source, and other metadata, and statistical significance testing of genome markers for user-defined groups.
+By supporting multiple \textit{in-silico} subtyping options, the platform can function as an online reference laboratory.
+Currently, the platform has been tested with XXX genome files and result storage for XX analyses modules.
+Future work will focus on expanding available analyses and supporting different species, which can be connected to the existing graph database without need for recalculation.
+To complement existing platforms such as IRIDA, analyses modules are individually packaged and easily integrated into Galaxy \cite{goecks2010galaxy} based platforms.
+The website and source-code are available at \url{https://superphy.github.io/}. \par
 

### 2017-08-03 paper-webserver.tex

@@
 % 1. brief: WGS is standard
-Whole genome sequencing (WGS) uses the entire genetic content of an organism, and is the new standard for bacterial surveillance \cite{ronholm2016navigating,lytsy2017time}, research into disease \cite{wang2014whole,yuen2015whole}, and clinical diagnoses \cite{willig2015whole,dewey2014clinical}.
+Whole genome sequencing (WGS) uses the entire genetic content of an organism, and is used for bacterial surveillance \cite{ronholm2016navigating,lytsy2017time}, research into disease \cite{wang2014whole,yuen2015whole}, and clinical diagnoses \cite{willig2015whole,dewey2014clinical}.
 % 2. Big Problem: but tools are for individual analysis
@@
 Typical bioinformatics software, such as RGI and Prokka, take single files as input, and integrated platforms, such as BAP and IRIDA, build workflows linking different analyses modules.
-BAP and IRIDA begin to solve "big-data" challenges by offering a hosted solution which computes results in real-time, and parallelizes analyses across computing resources.
-For WGS to further benefit surveillance, research, and clinical fields, analyses results should also be integrated and searchable; storage of results would avoid recomputation, persist data, and allow references to old analyses.
+BAP and IRIDA begin to solve "big-data" challenges by offering a hosted solution which computes results in real-time, and parallelizing analyses across computing resources.
+For WGS to further benefit surveillance, research, and clinical fields, analyses results should also be integrated and searchable; storage of results would avoid recomputation, enable data persistence, and allow references to old analyses.
 \par
@@
 We have previously developed Superphy \cite{whiteside2016superphy}, an online predictive genomics platform targeting \textit{E. coli}.
-Superphy integrates pre-computed results with domain-specific metadata to provide real-time analyses of epidemiology relations over thousands of genomes.
+Superphy integrates pre-computed results with domain-specific metadata to provide real-time analyses of epidemiology relations.
 While this tool has been useful for the thousands of pre-computed genomes in its database, the current pace of genome sequencing requires real-time predictive genomic analyses of tens of thousands of genomes, and the long term storage and referencing of these results, something that the original SuperPhy platform was incapable of.
 % 2. Why solving it is important (for Public health / research)
-WGS offers better resolution than traditional strain comparison methods, such as  pulsed-field gel electrophoresis (PFGE) \cite{ronholm2016navigating}, and is set to replace the functions of reference laboratories.
-Though the cost of developing, approving, and transforming existing workflows from wet-lab to sequence prediction approaches in time consuming and expensive \cite{koser2012routine}, performing real-time analyses and linkage to thousands of historical results is only possible through WGS.
+WGS offers improved resolution over traditional strain comparison methods, such as  pulsed-field gel electrophoresis (PFGE) \cite{ronholm2016navigating}, and is set to replace the functions of reference laboratories.
+Though the cost of developing, approving, and transforming existing workflows from wet-lab to sequence prediction approaches is time consuming and expensive \cite{koser2012routine}, platforms can only perform real-time analyses and linkage to thousands of historical results through leveraging WGS.
 % 3. how we solved it
 In this study, we present an update to the SuperPhy platform, called Spfy.
-Our update rewrites result storage with backing by a graph database and takes a modular approach to tool integration and distributes analyses over task queues, thereby allowing users to submit genomes and modules to run in parallel in real-time, and address code failures.
+The update rewrites result storage with backing by a graph database and takes a modular approach to tool integration and distributes analyses over task queues, thereby allowing users to submit genomes and modules to run in parallel in real-time, and address code failures.
 All results are stored as a series of linked nodes which enables the platform to build associations between results as they are generated. \par
@@
 % 4. benefits: rapid analyses in real-time -> huge comparisons, replace reference labs -> time & money saved, future work -> expand analyses, more genomes, more species
-Integrating task distribution with graph storage allows large-scale analyses, such as epidemiological associations between specific genotypes, biomarkers, host, source, and other metadata, and statistical significance testing of genome markers for user-defined groups.
-By supporting multiple \textit{in-silico} subtyping options, the platform can function as an online reference laboratory.
+Integrating task distribution with graph storage enables large-scale analyses, such as epidemiological associations between specific genotypes, biomarkers, host, source, and other metadata, and statistical significance testing of genome markers for user-defined groups.
+Subtyping options are ..., pan-genome generation ..., group comparisons via Fisher's, ML, .... for E.coli.
+By supporting multiple \textit{in-silico} subtyping options, the platform functions similar to a reference laboratory, with added support for big-data analyses.
 Currently, the platform has been tested with XXX genome files and result storage for XX analyses modules.
@@
 The website and source-code are available at \url{https://superphy.github.io/}. \par
+
+% end of new intro
+% below material is just old stuff, might or might not be used elsewhere
 

### 2017-08-03 paper-webserver.tex

@@
 % 1. brief: WGS is standard
-Whole genome sequencing (WGS) uses the entire genetic content of an organism, and is used for bacterial surveillance \cite{ronholm2016navigating,lytsy2017time}, research into disease \cite{wang2014whole,yuen2015whole}, and clinical diagnoses \cite{willig2015whole,dewey2014clinical}.
+Whole genome sequencing (WGS) uses the entire genetic content of an organism, and widely complements bacterial surveillance \cite{ronholm2016navigating,lytsy2017time}, research into disease \cite{wang2014whole,yuen2015whole}, and clinical diagnoses \cite{willig2015whole,dewey2014clinical}.
 % 2. Big Problem: but tools are for individual analysis

### 2017-08-04 paper-webserver.tex
ADD: subsections for materials and methods
CHANGE: dump mat and meth text into respective subsections
EDIT: few intro edits

@@
 Targeted software, such as the Resistance Gene Identifier (RGI) \cite{mcarthur2013comprehensive} for antimicrobial resistance (AMR) gene prediction, Prokka for bacterial genome annotation \cite{doi:10.1093/bioinformatics/btu153}, and integrated platforms, such as the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial} and the Integrated Rapid Infectious Disease Analysis (IRIDA) project \url{http://www.irida.ca/}, all leverage WGS data.
+% 3. 100,000 genomes, (lookup number for Enterobase, GenBank), how do we run analysis on it
+WGS generates data at a break-neck speed; for \textit{Escherichia coli} alone, the public genome databases EnteroBase \url{https://enterobase.warwick.ac.uk/} and GenBank \cite{doi:10.1093/nar/gks1195} have, respectively, 60,206 and 2,779,008 sequences uploaded.
 For bioinformatics software to reach true "big-data" goals, analyses modules must run in real-time and results must be integrated.
-% 3. 100,000 genomes, (lookup number for Enterobase, GenBank), how do we run analysis on it
-WGS generates data at a break-neck speed; for \textit{Escherichia coli} alone, the public genome databases EnteroBase \url{https://enterobase.warwick.ac.uk/} and GenBank \cite{doi:10.1093/nar/gks1195} have 60,206 and 2,779,008 sequences uploaded, respectively.
 Typical bioinformatics software, such as RGI and Prokka, take single files as input, and integrated platforms, such as BAP and IRIDA, build workflows linking different analyses modules.
-BAP and IRIDA begin to solve "big-data" challenges by offering a hosted solution which computes results in real-time, and parallelizing analyses across computing resources.
+BAP and IRIDA begin to solve big-data challenges by offering a hosted solution which computes results in real-time, and distributes analyses across computing resources.
 For WGS to further benefit surveillance, research, and clinical fields, analyses results should also be integrated and searchable; storage of results would avoid recomputation, enable data persistence, and allow references to old analyses.
@@
 In this study, we present an update to the SuperPhy platform, called Spfy.
-The update rewrites result storage with backing by a graph database and takes a modular approach to tool integration and distributes analyses over task queues, thereby allowing users to submit genomes and modules to run in parallel in real-time, and address code failures.
+The update rewrites result storage with backing by a graph database, takes a modular approach to tool integration, and distributes analyses over task queues, thereby allowing users to submit genomes and modules to run in parallel in real-time, and address code failures.
 All results are stored as a series of linked nodes which enables the platform to build associations between results as they are generated. \par
@@
 % 4. benefits: rapid analyses in real-time -> huge comparisons, replace reference labs -> time & money saved, future work -> expand analyses, more genomes, more species
-Integrating task distribution with graph storage enables large-scale analyses, such as epidemiological associations between specific genotypes, biomarkers, host, source, and other metadata, and statistical significance testing of genome markers for user-defined groups.
+By integrating task distribution with graph storage, Spfy enables large-scale analyses, such as epidemiological associations between specific genotypes, biomarkers, host, source, and other metadata, and statistical significance testing of genome markers for user-defined groups.
 Subtyping options are ..., pan-genome generation ..., group comparisons via Fisher's, ML, .... for E.coli.
@@
 Currently, the platform has been tested with XXX genome files and result storage for XX analyses modules.
-Future work will focus on expanding available analyses and supporting different species, which can be connected to the existing graph database without need for recalculation.
-To complement existing platforms such as IRIDA, analyses modules are individually packaged and easily integrated into Galaxy \cite{goecks2010galaxy} based platforms.
-The website and source-code are available at \url{https://superphy.github.io/}. \par
+Future work will focus on adding more analysis modules and supporting different species, which can be connected to the existing graph database without need for recalculation.
+To complement existing platforms such as IRIDA, modules are self-contained and can easily be integrated into Galaxy \cite{goecks2010galaxy} based platforms.
+The website and source code are available at \url{https://superphy.github.io/}. \par
 
@@
 
-We designed Spfy's front-end following the Material Design standards \url{https://material.io/}, released by Google.
-The user interface is implemented with the React JavaScript library \url{https://facebook.github.io/react/}, by Facebook, as a single-page application to allow efficient data-flow without reloading the website.
-Everything is compartmentalized within separate Docker containers where the front-end is networked, through Docker-Compose, to the back-end webserver.
+We present Spfy as an integrated predictive genomics toolkit for pathogens / \textit{E. coli} which leverages advances in web application development and deployment.
+
+which our analysis modules rely on.
+
+These changes enable real-time genome analysis and database storage of results for upwards of 5000 genomes.
+
+\subsection{Real-time analysis pipelines}
 
@@
 Containers are orchestrated through Docker-Compose allowing the entire platform: webservers, databases, and task workers, to be easily replicated by other researchers.
-We present Spfy as an integrated predictive genomics toolkit for pathogens / \textit{E. coli} which leverages advances in web application development and deployment.
 
-which our analysis modules rely on.
+\subsubsection{}
+
+\subsection{Task distribution and containerization}
+
+Everything is compartmentalized within separate Docker containers where the front-end is networked, through Docker-Compose, to the back-end webserver.
+
+Docker integration ensures that software dependencies, which typically must be manually installed \cite{doi:10.1093/bioinformatics/btu153,laing2010pan,inouye2014srst2}, are handled automatically.
+
+\subsection{Graph databases and ontologies}
+
 Graph databases focus on describing the relationships between different data, and is one of the emerging \cite{de2015trends} database types used for biological data.
@@
 
-Docker integration ensures that software dependencies, which typically must be manually installed \cite{doi:10.1093/bioinformatics/btu153,laing2010pan,inouye2014srst2}, are handled automatically.
-These changes enable real-time genome analysis and database storage of results for upwards of 5000 genomes.
+\subsection{Intuitive web design}
 
-\subsection{Materials subsection one}
+We designed Spfy's front-end following the Material Design standards \url{https://material.io/}, released by Google.
+The user interface is implemented with the React JavaScript library \url{https://facebook.github.io/react/}, by Facebook, as a single-page application to allow efficient data-flow without reloading the website.
 
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-
-\subsubsection{Materials subsubsection one.}
-
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text:
-\begin{align}
-\mathrm{LD}^r = \frac{\mathrm{LD}}{A_\mathrm{iso}}
-= 1.5 S \left( 3 \cos^2 \alpha_i - 1 \right)
-\end{align}
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-
-\subsection{Materials subsection two}
+\subsection{Continuous integration and testing}
 

### 2017-08-04 paper-webserver.tex

--- 2017-08-03 paper-webserver.tex
@@
 % 1. brief: WGS is standard
-Whole genome sequencing (WGS) uses the entire genetic content of an organism, and is used for bacterial surveillance \cite{ronholm2016navigating,lytsy2017time}, research into disease \cite{wang2014whole,yuen2015whole}, and clinical diagnoses \cite{willig2015whole,dewey2014clinical}.
+Whole genome sequencing (WGS) resolves the entire genetic content of an organism.
+WGS data can increase the resolution and sensitivity of bacterial surveillance \cite{ronholm2016navigating,lytsy2017time}, identification of potential disease mechanisms \cite{wang2014whole,yuen2015whole}, and clinical diagnoses \cite{willig2015whole,dewey2014clinical}.
 % 2. Big Problem: but tools are for individual analysis
 Targeted software, such as the Resistance Gene Identifier (RGI) \cite{mcarthur2013comprehensive} for antimicrobial resistance (AMR) gene prediction, Prokka for bacterial genome annotation \cite{doi:10.1093/bioinformatics/btu153}, and integrated platforms, such as the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial} and the Integrated Rapid Infectious Disease Analysis (IRIDA) project \url{http://www.irida.ca/}, all leverage WGS data.
-For bioinformatics software to reach true "big-data" goals, analyses modules must run in real-time and results must be integrated.
 % 3. 100,000 genomes, (lookup number for Enterobase, GenBank), how do we run analysis on it
 WGS generates data at a break-neck speed; for \textit{Escherichia coli} alone, the public genome databases EnteroBase \url{https://enterobase.warwick.ac.uk/} and GenBank \cite{doi:10.1093/nar/gks1195} have 60,206 and 2,779,008 sequences uploaded, respectively.
+To effectively exploit WGS "big-data" and maintain the rapid response time required by public health applications, one approach is to make results from WGS analysis integrated and progressive.
 Typical bioinformatics software, such as RGI and Prokka, take single files as input, and integrated platforms, such as BAP and IRIDA, build workflows linking different analyses modules.
-BAP and IRIDA begin to solve "big-data" challenges by offering a hosted solution which computes results in real-time, and parallelizing analyses across computing resources.
-For WGS to further benefit surveillance, research, and clinical fields, analyses results should also be integrated and searchable; storage of results would avoid recomputation, enable data persistence, and allow references to old analyses.
-\par
+BAP and IRIDA help address "big-data" challenges by offering a hosted solution which computes results in real-time, and parallelizing analyses across computing resources.
+While effective for self-contained workflows, many comparative analyses such as predictive genomics methods would benefit from a broad WGS reference base.
+To use vast amount of WGS data and maintain real-time results for comparative analyses, platforms would need store results in a way that avoids recomputation when adding new WGS data.
+A resource framework in which WGS results are integrated and searchable; whereby the storage of results avoids recomputation, persists, and allows for iterative, on-going learning, will expand the capacity of current comparative bioinformatics analyses.
+This increased capacity will further benefit surveillance, research, and clinical applications\par
 
@@
 % 2. Why solving it is important (for Public health / research)
-WGS offers improved resolution over traditional strain comparison methods, such as  pulsed-field gel electrophoresis (PFGE) \cite{ronholm2016navigating}, and is set to replace the functions of reference laboratories.
-Though the cost of developing, approving, and transforming existing workflows from wet-lab to sequence prediction approaches is time consuming and expensive \cite{koser2012routine}, platforms can only perform real-time analyses and linkage to thousands of historical results through leveraging WGS.
+WGS offers improved resolution over traditional strain comparison methods, such as pulsed-field gel electrophoresis (PFGE) \cite{ronholm2016navigating}.
+Though the cost of developing, approving, and transforming existing workflows from wet-lab to sequence prediction approaches is time consuming and expensive \cite{koser2012routine}, platforms can only perform real-time analyses and linkage to thousands of historical results by leveraging WGS.
 % 3. how we solved it

### 2017-08-04 paper-webserver.tex

--- 2017-08-04 paper-webserver.tex
@@
 % 1. brief: WGS is standard
-Whole genome sequencing (WGS) uses the entire genetic content of an organism, and widely complements bacterial surveillance \cite{ronholm2016navigating,lytsy2017time}, research into disease \cite{wang2014whole,yuen2015whole}, and clinical diagnoses \cite{willig2015whole,dewey2014clinical}.
+Whole genome sequencing (WGS) resolves the entire genetic content of an organism.
+WGS data can increase the resolution and sensitivity of bacterial surveillance \cite{ronholm2016navigating,lytsy2017time}, identification of potential disease mechanisms \cite{wang2014whole,yuen2015whole}, and clinical diagnoses \cite{willig2015whole,dewey2014clinical}.
 % 2. Big Problem: but tools are for individual analysis
@@
 WGS generates data at a break-neck speed; for \textit{Escherichia coli} alone, the public genome databases EnteroBase \url{https://enterobase.warwick.ac.uk/} and GenBank \cite{doi:10.1093/nar/gks1195} have, respectively, 60,206 and 2,779,008 sequences uploaded.
-For bioinformatics software to reach true "big-data" goals, analyses modules must run in real-time and results must be integrated.
+To effectively exploit WGS "big-data" and maintain the rapid response time required by public health applications, one approach is to make results from WGS analysis integrated and progressive.
 Typical bioinformatics software, such as RGI and Prokka, take single files as input, and integrated platforms, such as BAP and IRIDA, build workflows linking different analyses modules.
 BAP and IRIDA begin to solve big-data challenges by offering a hosted solution which computes results in real-time, and distributes analyses across computing resources.
-For WGS to further benefit surveillance, research, and clinical fields, analyses results should also be integrated and searchable; storage of results would avoid recomputation, enable data persistence, and allow references to old analyses.
-\par
+While effective for self-contained workflows, many comparative analyses such as predictive genomics methods would benefit from a broad WGS reference base.
+To use vast amount of WGS data and maintain real-time results for comparative analyses, platforms would need store results in a way that avoids recomputation when adding new WGS data.
+A resource framework in which WGS results are integrated and searchable; whereby the storage of results avoids recomputation, persists, and allows for iterative, on-going learning, will expand the capacity of current comparative bioinformatics analyses.
+This increased capacity will further benefit surveillance, research, and clinical applications\par
 

### 2017-08-05 paper-webserver.tex
ADD: basic outline for graph dbs and web design

@@
 
-\title{Article title}
+\title{Spfy: distributed predictive genomics of E.coli with graph based result linkage}
 
@@
 
-\section{MATERIALS AND METHODS}
+\section{METHODS AND IMPLEMENTATION}
 % outline
@@
 
-We present Spfy as an integrated predictive genomics toolkit for pathogens / \textit{E. coli} which leverages advances in web application development and deployment.
+% Inputs: more like going from file -> analysis -> graph db (not a hard subsection)
+Spfy takes assembled genomes as input, runs selected analyses modules, and converts the results into a graph structure for upload.
 
-which our analysis modules rely on.
+% Outputs: more like what we query from blazegraph and how we use it (not a hard subsection)
+Spfy queries the graph database to run analyses XYZ and build visualizations.
 
-These changes enable real-time genome analysis and database storage of results for upwards of 5000 genomes.
+\subsection{Graph databases and ontologies}
+% para
+% 1. spfy is built around graph technologies
+% 2. graph dbs store data as a series of links
+% 3. existing graph dbs and bioinformatics implementations
+% para
+% 1. how we structure our graph
+% 2. ontoogies used
+% 3. inferencing
+% 3 1/2. SPARQL queries
+% 4. results/timings
+
+Graph databases focus on describing the relationships between different data, and is one of the emerging \cite{de2015trends} database types used for biological data.
+Spfy
+
+Serotyping, VF, AMR predictions are computed within minutes, and results are efficiently stored within the graph.
+
+\subsection{Web design}
+% para
+% 1. goals: intuitive/familar, ease of use 
+% 2. design specs and different examples
+% 3. Google Material design
+% 4. some focus on card based design
+% para
+% 1. implementation: reactjs, react-md, ES6, JSX
+% 2. A to B: example of how users go from uploading files to results
+% 3. backwards compatibility with older browsers: transpile via Babel
+% 4. separation from Flask layer
+
+We designed Spfy's front-end following the Material Design standards \url{https://material.io/}, released by Google.
+The user interface is implemented with the React JavaScript library \url{https://facebook.github.io/react/}, by Facebook, as a single-page application to allow efficient data-flow without reloading the website.
 
@@
 Docker integration ensures that software dependencies, which typically must be manually installed \cite{doi:10.1093/bioinformatics/btu153,laing2010pan,inouye2014srst2}, are handled automatically.
-
-\subsection{Graph databases and ontologies}
-
-Graph databases focus on describing the relationships between different data, and is one of the emerging \cite{de2015trends} database types used for biological data.
-Spfy
-
-Serotyping, VF, AMR predictions are computed within minutes, and results are efficiently stored within the graph.
-
-\subsection{Intuitive web design}
-
-We designed Spfy's front-end following the Material Design standards \url{https://material.io/}, released by Google.
-The user interface is implemented with the React JavaScript library \url{https://facebook.github.io/react/}, by Facebook, as a single-page application to allow efficient data-flow without reloading the website.
 
@@
 
-\section{RESULTS}
+\subsection{Guides and documentation}
+
+\section{RESULTS AND DISCUSSION}
 
@@
 
-\section{DISCUSSION}
-
-\subsection{Discussion subsection one}
-
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text.
-
-\subsection{Discussion subsection two}
-
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text.
-
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-
-\subsection{Discussion subsection three}
-
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text.
-
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text.
-
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text.
-
 \section{CONCLUSION}

### 2017-08-09 paper-webserver.tex
ADD: last of the outlines for methods/implementation
DEL: old outline
DEL: NAR filler text
ADD: some thoughts on the results/discussion paras
ADD: similar bioinfo platforms to compare to

@@
 \section{METHODS AND IMPLEMENTATION}
-% outline
 % para
-% problem: predictive analylitcs in minutes - sol'n: RQ
-%	related: packaging of modules in conda
-% para (probably 2 paras)
-% problem: distribute tasks & automate installation - sol'n: docker
-% para(probably 2 for graphs)
-% problem: storage up 10,000+ genomes - sol'n: graph database
-% para
-% problem: responsive ui that's easy to navigate - sol'n: react & material design
-% shortpara
-% problem: CI/testing - sol'n Github, TravisCI, Pytest
+% goals: predictive analylitcs in minutes, storage up 50,000-100,000 genomes
 
@@
 % para
+% 0. Goals: big-data, everything linked, easy addition of new links
 % 1. spfy is built around graph technologies
@@
 % 3 1/2. SPARQL queries
-% 4. results/timings
+% 4. results/timings: scale
 
@@
 
+\subsection{Real-time analysis pipelines and task distribution}
 \subsection{Real-time analysis pipelines}
+% para
+% 1. goals: real-time, support for pipelines (linked modules)
+% 2. how pipelines have been handled in the past: Galaxy, other examples
+% 3. Python, RQ
+% para
+% 1. how we implemented RQ
+%	related: packaging of modules in conda
+% 2. different pipelines available
+% 3. how pipelines are started from the front-end
+% 4. result storage in graph db
 
@@
 
-\subsubsection{}
+\subsubsection{Task distribution and parallelization}
+% para
+% 1. goals: scale analyses to "big-data", error handling
+% 2. examples of different ways parallelization has been handled in the past
+% 3. how we handle parallelization with RQ
+% para
+% 1. how many tasks have we tested this with
+% 2. error handling: rq-dashboard, sentry
+% 3. why options like sentry are better than traditional logging: scales well to tons (big-data levels) of tasks, groups the same errors together, reporting via email
 
-\subsection{Task distribution and containerization}
+\subsection{Containerization}
+% para
+% 1. goals: why compartmentilizations
+% 2. examples of different virtualization technologies
+% 3. intro do docker
+% para
+% 1. how we implemented docker
+% 2. use of docker-compose
+% 3. how this lets us replicate worker containers and link everything together
 
@@
 
-Docker integration ensures that software dependencies, which typically must be manually installed \cite{doi:10.1093/bioinformatics/btu153,laing2010pan,inouye2014srst2}, are handled automatically.
+Docker integration ensures that software dependencies, which typically must be manually installed \cite{doi:10.1093/bioinformatics/btu153,laing2010pan,inouye2014srst2,naccache2014cloud}, are handled automatically.
 
 \subsection{Continuous integration and testing}
-
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text
-(see Figure \ref{NAR-fig1}).
-
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text.
-\begin{equation*}
-\mathrm{LD} \left( t \right) =
-\sum\limits_i
-a_i \exp \left( \frac{-t}{\tau_i} \right)
-\end{equation*}
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text.
+% para
+% 1. goals: why CI, testing is important
+% 2. examples of different testing platforms
+% 3. how we've implemented it, integration with github
 
@@
 \subsection{Guides and documentation}
+% para
+% 1. goals: use guide (ie. the README), dev guide
+% 2. how docs have been done in the past (in general)
+% 3. our use of ReadtheDocs
 
@@
 
-\subsection{Results subsection one}
+\subsection{Further containerizing analyses modules}
 
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text.
+% ex. in python, tox can be set to spin up docker containers for testing - see certbot
+% could spin up a new container for each analysis module to run see \cite{di2015impact} - neglible performance impact
+
+\subsection{Impact on Public Health Efforts}
+
+% para
+% focus on application
+
+\subsection{Deployment to Cloud-Computing Platforms}
+% how costs can be saved because using standard web tech means you can deploy to different cloud computing services
+
+\subsection{Comparison with other bioinformatic pipeline technologies}
+
+% namely galaxy
+
+\subsection{Comparison with similar bioinformatic pipelines}
+
+% para: \cite{naccache2014cloud}
+% must install a ton of deps and even create dbs manually https://github.com/chiulab/surpi
+% im not actually sure where they're getting the 'cloud-compatible' aspect of their paper from - its cloud compatible in that you deploy in the same manner as you would to bare-metal, but that's just like saying it 'runs on a computer'
+
+% para: \cite{aanensen2016whole}
+% makes some pretty pictures, but not really a pipeline
+
+% para: \cite{joensen2014real,thomsen2016bacterial}
+% we have a lot of similar directions to this, but just try to make it prettier, more friendly to use, and faster
 
@@
 
-\subsection{Results subsection two}
-
-Text.  Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text (see Table \ref{table:01}).
-
-Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text (see Figure \ref{NAR-fig2}a).
-
-Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-
 \begin{figure*}[t]
@@
 
-\subsection{Results subsection three}
-
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text.
-
 \section{CONCLUSION}
 
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text.
-
 \section{ACKNOWLEDGEMENTS}
-
-Text. Text. Text. Text. Text. Text. Text. Text. Text. Text. Text.
-Text. Text. Text. Text.
 

### 2017-08-11 paper-webserver.tex
ADD: matt's comments REMOVE: a couple of section headings
ADD: talking pts for Functionality section

@@
 Examination of PFGE bands are also subjective, difficult to share \cite{lytsy2017time}, and collative platforms such as PulseNet reported \cite{gilpin2014limitations} that even after collecting 72\% of \textit{Campylobacter jejuni} in a given year in Minnesota (673 cases), 87\% of isolates could not be linked by PFGE pattern.
-To further characterize strains ... serotyping, which detects the presence of specific cell surface antigens.
 Antimicrobial resistance testing, virulence factor testing ...
@@
 % Solution via software engineering
-Current programs that address these issues include the Comprehensive Antibiotic Resistance Database (CARD) \cite{mcarthur2013comprehensive}, which addresses data concerns by curating a set of known AMR genes.
+
 In addition, GenomeTrackr from the Food and Drug Administration (FDA) \cite{allard2016practical}; GenBank from the National Center for Biotechnology Information (NCBI) \cite{doi:10.1093/nar/gks1195}; and PulseNet from the Center for Disease Control and Prevention (CDC) \cite{swaminathan2001pulsenet}, all enable sharing of genome sequences and associated metadata, either openly or amongst member institutions.
@@
 %MLST software uses a similar reference-based approach \cite{larsen2012multilocus}.
-
-% rewrite of bigdata approach (BIGGER bigdata)
-Bacterial species, pan-genome ...Panseq \cite{laing2010pan}, Roary \cite{page2015roary}, chewBBACA \url{https://github.com/mickaelsilva/chewBBACA}, and Graphtyper \cite{Eggertsson148403} can compute these pan-genomic regions, and are freely available.
-These software programs are capable of processing X number of genomes per Y days at a cost of Z, which would take a A, B and C to do in a wet-lab.
 
@@
 
-\section{METHODS AND IMPLEMENTATION}
+% Some general comments
+% The NAR Database issue is more of a showcase then a rigorous exploration of software design choices.
+% Given this focus, i would suggest the following:
+% 1. Increase/highlight the discriptions of the functions and capabilities, maybe by adding a Functionality section
+% 2. In the Implmentation (or Methods) section, only give a cursory description of the layout and components. Don't need to provide too much justification
+% 3. Use the Results section to highlight the scope/size and speed. This can be short
+% 4.
+In the Discussion, this is where i would expand on the justications and reasons for specific design choices.
+Pick 2-3 main ones and discuss those (i.e.
+don't need to justify our choice of documentation software).
+Also compare with other software in Discussion.
+% 5. Add a conclusions section
+
+\section{FUNCTIONALITY}
+
+% Describe available functions in spfy
+% para covering everything
+Spfy performs reference laboratory tasks: O-antigen and H-antigen typing, Shiga-toxin (STX) typing,  virulence factor (VF) and antimicrobial resistance (AMR) gene determination, and related strain determination.
+Spfy also performs bioinformatics analyses: pangenome generation, statistical significance testing of genome markers for user-defined groups, and AMR predictions using support vector machines (SVMs).
+
+% para covering ectyper & RGI
+To further characterize strains ... serotyping, which detects the presence of specific cell surface antigens.
+O-antigen and H-antigen typing, and VF determination use our ECTyper \url{https://github.com/phac-nml/ecoli_serotyping} software.
+ECTYper compares the user-submitted genomes against a curated set of sequences, an approach originally implemented in the VirulenceFinder \cite{joensen2014real} tool.
+The Comprehensive Antibiotic Resistance Database (CARD) \cite{mcarthur2013comprehensive} maintains a set of known AMR genes and develops the Resistance Gene Identifier (RGI) program.
+Both ECTyper and RGI are based off of BLAST \cite{pmid2231712}.
+
+% para on phylotyper
+Stx typing looks at ...
+Phylotyper constructs phylogenetic trees and uses ancestral reconstruction to determine Stx-types ... others.
+The software can also be used to generate trees and determine closely related strains.
+
+% para on pangenome
+Bacterial species, pan-genome ...Panseq \cite{laing2010pan}, Roary \cite{page2015roary}, chewBBACA \url{https://github.com/mickaelsilva/chewBBACA}, and Graphtyper \cite{Eggertsson148403} can compute these pan-genomic regions, and are freely available.
+Spfy integrates panseq ...
+
+% storage of results
+The results from ECTyper, RGI, Phylotyper, and Panseq are converted into a graph structure and linked the the genomes they originated from.
+
+% use of stored results
+By retrieving results from the database, Spfy provides group comparisons using Fishers, and can build SVMs to predict AMR based off of the presence of particular pangenome regions.
+
+\section{IMPLEMENTATION}
 % para
-% goals: predictive analylitcs in minutes, storage up 50,000-100,000 genomes
+% goals: predictive analylitcs in minutes, storage up 50,000-100,000 genomes {}
 
@@
 
-\subsection{Graph databases and ontologies}
+\subsection{Data Storage}
 % para
@@
 % 1. spfy is built around graph technologies
-% 2. graph dbs store data as a series of links
-% 3. existing graph dbs and bioinformatics implementations
-% para
 % 1. how we structure our graph
@@
 % 3 1/2. SPARQL queries
-% 4. results/timings: scale
 
@@
 % 1. goals: intuitive/familar, ease of use 
-% 2. design specs and different examples
+% 2. design specs
 % 3. Google Material design
-% 4. some focus on card based design
-% para
 % 1. implementation: reactjs, react-md, ES6, JSX
-% 2. A to B: example of how users go from uploading files to results
-% 3. backwards compatibility with older browsers: transpile via Babel
 % 4. separation from Flask layer
@@
 
-\subsection{Real-time analysis pipelines and task distribution}
 \subsection{Real-time analysis pipelines}
@@
 %	related: packaging of modules in conda
-% 2. different pipelines available
-% 3. how pipelines are started from the front-end
-% 4. result storage in graph db
 
@@
 
-\subsubsection{Task distribution and parallelization}
 % para
 % 1. goals: scale analyses to "big-data", error handling
-% 2. examples of different ways parallelization has been handled in the past
 % 3. how we handle parallelization with RQ
@@
 
-\subsection{Containerization}
 % para
 % 1. goals: why compartmentilizations
-% 2. examples of different virtualization technologies
-% 3. intro do docker
-% para
 % 1. how we implemented docker
-% 2. use of docker-compose
 % 3. how this lets us replicate worker containers and link everything together
@@
 \subsection{Continuous integration and testing}
+% Keep this short
 % para
 % 1. goals: why CI, testing is important
-% 2. examples of different testing platforms
-% 3. how we've implemented it, integration with github
+% 2. how we've implemented it, integration with github
 
@@
 
-\subsection{Guides and documentation}
-% para
-% 1. goals: use guide (ie. the README), dev guide
-% 2. how docs have been done in the past (in general)
-% 3. our use of ReadtheDocs
+\section{RESULTS}
 
-\section{RESULTS AND DISCUSSION}
+% e.g.
+% database statistics
+% analysis run-time / throughput with different levels of parallelization
 
-\subsection{Further containerizing analyses modules}
+\section{DISCUSSION}
 
-% ex. in python, tox can be set to spin up docker containers for testing - see certbot
-% could spin up a new container for each analysis module to run see \cite{di2015impact} - neglible performance impact
+% Mention where users / developers can find documentation
 
@@
 % we have a lot of similar directions to this, but just try to make it prettier, more friendly to use, and faster
+
+\section{CONCLUSIONS}
 

### 2017-08-11 paper-webserver.tex
UPDATE: moved old material into relevant sections, deleted irrelevant material

@@
 % end of new intro
-% below material is just old stuff, might or might not be used elsewhere
-
-% costs of bacterial infections
-% We need to state the "big problem" up-front that we are working on, which is not really tracking foodborne illness.
-It is fast analyses of bacterial whole-genomes, providing reference-laboratory tests from in-silico data, predicting phenotype from genotype, and providing a linked database of known genomes that don't need to be re-computed.
-This can be used in the tracking of foodborne illness, but is more an application of the problem we are working on.
-The isolation and characterization of bacterial pathogens are critical for Public Health laboratories to rapidly respond to outbreaks, and to effectively monitor known and emerging pathogens through surveillance programs.
-Within Canada it is estimated that over 4,000 hospitalizations and 100 deaths occur due to bacterial infections \cite{thomas2015estimates}; in the United States, over 55,000 hospitalizations and 1,300 deaths occur per year \cite{scallan2011foodborne}.
-Within North America the total economic burden of these illnesses is estimated to be ACTUAL NUMBER \cite{todd1989costs,scharff2010health}.
-The bacterial species that are most frequently associated with severe disease, and cause the majority of this health burden in (Canada / NA/ Europe / Worldwide ) are ABC.
-%You could combine the Canada / US numbers into a North America number, and include world-wide data as well.
-Always avoid vacuous descriptors like "many", "various" -- use an actual quantifier.
-Eg.
-Most people prefer vanilla to chocolate.
-Fifty-one percent of people prefer vanilla to chocolate.
-Ninety-nine percent of people prefer vanilla to chocolate.
-
-% routinely monitor emerging threats, of which \textit{Salmonella enterica} and \textit{Escherichia coli} are amongst the two most common \cite{kozak2013foodborne}.
-In \textit{Salmonella enterica} \cite{bell2016recent}, \textit{Escherichia coli} \cite{fratamico2016advances}, and (others pathogens --> either state the other pathogens or leave it out entirely) \cite{ronholm2016navigating}
-
-% surveillance
-Until recently, Public-health agencies relied on laboratory tests such as XYZ to characterize bacterial isolates in outbreak and surveillance settings.
-The previous gold-standard in determining strain relatedness was pulsed-field gel electrophoresis (PFGE) {ronholm2016navigating}, which uses rare-cutting restriction enzymes to produce a unique banding pattern for each strain.
-However, in \textit{Enterococcus faecium}, PFGE has been shown to misclassify 9 of 132 isolates, when compared to whole-genome sequencing (WGS) based discrimination \cite{pinholt2015multiple}.
-In \textit{Klebsiella pneumoniae} \cite{marsh2015genomic}, \textit{Yersinia enterocolitica} \cite{gilpin2014limitations}, and \textit{Staphylococcus aureus} \cite{doi:10.1093/ofid/ofu096}, WGS was used to discriminate isolates after initial clustering by PFGE resulted in indistinguishable samples.
-Examination of PFGE bands are also subjective, difficult to share \cite{lytsy2017time}, and collative platforms such as PulseNet reported \cite{gilpin2014limitations} that even after collecting 72\% of \textit{Campylobacter jejuni} in a given year in Minnesota (673 cases), 87\% of isolates could not be linked by PFGE pattern.
-Antimicrobial resistance testing, virulence factor testing ...
-However, current efforts are focused on predictive genomics, where the relevant phenotypic information can be determined through examination of the whole-genome sequence.
-
-% potential of WGS
-, and as such can be used to evaluate the spread of outbreaks with better resolution and context than traditional methods \cite{ronholm2016navigating}.
-However, while these WGS-based methods can fully examine the underlying genetics, the expressed phenotype is more difficult to predict.
-For example, predictions involving reference databases are only as good as the databases themselves.
-
-% drawbacks - big data
-Many previous bioinformatics software programs have been developed \textit{ad hoc}, without the use of software engineering principles \cite{de2015trends}.
-Such tools were often script-based, with custom data formats, and only suitable for small collections of data \cite{de2015trends}.
-While this was acceptable for smaller analyses, bioinformatic pipelines utilizing WGS data are larger and involve linked dependencies, which require the application of systems engineering principles \cite{schatz2015biological}.
-Additionally, many subsets of biology now require the analyses of big-data, where the ability to perform computations in real-time, store data in flexible databases, and utilize a common application programming interface (API) linking resources are required \cite{swaminathan2016review}.
-
-% Solution via software engineering
-
-In addition, GenomeTrackr from the Food and Drug Administration (FDA) \cite{allard2016practical}; GenBank from the National Center for Biotechnology Information (NCBI) \cite{doi:10.1093/nar/gks1195}; and PulseNet from the Center for Disease Control and Prevention (CDC) \cite{swaminathan2001pulsenet}, all enable sharing of genome sequences and associated metadata, either openly or amongst member institutions.
-
-%By comparing a pre-collected set of reference gene sequences against a WGS result, bioinformatics software predicts wet-lab serotyping results \cite{whiteside2016superphy}, Virulence Factor (VF) results \cite{joensen2014real}, and antimicrobial resistance (AMR) results \cite{mcarthur2013comprehensive}.
-%MLST software uses a similar reference-based approach \cite{larsen2012multilocus}.
-
-While these programs are useful for computing novel analyses, storing results in a tiered structure, with rapid access available for retrieval while maintaining long-term data storage \cite{schatz2015biological}, is needed to prevent the constant re-computation of the same data, and to provide pre-computed analyses for common predictive genomic analyses needed by the public health, and research community \cite{de2015trends}.
-
-% why Superphy
-Currently, a number of organizations are creating user-friendly web services for common analyses of bacterial genomes.
-These include: NAME \cite{joensen2014real,larsen2012multilocus}, NAME \cite{liu2016construction}, and NAME \cite{hasman2015detection}, which do X, Y and Z respectively.
-
-While useful, leveraging big-data results from WGS is difficult due to the lack of bioinformatic standards and common infrastructure technologies.
-Integrating results from multiple genomes, and examining shared connections, is an area that currently requires more research \cite{fricke2014bacterial},
-
-% Superphy
-, which provided pre-computed, predictive genomic analyses for determining Shiga toxin subtype, AMR gene profiles, the presence of known virulence factors, and provides predictive biomarkers, both the presence / absence of genomic regions and single nucleotide variants, for bacterial sub-groups based on metadata such as serotype, host source, or location.
-
-Lastly, these services are containerized using Docker to mitigate the complexities of the underlying bioinformatic pipelines.
-All services are publicly available at \url{https://lfz.corefacility.ca/superphy/spfy/}.
 
@@
 % goals: predictive analylitcs in minutes, storage up 50,000-100,000 genomes {}
+% drawbacks - big data
+Many previous bioinformatics software programs have been developed \textit{ad hoc}, without the use of software engineering principles \cite{de2015trends}.
+Such tools were often script-based, with custom data formats, and only suitable for small collections of data \cite{de2015trends}.
+While this was acceptable for smaller analyses, bioinformatic pipelines utilizing WGS data are larger and involve linked dependencies, which require the application of systems engineering principles \cite{schatz2015biological}.
+Additionally, many subsets of biology now require the analyses of big-data, where the ability to perform computations in real-time, store data in flexible databases, and utilize a common application programming interface (API) linking resources are required \cite{swaminathan2016review}.
 
@@
 
-\begin{figure}[t]
-\begin{center}
-\includegraphics{NAR-fig1.eps}
-\end{center}
-\caption{Caption for figure within column.}
-\label{NAR-fig1}
-\end{figure}
+We developed Spfy using continuous integration (CI) techniques: all changes and update to the codebase is automatically ran against a series of tests ensuring functionality and backwards compatibility.
+We chose TravisCI \url{travisci.io} which offers free services for open-source projects.
+The individual tests use PyTest.
 
@@
 % database statistics
+We tested spfy with 55,353 samples genomes (267GB), storing both the entire genomes and results for all included analyses modules.
+The resulting database had XYZ nodes and XYZ edges, with XYZ object properties.
+This worked our to XYZ GB of data stored.
+
 % analysis run-time / throughput with different levels of parallelization
@@
 % Mention where users / developers can find documentation
+A guide to setting up Spfy is provided at \url{https://github.com/superphy/backend} and a developer guide is provided at \url{https://superphy.readthedocs.io/en/latest/}.
 
@@
 % focus on application
+The isolation and characterization of bacterial pathogens are critical for Public Health laboratories to rapidly respond to outbreaks, and to effectively monitor known and emerging pathogens through surveillance programs.
+Until recently, Public-health agencies relied on laboratory tests such as XYZ to characterize bacterial isolates in outbreak and surveillance settings.
+The previous gold-standard in determining strain relatedness was pulsed-field gel electrophoresis (PFGE) {ronholm2016navigating}, which uses rare-cutting restriction enzymes to produce a unique banding pattern for each strain.
+However, in \textit{Enterococcus faecium}, PFGE has been shown to misclassify 9 of 132 isolates, when compared to whole-genome sequencing (WGS) based discrimination \cite{pinholt2015multiple}.
+In \textit{Klebsiella pneumoniae} \cite{marsh2015genomic}, \textit{Yersinia enterocolitica} \cite{gilpin2014limitations}, and \textit{Staphylococcus aureus} \cite{doi:10.1093/ofid/ofu096}, WGS was used to discriminate isolates after initial clustering by PFGE resulted in indistinguishable samples.
+Examination of PFGE bands are also subjective, difficult to share \cite{lytsy2017time}, and collative platforms such as PulseNet reported \cite{gilpin2014limitations} that even after collecting 72\% of \textit{Campylobacter jejuni} in a given year in Minnesota (673 cases), 87\% of isolates could not be linked by PFGE pattern.
+Antimicrobial resistance testing, virulence factor testing ...
+However, current efforts are focused on predictive genomics, where the relevant phenotypic information can be determined through examination of the whole-genome sequence. 
+, and as such can be used to evaluate the spread of outbreaks with better resolution and context than traditional methods \cite{ronholm2016navigating}.
+
+% application: results similar to a wet-lab
+Spfy uses WGS results.
+After intial sequencing of new isolates, Spfy can be used in place of a traditional reference laborartory, to determine the O-type and H-type, Stx type, and all known VFs and AMR genes in real-time.
+These results can be shared with other ageencies and researchers over the internet.
+Futhermore, using Spfy's database of pre-processed genomes, Spfy can determine all strains a sample may be related to which is useful for ...
+Cost benefits ...
 
@@
 % how costs can be saved because using standard web tech means you can deploy to different cloud computing services
+Spfy's common infrastructure allows the platform to be deployed to multiple cloud computing services, without need to modify the code.
+Docker containerization is supported by Amazon AWS, Google Cloud, ...
+Neglibble impact of Docker containers on performance \cite{di2015impact}.
+Costs ...
 
@@
 % namely galaxy
+Other pipeline technologies such as Galaxy \cite{goecks2010galaxy}, Kepler ... \cite{de2015trends}.
+important to have rapid access available for retrieval while maintaining long-term data storage \cite{schatz2015biological}
+long term storage is needed to prevent the constant re-computation of the same data, and to provide pre-computed analyses for common predictive genomic analyses needed by the public health, and research community \cite{de2015trends}.
+Integrating results from multiple genomes, and examining shared connections, is an area that currently requires more research \cite{fricke2014bacterial},
 
@@
 \section{CONCLUSIONS}
-
-\begin{table}[b]
-\tableparts{%
-\caption{This is a table caption}
-\label{table:01}%
-}{%
-\begin{tabular*}{\columnwidth}{@{}lllll@{}}
-\toprule
-Col. head 1 & Col. head 2 & Col. head 3 & Col. head 4 & Col. head 5
-\\
-& (\%) & (s$^{-1}$) & (\%) & (s$^{-1}$)
-\\
-\colrule
-Row 1 & Row 1 & Row 1 & -- & --
-\\
-Row 2 & Row 2 & Row 2 & Row 2 & Row 2
-\\
-\botrule
-\end{tabular*}%
-}
-{This is a table footnote}
-\end{table}
-
-\begin{figure*}[t]
-\begin{center}
-\includegraphics{NAR-fig2.eps}
-\end{center}
-\caption{Caption for wide figure over two columns.
-\textbf{(a)} Left figure.
-\textbf{(b)} Right figure (see (a)).
-}
-\label{NAR-fig2}
-\end{figure*}
-
-\section{CONCLUSION}
 

### 2017-08-11 paper-webserver.tex
ADD: some restructuring

@@
 % goals: predictive analylitcs in minutes, storage up 50,000-100,000 genomes {}
-% drawbacks - big data
-Many previous bioinformatics software programs have been developed \textit{ad hoc}, without the use of software engineering principles \cite{de2015trends}.
-Such tools were often script-based, with custom data formats, and only suitable for small collections of data \cite{de2015trends}.
-While this was acceptable for smaller analyses, bioinformatic pipelines utilizing WGS data are larger and involve linked dependencies, which require the application of systems engineering principles \cite{schatz2015biological}.
-Additionally, many subsets of biology now require the analyses of big-data, where the ability to perform computations in real-time, store data in flexible databases, and utilize a common application programming interface (API) linking resources are required \cite{swaminathan2016review}.
 
@@
 
+% drawbacks - big data
+Many previous bioinformatics software programs have been developed \textit{ad hoc}, without the use of software engineering principles \cite{de2015trends}.
+Such tools were often script-based, with custom data formats, and only suitable for small collections of data \cite{de2015trends}.
+While this was acceptable for smaller analyses, bioinformatic pipelines utilizing WGS data are larger and involve linked dependencies, which require the application of systems engineering principles \cite{schatz2015biological}.
+Additionally, many subsets of biology now require the analyses of big-data, where the ability to perform computations in real-time, store data in flexible databases, and utilize a common application programming interface (API) linking resources are required \cite{swaminathan2016review}.
+
+One of the key goals in developing Spfy is to maintain instantaneity: modern websites have accustomed users to instant results.
+We attempt to use innovations in web development and bring a similar experience to Spfy as a predictive genomics platform for \textit{E. coli}.
 % Mention where users / developers can find documentation
-A guide to setting up Spfy is provided at \url{https://github.com/superphy/backend} and a developer guide is provided at \url{https://superphy.readthedocs.io/en/latest/}.
+Spfy's main documentation and codebase are provided at \url{https://github.com/superphy/backend} and a developer guide is provided at \url{https://superphy.readthedocs.io/en/latest/}.
 

### 2017-08-17 paper-webserver.tex
ADD: new first para for implementation

@@
 \section{IMPLEMENTATION}
-% para
-% goals: predictive analylitcs in minutes, storage up 50,000-100,000 genomes {}
-
-% Inputs: more like going from file -> analysis -> graph db (not a hard subsection)
-Spfy takes assembled genomes as input, runs selected analyses modules, and converts the results into a graph structure for upload.
-
-% Outputs: more like what we query from blazegraph and how we use it (not a hard subsection)
-Spfy queries the graph database to run analyses XYZ and build visualizations.
+% para: the semantic web
+Spfy is built around semantic web technologies which focuses on describing the relaions between different datum \cite{berners2001semantic}.
+In biological data, a semantic web focus describes individual data points by the type, for example as a genome, contiguous DNA sequence, or gene, and then link related data together in a queryable graph.
+Semantic web technlogies allow new, not previously described data to be seemlessly incorporated to the existing graph, and has been proposed as a common standard for the open sharing of data \cite{horrocks2005semantic}.
 

### 2017-08-17 paper-webserver.tex
ADD: second para for implementation

@@
 
+% para: semantic web in spfy
+In Spfy, the entirety of SuperPhy's previous code is replaced with methods for handling semantic web technologies.
+Generalized functions are used to convert the results from different analysis modules into a graph object, and that graph object is used to update the main graph database; information contained in the WGS file is also converted into a graph object and stored.
+The use of semantic web technologies allows results to be linked to the genomes they were computed from, and all data points to share a common datastructure that can be queried together.
+
 \subsection{Data Storage}
@@
 
-Graph databases focus on describing the relationships between different data, and is one of the emerging \cite{de2015trends} database types used for biological data.
+Graph databases focus on describing the relationships between different data points, and is one of the emerging \cite{de2015trends} database types used for biological data.
 Spfy

### 2017-08-17 paper-webserver.tex
ADD: 3rd intro para to implementation

@@
 The use of semantic web technologies allows results to be linked to the genomes they were computed from, and all data points to share a common datastructure that can be queried together.
+
+% para: intro to the spfy stack
+Spfy's server-side code is developed in Python and the website served to users is developed using the React JavaScript library.
+When users upload genomes for analysis: (i) they first make the upload through the ReactJS-based website. 
+The public web service accepts up to 200 MB of genome files (50 genomes compressed, or 120 genomes compressed).
+(ii) When files are uploaded, the user-selected analysis options are enqueued into the Redis Queue \url{http://python-rq.org/} task queue.
+Redis Queue consists of a Redis Database \url{https://redis.io/} and task queue workers which run as Python processes.
+(iii) The workers dequeue the various analyses and run them in parallel and results are stored back in the Redis database.
+(iv) Python functions are then used to parse the results and permanently store them in Blazegraph \url{https://www.blazegraph.com/} - a graph database.
+The entire Spfy platform is packaged as a series of Docker \url{https://www.docker.com/} containers and orchestrated using Docker Compose \url{https://docs.docker.com/compose/}.
 

### 2017-08-17 paper-webserver.tex
CHANGE: the functionality para should only cover the analysis modules, not data storage

@@
 \section{FUNCTIONALITY}
-
+% ONLY FOCUS ON THE ANALYSIS MODULES
 % Describe available functions in spfy
@@
 
-% storage of results
-The results from ECTyper, RGI, Phylotyper, and Panseq are converted into a graph structure and linked the the genomes they originated from.
+% group comparisons
+By retrieving results from the database, Spfy provides group comparisons using Fishers.
+...
 
-% use of stored results
-By retrieving results from the database, Spfy provides group comparisons using Fishers, and can build SVMs to predict AMR based off of the presence of particular pangenome regions.
+% Panpredic???
 

### 2017-08-17 paper-webserver.tex
ADD: something on data storage

@@
 
-Graph databases focus on describing the relationships between different data points, and is one of the emerging \cite{de2015trends} database types used for biological data.
-Spfy
-
-Serotyping, VF, AMR predictions are computed within minutes, and results are efficiently stored within the graph.
+Graph databases focus on describing the relationships between different data points, is one of the emerging \cite{de2015trends} database types used for biological data, and is one of the core building blocks of a semantic web stack \cite{horrocks2005semantic}.
+Spfy uses the RDFLib Python library \url{https://rdflib.readthedocs.org/} to represent all data meant for long-term storage.
+When serotyping, VF, AMR predictions, and pangenome generation tasks are completed, the results are stored within the graph database.
 

### 2017-08-18 paper-webserver.tex
ADD: more on data storage
ADD: another para on data storage

@@
 % 3 1/2. SPARQL queries
-
-Graph databases focus on describing the relationships between different data points, is one of the emerging \cite{de2015trends} database types used for biological data, and is one of the core building blocks of a semantic web stack \cite{horrocks2005semantic}.
+Graph databases focus on describing relationships between different data points, is one of the emerging \cite{de2015trends} database types used for biological data, and one of the core building blocks of a semantic web technology stack \cite{horrocks2005semantic}.
 Spfy uses the RDFLib Python library \url{https://rdflib.readthedocs.org/} to represent all data meant for long-term storage.
 When serotyping, VF, AMR predictions, and pangenome generation tasks are completed, the results are stored within the graph database.
+% note: below is a feature in development
+The permanent storage of results serve as a one-time cost, and allows population-wide analyses of all stored genomes; result storage also enables Spfy to avoid recomputation when the same analysis is re-run.
+
+% para
+% ontology
+Ontologies are indices which define types of data and the relations between them.
+Spfy describes genomes as a collection of contiguous sequences (contigs), and connects the predicted genes and pangenome regions to the contigs they were found on.
+(see Figure \ref{fig-ontology})
+Individual genes and pangenome regions are connected to all the contigs they are found on, such that it is possible to infer all the genomes which contain a particular target sequence.
+
+\begin{figure}[t]
+\begin{center}
+\includegraphics{images/spfy_ontology-1.svg}
+\end{center}
+\caption{Caption for figure within column.}
+\label{fig-ontology}
+\end{figure}
+
+blah...
 
@@
 Everything is compartmentalized within separate Docker containers where the front-end is networked, through Docker-Compose, to the back-end webserver.
+
+(see Figure \ref{fig-docker})
+
+\begin{figure}[t]
+\begin{center}
+\includegraphics{images/docker.svg}
+\end{center}
+\caption{Caption for figure within column.}
+\label{fig-docker}
+\end{figure}
 

### 2017-08-18 paper-webserver.tex
ADD: another para on data storage
ADD: mention  the FALDO and GenEpiO ontologies

@@
 Ontologies are indices which define types of data and the relations between them.
-Spfy describes genomes as a collection of contiguous sequences (contigs), and connects the predicted genes and pangenome regions to the contigs they were found on.
+By sharing ontologies, it is possible to link sources of data from different organizations together, and Spfy uses types defined in the  Genomic Epidemiology Application Ontology (GenEpiO) (\cite{griffiths2017context} and Feature Annotation Location Description Ontology (FALDO) \cite{bolleman2016faldo}.
+Spfy also connects genomes, contiguous sequences (contigs), pangenome regions, and predicted genes and alleles in novel ways; to allow the presence of various genes to be infered from a genome description, realtionship links are generically named.
 (see Figure \ref{fig-ontology})
-Individual genes and pangenome regions are connected to all the contigs they are found on, such that it is possible to infer all the genomes which contain a particular target sequence.
+Individual genes and pangenome regions are connected to all the contigs they are found on using this inferencing method, such that it is possible to determine all the genomes which contain a particular target sequence.
 
@@
 \end{figure}
-
-blah...
 

### 2017-08-18 paper-webserver.tex
ADD: main points for the web design para, need to rewrite though

@@
 
-We designed Spfy's front-end following the Material Design standards \url{https://material.io/}, released by Google.
-The user interface is implemented with the React JavaScript library \url{https://facebook.github.io/react/}, by Facebook, as a single-page application to allow efficient data-flow without reloading the website.
+The front-end website displayed to users is written using the React JavaScript library \url{https://facebook.github.io/react/} as a single-page application to allow efficient data-flow without reloading the website.
+To ensure a familiar user interface, we followed the Material Design specification \url{https://material.io/}, published by Google, and follow a card-based design.
+(see Figure \ref{fig-results})
+While data storage is graph-based, the results of various analysis modules are presented to users in a familiar tabular structure and available for download.
+(see Figure \ref{fig-tables})
+
+\begin{figure}[t]
+\begin{center}
+\includegraphics{images/results.png}
+\end{center}
+\caption{Caption for figure within column.}
+\label{fig-results}
+\end{figure}
+
+\begin{figure}[t]
+\begin{center}
+\includegraphics{images/tables.png}
+\end{center}
+\caption{Caption for figure within column.}
+\label{fig-tables}
+\end{figure}
 

### 2017-08-18 paper-webserver.tex
ADD: new para for pipelines

@@
 
-Spfy enables processing of thousands of genome sequences using multiple modules.
-By using task queue workers, enabled by the Python-based Redis Queue library \url{https://github.com/nvie/rq}, performance can quickly scale to available infrastructure.
-As a bioinformatics tool, we ensured reliability of the platform by integrating the open-source Sentry toolkit \url{https://github.com/getsentry/sentry} for real-time exception tracking.
-Containers are orchestrated through Docker-Compose allowing the entire platform: webservers, databases, and task workers, to be easily replicated by other researchers.
+Task queues are processes which schedule code exection across available computing infrastructure.
+In Spfy, the Python-based Redis Queue library \url{https://github.com/nvie/rq} is used to schedule analysis tasks and have them run asynchronously in response to user requests.
+When a user submits files for analysis or population-wide anlayses, separate tasks are enqueued at different priorities, depending on when users might expect the result to be returned.
+For example, population-wide analyses have a higher priority than bulk subtyping analyses, as we expect day-to-day web searches (for example, on Google) to respond instantly whereas a delay in subtyping is typical in similar web services such as the Center for Genomic Epidemiology Pipeline (CGE Pipeline) \cite{thomsen2016bacterial}.
+Spfy enables processing of thousands of genome sequences by using task queue workers running in parallel, which also allows performance to scale to available infrastructure.
+To ensure reliability of the platform, the open-source Sentry toolkit \url{https://github.com/getsentry/sentry} was integrated and used for real-time exception tracking.
 
@@
 
+Containers are orchestrated through Docker-Compose allowing the entire platform: webservers, databases, and task workers, to be easily replicated by other researchers.
 Everything is compartmentalized within separate Docker containers where the front-end is networked, through Docker-Compose, to the back-end webserver.

### 2017-08-21 paper-webserver.tex
ADD: on docker

@@
 
-Containers are orchestrated through Docker-Compose allowing the entire platform: webservers, databases, and task workers, to be easily replicated by other researchers.
-Everything is compartmentalized within separate Docker containers where the front-end is networked, through Docker-Compose, to the back-end webserver.
-
+The Spfy platform depends on a series of webservers, databases, and task workers and uses \url{https://www.docker.com/}, a virtualization software package, to run self-contained operating systems on the same host computer.
+Software packages are installed within the conatiners and the entire platform is networked together using Docker-Compose \url{https://docs.docker.com/compose/}.
 (see Figure \ref{fig-docker})
+Docker integration ensures that software dependencies, which typically must be manually installed \cite{doi:10.1093/bioinformatics/btu153,laing2010pan,inouye2014srst2,naccache2014cloud}, are handled automatically and that runtimes in one service are separated from another service.
 
@@
 \end{figure}
-
-Docker integration ensures that software dependencies, which typically must be manually installed \cite{doi:10.1093/bioinformatics/btu153,laing2010pan,inouye2014srst2,naccache2014cloud}, are handled automatically.
 

### 2017-08-22 paper-webserver.tex
ADD: on BAP

@@
 
-We developed Spfy using continuous integration (CI) techniques: all changes and update to the codebase is automatically ran against a series of tests ensuring functionality and backwards compatibility.
-We chose TravisCI \url{travisci.io} which offers free services for open-source projects.
-The individual tests use PyTest.
+TravisCI \url{travisci.io}, a continuous integration (CI) platform, is integrated into Spfy's Github repository \url{https://github.com/superphy/backend} and runs tests ensuring functionality and backwards compatibility with any changes to the codebase.
+The individual tests use PyTest \url{https://doc.pytest.org/} and are run within TravisCI's virtual environment.
+The current build status can be checked either on our Gtihub repository or at \url{https://travis-ci.org/superphy/backend}.
 
@@
 % we have a lot of similar directions to this, but just try to make it prettier, more friendly to use, and faster
+The Bacterial Analysis Platform (BAP), developed out of the Technical University of Denmark, provides and integrated analysis pipeline for bacterial WGS data, as a web service.
+BAP is novel in its combined approach to genome analysis as different programs, such as for VF and AMR determination, are included by default into the pipeline; Spfy provides similar functionallity with an expanded focus on integrated result storage and big-data analyses.
+In place of a MySQL database for storing the location of result files, Spfy parses result data into graph objects and integrates results into a persistant graph database.
+This enables Spfy to create large datasets where analysis results are linked to genomes, and to perform population-wide analyses.
+On run-times of analogous algorithmns for prediction tasks, Spfy performs at a similar speed to BAP, though Spfy does not provide genome assembly services.
+Spfy also uses a Python-based backend instead of PHP, and Spfy's use of Docker containerization and task queues is a novel as well.
 

### 2017-08-23 paper-webserver.tex
ADD: more on comparisons

@@
 % namely galaxy
-Other pipeline technologies such as Galaxy \cite{goecks2010galaxy}, Kepler ... \cite{de2015trends}.
+Other scientific workflow technologies such as Galaxy \cite{goecks2010galaxy}, Kepler \cite{ludascher2006scientific}, and Taverna \cite{oinn2004taverna} have been applied to bioinformatic tasks.
+Galaxy ...
+Taverna ...
+Kepler ...
+To approach the challenge of integrating different bioinformatics programs, Spfy instead uses technologies prevalent in common web services not necessarily related to scientific workflows.
+While we prefer to package individual bioinformatics programs using language-agnostic package managers such as Conda \url{https://conda.io/}, we do agree that suites of bioinformatics software benefits from having all dependencies installed and deployable in Docker containers \cite{di2015impact}.
 important to have rapid access available for retrieval while maintaining long-term data storage \cite{schatz2015biological}
@@
 This enables Spfy to create large datasets where analysis results are linked to genomes, and to perform population-wide analyses.
-On run-times of analogous algorithmns for prediction tasks, Spfy performs at a similar speed to BAP, though Spfy does not provide genome assembly services.
-Spfy also uses a Python-based backend instead of PHP, and Spfy's use of Docker containerization and task queues is a novel as well.
+On a per file basis, Spfy performs at a similar speed to BAP on predictive genomics tasks, though Spfy does not provide genome assembly services.
+Spfy also processes XXXX files over XX tasks in XXXX time by using a novel approach of distributing computations over a task queue and multiple Docker compartmentalized containers.
 

### 2017-08-23 paper-webserver.tex
CHANGE: integrate the cloud compatibility benefit with the comparison with galaxy

@@
 
-\subsection{Deployment to Cloud-Computing Platforms}
-% how costs can be saved because using standard web tech means you can deploy to different cloud computing services
-Spfy's common infrastructure allows the platform to be deployed to multiple cloud computing services, without need to modify the code.
-Docker containerization is supported by Amazon AWS, Google Cloud, ...
-Neglibble impact of Docker containers on performance \cite{di2015impact}.
-Costs ...
-
 \subsection{Comparison with other bioinformatic pipeline technologies}
@@
 To approach the challenge of integrating different bioinformatics programs, Spfy instead uses technologies prevalent in common web services not necessarily related to scientific workflows.
-While we prefer to package individual bioinformatics programs using language-agnostic package managers such as Conda \url{https://conda.io/}, we do agree that suites of bioinformatics software benefits from having all dependencies installed and deployable in Docker containers \cite{di2015impact}.
-important to have rapid access available for retrieval while maintaining long-term data storage \cite{schatz2015biological}
-long term storage is needed to prevent the constant re-computation of the same data, and to provide pre-computed analyses for common predictive genomic analyses needed by the public health, and research community \cite{de2015trends}.
-Integrating results from multiple genomes, and examining shared connections, is an area that currently requires more research \cite{fricke2014bacterial},
+While we prefer to package individual bioinformatics programs using language-agnostic package managers such as Conda \url{https://conda.io/}, we do agree that suites of bioinformatics software benefit from having all dependencies installed and deployable in Docker containers \cite{di2015impact}.
+These programs are connected to form a workflow using task queues; a design methodology used to offload long-running or non-instantaneous tasks over available computing infrastructure.
+Unlike platforms such as Galaxy which favor a drag-and-drop design for building workflows, Spfy favors a tightly coupled approach explicitedy linking together different bioinformatics programs within the task queue implemntation.
+In particular, Spfy uses asynchronous tasks queues which are tripped in response to user requests, thereby allowing immediate responsivity to users on the website, and processes generating human-readable results can be separated from processes handling long-term result storage.
+
+% standard web tech means you can deploy to different cloud computing services
+One of the key benefits of using more common-place technologies is the compatibility with other infrastructure resources.
+Docker containerization is widely supported by cloud computing services: Amazon Web Services (AWS) \url{https://aws.amazon.com/docker/}, Google Cloud Platform (GCloud) \url{https://cloud.google.com/container-engine/}, and Microsoft Azure \url{https://azure.microsoft.com/en-us/services/container-service/}, and self-hosted cloud computing technologies such as OpenStack \url{https://wiki.openstack.org/wiki/Docker}.
+% we may want to sign up for a free trial to give an example of this
+Spfy packages compute nodes, which manage a collection of task queue workers, as reproducible Docker containers which can be networked together to form a compute cluster.
+Docker containerization has a neglible impact on performance \cite{di2015impact}, and allows the platform to easily scale to demand.
+In contrast to Galaxy, Spfy is a targeted approach building a predictive analytics platform for \textit{E.
+coli} which leverages generic web technologies to favor a responsive, big-data design strategy tackling the challenge \cite{fricke2014bacterial} of integrating results from multiple genomes and examining shared connections.
 

### 2017-08-25 paper-webserver.tex
ADD: one line
ADD: general edits

@@
 Spfy also performs bioinformatics analyses: pangenome generation, statistical significance testing of genome markers for user-defined groups, and AMR predictions using support vector machines (SVMs).
+We introduce one new bioinformatics program, Ecoli-Serotyper (ECTyper) \url{https://github.com/phac-nml/ecoli_serotyping}, for determining surface-antigen type and associated VFs, and utilize existing software: the Resistance Gene Identifier (RGI) program \cite{mcarthur2013comprehensive}, Phylotyper \cite{whiteside2017phylotyper}, and Panseq \cite{laing2010pan}.
 
@@
 To further characterize strains ... serotyping, which detects the presence of specific cell surface antigens.
-O-antigen and H-antigen typing, and VF determination use our ECTyper \url{https://github.com/phac-nml/ecoli_serotyping} software.
+O-antigen and H-antigen typing, and VF determination use our ECTyper software.
 ECTYper compares the user-submitted genomes against a curated set of sequences, an approach originally implemented in the VirulenceFinder \cite{joensen2014real} tool.
-The Comprehensive Antibiotic Resistance Database (CARD) \cite{mcarthur2013comprehensive} maintains a set of known AMR genes and develops the Resistance Gene Identifier (RGI) program.
+The Comprehensive Antibiotic Resistance Database (CARD) \cite{mcarthur2013comprehensive} maintains a set of known AMR genes and develops the RGI program .
 Both ECTyper and RGI are based off of BLAST \cite{pmid2231712}.
@@
 % 3 1/2. SPARQL queries
-Graph databases focus on describing relationships between different data points, is one of the emerging \cite{de2015trends} database types used for biological data, and one of the core building blocks of a semantic web technology stack \cite{horrocks2005semantic}.
+Graph databases describe relationships between different data points, is one of the emerging \cite{de2015trends} database types used for biological data, and is one of the core building blocks of a semantic web technology stack \cite{horrocks2005semantic}.
 Spfy uses the RDFLib Python library \url{https://rdflib.readthedocs.org/} to represent all data meant for long-term storage.
@@
 
-The front-end website displayed to users is written using the React JavaScript library \url{https://facebook.github.io/react/} as a single-page application to allow efficient data-flow without reloading the website.
-To ensure a familiar user interface, we followed the Material Design specification \url{https://material.io/}, published by Google, and follow a card-based design.
+The front-end website is written using the React JavaScript library \url{https://facebook.github.io/react/} as a single-page application to allow efficient data-flow without reloading the website.
+To ensure a familiar user interface, we followed the Material Design specification \url{https://material.io/}, published by Google, surrounding a card-based design.
 (see Figure \ref{fig-results})
-While data storage is graph-based, the results of various analysis modules are presented to users in a familiar tabular structure and available for download.
+Both the task selection and results display follow this card-based design: while data storage is actually graph-based, the results of various analysis modules are presented to users in a familiar tabular structure and available for download.
 (see Figure \ref{fig-tables})
@@
 Task queues are processes which schedule code exection across available computing infrastructure.
-In Spfy, the Python-based Redis Queue library \url{https://github.com/nvie/rq} is used to schedule analysis tasks and have them run asynchronously in response to user requests.
+In Spfy, the Python-based Redis Queue library \url{https://github.com/nvie/rq} is used to manage analysis tasks and have them run asynchronously in response to user requests.
 When a user submits files for analysis or population-wide anlayses, separate tasks are enqueued at different priorities, depending on when users might expect the result to be returned.
-For example, population-wide analyses have a higher priority than bulk subtyping analyses, as we expect day-to-day web searches (for example, on Google) to respond instantly whereas a delay in subtyping is typical in similar web services such as the Center for Genomic Epidemiology Pipeline (CGE Pipeline) \cite{thomsen2016bacterial}.
+For example, population-wide analyses have a higher priority than bulk subtyping analyses, as we want search-response queries to respond instantly whereas a delay in subtyping is typical in similar web services such as the Center for Genomic Epidemiology Pipeline (CGE Pipeline) \cite{thomsen2016bacterial}.
 Spfy enables processing of thousands of genome sequences by using task queue workers running in parallel, which also allows performance to scale to available infrastructure.
@@
 
-TravisCI \url{travisci.io}, a continuous integration (CI) platform, is integrated into Spfy's Github repository \url{https://github.com/superphy/backend} and runs tests ensuring functionality and backwards compatibility with any changes to the codebase.
-The individual tests use PyTest \url{https://doc.pytest.org/} and are run within TravisCI's virtual environment.
-The current build status can be checked either on our Gtihub repository or at \url{https://travis-ci.org/superphy/backend}.
+TravisCI \url{travisci.io}, a continuous integration (CI) platform, is integrated into Spfy's Github repository \url{https://github.com/superphy/backend} and, with any changes to the codebase, runs tests for functionality and backwards compatibility.
+The individual tests use PyTest \url{https://doc.pytest.org/}, are run within TravisCI's virtual environment, and the current build status can be checked either on our Gtihub repository or at \url{https://travis-ci.org/superphy/backend}.
+CI is also used to automatically build Spfy's core Docker images, and upload them to Docker Hub \url{https://hub.docker.com/u/superphy/}.
 
@@
 
-% e.g.
-% database statistics
-We tested spfy with 55,353 samples genomes (267GB), storing both the entire genomes and results for all included analyses modules.
-The resulting database had XYZ nodes and XYZ edges, with XYZ object properties.
-This worked our to XYZ GB of data stored.
+Spfy was tested with 25,185 public \textit{E.
+coli} genomes, 5,353 genomes from GenBank and 19,832 genomes from Enterobase, and 55,353 generated samples genomes (267GB), storing both the entire sequences and results for all included analysis modules.
+The resulting database had XYZ nodes and XYZ edges, with XYZ object properties, which worked out to XYZ GB of data stored.
 
@@
 Other scientific workflow technologies such as Galaxy \cite{goecks2010galaxy}, Kepler \cite{ludascher2006scientific}, and Taverna \cite{oinn2004taverna} have been applied to bioinformatic tasks.
-Galaxy ...
+Galaxy aims to provide a reproducible compuation-based research environment which is accessible to individuals without programming knowledge.
 Taverna ...
@@
 coli} which leverages generic web technologies to favor a responsive, big-data design strategy tackling the challenge \cite{fricke2014bacterial} of integrating results from multiple genomes and examining shared connections.
+By packaging the individual bioinformatics software as Conda packages, they can be easily ported to Galaxy.
 
@@
 % we have a lot of similar directions to this, but just try to make it prettier, more friendly to use, and faster
-The Bacterial Analysis Platform (BAP), developed out of the Technical University of Denmark, provides and integrated analysis pipeline for bacterial WGS data, as a web service.
+The Bacterial Analysis Platform (BAP), developed out of the Technical University of Denmark, is the closest analogue to Spfy, and provides an integrated analysis pipeline for bacterial WGS data as a web service.
 BAP is novel in its combined approach to genome analysis as different programs, such as for VF and AMR determination, are included by default into the pipeline; Spfy provides similar functionallity with an expanded focus on integrated result storage and big-data analyses.

### 2017-08-25 paper-webserver.tex
EDIT: functionality section

@@
 % para covering ectyper & RGI
-To further characterize strains ... serotyping, which detects the presence of specific cell surface antigens.
-O-antigen and H-antigen typing, and VF determination use our ECTyper software.
-ECTYper compares the user-submitted genomes against a curated set of sequences, an approach originally implemented in the VirulenceFinder \cite{joensen2014real} tool.
-The Comprehensive Antibiotic Resistance Database (CARD) \cite{mcarthur2013comprehensive} maintains a set of known AMR genes and develops the RGI program .
-Both ECTyper and RGI are based off of BLAST \cite{pmid2231712}.
+To characterize strains, one approach is serotyping, which detects the presence of specific cell surface antigens.
+ECTyper uses a curated set of O-type and H-type specific gene sequences and runs the Basic Local Alignment Search Tool (BLAST) \cite{pmid2231712} against the query genome, before ranking the determined serotype in order of probability.
+A reference set of VFs sequences, obtained from literature review and available at \url{https://github.com/phac-nml/ecoli_vf}, is used with BLAST to determine associated VFs in a genome, an approach originally implemented in the VirulenceFinder program \cite{joensen2014real}
+We difer to the RGI tool, based off of the Comprehensive Antibiotic Resistance Database (CARD) \cite{mcarthur2013comprehensive}, maintains AMR gene determination.
 

### 2017-08-25 paper-webserver.tex
EDIT: more general edits

@@
 % para: the semantic web
-Spfy is built around semantic web technologies which focuses on describing the relaions between different datum \cite{berners2001semantic}.
-In biological data, a semantic web focus describes individual data points by the type, for example as a genome, contiguous DNA sequence, or gene, and then link related data together in a queryable graph.
+Spfy is built around semantic web technologies which describe the relaions between different datum \cite{berners2001semantic}.
+In biological data, a semantic web focus describes individual data points by its type, for example as a genome, contiguous DNA sequence, or gene, and then links related data together in a queryable graph.
 Semantic web technlogies allow new, not previously described data to be seemlessly incorporated to the existing graph, and has been proposed as a common standard for the open sharing of data \cite{horrocks2005semantic}.
@@
 % para: semantic web in spfy
-In Spfy, the entirety of SuperPhy's previous code is replaced with methods for handling semantic web technologies.
+In Spfy, the entirety of SuperPhy's previous code was replaced with methods for handling semantic web technologies.
 Generalized functions are used to convert the results from different analysis modules into a graph object, and that graph object is used to update the main graph database; information contained in the WGS file is also converted into a graph object and stored.
@@
 Spfy's server-side code is developed in Python and the website served to users is developed using the React JavaScript library.
-When users upload genomes for analysis: (i) they first make the upload through the ReactJS-based website.
-The public web service accepts up to 200 MB of genome files (50 genomes compressed, or 120 genomes compressed).
+When users upload genomes for analysis: (i) they first start the upload through the ReactJS-based website.
+The public web service accepts up to 200 MB of genome files (50 genomes uncompressed, or 120 genomes compressed).
 (ii) When files are uploaded, the user-selected analysis options are enqueued into the Redis Queue \url{http://python-rq.org/} task queue.
 Redis Queue consists of a Redis Database \url{https://redis.io/} and task queue workers which run as Python processes.
-(iii) The workers dequeue the various analyses and run them in parallel and results are stored back in the Redis database.
+(iii) The workers dequeue the various analyses, run them in parallel, and temporarily store results back in the Redis database.
 (iv) Python functions are then used to parse the results and permanently store them in Blazegraph \url{https://www.blazegraph.com/} - a graph database.
-The entire Spfy platform is packaged as a series of Docker \url{https://www.docker.com/} containers and orchestrated using Docker Compose \url{https://docs.docker.com/compose/}.
+The entire Spfy platform is packaged as a series of Docker \url{https://www.docker.com/} containers and connected together using Docker Compose \url{https://docs.docker.com/compose/}.
 
@@
 % ontology
-Ontologies are indices which define types of data and the relations between them.
-By sharing ontologies, it is possible to link sources of data from different organizations together, and Spfy uses types defined in the  Genomic Epidemiology Application Ontology (GenEpiO) (\cite{griffiths2017context} and Feature Annotation Location Description Ontology (FALDO) \cite{bolleman2016faldo}.
-Spfy also connects genomes, contiguous sequences (contigs), pangenome regions, and predicted genes and alleles in novel ways; to allow the presence of various genes to be infered from a genome description, realtionship links are generically named.
+The structure of the graph database is handled using ontologies: indices which define types of data and the relations between them.
+By sharing ontologies, it is possible to link databases from different organizations together, and Spfy uses types defined in the  Genomic Epidemiology Application Ontology (GenEpiO) \cite{griffiths2017context} and the Feature Annotation Location Description Ontology (FALDO) \cite{bolleman2016faldo}.
+Spfy also connects genomes, contiguous sequences (contigs), pangenome regions, and predicted genes and alleles using generic realtion edges, to allow the presence of various genes to be infered from a genome description.
 (see Figure \ref{fig-ontology})
@@
 (see Figure \ref{fig-results})
-Both the task selection and results display follow this card-based design: while data storage is actually graph-based, the results of various analysis modules are presented to users in a familiar tabular structure and available for download.
+Both the task selection and results page follow this card-based design: while data storage is actually graph-based, the results of various analysis modules are presented to users in a familiar tabular structure and available for download as .csv files.
 (see Figure \ref{fig-tables})
@@
 Task queues are processes which schedule code exection across available computing infrastructure.
-In Spfy, the Python-based Redis Queue library \url{https://github.com/nvie/rq} is used to manage analysis tasks and have them run asynchronously in response to user requests.
-When a user submits files for analysis or population-wide anlayses, separate tasks are enqueued at different priorities, depending on when users might expect the result to be returned.
-For example, population-wide analyses have a higher priority than bulk subtyping analyses, as we want search-response queries to respond instantly whereas a delay in subtyping is typical in similar web services such as the Center for Genomic Epidemiology Pipeline (CGE Pipeline) \cite{thomsen2016bacterial}.
+In Spfy, the Python-based Redis Queue library \url{https://github.com/nvie/rq} is used to manage analysis tasks and run them asynchronously in response to user requests.
+When a user submits files for analysis or requests population-wide comparisons, separate tasks are enqueued at different priorities, depending on our user experience goals.
+For example, population-wide analyses have a higher priority than bulk subtyping analyses, as we want search-response queries to return instantly whereas a delay in subtyping is typical in similar web services, such as the Center for Genomic Epidemiology Pipeline (CGE Pipeline) \cite{thomsen2016bacterial}.
 Spfy enables processing of thousands of genome sequences by using task queue workers running in parallel, which also allows performance to scale to available infrastructure.
-To ensure reliability of the platform, the open-source Sentry toolkit \url{https://github.com/getsentry/sentry} was integrated and used for real-time exception tracking.
+In addition, the open-source Sentry toolkit \url{https://github.com/getsentry/sentry} was integrated and used for real-time exception tracking and to ensure reliability of the platform.
 
@@
 
-The Spfy platform depends on a series of webservers, databases, and task workers and uses \url{https://www.docker.com/}, a virtualization software package, to run self-contained operating systems on the same host computer.
+The Spfy platform depends on a series of webservers, databases, and task workers and uses \url{https://www.docker.com/}, a virtualization technology, to run self-contained operating systems on the same host computer.
 Software packages are installed within the conatiners and the entire platform is networked together using Docker-Compose \url{https://docs.docker.com/compose/}.
 (see Figure \ref{fig-docker})
-Docker integration ensures that software dependencies, which typically must be manually installed \cite{doi:10.1093/bioinformatics/btu153,laing2010pan,inouye2014srst2,naccache2014cloud}, are handled automatically and that runtimes in one service are separated from another service.
+Docker integration ensures that software dependencies, which typically must be manually installed \cite{doi:10.1093/bioinformatics/btu153,laing2010pan,inouye2014srst2,naccache2014cloud}, are handled automatically, and that service runtimes are compartmentalized; this guarentees that code failures do not propagate to other services.
 

### 2017-10-27 paper-webserver.tex
Edit of current draft. Added TODO: throughout and some comments / changes.

@@
 
-\title{Spfy: distributed predictive genomics of E.coli with graph based result linkage}
+\title{Superphy: Real-time prediction of \textit{E. coli} phenotype using graph-based genome sequence storage and retrieval with the \textit{spfy} module}
 
@@
 % 1. brief: WGS is standard
-Whole genome sequencing (WGS) resolves the entire genetic content of an organism.
-WGS data can increase the resolution and sensitivity of bacterial surveillance \cite{ronholm2016navigating,lytsy2017time}, identification of potential disease mechanisms \cite{wang2014whole,yuen2015whole}, and clinical diagnoses \cite{willig2015whole,dewey2014clinical}.
+Whole genome sequencing (WGS) can in theory provide the entire genetic content of an organism.
+This unparalleled resolution and sensitivity has recently transformed public-health surveillance and outbreak response \cite{ronholm2016navigating,lytsy2017time}.
+Additionally, the identification of novel disease mechanisms \cite{wang2014whole,yuen2015whole}, and rapid clinical diagnoses and reference lab tests based on the specific mechanism of disease are now possible.
+\cite{willig2015whole,dewey2014clinical}.
+
 % 2. Big Problem: but tools are for individual analysis
-Targeted software, such as the Resistance Gene Identifier (RGI) \cite{mcarthur2013comprehensive} for antimicrobial resistance (AMR) gene prediction, Prokka for bacterial genome annotation \cite{doi:10.1093/bioinformatics/btu153}, and integrated platforms, such as the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial} and the Integrated Rapid Infectious Disease Analysis (IRIDA) project \url{http://www.irida.ca/}, all leverage WGS data.
+The rapid characterization and comparison of bacterial pathogens relies principally on the combination of outputs from multiple software programs that are targeted for specific applications.
+Examples include the Resistance Gene Identifier (RGI) \cite{mcarthur2013comprehensive} for antimicrobial resistance (AMR) gene prediction, Prokka for bacterial genome annotation \cite{doi:10.1093/bioinformatics/btu153}, and {ANOTHER EXAMPLE}.
+Comprehensive platforms that combine individual programs into a cohesive whole also exist.
+These include free platforms such as the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial}, the Integrated Rapid Infectious Disease Analysis (IRIDA) project \url{http://www.irida.ca/}, and PATRIC.
+Commercial applications, such as Bionumerics, which is used by PulseNet international for the analyses of WGS data in outbreak situations also exist, and offer support as well as accredited, standardized tests {REF}.
+
 % 3. 100,000 genomes, (lookup number for Enterobase, GenBank), how do we run analysis on it
-WGS generates data at a break-neck speed; for \textit{Escherichia coli} alone, the public genome databases EnteroBase \url{https://enterobase.warwick.ac.uk/} and GenBank \cite{doi:10.1093/nar/gks1195} have, respectively, 60,206 and 2,779,008 sequences uploaded.
-To effectively exploit WGS "big-data" and maintain the rapid response time required by public health applications, one approach is to make results from WGS analysis integrated and progressive.
-Typical bioinformatics software, such as RGI and Prokka, take single files as input, and integrated platforms, such as BAP and IRIDA, build workflows linking different analyses modules.
-BAP and IRIDA begin to solve big-data challenges by offering a hosted solution which computes results in real-time, and distributes analyses across computing resources.
-While effective for self-contained workflows, many comparative analyses such as predictive genomics methods would benefit from a broad WGS reference base.
-To use vast amount of WGS data and maintain real-time results for comparative analyses, platforms would need store results in a way that avoids recomputation when adding new WGS data.
-A resource framework in which WGS results are integrated and searchable; whereby the storage of results avoids recomputation, persists, and allows for iterative, on-going learning, will expand the capacity of current comparative bioinformatics analyses.
-This increased capacity will further benefit surveillance, research, and clinical applications\par
+WGS data for bacterial pathogens of public health importance have recently accumulated in public databases in the tens of thousands, with hundreds of thousands set to be available within the next few years.
+For \textit{Escherichia coli}, there are over sixty thousand publicly available genomes in EnteroBase \url{https://enterobase.warwick.ac.uk/} and X whole genomes in GenBank \cite{doi:10.1093/nar/gks1195}.
+
+To begin to solve this "big-data" problem, platforms such as BAP and IRIDA provide hosted solutions that compute results in real-time, and distribute analyses across computing resources.
+While effective for self-contained workflows, many of the comparative analyses that are run are broadly useful, and therefore computed multiple times.
+An effective method to mitigate the recomputing of analyses is to make the storage and retrieval of results part of the platform, and effectively linked to the organisms of interest with a standardized ontology.
+Such measures can help ensure the rapid response times required for public health applications, and allow results to be integrated and progressively updated as new data become available.
 
 % 1. Our specific problem, previous work
-We have previously developed Superphy \cite{whiteside2016superphy}, an online predictive genomics platform targeting \textit{E. coli}.
-Superphy integrates pre-computed results with domain-specific metadata to provide real-time analyses of epidemiology relations.
-While this tool has been useful for the thousands of pre-computed genomes in its database, the current pace of genome sequencing requires real-time predictive genomic analyses of tens of thousands of genomes, and the long term storage and referencing of these results, something that the original SuperPhy platform was incapable of.
+We have previously developed Superphy \cite{whiteside2016superphy}, an online predictive genomics platform targeting \textit{E.
+coli}.
+Superphy integrates pre-computed results with domain-specific knowledge to provide real-time exploration of publicly available genomes.
+While this tool has been useful for the thousands of pre-computed genomes in its database, the current pace of genome sequencing requires real-time predictive genomic analyses of tens-, and soon hundreds-of-thousands of genomes, and the long term storage and referencing of these results, something that the original SuperPhy platform was incapable of.
+
 % 2. Why solving it is important (for Public health / research)
-WGS offers improved resolution over traditional strain comparison methods, such as pulsed-field gel electrophoresis (PFGE) \cite{ronholm2016navigating}.
-Though the cost of developing, approving, and transforming existing workflows from wet-lab to sequence prediction approaches is time consuming and expensive \cite{koser2012routine}, platforms can only perform real-time analyses and linkage to thousands of historical results by leveraging WGS.
+- everything is being sequenced (surveillance / outbreak / research)
+- previously mentioned common analyses / want to leverage pre-computed results
+- WGS does not discard old methods, linkage to thousands of historical results by developing in-silico methods of traditional tests
+- need fast / standardized outbreak response
+- need fast / standardized in-silico reference lab
+- need fast / standardized storage and retrieval of results based on ontology
+- all known data can be leveraged, allows the most informed decisions possible
+- etc.
+
 % 3. how we solved it
 In this study, we present an update to the SuperPhy platform, called Spfy.
-The update rewrites result storage with backing by a graph database, takes a modular approach to tool integration, and distributes analyses over task queues, thereby allowing users to submit genomes and modules to run in parallel in real-time, and address code failures.
+The update rewrites the result storage by using a graph database, and takes a modular approach to tool integration, and distributes analyses over task queues; this allows users to submit genomes and modules to run in parallel in real-time.
+
+{needs to be beefed up, in language a biologist / public health worker would care about}
+
 All results are stored as a series of linked nodes which enables the platform to build associations between results as they are generated. \par
@@
 Subtyping options are ..., pan-genome generation ..., group comparisons via Fisher's, ML, .... for E.coli.
+
 By supporting multiple \textit{in-silico} subtyping options, the platform functions similar to a reference laboratory, with added support for big-data analyses.
 Currently, the platform has been tested with XXX genome files and result storage for XX analyses modules.
-Future work will focus on adding more analysis modules and supporting different species, which can be connected to the existing graph database without need for recalculation.
-To complement existing platforms such as IRIDA, modules are self-contained and can easily be integrated into Galaxy \cite{goecks2010galaxy} based platforms.
+
+Future work will focus on adding additional analyses modules, using machine learning and artificial neural networks to aid genotype to phenotype predictions, and supporting different species.
+While the integrated approach or storing and retrieving results provides enormous benefits, the developed analyses modules are self-contained and can easily be integrated into existing platforms such as Galaxy \cite{goecks2010galaxy}, and IRIDA.
+
 The website and source code are available at \url{https://superphy.github.io/}. \par
@@
 % para covering everything
-Spfy performs reference laboratory tasks: O-antigen and H-antigen typing, Shiga-toxin (STX) typing,  virulence factor (VF) and antimicrobial resistance (AMR) gene determination, and related strain determination.
+Spfy performs reference laboratory tasks: O-antigen and H-antigen typing using ectyper \url{https://github.com/phac-nml/ecoli_serotyping}, Shiga-toxin (STX) typing,  virulence factor (VF) and antimicrobial resistance (AMR) gene determination, and related strain determination.
+
 Spfy also performs bioinformatics analyses: pangenome generation, statistical significance testing of genome markers for user-defined groups, and AMR predictions using support vector machines (SVMs).
-We introduce one new bioinformatics program, Ecoli-Serotyper (ECTyper) \url{https://github.com/phac-nml/ecoli_serotyping}, for determining surface-antigen type and associated VFs, and utilize existing software: the Resistance Gene Identifier (RGI) program \cite{mcarthur2013comprehensive}, Phylotyper \cite{whiteside2017phylotyper}, and Panseq \cite{laing2010pan}.
 
-% para covering ectyper & RGI
-To characterize strains, one approach is serotyping, which detects the presence of specific cell surface antigens.
-ECTyper uses a curated set of O-type and H-type specific gene sequences and runs the Basic Local Alignment Search Tool (BLAST) \cite{pmid2231712} against the query genome, before ranking the determined serotype in order of probability.
-A reference set of VFs sequences, obtained from literature review and available at \url{https://github.com/phac-nml/ecoli_vf}, is used with BLAST to determine associated VFs in a genome, an approach originally implemented in the VirulenceFinder program \cite{joensen2014real}
-We difer to the RGI tool, based off of the Comprehensive Antibiotic Resistance Database (CARD) \cite{mcarthur2013comprehensive}, maintains AMR gene determination.
+TODO: List all of the below references in the above paragraphs.
+%We introduce one new bioinformatics program, Ecoli-Serotyper (ECTyper) , for determining surface-antigen type and associated VFs, and utilize existing software: the Resistance Gene Identifier (RGI) program \cite{mcarthur2013comprehensive}, Phylotyper \cite{whiteside2017phylotyper}, and Panseq \cite{laing2010pan}.
+
+% para covering ectyper & RGI -- don't really need this
+% RGI has its own paper and was in the previous SuperPhy paper
+% ectyper will get its own paper
 
@@
 % para: the semantic web
-Spfy is built around semantic web technologies which describe the relaions between different datum \cite{berners2001semantic}.
-In biological data, a semantic web focus describes individual data points by its type, for example as a genome, contiguous DNA sequence, or gene, and then links related data together in a queryable graph.
-Semantic web technlogies allow new, not previously described data to be seemlessly incorporated to the existing graph, and has been proposed as a common standard for the open sharing of data \cite{horrocks2005semantic}.
+Spfy is built around semantic web technologies which describe the relaions between different data \cite{berners2001semantic}.
+For biological data this means that individual data points such as genome, contiguous DNA sequence, or gene, are linked together in a queryable graph form.
+This allows novel data to be seamlessly incorporated into the existing graph, and has {TODO: what has been proposed specifically?} been proposed as a common standard for the open sharing of data \cite{horrocks2005semantic}.
 
 % para: semantic web in spfy
-In Spfy, the entirety of SuperPhy's previous code was replaced with methods for handling semantic web technologies.
-Generalized functions are used to convert the results from different analysis modules into a graph object, and that graph object is used to update the main graph database; information contained in the WGS file is also converted into a graph object and stored.
-The use of semantic web technologies allows results to be linked to the genomes they were computed from, and all data points to share a common datastructure that can be queried together.
+In Spfy, the results from all the analysis modules are converted into graph objects, which are used to update the main graph database. TODO: need information about ontology, terms etc. for metadata.
+
+The use of semantic web technologies allows results to be linked to the genomes they were computed from, and for all data points to share a common data structure that can be queried together.
 
 % para: intro to the spfy stack
-Spfy's server-side code is developed in Python and the website served to users is developed using the React JavaScript library.
-When users upload genomes for analysis: (i) they first start the upload through the ReactJS-based website.
+Spfy's server-side code is developed in Python and the website served to users using the React JavaScript library.
+
+For the addition of new data to the database, the following steps are taken:
+
+i) The upload begins through the ReactJS-based website, where user-defined analyses options are selected.
+The results of these chosen analyses are immediately reported to the user following their completion, while the remaining analyses are subsequently completed and stored in the database without interaction from the user.
 The public web service accepts up to 200 MB of genome files (50 genomes uncompressed, or 120 genomes compressed).
-(ii) When files are uploaded, the user-selected analysis options are enqueued into the Redis Queue \url{http://python-rq.org/} task queue.
+
+ii) User-selected analyses are enqueued into the Redis Queue \url{http://python-rq.org/} task queue.
 Redis Queue consists of a Redis Database \url{https://redis.io/} and task queue workers which run as Python processes.
-(iii) The workers dequeue the various analyses, run them in parallel, and temporarily store results back in the Redis database.
-(iv) Python functions are then used to parse the results and permanently store them in Blazegraph \url{https://www.blazegraph.com/} - a graph database.
-The entire Spfy platform is packaged as a series of Docker \url{https://www.docker.com/} containers and connected together using Docker Compose \url{https://docs.docker.com/compose/}.
+
+iii) The workers dequeue the analyses, run them in parallel, and temporarily store results in the Redis database.
+
+iv) Python functions parse the results and permanently store them in Blazegraph \url{https://www.blazegraph.com/}, the graph database used for Superphy.
+
+The entire Spfy update is packaged as a series of Docker \url{https://www.docker.com/} containers and connected together using Docker Compose \url{https://docs.docker.com/compose/}.
 
@@
 % 3 1/2. SPARQL queries
-Graph databases describe relationships between different data points, is one of the emerging \cite{de2015trends} database types used for biological data, and is one of the core building blocks of a semantic web technology stack \cite{horrocks2005semantic}.
-Spfy uses the RDFLib Python library \url{https://rdflib.readthedocs.org/} to represent all data meant for long-term storage.
-When serotyping, VF, AMR predictions, and pangenome generation tasks are completed, the results are stored within the graph database.
+Graph databases describe relationships between data points, and are one of the emerging \cite{de2015trends} database types for biological data; they are one of the core building blocks of semantic web technology \cite{horrocks2005semantic}.
+Spfy uses the RDFLib Python library \url{https://rdflib.readthedocs.org/} to represent all data, including the results of the analyses modules.
+
 % note: below is a feature in development
-The permanent storage of results serve as a one-time cost, and allows population-wide analyses of all stored genomes; result storage also enables Spfy to avoid recomputation when the same analysis is re-run.
+The permanent storage of results is as a one-time cost, and allows population-wide analyses of all stored genomes; result storage also enables Spfy to avoid recomputation when the same analysis is re-run.
 
@@
 
-Task queues are processes which schedule code exection across available computing infrastructure.
-In Spfy, the Python-based Redis Queue library \url{https://github.com/nvie/rq} is used to manage analysis tasks and run them asynchronously in response to user requests.
-When a user submits files for analysis or requests population-wide comparisons, separate tasks are enqueued at different priorities, depending on our user experience goals.
-For example, population-wide analyses have a higher priority than bulk subtyping analyses, as we want search-response queries to return instantly whereas a delay in subtyping is typical in similar web services, such as the Center for Genomic Epidemiology Pipeline (CGE Pipeline) \cite{thomsen2016bacterial}.
-Spfy enables processing of thousands of genome sequences by using task queue workers running in parallel, which also allows performance to scale to available infrastructure.
-In addition, the open-source Sentry toolkit \url{https://github.com/getsentry/sentry} was integrated and used for real-time exception tracking and to ensure reliability of the platform.
+TODO: The below was already covered above. No need to repeat. Either here or there, but combine.
+%% Task queues are processes which schedule code exection across available computing infrastructure.
+%% In Spfy, the Python-based Redis Queue library \url{https://github.com/nvie/rq} is used to manage analysis tasks and run them asynchronously in response to user requests.
+%% When a user submits files for analysis or requests population-wide comparisons, separate tasks are enqueued at different priorities, depending on our user experience goals.
+%% For example, population-wide analyses have a higher priority than bulk subtyping analyses, as we want search-response queries to return instantly whereas a delay in subtyping is typical in similar web services, such as the Center for Genomic Epidemiology Pipeline (CGE Pipeline) \cite{thomsen2016bacterial}.
+%% Spfy enables processing of thousands of genome sequences by using task queue workers running in parallel, which also allows performance to scale to available infrastructure.
+%% In addition, the open-source Sentry toolkit \url{https://github.com/getsentry/sentry} was integrated and used for real-time exception tracking and to ensure reliability of the platform.
 
@@
 
+TODO: bring the docker information from above down here.
 The Spfy platform depends on a series of webservers, databases, and task workers and uses \url{https://www.docker.com/}, a virtualization technology, to run self-contained operating systems on the same host computer.
@@
 (see Figure \ref{fig-docker})
-Docker integration ensures that software dependencies, which typically must be manually installed \cite{doi:10.1093/bioinformatics/btu153,laing2010pan,inouye2014srst2,naccache2014cloud}, are handled automatically, and that service runtimes are compartmentalized; this guarentees that code failures do not propagate to other services.
+Docker integration ensures that software dependencies, which typically must be manually installed \cite{doi:10.1093/bioinformatics/btu153,laing2010pan,inouye2014srst2,naccache2014cloud}, are handled automatically, and that service runtimes are compartmentalized; this guarantees that code failures do not propagate to other services.
 
@@
 
-TravisCI \url{travisci.io}, a continuous integration (CI) platform, is integrated into Spfy's Github repository \url{https://github.com/superphy/backend} and, with any changes to the codebase, runs tests for functionality and backwards compatibility.
-The individual tests use PyTest \url{https://doc.pytest.org/}, are run within TravisCI's virtual environment, and the current build status can be checked either on our Gtihub repository or at \url{https://travis-ci.org/superphy/backend}.
+TravisCI \url{travisci.io}, a continuous integration (CI) platform, is used to ensure that Spfy \url{https://github.com/superphy/backend} does not break with any changes to the codebase, and runs tests for functionality and backwards compatibility.
+The individual tests use PyTest \url{https://doc.pytest.org/}, and are run within TravisCI's virtual environment, and the current build status can be checked either on our Gtihub repository or at \url{https://travis-ci.org/superphy/backend}.
 CI is also used to automatically build Spfy's core Docker images, and upload them to Docker Hub \url{https://hub.docker.com/u/superphy/}.
@@
 \section{RESULTS}
-
+TODO: Update with all of Enterobase
 Spfy was tested with 25,185 public \textit{E.
@@
 
-One of the key goals in developing Spfy is to maintain instantaneity: modern websites have accustomed users to instant results.
-We attempt to use innovations in web development and bring a similar experience to Spfy as a predictive genomics platform for \textit{E. coli}.
+One of the key goals in developing Spfy is to maintain instantaneity, as modern websites have accustomed users to instant results.
+We attempt to use innovations in web development to bring a similar experience to Spfy as a predictive genomics platform for \textit{E. coli}.
 % Mention where users / developers can find documentation
@@
 The isolation and characterization of bacterial pathogens are critical for Public Health laboratories to rapidly respond to outbreaks, and to effectively monitor known and emerging pathogens through surveillance programs.
-Until recently, Public-health agencies relied on laboratory tests such as XYZ to characterize bacterial isolates in outbreak and surveillance settings.
-The previous gold-standard in determining strain relatedness was pulsed-field gel electrophoresis (PFGE) {ronholm2016navigating}, which uses rare-cutting restriction enzymes to produce a unique banding pattern for each strain.
-However, in \textit{Enterococcus faecium}, PFGE has been shown to misclassify 9 of 132 isolates, when compared to whole-genome sequencing (WGS) based discrimination \cite{pinholt2015multiple}.
-In \textit{Klebsiella pneumoniae} \cite{marsh2015genomic}, \textit{Yersinia enterocolitica} \cite{gilpin2014limitations}, and \textit{Staphylococcus aureus} \cite{doi:10.1093/ofid/ofu096}, WGS was used to discriminate isolates after initial clustering by PFGE resulted in indistinguishable samples.
-Examination of PFGE bands are also subjective, difficult to share \cite{lytsy2017time}, and collative platforms such as PulseNet reported \cite{gilpin2014limitations} that even after collecting 72\% of \textit{Campylobacter jejuni} in a given year in Minnesota (673 cases), 87\% of isolates could not be linked by PFGE pattern.
+Until recently, public-health agencies relied on laboratory tests such as XYZ to characterize bacterial isolates in outbreak and surveillance settings.
+
+TODO: The below has previously been covered by many papers, including our previous SuperPhy paper. We need to focus on the big-data aspect.
+%% The previous gold-standard in determining strain relatedness was pulsed-field gel electrophoresis (PFGE) {ronholm2016navigating}, which uses rare-cutting restriction enzymes to produce a unique banding pattern for each strain.
+%% However, in \textit{Enterococcus faecium}, PFGE has been shown to misclassify 9 of 132 isolates, when compared to whole-genome sequencing (WGS) based discrimination \cite{pinholt2015multiple}.
+%% In \textit{Klebsiella pneumoniae} \cite{marsh2015genomic}, \textit{Yersinia enterocolitica} \cite{gilpin2014limitations}, and \textit{Staphylococcus aureus} \cite{doi:10.1093/ofid/ofu096}, WGS was used to discriminate isolates after initial clustering by PFGE resulted in indistinguishable samples.
+%% Examination of PFGE bands are also subjective, difficult to share \cite{lytsy2017time}, and collative platforms such as PulseNet reported \cite{gilpin2014limitations} that even after collecting 72\% of \textit{Campylobacter jejuni} in a given year in Minnesota (673 c
+%% ases), 87\% of isolates could not be linked by PFGE pattern.
 Antimicrobial resistance testing, virulence factor testing ...
@@
 Galaxy aims to provide a reproducible compuation-based research environment which is accessible to individuals without programming knowledge.
+TODO: BAP and IRIDA, since they were the examples from above
 Taverna ...
@@
 % we have a lot of similar directions to this, but just try to make it prettier, more friendly to use, and faster
+
+TODO: The needs to be linked to the previous discussion above, as its absence was conspicuous.
 The Bacterial Analysis Platform (BAP), developed out of the Technical University of Denmark, is the closest analogue to Spfy, and provides an integrated analysis pipeline for bacterial WGS data as a web service.

### 2017-11-02 paper-webserver.tex
ADD: reference virulencefinder

@@
 The rapid characterization and comparison of bacterial pathogens relies principally on the combination of outputs from multiple software programs that are targeted for specific applications.
-Examples include the Resistance Gene Identifier (RGI) \cite{mcarthur2013comprehensive} for antimicrobial resistance (AMR) gene prediction, Prokka for bacterial genome annotation \cite{doi:10.1093/bioinformatics/btu153}, and {ANOTHER EXAMPLE}.
+Examples include the Resistance Gene Identifier (RGI) \cite{mcarthur2013comprehensive} for antimicrobial resistance (AMR) gene prediction, VirulenceFinder \cite{kleinheinz2014applying} virulence gene annotation, and Prokka for bacterial genome annotation with external tools \cite{doi:10.1093/bioinformatics/btu153}.
 Comprehensive platforms that combine individual programs into a cohesive whole also exist.

### 2017-11-02 paper-webserver.tex
START: on the beef

@@
 WGS data for bacterial pathogens of public health importance have recently accumulated in public databases in the tens of thousands, with hundreds of thousands set to be available within the next few years.
-For \textit{Escherichia coli}, there are over sixty thousand publicly available genomes in EnteroBase \url{https://enterobase.warwick.ac.uk/} and X whole genomes in GenBank \cite{doi:10.1093/nar/gks1195}.
+For \textit{Escherichia coli}, there are over sixty thousand publicly available genomes in EnteroBase \url{https://enterobase.warwick.ac.uk/} and three million whole genomes in GenBank \cite{doi:10.1093/nar/gks1195}.
 
@@
 In this study, we present an update to the SuperPhy platform, called Spfy.
-The update rewrites the result storage by using a graph database, and takes a modular approach to tool integration, and distributes analyses over task queues; this allows users to submit genomes and modules to run in parallel in real-time.
+Spfy enables large-scale genome processing in real-time and uniquely decomposses results into graph datastructures for further comparative genomics.
+Unlike traditional approaches which store results in separate data files or in a relational database, our graph datastructures are then integrated into a knowledge graph identifying relationships between all genomes sequenced in the past, present, and future.
+We present the use of graph databases as a solution to storing genome analysis results in a meaningful manner, and to allow retrospective comparisons as more genomes are sequenced or populations change.
+
+{I'm not sure if we should include that the same graph database can be expanded to cover different spcies}
 

### 2017-11-06 paper-webserver.tex
EDIT: some typing + merged the comparisons section

@@
 % 2. Why solving it is important (for Public health / research)
+WGS is becoming the norm for bacterial surveillance, outbreak tracing, and research \cite{ronholm2016navigating}.
+While pipelines exist \cite{joensen2014real} for running a variety of analyses on WGS data,
+In order to  Existing annotation and analysis software for WGS data represent units of compuatation; we can optimize WGS pipelines if these results are stored to avoid recomputation, and extend results by viewing them in the context of other analysis modules and genomes.
+ these computations can be 
 - everything is being sequenced (surveillance / outbreak / research)
@@
 Spfy enables large-scale genome processing in real-time and uniquely decomposses results into graph datastructures for further comparative genomics.
-Unlike traditional approaches which store results in separate data files or in a relational database, our graph datastructures are then integrated into a knowledge graph identifying relationships between all genomes sequenced in the past, present, and future.
+Unlike traditional approaches which store results in separate data files or in a relational database, the graph datastructures are then integrated into a knowledge graph identifying relationships between all genomes sequenced in the past, present, and future.
 We present the use of graph databases as a solution to storing genome analysis results in a meaningful manner, and to allow retrospective comparisons as more genomes are sequenced or populations change.
@@
 
-All results are stored as a series of linked nodes which enables the platform to build associations between results as they are generated. \par
+\par
 
@@
 
-The entire Spfy update is packaged as a series of Docker \url{https://www.docker.com/} containers and connected together using Docker Compose \url{https://docs.docker.com/compose/}.
+The entire Spfy update is packaged in Docker \url{https://www.docker.com/} containers and connected together using Docker Compose \url{https://docs.docker.com/compose/}.
 
@@
 % namely galaxy
-Other scientific workflow technologies such as Galaxy \cite{goecks2010galaxy}, Kepler \cite{ludascher2006scientific}, and Taverna \cite{oinn2004taverna} have been applied to bioinformatic tasks.
+Other scientific workflow technologies such as Galaxy \cite{goecks2010galaxy}, and bioinformatics pipelines IRIDA and BAP are used to run analysis modules on WGS data.
 Galaxy aims to provide a reproducible compuation-based research environment which is accessible to individuals without programming knowledge.
 TODO: BAP and IRIDA, since they were the examples from above
-Taverna ...
-Kepler ...
+IRIDA...
+BAP...
+
+The Bacterial Analysis Platform (BAP), developed out of the Technical University of Denmark, is the closest analogue to Spfy, and provides an integrated analysis pipeline for bacterial WGS data as a web service.
+BAP is novel in its combined approach to genome analysis as different programs, such as for VF and AMR determination, are included by default into the pipeline; Spfy provides similar functionallity with an expanded focus on integrated result storage and big-data analyses.
+In place of a MySQL database for storing the location of result files, Spfy parses result data into graph objects and integrates results into a persistant graph database.
+This enables Spfy to create large datasets where analysis results are linked to genomes, and to perform population-wide analyses.
+On a per file basis, Spfy performs at a similar speed to BAP on predictive genomics tasks, though Spfy does not provide genome assembly services.
+Spfy also processes XXXX files over XX tasks in XXXX time by using a novel approach of distributing computations over a task queue and multiple Docker compartmentalized containers.
+
 To approach the challenge of integrating different bioinformatics programs, Spfy instead uses technologies prevalent in common web services not necessarily related to scientific workflows.
@@
 By packaging the individual bioinformatics software as Conda packages, they can be easily ported to Galaxy.
-
-\subsection{Comparison with similar bioinformatic pipelines}
-
-% para: \cite{naccache2014cloud}
-% must install a ton of deps and even create dbs manually https://github.com/chiulab/surpi
-% im not actually sure where they're getting the 'cloud-compatible' aspect of their paper from - its cloud compatible in that you deploy in the same manner as you would to bare-metal, but that's just like saying it 'runs on a computer'
-
-% para: \cite{aanensen2016whole}
-% makes some pretty pictures, but not really a pipeline
-
-% para: \cite{joensen2014real,thomsen2016bacterial}
-% we have a lot of similar directions to this, but just try to make it prettier, more friendly to use, and faster
-
-TODO: The needs to be linked to the previous discussion above, as its absence was conspicuous.
-The Bacterial Analysis Platform (BAP), developed out of the Technical University of Denmark, is the closest analogue to Spfy, and provides an integrated analysis pipeline for bacterial WGS data as a web service.
-BAP is novel in its combined approach to genome analysis as different programs, such as for VF and AMR determination, are included by default into the pipeline; Spfy provides similar functionallity with an expanded focus on integrated result storage and big-data analyses.
-In place of a MySQL database for storing the location of result files, Spfy parses result data into graph objects and integrates results into a persistant graph database.
-This enables Spfy to create large datasets where analysis results are linked to genomes, and to perform population-wide analyses.
-On a per file basis, Spfy performs at a similar speed to BAP on predictive genomics tasks, though Spfy does not provide genome assembly services.
-Spfy also processes XXXX files over XX tasks in XXXX time by using a novel approach of distributing computations over a task queue and multiple Docker compartmentalized containers.
 

### 2017-11-10 paper-webserver.tex
EDIT: some cuts / changes here and ther

@@
 In this study, we present an update to the SuperPhy platform, called Spfy.
-Spfy enables large-scale genome processing in real-time and uniquely decomposses results into graph datastructures for further comparative genomics.
-Unlike traditional approaches which store results in separate data files or in a relational database, the graph datastructures are then integrated into a knowledge graph identifying relationships between all genomes sequenced in the past, present, and future.
+Spfy enables large-scale genome processing in real-time and uniquely decomposses results into graph data structures for further comparative genomics.
+Unlike traditional approaches which store results in separate data files or in a relational database, the graph data structures are then integrated into a knowledge graph identifying relationships between all genomes sequenced in the past, present, and future.
 We present the use of graph databases as a solution to storing genome analysis results in a meaningful manner, and to allow retrospective comparisons as more genomes are sequenced or populations change.
@@
 % para covering everything
-Spfy performs reference laboratory tasks: O-antigen and H-antigen typing using ectyper \url{https://github.com/phac-nml/ecoli_serotyping}, Shiga-toxin (STX) typing,  virulence factor (VF) and antimicrobial resistance (AMR) gene determination, and related strain determination.
+Spfy performs reference laboratory tasks: O-antigen typing, H-antigen typing, and virulence factor (VF) gene determination using ECTyper \url{https://github.com/phac-nml/ecoli_serotyping}, Shiga-toxin (STX) typing using Phylotyper \cite{whiteside2017phylotyper}, and antimicrobial resistance (AMR) gene determination using the Resistance Gene Identifier (RGI) program \cite{mcarthur2013comprehensive}.
 
-Spfy also performs bioinformatics analyses: pangenome generation, statistical significance testing of genome markers for user-defined groups, and AMR predictions using support vector machines (SVMs).
-
-TODO: List all of the below references in the above paragraphs.
-%We introduce one new bioinformatics program, Ecoli-Serotyper (ECTyper) , for determining surface-antigen type and associated VFs, and utilize existing software: the Resistance Gene Identifier (RGI) program \cite{mcarthur2013comprehensive}, Phylotyper \cite{whiteside2017phylotyper}, and Panseq \cite{laing2010pan}.
+Spfy also performs bioinformatics analyses: pangenome generation using Panseq \cite{laing2010pan}, statistical significance testing of genome markers for user-defined groups, and AMR predictions using support vector machines (SVMs).
 
@@
 iv) Python functions parse the results and permanently store them in Blazegraph \url{https://www.blazegraph.com/}, the graph database used for Superphy.
-
-The entire Spfy update is packaged in Docker \url{https://www.docker.com/} containers and connected together using Docker Compose \url{https://docs.docker.com/compose/}.
 
@@
 
-TODO: The below was already covered above. No need to repeat. Either here or there, but combine.
-%% Task queues are processes which schedule code exection across available computing infrastructure.
-%% In Spfy, the Python-based Redis Queue library \url{https://github.com/nvie/rq} is used to manage analysis tasks and run them asynchronously in response to user requests.
-%% When a user submits files for analysis or requests population-wide comparisons, separate tasks are enqueued at different priorities, depending on our user experience goals.
-%% For example, population-wide analyses have a higher priority than bulk subtyping analyses, as we want search-response queries to return instantly whereas a delay in subtyping is typical in similar web services, such as the Center for Genomic Epidemiology Pipeline (CGE Pipeline) \cite{thomsen2016bacterial}.
-%% Spfy enables processing of thousands of genome sequences by using task queue workers running in parallel, which also allows performance to scale to available infrastructure.
-%% In addition, the open-source Sentry toolkit \url{https://github.com/getsentry/sentry} was integrated and used for real-time exception tracking and to ensure reliability of the platform.
-
 % para
@@
 
-TODO: bring the docker information from above down here.
-The Spfy platform depends on a series of webservers, databases, and task workers and uses \url{https://www.docker.com/}, a virtualization technology, to run self-contained operating systems on the same host computer.
-Software packages are installed within the conatiners and the entire platform is networked together using Docker-Compose \url{https://docs.docker.com/compose/}.
+Docker \url{https://www.docker.com/} is a virtualization technology to run self-contained operating systems on the same host computer.
+The Spfy platform depends on a series of webservers, databases, and task workers and uses Docker to compartmentalize these services which are then networked together using Docker-Compose \url{https://docs.docker.com/compose/}.
 (see Figure \ref{fig-docker})

### 2017-11-13 paper-webserver.tex
EDIT: merged bullets into the into paras

@@
 Examples include the Resistance Gene Identifier (RGI) \cite{mcarthur2013comprehensive} for antimicrobial resistance (AMR) gene prediction, VirulenceFinder \cite{kleinheinz2014applying} virulence gene annotation, and Prokka for bacterial genome annotation with external tools \cite{doi:10.1093/bioinformatics/btu153}.
+In particular, RGI and VirulenceFinder represent a series of \textit{in-silico} methods which have been developed to mimick the results of traditional tests.
+These methods allow new WGS results to be viewed in the context of historical tests.
+
 Comprehensive platforms that combine individual programs into a cohesive whole also exist.
@@
 
-% 1. Our specific problem, previous work
+% 4. Our specific problem, previous work
 We have previously developed Superphy \cite{whiteside2016superphy}, an online predictive genomics platform targeting \textit{E.
@@
 
-% 2. Why solving it is important (for Public health / research)
-WGS is becoming the norm for bacterial surveillance, outbreak tracing, and research \cite{ronholm2016navigating}.
-While pipelines exist \cite{joensen2014real} for running a variety of analyses on WGS data,
-In order to  Existing annotation and analysis software for WGS data represent units of compuatation; we can optimize WGS pipelines if these results are stored to avoid recomputation, and extend results by viewing them in the context of other analysis modules and genomes.
- these computations can be 
-- everything is being sequenced (surveillance / outbreak / research)
-- previously mentioned common analyses / want to leverage pre-computed results
-- WGS does not discard old methods, linkage to thousands of historical results by developing in-silico methods of traditional tests
-- need fast / standardized outbreak response
-- need fast / standardized in-silico reference lab
-- need fast / standardized storage and retrieval of results based on ontology
-- all known data can be leveraged, allows the most informed decisions possible
-- etc.
+% ?. Why solving it is important (for Public health / research)
+% Merging points from here into the above sections (2,3,4) to avoid repetition
+% - everything is being sequenced (surveillance / outbreak / research)
+% - previously mentioned common analyses / want to leverage pre-computed results
+% - WGS does not discard old methods, linkage to thousands of historical results by developing in-silico methods of traditional tests
+% - need fast / standardized outbreak response
+% - need fast / standardized in-silico reference lab
+% - need fast / standardized storage and retrieval of results based on ontology
+% - all known data can be leveraged, allows the most informed decisions possible
+% - etc.
 
-% 3. how we solved it
+% 5. how we solved it
 In this study, we present an update to the SuperPhy platform, called Spfy.

### 2017-11-13 paper-webserver.tex
EDIT: less worcestershire sauce with the beef

@@
 In this study, we present an update to the SuperPhy platform, called Spfy.
-Spfy enables large-scale genome processing in real-time and uniquely decomposses results into graph data structures for further comparative genomics.
-Unlike traditional approaches which store results in separate data files or in a relational database, the graph data structures are then integrated into a knowledge graph identifying relationships between all genomes sequenced in the past, present, and future.
-We present the use of graph databases as a solution to storing genome analysis results in a meaningful manner, and to allow retrospective comparisons as more genomes are sequenced or populations change.
+Spfy aims to solve the problem of real-time analysis and merging of results from different analysis methods.
+By merging results and creating links between genome data, we are able to identifying relationships between all genomes sequenced in the past, present, and future.
+This graph-based result storage allows retrospective comparisons as more genomes are sequenced or populations change, and is flexible enough to accomadate new analysis methods as they are developed.
 
-{I'm not sure if we should include that the same graph database can be expanded to cover different spcies}
-
-{needs to be beefed up, in language a biologist / public health worker would care about}
-
-\par
-
-% 4. benefits: rapid analyses in real-time -> huge comparisons, replace reference labs -> time & money saved, future work -> expand analyses, more genomes, more species
-By integrating task distribution with graph storage, Spfy enables large-scale analyses, such as epidemiological associations between specific genotypes, biomarkers, host, source, and other metadata, and statistical significance testing of genome markers for user-defined groups.
-Subtyping options are ..., pan-genome generation ..., group comparisons via Fisher's, ML, .... for E.coli.
+% {needs to be beefed up, in language a biologist / public health worker would care about}
 
 By supporting multiple \textit{in-silico} subtyping options, the platform functions similar to a reference laboratory, with added support for big-data analyses.
+Subtyping options are ..., pan-genome generation ..., group comparisons via Fisher's, ML, ....
+for E.coli.
+These tasks are divided into subtasks, and distributed across a built-in task queue.
+Subtyping results are then decomposed into graph structures and stored within a larger graph database.
+By integrating task distribution with graph storage, Spfy enables large-scale analyses, such as epidemiological associations between specific genotypes, biomarkers, host, source, and other metadata, and statistical significance testing of genome markers for user-defined groups.
 Currently, the platform has been tested with XXX genome files and result storage for XX analyses modules.
@@
 
-The website and source code are available at \url{https://superphy.github.io/}. \par
+The website and source code are available at \url{https://lfz.corefacility.ca/superphy/spfy/}. \par
 

### 2017-11-16 paper-webserver.tex
STOP: about to merge into master, no real pt on a branch atm
EDIT: read-through of intro

@@
 The rapid characterization and comparison of bacterial pathogens relies principally on the combination of outputs from multiple software programs that are targeted for specific applications.
-Examples include the Resistance Gene Identifier (RGI) \cite{mcarthur2013comprehensive} for antimicrobial resistance (AMR) gene prediction, VirulenceFinder \cite{kleinheinz2014applying} virulence gene annotation, and Prokka for bacterial genome annotation with external tools \cite{doi:10.1093/bioinformatics/btu153}.
-In particular, RGI and VirulenceFinder represent a series of \textit{in-silico} methods which have been developed to mimick the results of traditional tests.
-These methods allow new WGS results to be viewed in the context of historical tests.
+Examples include the Resistance Gene Identifier (RGI) \cite{mcarthur2013comprehensive} for antimicrobial resistance (AMR) gene prediction, VirulenceFinder \cite{kleinheinz2014applying} for virulence factor (VF) annotation, and Prokka for bacterial genome annotation with external tools \cite{doi:10.1093/bioinformatics/btu153}.
+In particular, RGI and VirulenceFinder represent a series of \textit{in-silico} methods which have been developed to replicate the results of traditional wet-lab tests; this allows new WGS results to be viewed in the context of historical tests.
 
 Comprehensive platforms that combine individual programs into a cohesive whole also exist.
-These include free platforms such as the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial}, the Integrated Rapid Infectious Disease Analysis (IRIDA) project \url{http://www.irida.ca/}, and PATRIC.
-Commercial applications, such as Bionumerics, which is used by PulseNet international for the analyses of WGS data in outbreak situations also exist, and offer support as well as accredited, standardized tests {REF}.
+These include free platforms such as the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial}, the Integrated Rapid Infectious Disease Analysis (IRIDA) project \url{http://www.irida.ca/}, and PATRIC \cite{wattam2013patric}.
+Commercial applications, such as Bionumerics, which is used by PulseNet international for the analyses of WGS data in outbreak situations also exist, and offer support as well as accredited, standardized tests \cite{swaminathan2001pulsenet}.
 
@@
 Spfy aims to solve the problem of real-time analysis and merging of results from different analysis methods.
-By merging results and creating links between genome data, we are able to identifying relationships between all genomes sequenced in the past, present, and future.
+By merging results and creating links between genome data, Spfy is able to identify relationships between all genomes sequenced in the past, present, and future.
 This graph-based result storage allows retrospective comparisons as more genomes are sequenced or populations change, and is flexible enough to accomadate new analysis methods as they are developed.
@@
 By supporting multiple \textit{in-silico} subtyping options, the platform functions similar to a reference laboratory, with added support for big-data analyses.
-Subtyping options are ..., pan-genome generation ..., group comparisons via Fisher's, ML, ....
-for E.coli.
+Subtyping options for E.coli are O-antigen, H-antigen, Shiga-toxin 1 (Stx1), Shiga-toxin 2 (Stx2), and Intimin typing, VF and AMR annotation, pan-genome generation, group comparisons via Fisher's, and machine-learning (ML) based predictions for OmniLog AMR assays.
 These tasks are divided into subtasks, and distributed across a built-in task queue.
-Subtyping results are then decomposed into graph structures and stored within a larger graph database.
+Results are also decomposed into graph structures and stored within a larger graph database.
 By integrating task distribution with graph storage, Spfy enables large-scale analyses, such as epidemiological associations between specific genotypes, biomarkers, host, source, and other metadata, and statistical significance testing of genome markers for user-defined groups.
@@
 While the integrated approach or storing and retrieving results provides enormous benefits, the developed analyses modules are self-contained and can easily be integrated into existing platforms such as Galaxy \cite{goecks2010galaxy}, and IRIDA.
-
-The website and source code are available at \url{https://lfz.corefacility.ca/superphy/spfy/}. \par
+The website and source code are available at \url{https://lfz.corefacility.ca/superphy/spfy/}.
 
@@
 % ectyper will get its own paper
-
-% para on phylotyper
-Stx typing looks at ...
-Phylotyper constructs phylogenetic trees and uses ancestral reconstruction to determine Stx-types ... others.
-The software can also be used to generate trees and determine closely related strains.
+% phylotyper also has its own paper
 

### 2017-11-16 paper-webserver.tex
EDIT: read-through of the functionality and implenetation paras

@@
 % para covering everything
-Spfy performs reference laboratory tasks: O-antigen typing, H-antigen typing, and virulence factor (VF) gene determination using ECTyper \url{https://github.com/phac-nml/ecoli_serotyping}, Shiga-toxin (STX) typing using Phylotyper \cite{whiteside2017phylotyper}, and antimicrobial resistance (AMR) gene determination using the Resistance Gene Identifier (RGI) program \cite{mcarthur2013comprehensive}.
-
+Spfy performs reference laboratory tasks: O-antigen typing, H-antigen typing, and VF gene determination using ECTyper \url{https://github.com/phac-nml/ecoli_serotyping}, STX typing using Phylotyper \cite{whiteside2017phylotyper}, and AMR gene determination using the RGI program \cite{mcarthur2013comprehensive}.
 Spfy also performs bioinformatics analyses: pangenome generation using Panseq \cite{laing2010pan}, statistical significance testing of genome markers for user-defined groups, and AMR predictions using support vector machines (SVMs).
@@
 % phylotyper also has its own paper
-
-% para on pangenome
-Bacterial species, pan-genome ...Panseq \cite{laing2010pan}, Roary \cite{page2015roary}, chewBBACA \url{https://github.com/mickaelsilva/chewBBACA}, and Graphtyper \cite{Eggertsson148403} can compute these pan-genomic regions, and are freely available.
-Spfy integrates panseq ...
 
@@
 Spfy is built around semantic web technologies which describe the relaions between different data \cite{berners2001semantic}.
-For biological data this means that individual data points such as genome, contiguous DNA sequence, or gene, are linked together in a queryable graph form.
-This allows novel data to be seamlessly incorporated into the existing graph, and has {TODO: what has been proposed specifically?} been proposed as a common standard for the open sharing of data \cite{horrocks2005semantic}.
+For biological data, individual data points such as genome, contiguous DNA sequence, or gene, are linked together in a queryable graph structure.
+This allows novel data to be seamlessly incorporated into the existing graph, and has the use of graph databases for relational information has been proposed as a common standard for the open sharing of data \cite{horrocks2005semantic}.
 
 % para: semantic web in spfy
-In Spfy, the results from all the analysis modules are converted into graph objects, which are used to update the main graph database. TODO: need information about ontology, terms etc. for metadata.
-
+In Spfy, the results from all the analysis modules are converted into graph objects, which are used to update the main graph database.
+These graph objects, which consist of nodes and edges, must be annotated in order to attribute meaning to these data points.
+Spfy uses annotations from the GenEpiO \cite{griffiths2017context}, FALDO \cite{bolleman2016faldo}, and TypOn \cite{vaz2014typon} ontologies, as well as custom terminology to enable graph traversals of related datum.
 The use of semantic web technologies allows results to be linked to the genomes they were computed from, and for all data points to share a common data structure that can be queried together.

### 2017-11-16 paper-webserver.tex
CHANGE: merge the semantic web topics into the Data Storage subsection

@@
 \section{IMPLEMENTATION}
-% para: the semantic web
-Spfy is built around semantic web technologies which describe the relaions between different data \cite{berners2001semantic}.
-For biological data, individual data points such as genome, contiguous DNA sequence, or gene, are linked together in a queryable graph structure.
-This allows novel data to be seamlessly incorporated into the existing graph, and has the use of graph databases for relational information has been proposed as a common standard for the open sharing of data \cite{horrocks2005semantic}.
-
-% para: semantic web in spfy
-In Spfy, the results from all the analysis modules are converted into graph objects, which are used to update the main graph database.
-These graph objects, which consist of nodes and edges, must be annotated in order to attribute meaning to these data points.
-Spfy uses annotations from the GenEpiO \cite{griffiths2017context}, FALDO \cite{bolleman2016faldo}, and TypOn \cite{vaz2014typon} ontologies, as well as custom terminology to enable graph traversals of related datum.
-The use of semantic web technologies allows results to be linked to the genomes they were computed from, and for all data points to share a common data structure that can be queried together.
-
 % para: intro to the spfy stack
-Spfy's server-side code is developed in Python and the website served to users using the React JavaScript library.
-
-For the addition of new data to the database, the following steps are taken:
+Spfy's server-side code is developed in Python and the website served to users using the React JavaScript library. For the addition of new data to the database, the following steps are taken:
 
@@
 % 3 1/2. SPARQL queries
-Graph databases describe relationships between data points, and are one of the emerging \cite{de2015trends} database types for biological data; they are one of the core building blocks of semantic web technology \cite{horrocks2005semantic}.
-Spfy uses the RDFLib Python library \url{https://rdflib.readthedocs.org/} to represent all data, including the results of the analyses modules.
+% para: the semantic web
+Spfy is built around semantic web technologies which describe the relaions between different data \cite{berners2001semantic}.
+For biological data, individual data points such as genome, contiguous DNA sequence, or gene, are linked together in a queryable graph structure.
+This allows novel data to be seamlessly incorporated into the existing graph, and has the use of graph databases for relational information has been proposed as a common standard for the open sharing of data \cite{horrocks2005semantic}.
 
-% note: below is a feature in development
-The permanent storage of results is as a one-time cost, and allows population-wide analyses of all stored genomes; result storage also enables Spfy to avoid recomputation when the same analysis is re-run.
-
-% para
-% ontology
-The structure of the graph database is handled using ontologies: indices which define types of data and the relations between them.
-By sharing ontologies, it is possible to link databases from different organizations together, and Spfy uses types defined in the  Genomic Epidemiology Application Ontology (GenEpiO) \cite{griffiths2017context} and the Feature Annotation Location Description Ontology (FALDO) \cite{bolleman2016faldo}.
-Spfy also connects genomes, contiguous sequences (contigs), pangenome regions, and predicted genes and alleles using generic realtion edges, to allow the presence of various genes to be infered from a genome description.
-(see Figure \ref{fig-ontology})
-Individual genes and pangenome regions are connected to all the contigs they are found on using this inferencing method, such that it is possible to determine all the genomes which contain a particular target sequence.
+% para: semantic web in spfy
+In Spfy, the results from all the analysis modules are converted into graph objects, which are used to update the main graph database.
+These graph objects, which consist of nodes and edges, must be annotated in order to attribute meaning to these data points.
+Spfy uses annotations from the GenEpiO \cite{griffiths2017context}, FALDO \cite{bolleman2016faldo}, and TypOn \cite{vaz2014typon} ontologies, as well as custom terminology to enable graph traversals and related data to be inferred from the ontology (see Figure \ref{fig-ontology}).
+The use of semantic web technologies allows results to be linked to the genomes they were computed from, and for all data points to share a common data structure that can be queried together.
 
@@
 \end{figure}
+
+% note: below is a feature in development
+The permanent storage of results is as a one-time cost, and allows population-wide analyses of all stored genomes; result storage also enables Spfy to avoid recomputation when the same analysis is re-run.
+To identify analysis tasks, Spfy fingerprints requests by computing its SHA1 hash and tags the results in the graph database with this hash.
+Subsequent queries are then compared through fingerprinting, and results can be rebuilt from the graph by searching for matching fingerpirnts.
+
+For population analyses, Spfy searches the graph for all nodes annotated with the queried ontology term.
+Depending on the given analysis, this data is then typically vectorized into NumPy numerical arrays for use with the SciPy scientific computing library.
 

### 2017-11-18 paper-webserver.tex
ADD: new paragraph on group comparison in Functionality

@@
 % group comparisons
-By retrieving results from the database, Spfy provides group comparisons using Fishers.
-...
-
-% Panpredic???
+Knowledge graphs are a semantic method of linking data points by their relations.
+Spfy allows users to compare the results associated with particular genome types or metadata by storing all results in a graph structure.
+Comparisons can be made in real-time, as new genomes are submitted for analysis and the results stored in the existing graph.
+Any data type in the graph is valid for analysis, such as the presence or absence of pan-genome regions against determined serotypes or other subtyping options.
+In addition, Spfy supports the submission of user-specified metadata, such as location or source information.
 

### 2017-11-24 paper-webserver.tex
START: working on changes to the Discussion

@@
 While the integrated approach or storing and retrieving results provides enormous benefits, the developed analyses modules are self-contained and can easily be integrated into existing platforms such as Galaxy \cite{goecks2010galaxy}, and IRIDA.
-The website and source code are available at \url{https://lfz.corefacility.ca/superphy/spfy/}.
+The website is available at \url{https://lfz.corefacility.ca/superphy/spfy/}.
+Spfy's codebase is provided at \url{https://github.com/superphy/backend} and a developer guide is provided at \url{https://superphy.readthedocs.io/en/latest/}.
 
@@
 \section{DISCUSSION}
+% Gist, we address the the "Limitations" of \cite{de2015trends}.
+% Namely:
+% 1. Custom formats.
+% 2. Use of legacy tech (SQL) means youu can't accomadate "Big-Data" goals.
 
 % drawbacks - big data
-Many previous bioinformatics software programs have been developed \textit{ad hoc}, without the use of software engineering principles \cite{de2015trends}.
+Many previous bioinformatics software programs have been developed \textit{ad hoc}, with inividual researchers and laboratories developing software specific to their environment \cite{de2015trends}.
 Such tools were often script-based, with custom data formats, and only suitable for small collections of data \cite{de2015trends}.
@@
 
-One of the key goals in developing Spfy is to maintain instantaneity, as modern websites have accustomed users to instant results.
+One of the key goals in developing Spfy is to maintain instantaneity, as modern websites have accustomed users to immediate results.
 We attempt to use innovations in web development to bring a similar experience to Spfy as a predictive genomics platform for \textit{E. coli}.
-% Mention where users / developers can find documentation
-Spfy's main documentation and codebase are provided at \url{https://github.com/superphy/backend} and a developer guide is provided at \url{https://superphy.readthedocs.io/en/latest/}.
+The use of Docker containerization, task queues, and a graph database, were all chosen to support \textit{instantaneity}; modern analytics platforms must respond to user-requested analyses of big-data in the same efficiancy as old analyses of single files.
 

### 2017-11-24 paper-webserver.tex
DEL: PFGE

@@
 The isolation and characterization of bacterial pathogens are critical for Public Health laboratories to rapidly respond to outbreaks, and to effectively monitor known and emerging pathogens through surveillance programs.
-Until recently, public-health agencies relied on laboratory tests such as XYZ to characterize bacterial isolates in outbreak and surveillance settings.
+Until recently, public-health agencies relied on laboratory tests such as pulsed-field gel electrophoresis (PFGE) banding patterns {ronholm2016navigating}, XYZ to characterize bacterial isolates in outbreak and surveillance settings.
 
-TODO: The below has previously been covered by many papers, including our previous SuperPhy paper. We need to focus on the big-data aspect.
-%% The previous gold-standard in determining strain relatedness was pulsed-field gel electrophoresis (PFGE) {ronholm2016navigating}, which uses rare-cutting restriction enzymes to produce a unique banding pattern for each strain.
-%% However, in \textit{Enterococcus faecium}, PFGE has been shown to misclassify 9 of 132 isolates, when compared to whole-genome sequencing (WGS) based discrimination \cite{pinholt2015multiple}.
-%% In \textit{Klebsiella pneumoniae} \cite{marsh2015genomic}, \textit{Yersinia enterocolitica} \cite{gilpin2014limitations}, and \textit{Staphylococcus aureus} \cite{doi:10.1093/ofid/ofu096}, WGS was used to discriminate isolates after initial clustering by PFGE resulted in indistinguishable samples.
-%% Examination of PFGE bands are also subjective, difficult to share \cite{lytsy2017time}, and collative platforms such as PulseNet reported \cite{gilpin2014limitations} that even after collecting 72\% of \textit{Campylobacter jejuni} in a given year in Minnesota (673 c
-%% ases), 87\% of isolates could not be linked by PFGE pattern.
 Antimicrobial resistance testing, virulence factor testing ...

### 2017-11-27 paper-webserver.tex
ADD: some edits to the Impact of Public Health... section

@@
 The isolation and characterization of bacterial pathogens are critical for Public Health laboratories to rapidly respond to outbreaks, and to effectively monitor known and emerging pathogens through surveillance programs.
-Until recently, public-health agencies relied on laboratory tests such as pulsed-field gel electrophoresis (PFGE) banding patterns {ronholm2016navigating}, XYZ to characterize bacterial isolates in outbreak and surveillance settings.
+Until recently, public-health agencies relied on laboratory tests such as serotyping, pulsed-field gel electrophoresis (PFGE) banding patterns {ronholm2016navigating}, or Polymerase Chain Reaction (PCR)-based amplication to identify known VFs or AMR genes, along with other tests, to characterize bacterial isolates in outbreak and surveillance settings.
 
-Antimicrobial resistance testing, virulence factor testing ...
-However, current efforts are focused on predictive genomics, where the relevant phenotypic information can be determined through examination of the whole-genome sequence.
-, and as such can be used to evaluate the spread of outbreaks with better resolution and context than traditional methods \cite{ronholm2016navigating}.
+AMR testing, VF testing, and Shiga-toxin testing are of particular importance to surveillance efforts as the targets play a key role in a pathogen's lethality to humans.
+Current efforts are focused on predictive genomics, where the relevant phenotypic information can be determined through examination of the whole-genome sequence.
+WGS methods allow a single laboratory method - the sequencing step - to provide a variety of analysis options, such as AMR, VF, and Shiga-toxin results, and as such can be used to evaluate the spread of outbreaks with better resolution and context than traditional methods \cite{ronholm2016navigating}.
 
@@
 These results can be shared with other ageencies and researchers over the internet.
-Futhermore, using Spfy's database of pre-processed genomes, Spfy can determine all strains a sample may be related to which is useful for ...
-Cost benefits ...
+Futhermore, using Spfy's database of pre-processed genomes, Spfy can determine all strains a sample may be related to which is useful for monitoring the evolution of pathogens over time.
+A computational approach saves the time and cost associated with performing multiple tests per sample, and allows population comparisons which would not be possible with the current pace of WGS results.
 

### 2017-11-27 paper-webserver.tex
EDIT: some restructuring of the comparison section

@@
 The isolation and characterization of bacterial pathogens are critical for Public Health laboratories to rapidly respond to outbreaks, and to effectively monitor known and emerging pathogens through surveillance programs.
-Until recently, public-health agencies relied on laboratory tests such as serotyping, pulsed-field gel electrophoresis (PFGE) banding patterns {ronholm2016navigating}, or Polymerase Chain Reaction (PCR)-based amplication to identify known VFs or AMR genes, along with other tests, to characterize bacterial isolates in outbreak and surveillance settings.
+Until recently, public-health agencies relied on laboratory tests such as serotyping, pulsed-field gel electrophoresis (PFGE) banding patterns, or Polymerase Chain Reaction (PCR)-based amplication to identify known VFs or AMR genes, along with other tests \cite{ronholm2016navigating}, to characterize bacterial isolates in outbreak and surveillance settings.
 
@@
 \subsection{Comparison with other bioinformatic pipeline technologies}
+% trying to cut down on details that would be better suited for the "Implemntation" section.
 
@@
 Galaxy aims to provide a reproducible compuation-based research environment which is accessible to individuals without programming knowledge.
-TODO: BAP and IRIDA, since they were the examples from above
-IRIDA...
-BAP...
+Galaxy tackles the problem of linking different analysis software together, by defining interdependencies using a custom schema.
+A visual workflow editor is also provided for ease of use.
+IRIDA is build on top of the Galaxy framework, and adds prebuilt pipelines specific to bioinformatics uses, as well as sequence and result storage.
+IRIDA takes a project-based approach, with sequences stored per project, and results store linearly per sequence.
+Focus is given on refined controls for collaborating on projects, and common terms found in the GenEpiO ontology are used to describe results.
 
-The Bacterial Analysis Platform (BAP), developed out of the Technical University of Denmark, is the closest analogue to Spfy, and provides an integrated analysis pipeline for bacterial WGS data as a web service.
-BAP is novel in its combined approach to genome analysis as different programs, such as for VF and AMR determination, are included by default into the pipeline; Spfy provides similar functionallity with an expanded focus on integrated result storage and big-data analyses.
-In place of a MySQL database for storing the location of result files, Spfy parses result data into graph objects and integrates results into a persistant graph database.
-This enables Spfy to create large datasets where analysis results are linked to genomes, and to perform population-wide analyses.
+The Bacterial Analysis Platform (BAP), developed out of the Technical University of Denmark provides an integrated analysis pipeline for bacterial WGS data as a web service.
+BAP takes a combined approach to genome analysis including different programs, such as for VF and AMR determination, by default into the pipeline.
+A record of the different files generated by an analysis are stored in a relational database, and the record is queried to determine which files are returned to the user \cite{thomsen2016bacterial}.
+
+In Spfy, we take a more opionionated approach to result storage and data integration.
+The benefits of real-time analysis pipelines as found in BAP, and the use of the GenEpiO ontology as in IRIDA, are included in Spfy.
 On a per file basis, Spfy performs at a similar speed to BAP on predictive genomics tasks, though Spfy does not provide genome assembly services.
 Spfy also processes XXXX files over XX tasks in XXXX time by using a novel approach of distributing computations over a task queue and multiple Docker compartmentalized containers.
+In place of a relational database for storing the location of result files, Spfy parses result data into graph objects and integrates results into a persistant graph database.
+This enables Spfy to create large datasets where analysis results are linked to genomes, and to perform population-wide analyses.
+In all, Spfy provides similar functionallity with an expanded focus on integrated result storage and big-data analyses.
 

### 2017-11-27 paper-webserver.tex
FIX: mostly spelling

@@
 % para: the semantic web
-Spfy is built around semantic web technologies which describe the relaions between different data \cite{berners2001semantic}.
-For biological data, individual data points such as genome, contiguous DNA sequence, or gene, are linked together in a queryable graph structure.
+Spfy is built around semantic web technologies which describe the relations between different data \cite{berners2001semantic}.
+For biological data, individual data points such as genome, contiguous DNA sequence, or gene, are linked together in a query-able graph structure.
 This allows novel data to be seamlessly incorporated into the existing graph, and has the use of graph databases for relational information has been proposed as a common standard for the open sharing of data \cite{horrocks2005semantic}.
@@
 To identify analysis tasks, Spfy fingerprints requests by computing its SHA1 hash and tags the results in the graph database with this hash.
-Subsequent queries are then compared through fingerprinting, and results can be rebuilt from the graph by searching for matching fingerpirnts.
+Subsequent queries are then compared through fingerprinting, and results can be rebuilt from the graph by searching for matching fingerprints.
 
@@
 % drawbacks - big data
-Many previous bioinformatics software programs have been developed \textit{ad hoc}, with inividual researchers and laboratories developing software specific to their environment \cite{de2015trends}.
+Many previous bioinformatics software programs have been developed \textit{ad hoc}, with individual researchers and laboratories developing software specific to their environment \cite{de2015trends}.
 Such tools were often script-based, with custom data formats, and only suitable for small collections of data \cite{de2015trends}.
@@
 We attempt to use innovations in web development to bring a similar experience to Spfy as a predictive genomics platform for \textit{E. coli}.
-The use of Docker containerization, task queues, and a graph database, were all chosen to support \textit{instantaneity}; modern analytics platforms must respond to user-requested analyses of big-data in the same efficiancy as old analyses of single files.
+The use of Docker containerization, task queues, and a graph database, were all chosen to support \textit{instantaneity}; modern analytics platforms must respond to user-requested analyses of big-data in the same efficiency as old analyses of single files.
 
@@
 The isolation and characterization of bacterial pathogens are critical for Public Health laboratories to rapidly respond to outbreaks, and to effectively monitor known and emerging pathogens through surveillance programs.
-Until recently, public-health agencies relied on laboratory tests such as serotyping, pulsed-field gel electrophoresis (PFGE) banding patterns, or Polymerase Chain Reaction (PCR)-based amplication to identify known VFs or AMR genes, along with other tests \cite{ronholm2016navigating}, to characterize bacterial isolates in outbreak and surveillance settings.
+Until recently, public-health agencies relied on laboratory tests such as serotyping, pulsed-field gel electrophoresis (PFGE) banding patterns, or Polymerase Chain Reaction (PCR)-based amplification to identify known VFs or AMR genes, along with other tests \cite{ronholm2016navigating}, to characterize bacterial isolates in outbreak and surveillance settings.
 
@@
 Spfy uses WGS results.
-After intial sequencing of new isolates, Spfy can be used in place of a traditional reference laborartory, to determine the O-type and H-type, Stx type, and all known VFs and AMR genes in real-time.
-These results can be shared with other ageencies and researchers over the internet.
-Futhermore, using Spfy's database of pre-processed genomes, Spfy can determine all strains a sample may be related to which is useful for monitoring the evolution of pathogens over time.
+After initial sequencing of new isolates, Spfy can be used in place of a traditional reference laboratory, to determine the O-type and H-type, Stx type, and all known VFs and AMR genes in real-time.
+These results can be shared with other agencies and researchers over the internet.
+Furthermore, using Spfy's database of pre-processed genomes, Spfy can determine all strains a sample may be related to which is useful for monitoring the evolution of pathogens over time.
 A computational approach saves the time and cost associated with performing multiple tests per sample, and allows population comparisons which would not be possible with the current pace of WGS results.
@@
 Other scientific workflow technologies such as Galaxy \cite{goecks2010galaxy}, and bioinformatics pipelines IRIDA and BAP are used to run analysis modules on WGS data.
-Galaxy aims to provide a reproducible compuation-based research environment which is accessible to individuals without programming knowledge.
+Galaxy aims to provide a reproducible computation-based research environment which is accessible to individuals without programming knowledge.
 Galaxy tackles the problem of linking different analysis software together, by defining interdependencies using a custom schema.
@@
 
-In Spfy, we take a more opionionated approach to result storage and data integration.
+In Spfy, we take a more opinionated approach to result storage and data integration.
 The benefits of real-time analysis pipelines as found in BAP, and the use of the GenEpiO ontology as in IRIDA, are included in Spfy.
@@
 Spfy also processes XXXX files over XX tasks in XXXX time by using a novel approach of distributing computations over a task queue and multiple Docker compartmentalized containers.
-In place of a relational database for storing the location of result files, Spfy parses result data into graph objects and integrates results into a persistant graph database.
+In place of a relational database for storing the location of result files, Spfy parses result data into graph objects and integrates results into a persistent graph database.
 This enables Spfy to create large datasets where analysis results are linked to genomes, and to perform population-wide analyses.
-In all, Spfy provides similar functionallity with an expanded focus on integrated result storage and big-data analyses.
+In all, Spfy provides similar functionality with an expanded focus on integrated result storage and big-data analyses.
 
@@
 These programs are connected to form a workflow using task queues; a design methodology used to offload long-running or non-instantaneous tasks over available computing infrastructure.
-Unlike platforms such as Galaxy which favor a drag-and-drop design for building workflows, Spfy favors a tightly coupled approach explicitedy linking together different bioinformatics programs within the task queue implemntation.
+Unlike platforms such as Galaxy which favor a drag-and-drop design for building workflows, Spfy favors a tightly coupled approach explicitly linking together different bioinformatics programs within the task queue implementation.
 In particular, Spfy uses asynchronous tasks queues which are tripped in response to user requests, thereby allowing immediate responsivity to users on the website, and processes generating human-readable results can be separated from processes handling long-term result storage.
@@
 One of the key benefits of using more common-place technologies is the compatibility with other infrastructure resources.
-Docker containerization is widely supported by cloud computing services: Amazon Web Services (AWS) \url{https://aws.amazon.com/docker/}, Google Cloud Platform (GCloud) \url{https://cloud.google.com/container-engine/}, and Microsoft Azure \url{https://azure.microsoft.com/en-us/services/container-service/}, and self-hosted cloud computing technologies such as OpenStack \url{https://wiki.openstack.org/wiki/Docker}.
+Docker containerization is widely supported by cloud computing services: Amazon Web Services (AWS) \url{https://aws.amazon.com/docker/}, Google Cloud Platform (GCloud) \url{https://cloud.google.com/container-engine/}, and Microsoft Azure \url{https://azure.microsoft.com/en-us/services/container-service/}, and self-hosted cloud computing technologies such as OpenStack \url{https://wiki.openstack.org/wiki/Docker} all support Docker.
 % we may want to sign up for a free trial to give an example of this
 Spfy packages compute nodes, which manage a collection of task queue workers, as reproducible Docker containers which can be networked together to form a compute cluster.
-Docker containerization has a neglible impact on performance \cite{di2015impact}, and allows the platform to easily scale to demand.
+Docker containerization has a negligible impact on performance \cite{di2015impact}, and allows the platform to easily scale to demand.
 In contrast to Galaxy, Spfy is a targeted approach building a predictive analytics platform for \textit{E.

### 2017-11-28 paper-webserver.tex
EDIT: typos in first para

@@
 An effective method to mitigate the recomputing of analyses is to make the storage and retrieval of results part of the platform, and effectively linked to the organisms of interest with a standardized ontology.
-Such measures can help ensure the rapid response times required for public health applications, and allow results to be integrated and progressively updated as new data become available.
+Such measures can help ensure the rapid response times required for public health applications, and allow results to be integrated and progressively updated as new data becomes available.
 
@@
 By merging results and creating links between genome data, Spfy is able to identify relationships between all genomes sequenced in the past, present, and future.
-This graph-based result storage allows retrospective comparisons as more genomes are sequenced or populations change, and is flexible enough to accomadate new analysis methods as they are developed.
+This graph-based result storage allows retrospective comparisons as more genomes are sequenced or populations change, and is flexible to accommodate new analysis methods as they are developed.
 
@@
 % para covering everything
-Spfy performs reference laboratory tasks: O-antigen typing, H-antigen typing, and VF gene determination using ECTyper \url{https://github.com/phac-nml/ecoli_serotyping}, STX typing using Phylotyper \cite{whiteside2017phylotyper}, and AMR gene determination using the RGI program \cite{mcarthur2013comprehensive}.
+Spfy performs reference laboratory tasks: O-antigen typing, H-antigen typing, and VF gene determination using , STX typing using Phylotyper \cite{whiteside2017phylotyper}, and AMR gene determination using the RGI program \cite{mcarthur2013comprehensive}.
 Spfy also performs bioinformatics analyses: pangenome generation using Panseq \cite{laing2010pan}, statistical significance testing of genome markers for user-defined groups, and AMR predictions using support vector machines (SVMs).

### 2017-11-28 paper-webserver.tex
EDIT: readthrough of functionality section

@@
 % para covering everything
+
+% Need a sentance introducing the tools otherwise this sounds just like the intro section
 Spfy performs reference laboratory tasks: O-antigen typing, H-antigen typing, and VF gene determination using , STX typing using Phylotyper \cite{whiteside2017phylotyper}, and AMR gene determination using the RGI program \cite{mcarthur2013comprehensive}.
@@
 % group comparisons
-Knowledge graphs are a semantic method of linking data points by their relations.
+Knowledge graphs semantically link data points by relations defined in an ontology.
 Spfy allows users to compare the results associated with particular genome types or metadata by storing all results in a graph structure.
 Comparisons can be made in real-time, as new genomes are submitted for analysis and the results stored in the existing graph.
-Any data type in the graph is valid for analysis, such as the presence or absence of pan-genome regions against determined serotypes or other subtyping options.
+Any data type or relation in the graph is valid for analysis, such as the presence or absence of pan-genome regions against determined serotypes or other subtyping options.
 In addition, Spfy supports the submission of user-specified metadata, such as location or source information.
+Results are displayed on-screen and available for download.
 
@@
 % para: intro to the spfy stack
-Spfy's server-side code is developed in Python and the website served to users using the React JavaScript library. For the addition of new data to the database, the following steps are taken:
+Spfy's server-side code is developed in Python and the front-end website is developed using the React JavaScript library. For the addition of new data to the database, the following steps are taken:
 
@@
 The results of these chosen analyses are immediately reported to the user following their completion, while the remaining analyses are subsequently completed and stored in the database without interaction from the user.
-The public web service accepts up to 200 MB of genome files (50 genomes uncompressed, or 120 genomes compressed).
+The public web service accepts up to 200 MB of genome files (50 genomes uncompressed, or 120 genomes compressed) at a time.
 
@@
 % para: the semantic web
-Spfy is built around semantic web technologies which describe the relations between different data \cite{berners2001semantic}.
-For biological data, individual data points such as genome, contiguous DNA sequence, or gene, are linked together in a query-able graph structure.
+Semantic web technologies describe the relations between different data \cite{berners2001semantic}.
+For biological data, individual data points such as genome, contiguous DNA sequence, or gene, can be linked together in a query-able graph structure.
 This allows novel data to be seamlessly incorporated into the existing graph, and has the use of graph databases for relational information has been proposed as a common standard for the open sharing of data \cite{horrocks2005semantic}.
@@
 In Spfy, the results from all the analysis modules are converted into graph objects, which are used to update the main graph database.
-These graph objects, which consist of nodes and edges, must be annotated in order to attribute meaning to these data points.
-Spfy uses annotations from the GenEpiO \cite{griffiths2017context}, FALDO \cite{bolleman2016faldo}, and TypOn \cite{vaz2014typon} ontologies, as well as custom terminology to enable graph traversals and related data to be inferred from the ontology (see Figure \ref{fig-ontology}).
+These graph objects, which consist of nodes and edges, must be annotated with common terminology.
+Spfy uses annotations from the GenEpiO \cite{griffiths2017context}, FALDO \cite{bolleman2016faldo}, and TypOn \cite{vaz2014typon} ontologies, as well as custom terminology to enable graph traversals and related data points to be inferred from the ontology (see Figure \ref{fig-ontology}).
 The use of semantic web technologies allows results to be linked to the genomes they were computed from, and for all data points to share a common data structure that can be queried together.
@@
 (see Figure \ref{fig-results})
-Both the task selection and results page follow this card-based design: while data storage is actually graph-based, the results of various analysis modules are presented to users in a familiar tabular structure and available for download as .csv files.
+Both the task selection and result displays follow this card-based design: while data storage is graph-based, the results of various analysis modules are presented to users in a familiar tabular structure and available for download as .csv spreadsheet files.
 (see Figure \ref{fig-tables})
@@
 Docker \url{https://www.docker.com/} is a virtualization technology to run self-contained operating systems on the same host computer.
+lightweight...
 The Spfy platform depends on a series of webservers, databases, and task workers and uses Docker to compartmentalize these services which are then networked together using Docker-Compose \url{https://docs.docker.com/compose/}.
 (see Figure \ref{fig-docker})
-Docker integration ensures that software dependencies, which typically must be manually installed \cite{doi:10.1093/bioinformatics/btu153,laing2010pan,inouye2014srst2,naccache2014cloud}, are handled automatically, and that service runtimes are compartmentalized; this guarantees that code failures do not propagate to other services.
+Docker integration ensures that software dependencies, which are typically manually installed \cite{doi:10.1093/bioinformatics/btu153,laing2010pan,inouye2014srst2,naccache2014cloud}, are instead handled automatically.
+With compartmentalization of service runtimes, code failures can not propagate to other services.
 

### 2017-11-30 paper-webserver.tex
ADD: connecting sentances

@@
 % Need a sentance introducing the tools otherwise this sounds just like the intro section
-Spfy performs reference laboratory tasks: O-antigen typing, H-antigen typing, and VF gene determination using , STX typing using Phylotyper \cite{whiteside2017phylotyper}, and AMR gene determination using the RGI program \cite{mcarthur2013comprehensive}.
-Spfy also performs bioinformatics analyses: pangenome generation using Panseq \cite{laing2010pan}, statistical significance testing of genome markers for user-defined groups, and AMR predictions using support vector machines (SVMs).
+Prebuilt tools are available for a variety of reference laboratory tasks.
+Spfy performs reference laboratory tasks: O-antigen typing, H-antigen typing, and VF gene determination using ECtyper \url{https://github.com/phac-nml/ecoli_serotyping}, STX typing using Phylotyper \cite{whiteside2017phylotyper}, and AMR gene determination using the RGI program \cite{mcarthur2013comprehensive}.
+Spfy also performs bioinformatics analyses: pangenome generation using Panseq \cite{laing2010pan}, statistical significance testing of genome markers for user-defined groups using SciPy \cite{jones2014scipy}, and support vector machine (SVM)-backed AMR predictions using Scikit-learn \cite{pedregosa2011scikit}.
+These modules are used for the data-generation aspects of Spfy.
 
@@
 
-\subsection{Real-time analysis pipelines}
+\subsection{Service Virtualization}
 % para
@@
 
-Docker \url{https://www.docker.com/} is a virtualization technology to run self-contained operating systems on the same host computer.
-lightweight...
+% cost of docker < a full VM
+% it's not quite a full OS
+Docker \url{https://www.docker.com/} is a virtualization technology to simulate self-contained operating systems on the same host computer, without the overhead of full hardware virtualization \cite{felter2015updated}.
 The Spfy platform depends on a series of webservers, databases, and task workers and uses Docker to compartmentalize these services which are then networked together using Docker-Compose \url{https://docs.docker.com/compose/}.

### 2017-11-30 paper-webserver.tex
EDIT: readthrough up until CI

--- 2017-11-28 paper-webserver.tex
@@
 Whole genome sequencing (WGS) can in theory provide the entire genetic content of an organism.
-This unparalleled resolution and sensitivity has recently transformed public-health surveillance and outbreak response \cite{ronholm2016navigating,lytsy2017time}.
-Additionally, the identification of novel disease mechanisms \cite{wang2014whole,yuen2015whole}, and rapid clinical diagnoses and reference lab tests based on the specific mechanism of disease are now possible.
-\cite{willig2015whole,dewey2014clinical}.
+This unparalleled resolution and sensitivity has recently transformed public-health surveillance and outbreak response \citep{ronholm2016navigating,lytsy2017time}.
+Additionally, the identification of novel disease mechanisms \citep{wang2014whole,yuen2015whole}, and rapid clinical diagnoses and reference lab tests based on the specific mechanism of disease are now possible.
+\citep{willig2015whole,dewey2014clinical}.
 
@@
 The rapid characterization and comparison of bacterial pathogens relies principally on the combination of outputs from multiple software programs that are targeted for specific applications.
-Examples include the Resistance Gene Identifier (RGI) \cite{mcarthur2013comprehensive} for antimicrobial resistance (AMR) gene prediction, VirulenceFinder \cite{kleinheinz2014applying} for virulence factor (VF) annotation, and Prokka for bacterial genome annotation with external tools \cite{doi:10.1093/bioinformatics/btu153}.
+Examples include the Resistance Gene Identifier (RGI) \citep{mcarthur2013comprehensive} for antimicrobial resistance (AMR) gene prediction, VirulenceFinder \citep{kleinheinz2014applying} for virulence factor (VF) annotation, and Prokka for bacterial genome annotation with external tools \citep{doi:10.1093/bioinformatics/btu153}.
 In particular, RGI and VirulenceFinder represent a series of \textit{in-silico} methods which have been developed to replicate the results of traditional wet-lab tests; this allows new WGS results to be viewed in the context of historical tests.
@@
 Comprehensive platforms that combine individual programs into a cohesive whole also exist.
-These include free platforms such as the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial}, the Integrated Rapid Infectious Disease Analysis (IRIDA) project \url{http://www.irida.ca/}, and PATRIC \cite{wattam2013patric}.
-Commercial applications, such as Bionumerics, which is used by PulseNet international for the analyses of WGS data in outbreak situations also exist, and offer support as well as accredited, standardized tests \cite{swaminathan2001pulsenet}.
+These include free platforms such as the Bacterium Analysis Pipeline (BAP) \citep{thomsen2016bacterial}, the Integrated Rapid Infectious Disease Analysis (IRIDA) project \url{http://www.irida.ca/}, and PATRIC \citep{wattam2013patric}.
+Commercial applications, such as Bionumerics, which is used by PulseNet international for the analyses of WGS data in outbreak situations also exist, and offer support as well as accredited, standardized tests \citep{swaminathan2001pulsenet}.
 
@@
 WGS data for bacterial pathogens of public health importance have recently accumulated in public databases in the tens of thousands, with hundreds of thousands set to be available within the next few years.
-For \textit{Escherichia coli}, there are over sixty thousand publicly available genomes in EnteroBase \url{https://enterobase.warwick.ac.uk/} and three million whole genomes in GenBank \cite{doi:10.1093/nar/gks1195}.
+For \textit{Escherichia coli}, there are over sixty thousand publicly available genomes in EnteroBase \url{https://enterobase.warwick.ac.uk/} and three million whole genomes in GenBank \citep{doi:10.1093/nar/gks1195}.
 
@@
 % 4. Our specific problem, previous work
-We have previously developed Superphy \cite{whiteside2016superphy}, an online predictive genomics platform targeting \textit{E.
+We have previously developed Superphy \citep{whiteside2016superphy}, an online predictive genomics platform targeting \textit{E.
 coli}.
@@
 Spfy aims to solve the problem of real-time analysis and merging of results from different analysis methods.
-By merging results and creating links between genome data, Spfy is able to identify relationships between all genomes sequenced in the past, present, and future.
+By merging results and resolving connections between genome data, Spfy is able to identify relationships between all genomes sequenced in the past, present, and future.
 This graph-based result storage allows retrospective comparisons as more genomes are sequenced or populations change, and is flexible to accommodate new analysis methods as they are developed.
@@
 
-By supporting multiple \textit{in-silico} subtyping options, the platform functions similar to a reference laboratory, with added support for big-data analyses.
+By supporting multiple \textit{in-silico} subtyping options, the platform functions similar to a reference laboratory, while adding support for big-data analyses.
 Subtyping options for E.coli are O-antigen, H-antigen, Shiga-toxin 1 (Stx1), Shiga-toxin 2 (Stx2), and Intimin typing, VF and AMR annotation, pan-genome generation, group comparisons via Fisher's, and machine-learning (ML) based predictions for OmniLog AMR assays.
@@
 Results are also decomposed into graph structures and stored within a larger graph database.
-By integrating task distribution with graph storage, Spfy enables large-scale analyses, such as epidemiological associations between specific genotypes, biomarkers, host, source, and other metadata, and statistical significance testing of genome markers for user-defined groups.
+By integrating task distribution with graph storage, Spfy enables large-scale analyses, such as epidemiological associations between specific genotypes, biomarkers, host, source, and other metadata, and statistical significance testing of genome markers for user-defined groups, in real-time.
 Currently, the platform has been tested with XXX genome files and result storage for XX analyses modules.
@@
 Future work will focus on adding additional analyses modules, using machine learning and artificial neural networks to aid genotype to phenotype predictions, and supporting different species.
-While the integrated approach or storing and retrieving results provides enormous benefits, the developed analyses modules are self-contained and can easily be integrated into existing platforms such as Galaxy \cite{goecks2010galaxy}, and IRIDA.
+While the integrated approach of storing and retrieving results provides enormous benefits, the developed analyses modules are self-contained and can easily be integrated into existing platforms such as Galaxy \citep{goecks2010galaxy}, and IRIDA.
 The website is available at \url{https://lfz.corefacility.ca/superphy/spfy/}.
@@
 % Need a sentance introducing the tools otherwise this sounds just like the intro section
-Spfy performs reference laboratory tasks: O-antigen typing, H-antigen typing, and VF gene determination using , STX typing using Phylotyper \cite{whiteside2017phylotyper}, and AMR gene determination using the RGI program \cite{mcarthur2013comprehensive}.
-Spfy also performs bioinformatics analyses: pangenome generation using Panseq \cite{laing2010pan}, statistical significance testing of genome markers for user-defined groups, and AMR predictions using support vector machines (SVMs).
+Prebuilt tools are available for a variety of reference laboratory tasks.
+Spfy performs: O-antigen typing, H-antigen typing, and VF gene determination using ECtyper \url{https://github.com/phac-nml/ecoli\_serotyping}, STX typing using Phylotyper \citep{whiteside2017phylotyper}, and AMR gene determination using the RGI program \citep{mcarthur2013comprehensive}.
+Spfy also performs bioinformatics analyses: pangenome generation using Panseq \citep{laing2010pan}, statistical significance testing of genome markers for user-defined groups using SciPy \citep{jones2014scipy}, and support vector machine (SVM)-backed AMR predictions using Scikit-learn \citep{pedregosa2011scikit}.
+For the larger population comparisons, we store and aggregate the results from individual genome analyses.
 
@@
 % group comparisons
-Knowledge graphs semantically link data points by relations defined in an ontology.
+Knowledge graphs semantically link data points using relations defined in an ontology.
 Spfy allows users to compare the results associated with particular genome types or metadata by storing all results in a graph structure.
@@
 The results of these chosen analyses are immediately reported to the user following their completion, while the remaining analyses are subsequently completed and stored in the database without interaction from the user.
-The public web service accepts up to 200 MB of genome files (50 genomes uncompressed, or 120 genomes compressed) at a time.
+The public web service accepts up to 200 MB of genome files (50 genomes uncompressed, or 120 genomes compressed) at a time, though an unlimited amount of data can be submitted locally.
 
@@
 % para: the semantic web
-Semantic web technologies describe the relations between different data \cite{berners2001semantic}.
+Semantic web technologies describe the relations between different data \citep{berners2001semantic}.
 For biological data, individual data points such as genome, contiguous DNA sequence, or gene, can be linked together in a query-able graph structure.
-This allows novel data to be seamlessly incorporated into the existing graph, and has the use of graph databases for relational information has been proposed as a common standard for the open sharing of data \cite{horrocks2005semantic}.
+This allows novel data to be seamlessly incorporated into the existing graph, and has the use of graph databases for relational information has been proposed as a common standard for the open sharing of data \citep{horrocks2005semantic}.
 
@@
 These graph objects, which consist of nodes and edges, must be annotated with common terminology.
-Spfy uses annotations from the GenEpiO \cite{griffiths2017context}, FALDO \cite{bolleman2016faldo}, and TypOn \cite{vaz2014typon} ontologies, as well as custom terminology to enable graph traversals and related data points to be inferred from the ontology (see Figure \ref{fig-ontology}).
+Spfy uses annotations from the GenEpiO \citep{griffiths2017context}, FALDO \citep{bolleman2016faldo}, and TypOn \citep{vaz2014typon} ontologies, as well as custom terminology to enable graph traversals and related data points to be inferred from the ontology (see Figure \ref{fig-ontology}).
 The use of semantic web technologies allows results to be linked to the genomes they were computed from, and for all data points to share a common data structure that can be queried together.
@@
 For population analyses, Spfy searches the graph for all nodes annotated with the queried ontology term.
-Depending on the given analysis, this data is then typically vectorized into NumPy numerical arrays for use with the SciPy scientific computing library.
+Depending on the given analysis, this data is then vectorized into NumPy numerical arrays for use with the SciPy scientific computing library, or the Scikit-learn machine learning library.
 
@@
 
-Docker \url{https://www.docker.com/} is a virtualization technology to run self-contained operating systems on the same host computer.
-lightweight...
+% cost of docker < a full VM
+% it's not quite a full OS
+Docker \url{https://www.docker.com/} is a virtualization technology to simulate self-contained operating systems on the same host computer, without the overhead of full hardware virtualization \citep{felter2015updated}.
 The Spfy platform depends on a series of webservers, databases, and task workers and uses Docker to compartmentalize these services which are then networked together using Docker-Compose \url{https://docs.docker.com/compose/}.
 (see Figure \ref{fig-docker})
-Docker integration ensures that software dependencies, which are typically manually installed \cite{doi:10.1093/bioinformatics/btu153,laing2010pan,inouye2014srst2,naccache2014cloud}, are instead handled automatically.
-With compartmentalization of service runtimes, code failures can not propagate to other services.
+Docker integration ensures that software dependencies, which are typically manually installed \citep{doi:10.1093/bioinformatics/btu153,laing2010pan,inouye2014srst2,naccache2014cloud}, are instead handled automatically.
+With compartmentalization of service runtimes, code failures do not propagate to other services.
 
@@
 TODO: Update with all of Enterobase
-Spfy was tested with 25,185 public \textit{E.
-coli} genomes, 5,353 genomes from GenBank and 19,832 genomes from Enterobase, and 55,353 generated samples genomes (267GB), storing both the entire sequences and results for all included analysis modules.
+Spfy was tested with 59,5323 public \textit{E.
+coli} genomes, 5,353 genomes from GenBank and 54,181 genomes from Enterobase totaling 596 GB, storing both the entire sequences and results for all included analysis modules.
 The resulting database had XYZ nodes and XYZ edges, with XYZ object properties, which worked out to XYZ GB of data stored.
@@
 \section{DISCUSSION}
-% Gist, we address the the "Limitations" of \cite{de2015trends}.
+% Gist, we address the the "Limitations" of \citep{de2015trends}.
 % Namely:
@@
 % drawbacks - big data
-Many previous bioinformatics software programs have been developed \textit{ad hoc}, with individual researchers and laboratories developing software specific to their environment \cite{de2015trends}.
-Such tools were often script-based, with custom data formats, and only suitable for small collections of data \cite{de2015trends}.
-While this was acceptable for smaller analyses, bioinformatic pipelines utilizing WGS data are larger and involve linked dependencies, which require the application of systems engineering principles \cite{schatz2015biological}.
-Additionally, many subsets of biology now require the analyses of big-data, where the ability to perform computations in real-time, store data in flexible databases, and utilize a common application programming interface (API) linking resources are required \cite{swaminathan2016review}.
+Many previous bioinformatics software programs have been developed \textit{ad hoc}, with individual researchers and laboratories developing software specific to their environment \citep{de2015trends}.
+Such tools were often script-based, with custom data formats, and only suitable for small collections of data \citep{de2015trends}.
+While this was acceptable for smaller analyses, bioinformatic pipelines utilizing WGS data are larger and involve linked dependencies, which require the application of systems engineering principles \citep{schatz2015biological}.
+Additionally, many subsets of biology now require the analyses of big-data, where the ability to perform computations in real-time, store data in flexible databases, and utilize a common application programming interface (API) linking resources are required \citep{swaminathan2016review}.
 
@@
 The isolation and characterization of bacterial pathogens are critical for Public Health laboratories to rapidly respond to outbreaks, and to effectively monitor known and emerging pathogens through surveillance programs.
-Until recently, public-health agencies relied on laboratory tests such as serotyping, pulsed-field gel electrophoresis (PFGE) banding patterns, or Polymerase Chain Reaction (PCR)-based amplification to identify known VFs or AMR genes, along with other tests \cite{ronholm2016navigating}, to characterize bacterial isolates in outbreak and surveillance settings.
+Until recently, public-health agencies relied on laboratory tests such as serotyping, pulsed-field gel electrophoresis (PFGE) banding patterns, or Polymerase Chain Reaction (PCR)-based amplification to identify known VFs or AMR genes, along with other tests \citep{ronholm2016navigating}, to characterize bacterial isolates in outbreak and surveillance settings.
 
@@
 Current efforts are focused on predictive genomics, where the relevant phenotypic information can be determined through examination of the whole-genome sequence.
-WGS methods allow a single laboratory method - the sequencing step - to provide a variety of analysis options, such as AMR, VF, and Shiga-toxin results, and as such can be used to evaluate the spread of outbreaks with better resolution and context than traditional methods \cite{ronholm2016navigating}.
+WGS methods allow a single laboratory method - the sequencing step - to provide a variety of analysis options, such as AMR, VF, and Shiga-toxin results, and as such can be used to evaluate the spread of outbreaks with better resolution and context than traditional methods \citep{ronholm2016navigating}.
 
@@
 % namely galaxy
-Other scientific workflow technologies such as Galaxy \cite{goecks2010galaxy}, and bioinformatics pipelines IRIDA and BAP are used to run analysis modules on WGS data.
+Other scientific workflow technologies such as Galaxy \citep{goecks2010galaxy}, and bioinformatics pipelines IRIDA and BAP are used to run analysis modules on WGS data.
 Galaxy aims to provide a reproducible computation-based research environment which is accessible to individuals without programming knowledge.
@@
 BAP takes a combined approach to genome analysis including different programs, such as for VF and AMR determination, by default into the pipeline.
-A record of the different files generated by an analysis are stored in a relational database, and the record is queried to determine which files are returned to the user \cite{thomsen2016bacterial}.
+A record of the different files generated by an analysis are stored in a relational database, and the record is queried to determine which files are returned to the user \citep{thomsen2016bacterial}.
 
@@
 To approach the challenge of integrating different bioinformatics programs, Spfy instead uses technologies prevalent in common web services not necessarily related to scientific workflows.
-While we prefer to package individual bioinformatics programs using language-agnostic package managers such as Conda \url{https://conda.io/}, we do agree that suites of bioinformatics software benefit from having all dependencies installed and deployable in Docker containers \cite{di2015impact}.
+While we prefer to package individual bioinformatics programs using language-agnostic package managers such as Conda \url{https://conda.io/}, we do agree that suites of bioinformatics software benefit from having all dependencies installed and deployable in Docker containers \citep{di2015impact}.
 These programs are connected to form a workflow using task queues; a design methodology used to offload long-running or non-instantaneous tasks over available computing infrastructure.
@@
 Spfy packages compute nodes, which manage a collection of task queue workers, as reproducible Docker containers which can be networked together to form a compute cluster.
-Docker containerization has a negligible impact on performance \cite{di2015impact}, and allows the platform to easily scale to demand.
+Docker containerization has a negligible impact on performance \citep{di2015impact}, and allows the platform to easily scale to demand.
 In contrast to Galaxy, Spfy is a targeted approach building a predictive analytics platform for \textit{E.
-coli} which leverages generic web technologies to favor a responsive, big-data design strategy tackling the challenge \cite{fricke2014bacterial} of integrating results from multiple genomes and examining shared connections.
+coli} which leverages generic web technologies to favor a responsive, big-data design strategy tackling the challenge \citep{fricke2014bacterial} of integrating results from multiple genomes and examining shared connections.
 By packaging the individual bioinformatics software as Conda packages, they can be easily ported to Galaxy.

### 2017-11-30 paper-webserver.tex
STOP: need to spend some time chopping through the discussion

@@
 Spfy was tested with 59,5323 public \textit{E.
-coli} genomes, 5,353 genomes from GenBank and 54,181 genomes from Enterobase totaling 596 GB, storing both the entire sequences and results for all included analysis modules.
-The resulting database had XYZ nodes and XYZ edges, with XYZ object properties, which worked out to XYZ GB of data stored.
+coli} genomes, 5,353 genomes from GenBank and 54,181 genomes from Enterobase (\~596 GB), storing both the entire sequences and results for all included analysis modules.
+The resulting database had XYZ million nodes and XYZ million edges, with XYZ object properties, which worked out to XYZ TB of data stored.
 
@@
 We attempt to use innovations in web development to bring a similar experience to Spfy as a predictive genomics platform for \textit{E. coli}.
-The use of Docker containerization, task queues, and a graph database, were all chosen to support \textit{instantaneity}; modern analytics platforms must respond to user-requested analyses of big-data in the same efficiency as old analyses of single files.
+The use of Docker containerization, task queues, and a graph database, were all chosen to support instantaneity; modern analytics platforms must respond to user-requested analyses of big-data in the same efficiency as old analyses of single files.
 
@@
 
-AMR testing, VF testing, and Shiga-toxin testing are of particular importance to surveillance efforts as the targets play a key role in a pathogen's lethality to humans.
+AMR testing, VF testing, and Shiga-toxin testing are of particular importance to surveillance efforts due to their role in a pathogen's lethality.
 Current efforts are focused on predictive genomics, where the relevant phenotypic information can be determined through examination of the whole-genome sequence.
@@
 Furthermore, using Spfy's database of pre-processed genomes, Spfy can determine all strains a sample may be related to which is useful for monitoring the evolution of pathogens over time.
-A computational approach saves the time and cost associated with performing multiple tests per sample, and allows population comparisons which would not be possible with the current pace of WGS results.
+A computational approach saves the time and cost associated with performing multiple tests per sample, and allows population comparisons to keep pace the current generation WGS data.
 
@@
 % namely galaxy
-Other scientific workflow technologies such as Galaxy \citep{goecks2010galaxy}, and bioinformatics pipelines IRIDA and BAP are used to run analysis modules on WGS data.
-Galaxy aims to provide a reproducible computation-based research environment which is accessible to individuals without programming knowledge.
+Other scientific workflow technologies such as Galaxy \citep{goecks2010galaxy}, and bioinformatics pipelines IRIDA and BAP also run analysis modules on WGS data.
+Galaxy aims to provide a reproducible, computation-based research environment which is accessible to individuals without programming knowledge.
 Galaxy tackles the problem of linking different analysis software together, by defining interdependencies using a custom schema.
@@
 IRIDA is build on top of the Galaxy framework, and adds prebuilt pipelines specific to bioinformatics uses, as well as sequence and result storage.
-IRIDA takes a project-based approach, with sequences stored per project, and results store linearly per sequence.
-Focus is given on refined controls for collaborating on projects, and common terms found in the GenEpiO ontology are used to describe results.
+IRIDA takes a project-based approach, with sequences stored per project, and results stored linearly per sequence.
+IRIDA adds Controls for collaborating on projects, and uses common terms found in the GenEpiO ontology to describe results.
 
-The Bacterial Analysis Platform (BAP), developed out of the Technical University of Denmark provides an integrated analysis pipeline for bacterial WGS data as a web service.
+The Bacterial Analysis Platform (BAP), developed out of the Technical University of Denmark, provides an integrated analysis pipeline for bacterial WGS data as a web service.
 BAP takes a combined approach to genome analysis including different programs, such as for VF and AMR determination, by default into the pipeline.

### 2017-12-05 paper-proposal.tex
ADD: where the manuscript fits into the proposal

--- 2017-11-30 paper-webserver.tex
@@
-\documentclass[a4,center,fleqn]{NAR}
-
-% Enter dates of publication
-\copyrightyear{2008}
-\pubdate{31 July 2009}
-\pubyear{2009}
-\jvolume{37}
-\jissue{12}
-
 \bibliography{paper-webserver.bib}
 
-%\articlesubtype{This is the article type (optional)}
+\begin{document}
+%contain, at the top, the following affirmative statement.
+"This website is free and open to all users and there is no login requirement." Additionally, any third party software employed by the website that has more restrictive usage terms must be listed.
+This website is free and open to all users and there is no login requirement.
+The code for this webserver and all third party software used by this website is available under the open-source Apache 2.0 or BSD 3-clause licenses.
 
-\begin{document}
+% include the website address; website name; and the names, affiliations, and email addresses of all authors.
 
-\title{Superphy: Real-time prediction of \textit{E. coli} phenotype using graph-based genome sequence storage and retrieval with the \textit{spfy} module}
-
-\author{%
-Corresponding Author\,$^{1,*}$,
-First Co-Author\,$^{2}$
-and Second Co-Author\,$^2$%
-\footnote{To whom correspondence should be addressed.
-Tel: +44 000 0000000; Fax: +44 000 0000000; Email: xxx@yyyy.ac.zz}}
-
-\address{%
-$^{1}$Affiliation of Corresponding Author
-and
-$^{2}$Affiliation of Both Co-Authors}
-% Affiliation must include:
-% Department name, institution name, full road and district address,
-% state, Zip or postal code, country
-
-\history{%
-Received January 1, 2009;
-Revised February 1, 2009;
-Accepted March 1, 2009}
-
-\maketitle
-
-\begin{abstract}
-
-\end{abstract}
-
-\section{Introduction}
-% new outline
-% para
-% 1. brief: WGS is standard
-% 2. Big Problem: but tools are for individual analysis
-% 3. 100,000 genomes, (lookup number for Enterobase, GenBank), how do we run analysis on it
-% 4. previous methods (Galaxy/IRIDA: real-time, but no storage of results, other website examples - Denmark?), what problems they addressed
-% 5. Problems that remain / lack of result storage means: recomputation, lost data, can't reference old analyses, not suited for "big data"
-% 6. Solutions in general: store and increment, huga "big data" analyses, parallelization /queues
-% para
-% 1. Our specific problem, previous work
-% 2. Why solving it is important (for Public health / research)
-% 3. how we solved it
-% para
-% 4. benefits: rapid analyses in real-time -> huge comparisons, replace reference labs -> time & money saved, future work -> expand analyses, more genomes, more species
-% 5. analyses modules -> conda -> IRIDA/Galaxy
-% 6. short snippit on website link & github link
-
-% 1. brief: WGS is standard
-Whole genome sequencing (WGS) can in theory provide the entire genetic content of an organism.
-This unparalleled resolution and sensitivity has recently transformed public-health surveillance and outbreak response \citep{ronholm2016navigating,lytsy2017time}.
-Additionally, the identification of novel disease mechanisms \citep{wang2014whole,yuen2015whole}, and rapid clinical diagnoses and reference lab tests based on the specific mechanism of disease are now possible.
-\citep{willig2015whole,dewey2014clinical}.
-
-% 2. Big Problem: but tools are for individual analysis
-The rapid characterization and comparison of bacterial pathogens relies principally on the combination of outputs from multiple software programs that are targeted for specific applications.
-Examples include the Resistance Gene Identifier (RGI) \citep{mcarthur2013comprehensive} for antimicrobial resistance (AMR) gene prediction, VirulenceFinder \citep{kleinheinz2014applying} for virulence factor (VF) annotation, and Prokka for bacterial genome annotation with external tools \citep{doi:10.1093/bioinformatics/btu153}.
-In particular, RGI and VirulenceFinder represent a series of \textit{in-silico} methods which have been developed to replicate the results of traditional wet-lab tests; this allows new WGS results to be viewed in the context of historical tests.
-
-Comprehensive platforms that combine individual programs into a cohesive whole also exist.
-These include free platforms such as the Bacterium Analysis Pipeline (BAP) \citep{thomsen2016bacterial}, the Integrated Rapid Infectious Disease Analysis (IRIDA) project \url{http://www.irida.ca/}, and PATRIC \citep{wattam2013patric}.
-Commercial applications, such as Bionumerics, which is used by PulseNet international for the analyses of WGS data in outbreak situations also exist, and offer support as well as accredited, standardized tests \citep{swaminathan2001pulsenet}.
-
-% 3. 100,000 genomes, (lookup number for Enterobase, GenBank), how do we run analysis on it
-WGS data for bacterial pathogens of public health importance have recently accumulated in public databases in the tens of thousands, with hundreds of thousands set to be available within the next few years.
-For \textit{Escherichia coli}, there are over sixty thousand publicly available genomes in EnteroBase \url{https://enterobase.warwick.ac.uk/} and three million whole genomes in GenBank \citep{doi:10.1093/nar/gks1195}.
-
-To begin to solve this "big-data" problem, platforms such as BAP and IRIDA provide hosted solutions that compute results in real-time, and distribute analyses across computing resources.
-While effective for self-contained workflows, many of the comparative analyses that are run are broadly useful, and therefore computed multiple times.
-An effective method to mitigate the recomputing of analyses is to make the storage and retrieval of results part of the platform, and effectively linked to the organisms of interest with a standardized ontology.
-Such measures can help ensure the rapid response times required for public health applications, and allow results to be integrated and progressively updated as new data becomes available.
-
-% 4. Our specific problem, previous work
+% include a notification if this is an update from a previous publication in the Web Server issue, and in that case, include an estimate of the number of users and the number of citations.
+% For web servers, or essentially similar web servers, that have been the subject of a previous publication, including publication in journals other than NAR, there is a minimum two-year interval before re-publication in the Web Server Issue.
 We have previously developed Superphy \citep{whiteside2016superphy}, an online predictive genomics platform targeting \textit{E.
@@
 
-% ?. Why solving it is important (for Public health / research)
-% Merging points from here into the above sections (2,3,4) to avoid repetition
-% - everything is being sequenced (surveillance / outbreak / research)
-% - previously mentioned common analyses / want to leverage pre-computed results
-% - WGS does not discard old methods, linkage to thousands of historical results by developing in-silico methods of traditional tests
-% - need fast / standardized outbreak response
-% - need fast / standardized in-silico reference lab
-% - need fast / standardized storage and retrieval of results based on ontology
-% - all known data can be leveraged, allows the most informed decisions possible
-% - etc.
-
-% 5. how we solved it
+% IF THE WEBSITE IMPLEMENTS A META-SERVER OR COMPUTATIONAL WORKFLOW, the summary MUST describe 1) significant added value beyond the simple chaining together of existing third party software or the calculation of a consensus prediction from third party predictors and classifiers; and at least one of the following: 2) how user time for data gathering and multi-step analysis is significantly reduced, or 3) how the website offers significantly enhanced display of the data and results.
 In this study, we present an update to the SuperPhy platform, called Spfy.
@@
 
-% {needs to be beefed up, in language a biologist / public health worker would care about}
-
-By supporting multiple \textit{in-silico} subtyping options, the platform functions similar to a reference laboratory, while adding support for big-data analyses.
-Subtyping options for E.coli are O-antigen, H-antigen, Shiga-toxin 1 (Stx1), Shiga-toxin 2 (Stx2), and Intimin typing, VF and AMR annotation, pan-genome generation, group comparisons via Fisher's, and machine-learning (ML) based predictions for OmniLog AMR assays.
-These tasks are divided into subtasks, and distributed across a built-in task queue.
-Results are also decomposed into graph structures and stored within a larger graph database.
-By integrating task distribution with graph storage, Spfy enables large-scale analyses, such as epidemiological associations between specific genotypes, biomarkers, host, source, and other metadata, and statistical significance testing of genome markers for user-defined groups, in real-time.
-Currently, the platform has been tested with XXX genome files and result storage for XX analyses modules.
-
-Future work will focus on adding additional analyses modules, using machine learning and artificial neural networks to aid genotype to phenotype predictions, and supporting different species.
-While the integrated approach of storing and retrieving results provides enormous benefits, the developed analyses modules are self-contained and can easily be integrated into existing platforms such as Galaxy \citep{goecks2010galaxy}, and IRIDA.
-The website is available at \url{https://lfz.corefacility.ca/superphy/spfy/}.
-Spfy's codebase is provided at \url{https://github.com/superphy/backend} and a developer guide is provided at \url{https://superphy.readthedocs.io/en/latest/}.
-
-% end of new intro
-
-% **************************************************************
-% Keep this command to avoid text of first page running into the
-% first page footnotes
-\enlargethispage{-65.1pt}
-% **************************************************************
-
-% Some general comments
-% The NAR Database issue is more of a showcase then a rigorous exploration of software design choices.
-% Given this focus, i would suggest the following:
-% 1. Increase/highlight the discriptions of the functions and capabilities, maybe by adding a Functionality section
-% 2. In the Implmentation (or Methods) section, only give a cursory description of the layout and components. Don't need to provide too much justification
-% 3. Use the Results section to highlight the scope/size and speed. This can be short
-% 4.
-In the Discussion, this is where i would expand on the justications and reasons for specific design choices.
-Pick 2-3 main ones and discuss those (i.e.
-don't need to justify our choice of documentation software).
-Also compare with other software in Discussion.
-% 5. Add a conclusions section
-
-\section{FUNCTIONALITY}
-% ONLY FOCUS ON THE ANALYSIS MODULES
-% Describe available functions in spfy
-% para covering everything
-
-% Need a sentance introducing the tools otherwise this sounds just like the intro section
-Prebuilt tools are available for a variety of reference laboratory tasks.
-Spfy performs: O-antigen typing, H-antigen typing, and VF gene determination using ECtyper \url{https://github.com/phac-nml/ecoli\_serotyping}, STX typing using Phylotyper \citep{whiteside2017phylotyper}, and AMR gene determination using the RGI program \citep{mcarthur2013comprehensive}.
-Spfy also performs bioinformatics analyses: pangenome generation using Panseq \citep{laing2010pan}, statistical significance testing of genome markers for user-defined groups using SciPy \citep{jones2014scipy}, and support vector machine (SVM)-backed AMR predictions using Scikit-learn \citep{pedregosa2011scikit}.
-For the larger population comparisons, we store and aggregate the results from individual genome analyses.
-
-% para covering ectyper & RGI -- don't really need this
-% RGI has its own paper and was in the previous SuperPhy paper
-% ectyper will get its own paper
-% phylotyper also has its own paper
-
-% group comparisons
-Knowledge graphs semantically link data points using relations defined in an ontology.
-Spfy allows users to compare the results associated with particular genome types or metadata by storing all results in a graph structure.
-Comparisons can be made in real-time, as new genomes are submitted for analysis and the results stored in the existing graph.
-Any data type or relation in the graph is valid for analysis, such as the presence or absence of pan-genome regions against determined serotypes or other subtyping options.
-In addition, Spfy supports the submission of user-specified metadata, such as location or source information.
-Results are displayed on-screen and available for download.
-
-\section{IMPLEMENTATION}
-% para: intro to the spfy stack
-Spfy's server-side code is developed in Python and the front-end website is developed using the React JavaScript library. For the addition of new data to the database, the following steps are taken:
-
-i) The upload begins through the ReactJS-based website, where user-defined analyses options are selected.
-The results of these chosen analyses are immediately reported to the user following their completion, while the remaining analyses are subsequently completed and stored in the database without interaction from the user.
-The public web service accepts up to 200 MB of genome files (50 genomes uncompressed, or 120 genomes compressed) at a time, though an unlimited amount of data can be submitted locally.
-
-ii) User-selected analyses are enqueued into the Redis Queue \url{http://python-rq.org/} task queue.
-Redis Queue consists of a Redis Database \url{https://redis.io/} and task queue workers which run as Python processes.
-
-iii) The workers dequeue the analyses, run them in parallel, and temporarily store results in the Redis database.
-
-iv) Python functions parse the results and permanently store them in Blazegraph \url{https://www.blazegraph.com/}, the graph database used for Superphy.
-
-\subsection{Data Storage}
-% para
-% 0. Goals: big-data, everything linked, easy addition of new links
-% 1. spfy is built around graph technologies
-% 1. how we structure our graph
-% 2. ontoogies used
-% 3. inferencing
-% 3 1/2. SPARQL queries
-% para: the semantic web
-Semantic web technologies describe the relations between different data \citep{berners2001semantic}.
-For biological data, individual data points such as genome, contiguous DNA sequence, or gene, can be linked together in a query-able graph structure.
-This allows novel data to be seamlessly incorporated into the existing graph, and has the use of graph databases for relational information has been proposed as a common standard for the open sharing of data \citep{horrocks2005semantic}.
-
-% para: semantic web in spfy
-In Spfy, the results from all the analysis modules are converted into graph objects, which are used to update the main graph database.
-These graph objects, which consist of nodes and edges, must be annotated with common terminology.
-Spfy uses annotations from the GenEpiO \citep{griffiths2017context}, FALDO \citep{bolleman2016faldo}, and TypOn \citep{vaz2014typon} ontologies, as well as custom terminology to enable graph traversals and related data points to be inferred from the ontology (see Figure \ref{fig-ontology}).
-The use of semantic web technologies allows results to be linked to the genomes they were computed from, and for all data points to share a common data structure that can be queried together.
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics{images/spfy_ontology-1.svg}
-\end{center}
-\caption{Caption for figure within column.}
-\label{fig-ontology}
-\end{figure}
-
-% note: below is a feature in development
-The permanent storage of results is as a one-time cost, and allows population-wide analyses of all stored genomes; result storage also enables Spfy to avoid recomputation when the same analysis is re-run.
-To identify analysis tasks, Spfy fingerprints requests by computing its SHA1 hash and tags the results in the graph database with this hash.
-Subsequent queries are then compared through fingerprinting, and results can be rebuilt from the graph by searching for matching fingerprints.
-
-For population analyses, Spfy searches the graph for all nodes annotated with the queried ontology term.
-Depending on the given analysis, this data is then vectorized into NumPy numerical arrays for use with the SciPy scientific computing library, or the Scikit-learn machine learning library.
-
-\subsection{Web design}
-% para
-% 1. goals: intuitive/familar, ease of use
-% 2. design specs
-% 3. Google Material design
-% 1. implementation: reactjs, react-md, ES6, JSX
-% 4. separation from Flask layer
-
-The front-end website is written using the React JavaScript library \url{https://facebook.github.io/react/} as a single-page application to allow efficient data-flow without reloading the website.
-To ensure a familiar user interface, we followed the Material Design specification \url{https://material.io/}, published by Google, surrounding a card-based design.
-(see Figure \ref{fig-results})
-Both the task selection and result displays follow this card-based design: while data storage is graph-based, the results of various analysis modules are presented to users in a familiar tabular structure and available for download as .csv spreadsheet files.
-(see Figure \ref{fig-tables})
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics{images/results.png}
-\end{center}
-\caption{Caption for figure within column.}
-\label{fig-results}
-\end{figure}
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics{images/tables.png}
-\end{center}
-\caption{Caption for figure within column.}
-\label{fig-tables}
-\end{figure}
-
-\subsection{Service Virtualization}
-% para
-% 1. goals: real-time, support for pipelines (linked modules)
-% 2. how pipelines have been handled in the past: Galaxy, other examples
-% 3. Python, RQ
-% para
-% 1. how we implemented RQ
-%	related: packaging of modules in conda
-
-% para
-% 1. goals: scale analyses to "big-data", error handling
-% 3. how we handle parallelization with RQ
-% para
-% 1. how many tasks have we tested this with
-% 2. error handling: rq-dashboard, sentry
-% 3. why options like sentry are better than traditional logging: scales well to tons (big-data levels) of tasks, groups the same errors together, reporting via email
-
-% para
-% 1. goals: why compartmentilizations
-% 1. how we implemented docker
-% 3. how this lets us replicate worker containers and link everything together
-
-% cost of docker < a full VM
-% it's not quite a full OS
-Docker \url{https://www.docker.com/} is a virtualization technology to simulate self-contained operating systems on the same host computer, without the overhead of full hardware virtualization \citep{felter2015updated}.
-The Spfy platform depends on a series of webservers, databases, and task workers and uses Docker to compartmentalize these services which are then networked together using Docker-Compose \url{https://docs.docker.com/compose/}.
-(see Figure \ref{fig-docker})
-Docker integration ensures that software dependencies, which are typically manually installed \citep{doi:10.1093/bioinformatics/btu153,laing2010pan,inouye2014srst2,naccache2014cloud}, are instead handled automatically.
-With compartmentalization of service runtimes, code failures do not propagate to other services.
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics{images/docker.svg}
-\end{center}
-\caption{Caption for figure within column.}
-\label{fig-docker}
-\end{figure}
-
-\subsection{Continuous integration and testing}
-% Keep this short
-% para
-% 1. goals: why CI, testing is important
-% 2. how we've implemented it, integration with github
-
-TravisCI \url{travisci.io}, a continuous integration (CI) platform, is used to ensure that Spfy \url{https://github.com/superphy/backend} does not break with any changes to the codebase, and runs tests for functionality and backwards compatibility.
-The individual tests use PyTest \url{https://doc.pytest.org/}, and are run within TravisCI's virtual environment, and the current build status can be checked either on our Gtihub repository or at \url{https://travis-ci.org/superphy/backend}.
-CI is also used to automatically build Spfy's core Docker images, and upload them to Docker Hub \url{https://hub.docker.com/u/superphy/}.
-
-\section{RESULTS}
-TODO: Update with all of Enterobase
+% provide descriptions of the input data, the output, and the processing method; complete citations for previous publications of the method or the web server; and two to four keywords.
+Additionally, authors must indicate how long the server has been running, the number of inputs analyzed during testing, and an estimate of the number of individuals outside of the authors' group who have been involved in the testing.
 Spfy was tested with 59,5323 public \textit{E.
@@
 % analysis run-time / throughput with different levels of parallelization
-
-\section{DISCUSSION}
-% Gist, we address the the "Limitations" of \citep{de2015trends}.
-% Namely:
-% 1. Custom formats.
-% 2. Use of legacy tech (SQL) means youu can't accomadate "Big-Data" goals.
-
-% drawbacks - big data
-Many previous bioinformatics software programs have been developed \textit{ad hoc}, with individual researchers and laboratories developing software specific to their environment \citep{de2015trends}.
-Such tools were often script-based, with custom data formats, and only suitable for small collections of data \citep{de2015trends}.
-While this was acceptable for smaller analyses, bioinformatic pipelines utilizing WGS data are larger and involve linked dependencies, which require the application of systems engineering principles \citep{schatz2015biological}.
-Additionally, many subsets of biology now require the analyses of big-data, where the ability to perform computations in real-time, store data in flexible databases, and utilize a common application programming interface (API) linking resources are required \citep{swaminathan2016review}.
-
-One of the key goals in developing Spfy is to maintain instantaneity, as modern websites have accustomed users to immediate results.
-We attempt to use innovations in web development to bring a similar experience to Spfy as a predictive genomics platform for \textit{E. coli}.
-The use of Docker containerization, task queues, and a graph database, were all chosen to support instantaneity; modern analytics platforms must respond to user-requested analyses of big-data in the same efficiency as old analyses of single files.
-
-\subsection{Impact on Public Health Efforts}
-
-% para
-% focus on application
-The isolation and characterization of bacterial pathogens are critical for Public Health laboratories to rapidly respond to outbreaks, and to effectively monitor known and emerging pathogens through surveillance programs.
-Until recently, public-health agencies relied on laboratory tests such as serotyping, pulsed-field gel electrophoresis (PFGE) banding patterns, or Polymerase Chain Reaction (PCR)-based amplification to identify known VFs or AMR genes, along with other tests \citep{ronholm2016navigating}, to characterize bacterial isolates in outbreak and surveillance settings.
-
-AMR testing, VF testing, and Shiga-toxin testing are of particular importance to surveillance efforts due to their role in a pathogen's lethality.
-Current efforts are focused on predictive genomics, where the relevant phenotypic information can be determined through examination of the whole-genome sequence.
-WGS methods allow a single laboratory method - the sequencing step - to provide a variety of analysis options, such as AMR, VF, and Shiga-toxin results, and as such can be used to evaluate the spread of outbreaks with better resolution and context than traditional methods \citep{ronholm2016navigating}.
-
-% application: results similar to a wet-lab
-Spfy uses WGS results.
-After initial sequencing of new isolates, Spfy can be used in place of a traditional reference laboratory, to determine the O-type and H-type, Stx type, and all known VFs and AMR genes in real-time.
-These results can be shared with other agencies and researchers over the internet.
-Furthermore, using Spfy's database of pre-processed genomes, Spfy can determine all strains a sample may be related to which is useful for monitoring the evolution of pathogens over time.
-A computational approach saves the time and cost associated with performing multiple tests per sample, and allows population comparisons to keep pace the current generation WGS data.
-
-\subsection{Comparison with other bioinformatic pipeline technologies}
-% trying to cut down on details that would be better suited for the "Implemntation" section.
-
-% namely galaxy
-Other scientific workflow technologies such as Galaxy \citep{goecks2010galaxy}, and bioinformatics pipelines IRIDA and BAP also run analysis modules on WGS data.
-Galaxy aims to provide a reproducible, computation-based research environment which is accessible to individuals without programming knowledge.
-Galaxy tackles the problem of linking different analysis software together, by defining interdependencies using a custom schema.
-A visual workflow editor is also provided for ease of use.
-IRIDA is build on top of the Galaxy framework, and adds prebuilt pipelines specific to bioinformatics uses, as well as sequence and result storage.
-IRIDA takes a project-based approach, with sequences stored per project, and results stored linearly per sequence.
-IRIDA adds Controls for collaborating on projects, and uses common terms found in the GenEpiO ontology to describe results.
-
-The Bacterial Analysis Platform (BAP), developed out of the Technical University of Denmark, provides an integrated analysis pipeline for bacterial WGS data as a web service.
-BAP takes a combined approach to genome analysis including different programs, such as for VF and AMR determination, by default into the pipeline.
-A record of the different files generated by an analysis are stored in a relational database, and the record is queried to determine which files are returned to the user \citep{thomsen2016bacterial}.
-
-In Spfy, we take a more opinionated approach to result storage and data integration.
-The benefits of real-time analysis pipelines as found in BAP, and the use of the GenEpiO ontology as in IRIDA, are included in Spfy.
-On a per file basis, Spfy performs at a similar speed to BAP on predictive genomics tasks, though Spfy does not provide genome assembly services.
-Spfy also processes XXXX files over XX tasks in XXXX time by using a novel approach of distributing computations over a task queue and multiple Docker compartmentalized containers.
-In place of a relational database for storing the location of result files, Spfy parses result data into graph objects and integrates results into a persistent graph database.
-This enables Spfy to create large datasets where analysis results are linked to genomes, and to perform population-wide analyses.
-In all, Spfy provides similar functionality with an expanded focus on integrated result storage and big-data analyses.
-
-To approach the challenge of integrating different bioinformatics programs, Spfy instead uses technologies prevalent in common web services not necessarily related to scientific workflows.
-While we prefer to package individual bioinformatics programs using language-agnostic package managers such as Conda \url{https://conda.io/}, we do agree that suites of bioinformatics software benefit from having all dependencies installed and deployable in Docker containers \citep{di2015impact}.
-These programs are connected to form a workflow using task queues; a design methodology used to offload long-running or non-instantaneous tasks over available computing infrastructure.
-Unlike platforms such as Galaxy which favor a drag-and-drop design for building workflows, Spfy favors a tightly coupled approach explicitly linking together different bioinformatics programs within the task queue implementation.
-In particular, Spfy uses asynchronous tasks queues which are tripped in response to user requests, thereby allowing immediate responsivity to users on the website, and processes generating human-readable results can be separated from processes handling long-term result storage.
-
-% standard web tech means you can deploy to different cloud computing services
-One of the key benefits of using more common-place technologies is the compatibility with other infrastructure resources.
-Docker containerization is widely supported by cloud computing services: Amazon Web Services (AWS) \url{https://aws.amazon.com/docker/}, Google Cloud Platform (GCloud) \url{https://cloud.google.com/container-engine/}, and Microsoft Azure \url{https://azure.microsoft.com/en-us/services/container-service/}, and self-hosted cloud computing technologies such as OpenStack \url{https://wiki.openstack.org/wiki/Docker} all support Docker.
-% we may want to sign up for a free trial to give an example of this
-Spfy packages compute nodes, which manage a collection of task queue workers, as reproducible Docker containers which can be networked together to form a compute cluster.
-Docker containerization has a negligible impact on performance \citep{di2015impact}, and allows the platform to easily scale to demand.
-In contrast to Galaxy, Spfy is a targeted approach building a predictive analytics platform for \textit{E.
-coli} which leverages generic web technologies to favor a responsive, big-data design strategy tackling the challenge \citep{fricke2014bacterial} of integrating results from multiple genomes and examining shared connections.
-By packaging the individual bioinformatics software as Conda packages, they can be easily ported to Galaxy.
-
-\section{CONCLUSIONS}
-
-\section{ACKNOWLEDGEMENTS}
-
-\subsubsection{Conflict of interest statement.} None declared.
-\newpage
-
-\begin{thebibliography}{4}
-
-% Format for article
-\bibitem{1}
-Author,A.B. and Author,C. (1992)
-Article title.
-\textit{Abbreviated Journal Name}, \textbf{5}, 300--330.
-
-% Format for book
-\bibitem{2}
-Author,D., Author,E.F. and Author,G. (1995)
-\textit{Book Title}.
-Publisher Name, Publisher Address.
-
-% Format for chapter in book
-\bibitem{3}
-Author,H. and Author,I. (2005)
-Chapter title.
-In
-Editor,A. and Editor,B. (eds),
-\textit{Book Title},
-Publisher Name, Publisher Address,
-pp.\ 60--80.
-
-% Another article
-\bibitem{4}
-Author,Y. and Author,Z. (2002)
-Article title.
-\textit{Abbreviated Journal Name}, \textbf{53}, 500--520.
-
-\end{thebibliography}
-
-\end{document}

### 2017-12-05 paper-proposal.tex
STOP: copy and pasting

@@
 This website is free and open to all users and there is no login requirement.
-The code for this webserver and all third party software used by this website is available under the open-source Apache 2.0 or BSD 3-clause licenses.
+The code for this webserver and all third party software used by this website are available under the open-source Apache 2.0 or BSD 3-clause licenses.
 
 % include the website address; website name; and the names, affiliations, and email addresses of all authors.
+The website is available at \url{https://lfz.corefacility.ca/superphy/spfy/}.
+Spfy's codebase is provided at \url{https://github.com/superphy/backend} and a developer guide is provided at \url{https://superphy.readthedocs.io/en/latest/}.
 

### 2017-12-05 paper-proposal.tex
STOP: going to start with spfy first

@@
 The website is available at \url{https://lfz.corefacility.ca/superphy/spfy/}.
-Spfy's codebase is provided at \url{https://github.com/superphy/backend} and a developer guide is provided at \url{https://superphy.readthedocs.io/en/latest/}.
+Spfy's codebase is provided at \url{https://github.com/superphy/backend} and the documentation at \url{https://superphy.readthedocs.io/en/latest/}.
 
@@
 Superphy integrates pre-computed results with domain-specific knowledge to provide real-time exploration of publicly available genomes.
+Our update, called Spfy, adds real-time analysis and merging of results from different analysis methods
+
 While this tool has been useful for the thousands of pre-computed genomes in its database, the current pace of genome sequencing requires real-time predictive genomic analyses of tens-, and soon hundreds-of-thousands of genomes, and the long term storage and referencing of these results, something that the original SuperPhy platform was incapable of.
@@
 % IF THE WEBSITE IMPLEMENTS A META-SERVER OR COMPUTATIONAL WORKFLOW, the summary MUST describe 1) significant added value beyond the simple chaining together of existing third party software or the calculation of a consensus prediction from third party predictors and classifiers; and at least one of the following: 2) how user time for data gathering and multi-step analysis is significantly reduced, or 3) how the website offers significantly enhanced display of the data and results.
-In this study, we present an update to the SuperPhy platform, called Spfy.
-Spfy aims to solve the problem of real-time analysis and merging of results from different analysis methods.
 By merging results and resolving connections between genome data, Spfy is able to identify relationships between all genomes sequenced in the past, present, and future.

### 2017-12-06 paper-proposal.tex
ADD: new structure and DEL: old matter

@@
 This website is free and open to all users and there is no login requirement.
-The code for this webserver and all third party software used by this website are available under the open-source Apache 2.0 or BSD 3-clause licenses.
+The code for this webserver and all third party software used by this website are available under the open-source Apache 2.0, BSD 3-clause, or similar licenses.
 
@@
 % For web servers, or essentially similar web servers, that have been the subject of a previous publication, including publication in journals other than NAR, there is a minimum two-year interval before re-publication in the Web Server Issue.
-We have previously developed Superphy \citep{whiteside2016superphy}, an online predictive genomics platform targeting \textit{E.
-coli}.
-Superphy integrates pre-computed results with domain-specific knowledge to provide real-time exploration of publicly available genomes.
-Our update, called Spfy, adds real-time analysis and merging of results from different analysis methods
-
-While this tool has been useful for the thousands of pre-computed genomes in its database, the current pace of genome sequencing requires real-time predictive genomic analyses of tens-, and soon hundreds-of-thousands of genomes, and the long term storage and referencing of these results, something that the original SuperPhy platform was incapable of.
-
+Our proposal covers an update to Superphy \citep{whiteside2016superphy}, an online predictive genomics platform targeting \textit{E. coli}.
+The update, called Spfy, demonstrates a novel approach to result storage and retrieval for computational workflows.
+Many of the comparative analyses that are run on current computational workflows chain different analysis software, but lack storage methods which examine the relation between results.
 % IF THE WEBSITE IMPLEMENTS A META-SERVER OR COMPUTATIONAL WORKFLOW, the summary MUST describe 1) significant added value beyond the simple chaining together of existing third party software or the calculation of a consensus prediction from third party predictors and classifiers; and at least one of the following: 2) how user time for data gathering and multi-step analysis is significantly reduced, or 3) how the website offers significantly enhanced display of the data and results.
-By merging results and resolving connections between genome data, Spfy is able to identify relationships between all genomes sequenced in the past, present, and future.
-This graph-based result storage allows retrospective comparisons as more genomes are sequenced or populations change, and is flexible to accommodate new analysis methods as they are developed.
+Instead, we can store output from different software together, in the context of a genome graph and thus link analysis results with epidemiological data.
+In Spfy, we demonstrate the use knowledge graphs and a persistant graph database to store results from a variety of subtyping modules for \textit{E. coli} and integrate external sample information.
+This approach scales to the large number of WGS data being accumulated, and allows comparative analysis between different analysis software.
 
@@
 Spfy was tested with 59,5323 public \textit{E.
-coli} genomes, 5,353 genomes from GenBank and 54,181 genomes from Enterobase (\~596 GB), storing both the entire sequences and results for all included analysis modules.
+coli} assembled genomes, 5,353 genomes from GenBank and 54,181 genomes from Enterobase (\~596 GB), storing both the entire sequences and results for all included analysis modules.
 The resulting database had XYZ million nodes and XYZ million edges, with XYZ object properties, which worked out to XYZ TB of data stored.
+% up-time
+
+% collaborators
+
+% existing
+Existing scientific workflow technologies such as Galaxy \citep{goecks2010galaxy}, and pipelines such as Integrated Rapid Infectious Disease Analysis (IRIDA) \url{http://www.irida.ca/} and the Bacterium Analysis Pipeline (BAP) \citep{thomsen2016bacterial} are well established.
+% data integration
+Results from these workflows are stored in tables, or the file is saved on disk.
+Because output from different analysis software can have different structures and terminology, formats must be converted before performing any sort of comparison.
+Without a unified structure, the number of conversions quickly become impractical broad usage.
+% ontologies
+We did not want to invent a new standard which would simply add to the number of current ones.
+Instead, Spfy uses annotations from the GenEpiO \citep{griffiths2017context}, FALDO \citep{bolleman2016faldo}, and TypOn \citep{vaz2014typon} ontologies which describe biological data.
+
+is flexible to accommodate new analysis methods as they are developed.
 

### 2017-12-06 paper-webserver.tex
ADD: new structure and DEL: old matter

@@
 % namely galaxy
-Other scientific workflow technologies such as Galaxy \citep{goecks2010galaxy}, and bioinformatics pipelines IRIDA and BAP also run analysis modules on WGS data.
+Scientific workflow technologies such as Galaxy \citep{goecks2010galaxy}, and bioinformatics pipelines IRIDA and BAP run analysis modules on WGS data.
 Galaxy aims to provide a reproducible, computation-based research environment which is accessible to individuals without programming knowledge.

### 2017-12-07 paper-proposal.tex
STOP: not sure i want to focus too much on the subtyping in our proposal

@@
 Our proposal covers an update to Superphy \citep{whiteside2016superphy}, an online predictive genomics platform targeting \textit{E. coli}.
-The update, called Spfy, demonstrates a novel approach to result storage and retrieval for computational workflows.
+The update, called Spfy, uses graph datastructures to store and retrieve results for computational workflows.
+We demonstrate how graph datastructures scale to the large number of WGS data being accumulated, and allows statistical significance testing between different analysis software.
+To build these graphs, Spfy first provides real-time subtyping through a ReactJS-based website, where user-defined analyses options are selected. 
+The chosen analyses are ran in parallel using task queues and Docker containerization, and the results immediately displayed to the user following their completion.
+Subtyping options are O-antigen, H-antigen, Shiga-toxin 1 (Stx1), Shiga-toxin 2 (Stx2), and Intimin typing, VF and AMR annotation, and pan-genome generation.
+
 Many of the comparative analyses that are run on current computational workflows chain different analysis software, but lack storage methods which examine the relation between results.
@@
 In Spfy, we demonstrate the use knowledge graphs and a persistant graph database to store results from a variety of subtyping modules for \textit{E. coli} and integrate external sample information.
-This approach scales to the large number of WGS data being accumulated, and allows comparative analysis between different analysis software.
+This approach
 

### 2017-12-07 paper-webserver.tex
STOP: not sure i want to focus too much on the subtyping in our proposal

@@
 i) The upload begins through the ReactJS-based website, where user-defined analyses options are selected.
-The results of these chosen analyses are immediately reported to the user following their completion, while the remaining analyses are subsequently completed and stored in the database without interaction from the user.
+The results of these chosen analyses are immediately dispalyed to the user following their completion, while the remaining analyses are subsequently completed and stored in the database without interaction from the user.
 The public web service accepts up to 200 MB of genome files (50 genomes uncompressed, or 120 genomes compressed) at a time, though an unlimited amount of data can be submitted locally.

### 2017-12-07 paper-proposal.tex
ADD: more on the integrated approach

--- 2017-12-06 paper-proposal.tex
@@
 Our proposal covers an update to Superphy \citep{whiteside2016superphy}, an online predictive genomics platform targeting \textit{E. coli}.
-The update, called Spfy, demonstrates a novel approach to result storage and retrieval for computational workflows.
-Many of the comparative analyses that are run on current computational workflows chain different analysis software, but lack storage methods which examine the relation between results.
+The update, called Spfy, offers real-time subtyping options for \textit{E. coli} and uses graph datastructures to store and retrieve results for additional analyses.
+Many of the comparative analyses that are run on current workflows chain different analysis software, but lack storage methods which understand the relation between results.
 % IF THE WEBSITE IMPLEMENTS A META-SERVER OR COMPUTATIONAL WORKFLOW, the summary MUST describe 1) significant added value beyond the simple chaining together of existing third party software or the calculation of a consensus prediction from third party predictors and classifiers; and at least one of the following: 2) how user time for data gathering and multi-step analysis is significantly reduced, or 3) how the website offers significantly enhanced display of the data and results.
-Instead, we can store output from different software together, in the context of a genome graph and thus link analysis results with epidemiological data.
-In Spfy, we demonstrate the use knowledge graphs and a persistant graph database to store results from a variety of subtyping modules for \textit{E. coli} and integrate external sample information.
-This approach scales to the large number of WGS data being accumulated, and allows comparative analysis between different analysis software.
+By making the storage and retrieval of results part of the platform, and effectively linked to the organisms of interest with a standardized ontology, we can mitigate the recomputing of analyses.
+Integrated data storage will be necessary as WGS data for bacterial pathogens of public health importance have recently accumulated in public databases in the tens of thousands, with hundreds of thousands set to be available within the next few years.
+
+% something demonstrating retrival of data (group comparisons)
+By integrating workflows with result storage, the platform can perform statistical significance testing between any data points in the graph, even as new genomes are submitted for analysis.
+Any data type or relation in the graph is valid for analysis, such as the presence or absence of pan-genome regions against determined serotypes or other subtyping options.
 

### 2017-12-07 paper-proposal.tex
DEL: removed excess parts of the group comparisons

--- 2017-12-06 paper-proposal.tex
@@
 Our proposal covers an update to Superphy \citep{whiteside2016superphy}, an online predictive genomics platform targeting \textit{E. coli}.
-The update, called Spfy, demonstrates a novel approach to result storage and retrieval for computational workflows.
-Many of the comparative analyses that are run on current computational workflows chain different analysis software, but lack storage methods which examine the relation between results.
+The update, called Spfy, offers real-time subtyping options and uses graph datastructures to store and retrieve results for additional analyses.
+Many of the comparative analyses that are run on current workflows chain different analysis software, but lack storage methods which understand the relations between results.
 % IF THE WEBSITE IMPLEMENTS A META-SERVER OR COMPUTATIONAL WORKFLOW, the summary MUST describe 1) significant added value beyond the simple chaining together of existing third party software or the calculation of a consensus prediction from third party predictors and classifiers; and at least one of the following: 2) how user time for data gathering and multi-step analysis is significantly reduced, or 3) how the website offers significantly enhanced display of the data and results.
-Instead, we can store output from different software together, in the context of a genome graph and thus link analysis results with epidemiological data.
-In Spfy, we demonstrate the use knowledge graphs and a persistant graph database to store results from a variety of subtyping modules for \textit{E. coli} and integrate external sample information.
-This approach scales to the large number of WGS data being accumulated, and allows comparative analysis between different analysis software.
+By making the storage and retrieval of results part of the platform, and effectively linked to the organisms of interest with a standardized ontology, we can mitigate the recomputing of analyses.
+We can also perform comparitive analyses between any data points in the pipeline.
+Integrated data storage will be necessary as WGS data for bacterial pathogens of public health importance have accumulated in public databases in the tens of thousands, with hundreds of thousands set to be available within the next few years.
 

### 2017-12-07 paper-proposal.tex
ADD: implementation text

@@
 \begin{document}
-%contain, at the top, the following affirmative statement.
-"This website is free and open to all users and there is no login requirement." Additionally, any third party software employed by the website that has more restrictive usage terms must be listed.
+% "contain, at the top, the following affirmative statement.
+"This website is free and open to all users and there is no login requirement." Additionally, any third party software employed by the website that has more restrictive usage terms must be listed."
 This website is free and open to all users and there is no login requirement.
@@
 
-% include the website address; website name; and the names, affiliations, and email addresses of all authors.
+% "include the website address; website name; and the names, affiliations, and email addresses of all authors."
 The website is available at \url{https://lfz.corefacility.ca/superphy/spfy/}.
@@
 
-% include a notification if this is an update from a previous publication in the Web Server issue, and in that case, include an estimate of the number of users and the number of citations.
-% For web servers, or essentially similar web servers, that have been the subject of a previous publication, including publication in journals other than NAR, there is a minimum two-year interval before re-publication in the Web Server Issue.
+% MAIN CONTENT
+% "include a notification if this is an update from a previous publication in the Web Server issue, and in that case, include an estimate of the number of users and the number of citations."
+% "For web servers, or essentially similar web servers, that have been the subject of a previous publication, including publication in journals other than NAR, there is a minimum two-year interval before re-publication in the Web Server Issue."
 Our proposal covers an update to Superphy \citep{whiteside2016superphy}, an online predictive genomics platform targeting \textit{E. coli}.
-The update, called Spfy, offers real-time subtyping options and uses graph datastructures to store and retrieve results for additional analyses.
-Many of the comparative analyses that are run on current workflows chain different analysis software, but lack storage methods which understand the relations between results.
-% IF THE WEBSITE IMPLEMENTS A META-SERVER OR COMPUTATIONAL WORKFLOW, the summary MUST describe 1) significant added value beyond the simple chaining together of existing third party software or the calculation of a consensus prediction from third party predictors and classifiers; and at least one of the following: 2) how user time for data gathering and multi-step analysis is significantly reduced, or 3) how the website offers significantly enhanced display of the data and results.
+The update, called Spfy, adds real-time subtyping options and uses graph datastructures to store and retrieve results for additional analyses.
+% I'm unsure if we should add more about the subtyping options. For example, see:
+% https://github.com/superphy/paper_platform/commit/c017b1e022d310e16a1433af9d58a73e9550a401
+Many of the comparative analyses that are run on current workflows chain different software, but lack storage methods which understand the relations between results.
+% "IF THE WEBSITE IMPLEMENTS A META-SERVER OR COMPUTATIONAL WORKFLOW, the summary MUST describe 1) significant added value beyond the simple chaining together of existing third party software or the calculation of a consensus prediction from third party predictors and classifiers; and at least one of the following: 2) how user time for data gathering and multi-step analysis is significantly reduced, or 3) how the website offers significantly enhanced display of the data and results."
 By making the storage and retrieval of results part of the platform, and effectively linked to the organisms of interest with a standardized ontology, we can mitigate the recomputing of analyses.
 We can also perform comparitive analyses between any data points in the pipeline.
-Integrated data storage will be necessary as WGS data for bacterial pathogens of public health importance have accumulated in public databases in the tens of thousands, with hundreds of thousands set to be available within the next few years.
+Integrated data storage will be necessary as whole genome sequencing (WGS) data for bacterial pathogens of public health importance have accumulated in public databases in the tens of thousands, with hundreds of thousands set to be available within the next few years.
 
-% provide descriptions of the input data, the output, and the processing method; complete citations for previous publications of the method or the web server; and two to four keywords.
-Additionally, authors must indicate how long the server has been running, the number of inputs analyzed during testing, and an estimate of the number of individuals outside of the authors' group who have been involved in the testing.
+% COMPARED TO EXISTING PLATFORMS
+% This aims to be more of an implementation paragraph.
+Existing scientific workflow technologies such as Galaxy \citep{goecks2010galaxy}, and pipelines such as the Bacterium Analysis Pipeline (BAP) \citep{thomsen2016bacterial} and the Integrated Rapid Infectious Disease Analysis (IRIDA) platform \url{http://www.irida.ca/} help automate the use of WGS data for public-health surveillance.
+% data integration
+Like IRIDA and BAP, Spfy automates workflows for users, and like Galaxy, Spfy uses task queues to distribute selected analysis.
+To these concepts, we add the use of Docker containerization for task queue workers, thus allowing anaylsis software to safely run in parallel.
+The main graph database uses annotations from the GenEpiO \citep{griffiths2017context}, FALDO \citep{bolleman2016faldo}, and TypOn \citep{vaz2014typon} ontologies which describe biological data.
+The entire platform is packaged using Docker-Compose, and can be rebuilt with a simple command.
+
+% STATISTICS{}
+% "provide descriptions of the input data, the output, and the processing method; complete citations for previous publications of the method or the web server; and two to four keywords.
+Additionally, authors must indicate how long the server has been running, the number of inputs analyzed during testing, and an estimate of the number of individuals outside of the authors' group who have been involved in the testing."
 Spfy was tested with 59,5323 public \textit{E.
 coli} assembled genomes, 5,353 genomes from GenBank and 54,181 genomes from Enterobase (\~596 GB), storing both the entire sequences and results for all included analysis modules.
-The resulting database had XYZ million nodes and XYZ million edges, with XYZ object properties, which worked out to XYZ TB of data stored.
+The resulting database had XYZ million nodes and XYZ million edges, with XYZ object properties, which worked out to X TB of data stored.
 % up-time
@@
 
-% existing
-Existing scientific workflow technologies such as Galaxy \citep{goecks2010galaxy}, and pipelines such as Integrated Rapid Infectious Disease Analysis (IRIDA) \url{http://www.irida.ca/} and the Bacterium Analysis Pipeline (BAP) \citep{thomsen2016bacterial} are well established.
-% data integration
-Results from these workflows are stored in tables, or the file is saved on disk.
-Because output from different analysis software can have different structures and terminology, formats must be converted before performing any sort of comparison.
-Without a unified structure, the number of conversions quickly become impractical broad usage.
-% ontologies
-We did not want to invent a new standard which would simply add to the number of current ones.
-Instead, Spfy uses annotations from the GenEpiO \citep{griffiths2017context}, FALDO \citep{bolleman2016faldo}, and TypOn \citep{vaz2014typon} ontologies which describe biological data.
-
-is flexible to accommodate new analysis methods as they are developed.
-
 % analysis run-time / throughput with different levels of parallelization

### 2017-12-07 paper-proposal.tex
ADD: example pdf

@@
+\documentclass{article}
+\usepackage[parfill]{parskip}
+\usepackage[backend=bibtex,style=numeric-comp]{biblatex}
 \bibliography{paper-webserver.bib}
@@
 % "For web servers, or essentially similar web servers, that have been the subject of a previous publication, including publication in journals other than NAR, there is a minimum two-year interval before re-publication in the Web Server Issue."
-Our proposal covers an update to Superphy \citep{whiteside2016superphy}, an online predictive genomics platform targeting \textit{E. coli}.
+Our proposal covers an update to Superphy \cite{whiteside2016superphy}, an online predictive genomics platform targeting \textit{E. coli}.
 The update, called Spfy, adds real-time subtyping options and uses graph datastructures to store and retrieve results for additional analyses.
@@
 By making the storage and retrieval of results part of the platform, and effectively linked to the organisms of interest with a standardized ontology, we can mitigate the recomputing of analyses.
-We can also perform comparitive analyses between any data points in the pipeline.
+We can also perform comparitive analyses between any data points generated in the pipeline.
 Integrated data storage will be necessary as whole genome sequencing (WGS) data for bacterial pathogens of public health importance have accumulated in public databases in the tens of thousands, with hundreds of thousands set to be available within the next few years.
@@
 % This aims to be more of an implementation paragraph.
-Existing scientific workflow technologies such as Galaxy \citep{goecks2010galaxy}, and pipelines such as the Bacterium Analysis Pipeline (BAP) \citep{thomsen2016bacterial} and the Integrated Rapid Infectious Disease Analysis (IRIDA) platform \url{http://www.irida.ca/} help automate the use of WGS data for public-health surveillance.
+Existing scientific workflow technologies such as Galaxy \cite{goecks2010galaxy}, and pipelines such as the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial} and the Integrated Rapid Infectious Disease Analysis (IRIDA) platform \url{http://www.irida.ca/} help automate the use of WGS data for public-health surveillance.
 % data integration
@@
 To these concepts, we add the use of Docker containerization for task queue workers, thus allowing anaylsis software to safely run in parallel.
-The main graph database uses annotations from the GenEpiO \citep{griffiths2017context}, FALDO \citep{bolleman2016faldo}, and TypOn \citep{vaz2014typon} ontologies which describe biological data.
+The main graph database uses annotations from the GenEpiO \cite{griffiths2017context}, FALDO \cite{bolleman2016faldo}, and TypOn \cite{vaz2014typon} ontologies which describe biological data.
 The entire platform is packaged using Docker-Compose, and can be rebuilt with a simple command.
@@
 % analysis run-time / throughput with different levels of parallelization
+\end{document}

### 2017-12-07 paper-proposal.tex
EDIT: some cleanup

@@
 This website is free and open to all users and there is no login requirement.
-The code for this webserver and all third party software used by this website are available under the open-source Apache 2.0, BSD 3-clause, or similar licenses.
+The code for this webserver, and all third party software used, are available under the open-source Apache 2.0, BSD 3-clause, or similar licenses.
 
@@
 The website is available at \url{https://lfz.corefacility.ca/superphy/spfy/}.
-Spfy's codebase is provided at \url{https://github.com/superphy/backend} and the documentation at \url{https://superphy.readthedocs.io/en/latest/}.
+Spfy's code is provided at \url{https://github.com/superphy/backend} and documentation at \url{https://superphy.readthedocs.io/en/latest/}.
 
@@
 By making the storage and retrieval of results part of the platform, and effectively linked to the organisms of interest with a standardized ontology, we can mitigate the recomputing of analyses.
-We can also perform comparitive analyses between any data points generated in the pipeline.
-Integrated data storage will be necessary as whole genome sequencing (WGS) data for bacterial pathogens of public health importance have accumulated in public databases in the tens of thousands, with hundreds of thousands set to be available within the next few years.
+Unlike other methods, graph storage enables comparitive analyses between results generated from any of platform's modules.
+Integrated data storage will be necessary as whole genome sequencing (WGS) data for bacterial pathogens have accumulated in public databases in the tens of thousands, with hundreds of thousands set to be available within the next few years.
 
@@
 Like IRIDA and BAP, Spfy automates workflows for users, and like Galaxy, Spfy uses task queues to distribute selected analysis.
+File uploads begin through the ReactJS-based website, where user-defined analyses options are selected.
 To these concepts, we add the use of Docker containerization for task queue workers, thus allowing anaylsis software to safely run in parallel.
 The main graph database uses annotations from the GenEpiO \cite{griffiths2017context}, FALDO \cite{bolleman2016faldo}, and TypOn \cite{vaz2014typon} ontologies which describe biological data.
-The entire platform is packaged using Docker-Compose, and can be rebuilt with a simple command.
+The entire platform is packaged using Docker-Compose, and can be recreated with a simple command.
 

### 2017-12-07 paper-proposal.tex
Good job. I think we should focus on the "big and small problem" in the introduction, then lead into "in this study ..."  and how we solved the problems, and finally conclude with the benefits spfy offers.

--- 2017-12-07 paper-proposal.tex
@@
 \begin{document}
-% "contain, at the top, the following affirmative statement.
-"This website is free and open to all users and there is no login requirement." Additionally, any third party software employed by the website that has more restrictive usage terms must be listed."
+%contain, at the top, the following affirmative statement.
+"This website is free and open to all users and there is no login requirement." Additionally, any third party software employed by the website that has more restrictive usage terms must be listed.
 This website is free and open to all users and there is no login requirement.
@@
 
-% "include the website address; website name; and the names, affiliations, and email addresses of all authors."
+% include the website address; website name; and the names, affiliations, and email addresses of all authors.
 The website is available at \url{https://lfz.corefacility.ca/superphy/spfy/}.
@@
 
-% MAIN CONTENT
-% "include a notification if this is an update from a previous publication in the Web Server issue, and in that case, include an estimate of the number of users and the number of citations."
-% "For web servers, or essentially similar web servers, that have been the subject of a previous publication, including publication in journals other than NAR, there is a minimum two-year interval before re-publication in the Web Server Issue."
-Our proposal covers an update to Superphy \citep{whiteside2016superphy}, an online predictive genomics platform targeting \textit{E. coli}.
-The update, called Spfy, adds real-time subtyping options and uses graph datastructures to store and retrieve results for additional analyses.
-% I'm unsure if we should add more about the subtyping options. For example, see:
-% https://github.com/superphy/paper_platform/commit/c017b1e022d310e16a1433af9d58a73e9550a401
-Many of the comparative analyses that are run on current workflows chain different software, but lack storage methods which understand the relations between results.
-% "IF THE WEBSITE IMPLEMENTS A META-SERVER OR COMPUTATIONAL WORKFLOW, the summary MUST describe 1) significant added value beyond the simple chaining together of existing third party software or the calculation of a consensus prediction from third party predictors and classifiers; and at least one of the following: 2) how user time for data gathering and multi-step analysis is significantly reduced, or 3) how the website offers significantly enhanced display of the data and results."
-By making the storage and retrieval of results part of the platform, and effectively linked to the organisms of interest with a standardized ontology, we can mitigate the recomputing of analyses.
-We can also perform comparitive analyses between any data points in the pipeline.
-Integrated data storage will be necessary as whole genome sequencing (WGS) data for bacterial pathogens of public health importance have accumulated in public databases in the tens of thousands, with hundreds of thousands set to be available within the next few years.
+% include a notification if this is an update from a previous publication in the Web Server issue, and in that case, include an estimate of the number of users and the number of citations.
+% For web servers, or essentially similar web servers, that have been the subject of a previous publication, including publication in journals other than NAR, there is a minimum two-year interval before re-publication in the Web Server Issue.
+Our proposal covers an update to Superphy \citep{whiteside2016superphy}, an online predictive genomics platform targeting \textit{Escherichia coli}.
+The update, called Spfy, uses graph datastructures to store and retrieve results for computational workflows.
+We demonstrate the ability of graph data structures to scale to the [approximate number, eg.
+greater than 50,000]l of whole-genome sequences  accumulated so far, and show the ability to scale to X genomes.
 
-% COMPARED TO EXISTING PLATFORMS
-% This aims to be more of an implementation paragraph.
-Existing scientific workflow technologies such as Galaxy \citep{goecks2010galaxy}, and pipelines such as the Bacterium Analysis Pipeline (BAP) \citep{thomsen2016bacterial} and the Integrated Rapid Infectious Disease Analysis (IRIDA) platform \url{http://www.irida.ca/} help automate the use of WGS data for public-health surveillance.
-% data integration
-Like IRIDA and BAP, Spfy automates workflows for users, and like Galaxy, Spfy uses task queues to distribute selected analysis.
-To these concepts, we add the use of Docker containerization for task queue workers, thus allowing anaylsis software to safely run in parallel.
-The main graph database uses annotations from the GenEpiO \citep{griffiths2017context}, FALDO \citep{bolleman2016faldo}, and TypOn \citep{vaz2014typon} ontologies which describe biological data.
-The entire platform is packaged using Docker-Compose, and can be rebuilt with a simple command.
+Spfy provides real-time subtyping through a ReactJS-based website, where user-defined analyses options are selected.
+Analyses are run in parallel using task queues and Docker containerization, and the results are immediately displayed to the user following their completion.
+Subtyping options include O-antigen, H-antigen, Shiga-toxin 1 (Stx1), Shiga-toxin 2 (Stx2), and Intimin typing.
+Reference-lab type tests include virulence factor and anti-microbial resistance annotation.
+All genomes are analyzed withing the pan-genome framework of \textit{E.
+coli}.
 
-% STATISTICS{}
-% "provide descriptions of the input data, the output, and the processing method; complete citations for previous publications of the method or the web server; and two to four keywords.
-Additionally, authors must indicate how long the server has been running, the number of inputs analyzed during testing, and an estimate of the number of individuals outside of the authors' group who have been involved in the testing."
-Spfy was tested with 59,5323 public \textit{E.
-coli} assembled genomes, 5,353 genomes from GenBank and 54,181 genomes from Enterobase (\~596 GB), storing both the entire sequences and results for all included analysis modules.
-The resulting database had XYZ million nodes and XYZ million edges, with XYZ object properties, which worked out to X TB of data stored.
+% IF THE WEBSITE IMPLEMENTS A META-SERVER OR COMPUTATIONAL WORKFLOW, the summary MUST describe 1) significant added value beyond the simple chaining together of existing third party software or the calculation of a consensus prediction from third party predictors and classifiers; and at least one of the following: 2) how user time for data gathering and multi-step analysis is significantly reduced, or 3) how the website offers significantly enhanced display of the data and results.
+Current comparative computational workflows chain different analysis software, but lack storage and retrieval methods for generated results.
+
+Within Spfy, we store the output from every analysis, and link them together in the context of a genome graph.
+This graph also stores metadata for each genome, facilitating inquiries ranging from population genomics to epidemiological investigations.
+
+% provide descriptions of the input data, the output, and the processing method; complete citations for previous publications of the method or the web server; and two to four keywords.
+Additionally, authors must indicate how long the server has been running, the number of inputs analyzed during testing, and an estimate of the number of individuals outside of the authors' group who have been involved in the testing.
+Spfy currently contains 59,5323 public \textit{E.
+coli} assembled genomes, 5,353 genomes from GenBank and 54,181 genomes from Enterobase (\~596 GB), storing the genome sequences and results from all analyses.
+
+The resulting database had XYZ million nodes and XYZ million edges, with XYZ object properties, which worked out to XYZ TB of data stored.
 % up-time
@@
 
+% existing
+Existing scientific workflow technologies such as Galaxy \citep{goecks2010galaxy}, and pipelines such as the Integrated Rapid Infectious Disease Analysis (IRIDA) \url{http://www.irida.ca/} and the Bacterium Analysis Pipeline (BAP) \citep{thomsen2016bacterial} are well established.
+% data integration
+Results from these workflows are stored in tables, or the file is saved on disk.
+Because output from different analyses are structured differently using distinct terminology, formats must be converted before they can be compared.
+Without a unified structure, these conversions quickly become impractical for broad usage.
+% ontologies
+To avoid proliferating ontologies, and to allow Spfy to integrate with existing ones, annotations from the GenEpiO \citep{griffiths2017context}, FALDO \citep{bolleman2016faldo}, and TypOn \citep{vaz2014typon} ontologies are used to describe biological data.
+
 % analysis run-time / throughput with different levels of parallelization

### 2017-12-17 paper-proposal.tex
edit

--- 2017-12-07 paper-proposal.tex
@@
 % https://github.com/superphy/paper_platform/commit/c017b1e022d310e16a1433af9d58a73e9550a401
-Many of the comparative analyses that are run on current workflows chain different software, but lack storage methods which understand the relations between results.
+Many of the comparative genomics analysis workflows chain different software, but lack storage methods which capture the relations between results.
 % "IF THE WEBSITE IMPLEMENTS A META-SERVER OR COMPUTATIONAL WORKFLOW, the summary MUST describe 1) significant added value beyond the simple chaining together of existing third party software or the calculation of a consensus prediction from third party predictors and classifiers; and at least one of the following: 2) how user time for data gathering and multi-step analysis is significantly reduced, or 3) how the website offers significantly enhanced display of the data and results."
-By making the storage and retrieval of results part of the platform, and effectively linked to the organisms of interest with a standardized ontology, we can mitigate the recomputing of analyses.
-Unlike other methods, graph storage enables comparitive analyses between results generated from any of platform's modules.
+By making the storage and retrieval of results part of the platform, with data effectively linked to the organisms of interest through a standardized ontology, we can mitigate the recomputing of analyses.
+Graph storage enables direct comparison of analyses results generated from any of platform's modules.
 Integrated data storage will be necessary as whole genome sequencing (WGS) data for bacterial pathogens have accumulated in public databases in the tens of thousands, with hundreds of thousands set to be available within the next few years.

### 2017-12-17 paper-proposal.tex

@@
 The code for this webserver, and all third party software used, are available under the open-source Apache 2.0, BSD 3-clause, or similar licenses.
+\para
 
@@
 Spfy's code is provided at \url{https://github.com/superphy/backend} and documentation at \url{https://superphy.readthedocs.io/en/latest/}.
+\para
 
@@
 % "For web servers, or essentially similar web servers, that have been the subject of a previous publication, including publication in journals other than NAR, there is a minimum two-year interval before re-publication in the Web Server Issue."
-Our proposal covers an update to Superphy \cite{whiteside2016superphy}, an online predictive genomics platform targeting \textit{E. coli}.
-The update, called Spfy, adds real-time subtyping options and uses graph datastructures to store and retrieve results for additional analyses.
+Our proposal covers an update to Superphy \citep{whiteside2016superphy}, an online predictive genomics platform targeting \textit{Escherichia coli}.
+The update, called Spfy, uses graph datastructures to store and retrieve results for computational workflows.
+We demonstrate the ability of graph data structures to scale to the [approximate number, eg.
+greater than 50,000]l of whole-genome sequences  accumulated so far, and show the ability to scale to X genomes.
 % I'm unsure if we should add more about the subtyping options. For example, see:
 % https://github.com/superphy/paper_platform/commit/c017b1e022d310e16a1433af9d58a73e9550a401
-Many of the comparative genomics analysis workflows chain different software, but lack storage methods which capture the relations between results.
+Current comparative computational workflows chain different analysis software, but lack storage and retrieval methods for generated results.
 % "IF THE WEBSITE IMPLEMENTS A META-SERVER OR COMPUTATIONAL WORKFLOW, the summary MUST describe 1) significant added value beyond the simple chaining together of existing third party software or the calculation of a consensus prediction from third party predictors and classifiers; and at least one of the following: 2) how user time for data gathering and multi-step analysis is significantly reduced, or 3) how the website offers significantly enhanced display of the data and results."
 By making the storage and retrieval of results part of the platform, with data effectively linked to the organisms of interest through a standardized ontology, we can mitigate the recomputing of analyses.
-Graph storage enables direct comparison of analyses results generated from any of platform's modules.
+Within Spfy, we store the output from every analysis, and link them together in the context of a genome graph.
+This graph also stores metadata for each genome, facilitating inquiries ranging from population genomics to epidemiological investigations.
 Integrated data storage will be necessary as whole genome sequencing (WGS) data for bacterial pathogens have accumulated in public databases in the tens of thousands, with hundreds of thousands set to be available within the next few years.
+\para
+
+% STATISTICS{}
+% "provide descriptions of the input data, the output, and the processing method; complete citations for previous publications of the method or the web server; and two to four keywords.
+Additionally, authors must indicate how long the server has been running, the number of inputs analyzed during testing, and an estimate of the number of individuals outside of the authors' group who have been involved in the testing."
+Spfy was tested with 59,5323 public \textit{E.
+coli} assembled genomes, 5,353 genomes from GenBank and 54,181 genomes from Enterobase (\~596 GB), storing both the entire sequences and results for all included analysis modules.
+Spfy provides real-time subtyping, and the results are immediately displayed to the user following their completion.
+Subtyping options include O-antigen, H-antigen, Shiga-toxin 1 (Stx1), Shiga-toxin 2 (Stx2), and Intimin typing.
+Reference-lab type tests include virulence factor and anti-microbial resistance annotation.
+All genomes are analyzed withing the pan-genome framework of \textit{E.
+coli}.
+The resulting database had XYZ million nodes and XYZ million edges, with XYZ object properties, which worked out to X TB of data stored. \para
 
@@
 To these concepts, we add the use of Docker containerization for task queue workers, thus allowing anaylsis software to safely run in parallel.
-The main graph database uses annotations from the GenEpiO \cite{griffiths2017context}, FALDO \cite{bolleman2016faldo}, and TypOn \cite{vaz2014typon} ontologies which describe biological data.
-The entire platform is packaged using Docker-Compose, and can be recreated with a simple command.
+To avoid proliferating ontologies, and to allow Spfy to integrate with existing ones, annotations from the GenEpiO \citep{griffiths2017context}, FALDO \citep{bolleman2016faldo}, and TypOn \citep{vaz2014typon} ontologies are used to describe biological data.
+The entire platform is packaged using Docker-Compose, and can be recreated with a simple command. \para
 
-% STATISTICS{}
-% "provide descriptions of the input data, the output, and the processing method; complete citations for previous publications of the method or the web server; and two to four keywords.
-Additionally, authors must indicate how long the server has been running, the number of inputs analyzed during testing, and an estimate of the number of individuals outside of the authors' group who have been involved in the testing."
-Spfy was tested with 59,5323 public \textit{E.
-coli} assembled genomes, 5,353 genomes from GenBank and 54,181 genomes from Enterobase (\~596 GB), storing both the entire sequences and results for all included analysis modules.
-The resulting database had XYZ million nodes and XYZ million edges, with XYZ object properties, which worked out to X TB of data stored.
 % up-time
@@
 % analysis run-time / throughput with different levels of parallelization
+\para
 \end{document}

### 2017-12-17 paper-webserver.tex

--- 2017-12-06 paper-webserver.tex
@@
 Spfy performs: O-antigen typing, H-antigen typing, and VF gene determination using ECtyper \url{https://github.com/phac-nml/ecoli\_serotyping}, STX typing using Phylotyper \citep{whiteside2017phylotyper}, and AMR gene determination using the RGI program \citep{mcarthur2013comprehensive}.
-Spfy also performs bioinformatics analyses: pangenome generation using Panseq \citep{laing2010pan}, statistical significance testing of genome markers for user-defined groups using SciPy \citep{jones2014scipy}, and support vector machine (SVM)-backed AMR predictions using Scikit-learn \citep{pedregosa2011scikit}.
+Spfy also performs bioinformatics analyses: pangenome generation using Panseq \citep{laing2010pan}, statistical significance testing using SciPy \citep{jones2014scipy}, and support vector machine (SVM)-backed AMR predictions using Scikit-learn \citep{pedregosa2011scikit}.
 For the larger population comparisons, we store and aggregate the results from individual genome analyses.

### 2017-12-18 paper-proposal.tex
EDIT: typo

@@
 Our proposal covers an update to Superphy \citep{whiteside2016superphy}, an online predictive genomics platform targeting \textit{Escherichia coli}.
-The update, called Spfy, uses graph datastructures to store and retrieve results for computational workflows.
+The update, called Spfy, uses graph data structures to store and retrieve results for computational workflows.
 We demonstrate the ability of graph data structures to scale to the [approximate number, eg.
@@
 File uploads begin through the ReactJS-based website, where user-defined analyses options are selected.
-To these concepts, we add the use of Docker containerization for task queue workers, thus allowing anaylsis software to safely run in parallel.
+To these concepts, we add the use of Docker containerization for task queue workers, thus allowing analysis software to safely run in parallel.
 To avoid proliferating ontologies, and to allow Spfy to integrate with existing ones, annotations from the GenEpiO \citep{griffiths2017context}, FALDO \citep{bolleman2016faldo}, and TypOn \citep{vaz2014typon} ontologies are used to describe biological data.

### 2017-12-18 paper-proposal.tex
ADD: re matts suggestions

@@
 This graph also stores metadata for each genome, facilitating inquiries ranging from population genomics to epidemiological investigations.
-Integrated data storage will be necessary as whole genome sequencing (WGS) data for bacterial pathogens have accumulated in public databases in the tens of thousands, with hundreds of thousands set to be available within the next few years.
+Integrated data storage will be necessary as whole genome sequencing data for bacterial pathogens have accumulated in public databases in the tens of thousands, with hundreds of thousands set to be available within the next few years.
 \para
@@
 All genomes are analyzed withing the pan-genome framework of \textit{E.
-coli}.
+coli}, and results from different analysis software can be grouped back to the source genome.
 The resulting database had XYZ million nodes and XYZ million edges, with XYZ object properties, which worked out to X TB of data stored. \para
@@
 To these concepts, we add the use of Docker containerization for task queue workers, thus allowing analysis software to safely run in parallel.
+% re: Matt "2. When comparing to IRIDA, BAP etc., can mention some differences with Superphy, e.g. the storage of interim result data that allows downstream integrated analysis"
+For result storage, existing workflow technologies use relational tables \cite{goecks2010galaxy}, or store resulting files to disk \cite{thomsen2016bacterial}.
+Because output from different analyses are structured differently using distinct terminology, formats must be converted before they can be compared.
+Without a unified structure, these conversions quickly become impractical for broad usage.
+Graph-based storage solves this problem.
+% as before
 To avoid proliferating ontologies, and to allow Spfy to integrate with existing ones, annotations from the GenEpiO \citep{griffiths2017context}, FALDO \citep{bolleman2016faldo}, and TypOn \citep{vaz2014typon} ontologies are used to describe biological data.

### 2017-12-21 paper-proposal.tex
EDIT: one suggested word change

@@
 The update, called Spfy, uses graph data structures to store and retrieve results for computational workflows.
-We demonstrate the ability of graph data structures to scale to the [approximate number, eg.
-greater than 50,000]l of whole-genome sequences accumulated so far, and show the ability to scale to X genomes.
+We demonstrate the ability of graph data structures to manage the ~60,000 whole-genome sequences of \textit{E. coli} accumulated so far, and show the ability to scale to X genomes.
 % I'm unsure if we should add more about the subtyping options. For example, see:
@@
 By making the storage and retrieval of results part of the platform, with data effectively linked to the organisms of interest through a standardized ontology, we can mitigate the recomputing of analyses.
-Within Spfy, we store the output from every analysis, and link them together in the context of a genome graph.
+Within Spfy, we store the output from all analyses, and link them together in the context of a genome graph.
 This graph also stores metadata for each genome, facilitating inquiries ranging from population genomics to epidemiological investigations.
@@
 Spfy was tested with 59,5323 public \textit{E.
-coli} assembled genomes, 5,353 genomes from GenBank and 54,181 genomes from Enterobase (\~596 GB), storing both the entire sequences and results for all included analysis modules.
+coli} assembled genomes, 5,353 genomes from GenBank and 54,181 genomes \note{are the GenBank genomes not included in the Enterobase set?} from Enterobase (\~596 GB), storing every sequence and results for all included analysis modules.
 Spfy provides real-time subtyping, and the results are immediately displayed to the user following their completion.
 Subtyping options include O-antigen, H-antigen, Shiga-toxin 1 (Stx1), Shiga-toxin 2 (Stx2), and Intimin typing.
-Reference-lab type tests include virulence factor and anti-microbial resistance annotation.
-All genomes are analyzed withing the pan-genome framework of \textit{E.
-coli}, and results from different analysis software can be grouped back to the source genome.
+Reference-lab tests include virulence factor and anti-microbial resistance annotation.
+All genomes are analyzed within the pan-genome framework of \textit{E.
+coli}, and results from all analyses are automatically associated with the source genome.
 The resulting database had XYZ million nodes and XYZ million edges, with XYZ object properties, which worked out to X TB of data stored. \para
@@
 For result storage, existing workflow technologies use relational tables \cite{goecks2010galaxy}, or store resulting files to disk \cite{thomsen2016bacterial}.
-Because output from different analyses are structured differently using distinct terminology, formats must be converted before they can be compared.
+Because output from these programs is user-specific or transitory, results from identical comparisons are often recomputed.
+Additionally, output from different analyses are structured using distinct terminology and formats, which must be converted before they can be compared.
 Without a unified structure, these conversions quickly become impractical for broad usage.
-Graph-based storage solves this problem.
+Graph-based storage of all results solves these problems.
 % as before

### 2017-12-21 paper-proposal.tex
DEBUG: testing latex build

@@
 The code for this webserver, and all third party software used, are available under the open-source Apache 2.0, BSD 3-clause, or similar licenses.
-\para
+\par
 
@@
 Spfy's code is provided at \url{https://github.com/superphy/backend} and documentation at \url{https://superphy.readthedocs.io/en/latest/}.
-\para
+\par
 
@@
 Integrated data storage will be necessary as whole genome sequencing data for bacterial pathogens have accumulated in public databases in the tens of thousands, with hundreds of thousands set to be available within the next few years.
-\para
+\par
 
@@
 coli}, and results from all analyses are automatically associated with the source genome.
-The resulting database had XYZ million nodes and XYZ million edges, with XYZ object properties, which worked out to X TB of data stored. \para
+The resulting database had XYZ million nodes and XYZ million edges, with XYZ object properties, which worked out to X TB of data stored. \par
 
@@
 To avoid proliferating ontologies, and to allow Spfy to integrate with existing ones, annotations from the GenEpiO \citep{griffiths2017context}, FALDO \citep{bolleman2016faldo}, and TypOn \citep{vaz2014typon} ontologies are used to describe biological data.
-The entire platform is packaged using Docker-Compose, and can be recreated with a simple command. \para
+The entire platform is packaged using Docker-Compose, and can be recreated with a simple command. \par
 
 % up-time
+Superphy has been up since early 2016 and the Spfy update has been up since May 2017.
+The server accepts assembled \textit{E.
+coli} genomes with the \textit{.fasta} or \textit{.fna} extensions.
+Submissions are checked against a reference set of \textit{E.
+coli} gene sequences before running analyses.
+Outputs are displayed on the website in tables and can be downloaded as \textit{.csv} files.
+\par
 
-% collaborators
-
-% analysis run-time / throughput with different levels of parallelization
-\para
+% "and two to four keywords."
+Keywords: Bioinformatics, Graph database, Real-time analysis
 \end{document}

### 2017-12-21 paper-proposal.tex
FIX: im good w these  margins
ADD: current numbers + a pdf

@@
 \usepackage[backend=bibtex,style=numeric-comp]{biblatex}
+\usepackage[letterpaper, portrait, margin=1in]{geometry}
 \bibliography{paper-webserver.bib}
@@
 % "For web servers, or essentially similar web servers, that have been the subject of a previous publication, including publication in journals other than NAR, there is a minimum two-year interval before re-publication in the Web Server Issue."
-Our proposal covers an update to Superphy \citep{whiteside2016superphy}, an online predictive genomics platform targeting \textit{Escherichia coli}.
+Our proposal covers an update to Superphy \cite{whiteside2016superphy}, an online predictive genomics platform targeting \textit{Escherichia coli}.
 The update, called Spfy, uses graph data structures to store and retrieve results for computational workflows.
@@
 Additionally, authors must indicate how long the server has been running, the number of inputs analyzed during testing, and an estimate of the number of individuals outside of the authors' group who have been involved in the testing."
-Spfy was tested with 59,5323 public \textit{E.
-coli} assembled genomes, 5,353 genomes from GenBank and 54,181 genomes \note{are the GenBank genomes not included in the Enterobase set?} from Enterobase (\~596 GB), storing every sequence and results for all included analysis modules.
+Spfy was tested with 4,622 public \textit{E. coli} assembled genomes from Enterobase, storing every sequence and results for all included analysis modules.
 Spfy provides real-time subtyping, and the results are immediately displayed to the user following their completion.
@@
 coli}, and results from all analyses are automatically associated with the source genome.
-The resulting database had XYZ million nodes and XYZ million edges, with XYZ object properties, which worked out to X TB of data stored. \par
+The resulting database had 1,333 nodes and 683,666 leaves, with 374,836,872 object properties. \par
 
@@
 % as before
-To avoid proliferating ontologies, and to allow Spfy to integrate with existing ones, annotations from the GenEpiO \citep{griffiths2017context}, FALDO \citep{bolleman2016faldo}, and TypOn \citep{vaz2014typon} ontologies are used to describe biological data.
+To avoid proliferating ontologies, and to allow Spfy to integrate with existing ones, annotations from the GenEpiO \cite{griffiths2017context}, FALDO \cite{bolleman2016faldo}, and TypOn \cite{vaz2014typon} ontologies are used to describe biological data.
 The entire platform is packaged using Docker-Compose, and can be recreated with a simple command. \par
@@
 % up-time
-Superphy has been up since early 2016 and the Spfy update has been up since May 2017.
+The Spfy update has been up since May 2017 and Superphy has been up since early 2016.
 The server accepts assembled \textit{E.

### 2017-12-21 paper-proposal.tex
ADD: tens of thousnads + downstream

@@
 The update, called Spfy, uses graph data structures to store and retrieve results for computational workflows.
-We demonstrate the ability of graph data structures to manage the ~60,000 whole-genome sequences of \textit{E. coli} accumulated so far, and show the ability to scale to X genomes.
+We demonstrate the ability of graph data structures to manage the tens of thousands of whole-genome \textit{E. coli} sequences accumulated so far, and use graphs for further downstream processing.
 % I'm unsure if we should add more about the subtyping options. For example, see:

### 2017-12-21 paper-proposal.tex
final edits excluding number of genomes.

--- 2017-12-21 paper-proposal.tex
@@
 Our proposal covers an update to Superphy \cite{whiteside2016superphy}, an online predictive genomics platform targeting \textit{Escherichia coli}.
-The update, called Spfy, uses graph data structures to store and retrieve results for computational workflows.
-We demonstrate the ability of graph data structures to manage the ~60,000 whole-genome sequences of \textit{E. coli} accumulated so far, and show the ability to scale to X genomes.
+The update, called Spfy, uses graph data structures to store and retrieve results for computational workflows, facilitating the management and querying of tens of thousands of whole-genome \textit{E.
+coli} sequences, and efficient downstream processing.
 % I'm unsure if we should add more about the subtyping options. For example, see:
@@
 By making the storage and retrieval of results part of the platform, with data effectively linked to the organisms of interest through a standardized ontology, we can mitigate the recomputing of analyses.
-Within Spfy, we store the output from all analyses, and link them together in the context of a genome graph.
+Within Spfy, the output from all analyses is stored, and linked together in the context of a genome graph.
 This graph also stores metadata for each genome, facilitating inquiries ranging from population genomics to epidemiological investigations.
-Integrated data storage will be necessary as whole genome sequencing data for bacterial pathogens have accumulated in public databases in the tens of thousands, with hundreds of thousands set to be available within the next few years.
+Integrated data storage will be necessary as publicly available whole genome sequencing data for bacterial pathogens currently numbers in the tens of thousands, with hundreds of thousands set to be available within the next few years.
 \par
@@
 Spfy provides real-time subtyping, and the results are immediately displayed to the user following their completion.
-Subtyping options include O-antigen, H-antigen, Shiga-toxin 1 (Stx1), Shiga-toxin 2 (Stx2), and Intimin typing.
+Subtyping options include O-antigen, H-antigen, Shiga-toxin 1, Shiga-toxin 2, and Intimin typing.
 Reference-lab tests include virulence factor and anti-microbial resistance annotation.
@@
 File uploads begin through the ReactJS-based website, where user-defined analyses options are selected.
-To these concepts, we add the use of Docker containerization for task queue workers, thus allowing analysis software to safely run in parallel.
+To these concepts we add Docker containerization for task queue workers, allowing analysis software to safely run in parallel.
 % re: Matt "2. When comparing to IRIDA, BAP etc., can mention some differences with Superphy, e.g. the storage of interim result data that allows downstream integrated analysis"

### 2017-12-21 paper-proposal.tex
ADD: NAR layout + citations

--- 2017-12-21 paper-proposal.tex
@@
-\documentclass{article}
+\documentclass[a4,center,fleqn]{NAR}
+\RequirePackage{hyperref}
+\usepackage{lscape}
+\usepackage{graphicx}
+\usepackage{pdflscape}
 \usepackage[parfill]{parskip}
-\usepackage[backend=bibtex,style=numeric-comp]{biblatex}
-\usepackage[letterpaper, portrait, margin=1in]{geometry}
-\bibliography{paper-webserver.bib}
+% \usepackage[backend=bibtex,style=numeric-comp]{biblatex}
+% \usepackage[letterpaper, portrait, margin=1in]{geometry}
 
 \begin{document}
+
+\title{Spfy: bringing real-time, big data batch analyses of E. coli to SuperPhy}
+
+\author{%
+Kevin K Le\,$^{1,*}$,
+Matthew D Whiteside\,$^{1}$,
+Chad R Laing\,$^1$,
+and Victor PJ Gannon\,$^1$%
+\footnote{To whom correspondence should be addressed.
+Tel: +1 403-382-5516; Fax: +1 403-381-1202; Email: chad.r.laing@phac-aspc.gc.ca}}
+
+\address{%
+$^{1}$Laboratory for Foodborne Zoonoses, Public Health Agency of Canada, Twp Rd 9-1, Lethbridge, AB, T1J 3Z4, Canada}
+% Affiliation must include:
+% Department name, institution name, full road and district address,
+% state, Zip or postal code, country
+
+\maketitle
+
 % "contain, at the top, the following affirmative statement.
@@
 % "and two to four keywords."
-Keywords: Bioinformatics, Graph database, Real-time analysis
+Keywords: Comparative genomics analysis, Epidemiology, Microbial genomics
+
+\newpage
+
+\begin{thebibliography}{4}
+
+\bibitem{whiteside2016superphy}
+Whiteside, M.D., Laing, C.R., Manji, A., Kruczkiewicz, P., Taboada, E.N., and Gannon, V.P.J. (2016)
+SuperPhy: predictive genomics for the bacterial pathogen Escherichia coli.
+\textit{BMC Microbiol}, \textbf{16}, 65.
+
+\bibitem{goecks2010galaxy}
+Goecks, J., Nekrutenko, A., Taylor, J. (2010)
+Galaxy: a comprehensive approach for supporting accessible, reproducible, and transparent computational research in the life sciences.
+\textit{Genome Biol}, \textbf{11}, R86.
+
+\bibitem{thomsen2016bacterial}
+Thomsen, M.C.F., Ahrenfeldt, J., Cisneros, J.L.B., Jurtz, V., Larsen, M.V., Hasman, H., Aarestrup, F.M., and Lund, O. (2016)
+A bacterial analysis platform: an integrated system for analysing bacterial whole genome sequencing data for clinical diagnostics and surveillance.
+\textit{PloS One}, \textbf{11}, e0157718.
+
+\bibitem{griffiths2017context}
+Griffiths, E., Dooley, D., Graham, M., Van Domselaar, G., Brinkman, F.S.L., and Hsiao, W.W.L. (2017)
+Context Is Everything: Harmonization of Critical Food Microbiology Descriptors and Metadata for Improved Food Safety and Surveillance
+\textit{Front Microbiol}, \textbf{8}, 1068.
+
+\bibitem{bolleman2016faldo}
+Bolleman, J.T., Mungall, C.J., Strozzi, F., Baran, J., Dumontier, M., Bonnal, R.J.P., Buels, R., Hoehndorf, R., Fujisawa, T., Katayama, T., et al. (2016)
+FALDO: a semantic standard for describing the location of nucleotide and protein feature annotation.
+\textit{BMC J Biomed Sem}, \textbf{7}, 1068.
+
+\bibitem{vaz2014typon}
+Vaz, C., Francisco, A.P., Silva, M., Jolley, K.A., Bray, J.E., Puseele, H., Rothganger, J., Ramirez, M., Carri\c{c}o, J.A. (2014)
+TypOn: the microbial typing ontology
+\textit{BMC J Biomed Sem}, \textbf{5}, 43.
+
+\end{thebibliography}
+
 \end{document}

### 2017-12-21 paper-proposal.tex

@@
 Our proposal covers an update to Superphy \cite{whiteside2016superphy}, an online predictive genomics platform targeting \textit{Escherichia coli}.
-The update, called Spfy, uses graph data structures to store and retrieve results for computational workflows.
-We demonstrate the ability of graph data structures to manage the tens of thousands of whole-genome \textit{E. coli} sequences accumulated so far, and use graphs for further downstream processing.
+The update, called Spfy, uses graph data structures to store and retrieve results for computational workflows, facilitating the management and querying of tens of thousands of whole-genome \textit{E.
+coli} sequences, and efficient downstream processing.
 % I'm unsure if we should add more about the subtyping options. For example, see:

### 2017-12-21 paper-proposal.tex
EDIT: add graph database as keyword
ADD: hardcoded a pdf

@@
-\documentclass[a4,center,fleqn]{NAR}
+\documentclass[a4paper, onecolumn]{NAR}
 \RequirePackage{hyperref}
@@
 \author{%
-Kevin K Le\,$^{1,*}$,
+Kevin K Le\,$^{1, *}$,
 Matthew D Whiteside\,$^{1}$,
-Chad R Laing\,$^1$,
-and Victor PJ Gannon\,$^1$%
+Chad R Laing\,$^{1}$,
+and Victor PJ Gannon\,$^{1}$
 \footnote{To whom correspondence should be addressed.
@@
 % "and two to four keywords."
-Keywords: Comparative genomics analysis, Epidemiology, Microbial genomics
+Keywords: Comparative genomics analysis, Epidemiology, Microbial genomics, Graph database
 

### 2017-12-21 paper-proposal_raw.tex
ADD: hardcoded a pdf

--- 2017-12-21 paper-proposal.tex
@@
-\documentclass[a4paper, onecolumn]{NAR}
+\documentclass{article}
+
 \RequirePackage{hyperref}
-\usepackage{lscape}
-\usepackage{graphicx}
-\usepackage{pdflscape}
+
 \usepackage[parfill]{parskip}
-% \usepackage[backend=bibtex,style=numeric-comp]{biblatex}
-% \usepackage[letterpaper, portrait, margin=1in]{geometry}
+\usepackage[letterpaper, portrait, margin=1in]{geometry}
 
@@
 
-\title{Spfy: bringing real-time, big data batch analyses of E. coli to SuperPhy}
+Spfy: bringing real-time, big data batch analyses of E. coli to SuperPhy
 
-\author{%
-Kevin K Le\,$^{1, *}$,
+Kevin K Le\,$^{1}$,
 Matthew D Whiteside\,$^{1}$,
@@
 and Victor PJ Gannon\,$^{1}$
-\footnote{To whom correspondence should be addressed.
-Tel: +1 403-382-5516; Fax: +1 403-381-1202; Email: chad.r.laing@phac-aspc.gc.ca}}
 
-\address{%
-$^{1}$Laboratory for Foodborne Zoonoses, Public Health Agency of Canada, Twp Rd 9-1, Lethbridge, AB, T1J 3Z4, Canada}
-% Affiliation must include:
-% Department name, institution name, full road and district address,
-% state, Zip or postal code, country
+$^{1}$Laboratory for Foodborne Zoonoses, Public Health Agency of Canada, Twp Rd 9-1, Lethbridge, AB, T1J 3Z4, Canada
 
-\maketitle
+Tel: +1 403-382-5516; Fax: +1 403-381-1202; Email: chad.r.laing@phac-aspc.gc.ca
 
-% "contain, at the top, the following affirmative statement.
-"This website is free and open to all users and there is no login requirement." Additionally, any third party software employed by the website that has more restrictive usage terms must be listed."
 This website is free and open to all users and there is no login requirement.
 The code for this webserver, and all third party software used, are available under the open-source Apache 2.0, BSD 3-clause, or similar licenses.
-\par
 
-% "include the website address; website name; and the names, affiliations, and email addresses of all authors."
 The website is available at \url{https://lfz.corefacility.ca/superphy/spfy/}.
-Spfy's code is provided at \url{https://github.com/superphy/backend} and documentation at \url{https://superphy.readthedocs.io/en/latest/}.
-\par
+Spfy’s code is provided at \url{https://github.com/superphy/backend} and documentation at \url{https://superphy.readthedocs.io/en/latest/}.
 
-% MAIN CONTENT
-% "include a notification if this is an update from a previous publication in the Web Server issue, and in that case, include an estimate of the number of users and the number of citations."
-% "For web servers, or essentially similar web servers, that have been the subject of a previous publication, including publication in journals other than NAR, there is a minimum two-year interval before re-publication in the Web Server Issue."
-Our proposal covers an update to Superphy \cite{whiteside2016superphy}, an online predictive genomics platform targeting \textit{Escherichia coli}.
-The update, called Spfy, uses graph data structures to store and retrieve results for computational workflows, facilitating the management and querying of tens of thousands of whole-genome \textit{E.
-coli} sequences, and efficient downstream processing.
-% I'm unsure if we should add more about the subtyping options. For example, see:
-% https://github.com/superphy/paper_platform/commit/c017b1e022d310e16a1433af9d58a73e9550a401
+Our proposal covers an update to Superphy (1), an online predictive genomics platform targeting Escherichia coli.
+The update, called Spfy, uses graph data structures to store and retrieve results for computational workflows, facilitating the management and querying of tens of thousands of whole-genome E.
+coli sequences, and efficient downstream processing.
 Current comparative computational workflows chain different analysis software, but lack storage and retrieval methods for generated results.
-% "IF THE WEBSITE IMPLEMENTS A META-SERVER OR COMPUTATIONAL WORKFLOW, the summary MUST describe 1) significant added value beyond the simple chaining together of existing third party software or the calculation of a consensus prediction from third party predictors and classifiers; and at least one of the following: 2) how user time for data gathering and multi-step analysis is significantly reduced, or 3) how the website offers significantly enhanced display of the data and results."
 By making the storage and retrieval of results part of the platform, with data effectively linked to the organisms of interest through a standardized ontology, we can mitigate the recomputing of analyses.
@@
 Integrated data storage will be necessary as publicly available whole genome sequencing data for bacterial pathogens currently numbers in the tens of thousands, with hundreds of thousands set to be available within the next few years.
-\par
 
-% STATISTICS{}
-% "provide descriptions of the input data, the output, and the processing method; complete citations for previous publications of the method or the web server; and two to four keywords.
-Additionally, authors must indicate how long the server has been running, the number of inputs analyzed during testing, and an estimate of the number of individuals outside of the authors' group who have been involved in the testing."
-Spfy was tested with 4,622 public \textit{E. coli} assembled genomes from Enterobase, storing every sequence and results for all included analysis modules.
+Spfy was tested with 4,622 public E.
+coli assembled genomes from Enterobase, storing every sequence and results for all included analysis modules.
 Spfy provides real-time subtyping, and the results are immediately displayed to the user following their completion.
-Subtyping options include O-antigen, H-antigen, Shiga-toxin 1, Shiga-toxin 2, and Intimin typing.
+Subtyping options include O-antigen, H- antigen, Shiga-toxin 1, Shiga-toxin 2, and Intimin typing.
 Reference-lab tests include virulence factor and anti-microbial resistance annotation.
-All genomes are analyzed within the pan-genome framework of \textit{E.
-coli}, and results from all analyses are automatically associated with the source genome.
-The resulting database had 1,333 nodes and 683,666 leaves, with 374,836,872 object properties. \par
+All genomes are analyzed within the pan-genome framework of E.
+coli, and results from all analyses are automatically associated with the source genome.
+The resulting database had 1,333 nodes and 683,666 leaves, with 374,836,872 object properties.
 
-% COMPARED TO EXISTING PLATFORMS
-% This aims to be more of an implementation paragraph.
-Existing scientific workflow technologies such as Galaxy \cite{goecks2010galaxy}, and pipelines such as the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial} and the Integrated Rapid Infectious Disease Analysis (IRIDA) platform \url{http://www.irida.ca/} help automate the use of WGS data for public-health surveillance.
-% data integration
+Existing scientific workflow technologies such as Galaxy (2), and pipelines such as the Bacterium Analysis Pipeline (BAP) (3) and the Integrated Rapid Infectious Disease Analysis (IRIDA) platform \url{http://www.irida.ca/} help automate the use of WGS data for public-health surveillance.
 Like IRIDA and BAP, Spfy automates workflows for users, and like Galaxy, Spfy uses task queues to distribute selected analysis.
-File uploads begin through the ReactJS-based website, where user-defined analyses options are selected.
+File uploads begin through the ReactJS- based website, where user-defined analyses options are selected.
 To these concepts we add Docker containerization for task queue workers, allowing analysis software to safely run in parallel.
-% re: Matt "2. When comparing to IRIDA, BAP etc., can mention some differences with Superphy, e.g. the storage of interim result data that allows downstream integrated analysis"
-For result storage, existing workflow technologies use relational tables \cite{goecks2010galaxy}, or store resulting files to disk \cite{thomsen2016bacterial}.
+For result storage, existing workflow technologies use relational tables (2), or store resulting files to disk (3).
 Because output from these programs is user-specific or transitory, results from identical comparisons are often recomputed.
@@
 Graph-based storage of all results solves these problems.
-% as before
-To avoid proliferating ontologies, and to allow Spfy to integrate with existing ones, annotations from the GenEpiO \cite{griffiths2017context}, FALDO \cite{bolleman2016faldo}, and TypOn \cite{vaz2014typon} ontologies are used to describe biological data.
-The entire platform is packaged using Docker-Compose, and can be recreated with a simple command. \par
+To avoid proliferating ontologies, and to allow Spfy to integrate with existing ones, annotations from the GenEpiO (4), FALDO (5), and TypOn (6) ontologies are used to describe biological data.
+The entire platform is packaged using Docker-Compose, and can be recreated with a simple command.
 
-% up-time
 The Spfy update has been up since May 2017 and Superphy has been up since early 2016.
-The server accepts assembled \textit{E.
-coli} genomes with the \textit{.fasta} or \textit{.fna} extensions.
-Submissions are checked against a reference set of \textit{E.
-coli} gene sequences before running analyses.
-Outputs are displayed on the website in tables and can be downloaded as \textit{.csv} files.
-\par
+The server accepts assembled E.
+coli genomes with the .fasta or .fna extensions.
+Submissions are checked against a reference set of E.
+coli gene sequences before running analyses.
+Outputs are displayed on the website in tables and can be downloaded as .csv files.
 
-% "and two to four keywords."
 Keywords: Comparative genomics analysis, Epidemiology, Microbial genomics, Graph database
@@
 
-\begin{thebibliography}{4}
+REFERENCES
 
-\bibitem{whiteside2016superphy}
+1.
 Whiteside, M.D., Laing, C.R., Manji, A., Kruczkiewicz, P., Taboada, E.N., and Gannon, V.P.J. (2016)
@@
 
-\bibitem{goecks2010galaxy}
+2.
 Goecks, J., Nekrutenko, A., Taylor, J. (2010)
@@
 
-\bibitem{thomsen2016bacterial}
+3.
 Thomsen, M.C.F., Ahrenfeldt, J., Cisneros, J.L.B., Jurtz, V., Larsen, M.V., Hasman, H., Aarestrup, F.M., and Lund, O. (2016)
@@
 
-\bibitem{griffiths2017context}
+4.
 Griffiths, E., Dooley, D., Graham, M., Van Domselaar, G., Brinkman, F.S.L., and Hsiao, W.W.L. (2017)
@@
 
-\bibitem{bolleman2016faldo}
+5.
 Bolleman, J.T., Mungall, C.J., Strozzi, F., Baran, J., Dumontier, M., Bonnal, R.J.P., Buels, R., Hoehndorf, R., Fujisawa, T., Katayama, T., et al. (2016)
@@
 
-\bibitem{vaz2014typon}
+6.
 Vaz, C., Francisco, A.P., Silva, M., Jolley, K.A., Bray, J.E., Puseele, H., Rothganger, J., Ramirez, M., Carri\c{c}o, J.A. (2014)
@@
 
-\end{thebibliography}
-
 \end{document}

### 2017-12-22 paper-proposal.tex
EDIT: affil
EDIT: authors, email

@@
 Matthew D Whiteside\,$^{1}$,
-Chad R Laing\,$^{1}$,
-and Victor PJ Gannon\,$^{1}$
+James Hopkins\,$^{1}$,
+Victor PJ Gannon\,$^{1}$,
+and Chad R Laing\,$^{1}$}
 \footnote{To whom correspondence should be addressed.
-Tel: +1 403-382-5516; Fax: +1 403-381-1202; Email: chad.r.laing@phac-aspc.gc.ca}}
+Tel: +1 403-382-5516; Fax: +1 403-381-1202; Email: chad.laing@canada.ca}
 
 \address{%
-$^{1}$Laboratory for Foodborne Zoonoses, Public Health Agency of Canada, Twp Rd 9-1, Lethbridge, AB, T1J 3Z4, Canada}
+$^{1}$National Microbiology Laboratory at Lethbridge, Public Health Agency of Canada, Twp Rd 9-1, Lethbridge, AB, T1J 3Z4, Canada}
 % Affiliation must include:

### 2017-12-22 paper-proposal_raw.tex
EDIT: affil
EDIT: authors, email

@@
 
-Kevin K Le\,$^{1}$,
+Kevin K Le\,$^{1, *}$,
 Matthew D Whiteside\,$^{1}$,
-Chad R Laing\,$^{1}$,
-and Victor PJ Gannon\,$^{1}$
+James Hopkins\,$^{1}$,
+Victor PJ Gannon\,$^{1}$,
+and Chad R Laing\,$^{1}$
 
-$^{1}$Laboratory for Foodborne Zoonoses, Public Health Agency of Canada, Twp Rd 9-1, Lethbridge, AB, T1J 3Z4, Canada
+$^{1}$National Microbiology Laboratory at Lethbridge, Public Health Agency of Canada, Twp Rd 9-1, Lethbridge, AB, T1J 3Z4, Canada
 
-Tel: +1 403-382-5516; Fax: +1 403-381-1202; Email: chad.r.laing@phac-aspc.gc.ca
+Tel: +1 403-382-5516; Fax: +1 403-381-1202; Email: chad.laing@canada.ca
 

### 2018-01-15 paper-webserver.tex
START: focus on spfy

@@
 
-\title{Superphy: Real-time prediction of \textit{E. coli} phenotype using graph-based genome sequence storage and retrieval with the \textit{spfy} module}
+\title{Spfy: a graph-based genome sequence database for real-time prediction of \textit{E. coli} phenotypes}
 
@@
 \begin{abstract}
+
+Spfy, uses graph data structures to store and retrieve results for computational workflows, facilitating the management and querying of tens of thousands of whole-genome \textit{E.
+coli} sequences, and efficient downstream processing.
+Current comparative computational workflows chain different analysis software, but lack storage and retrieval methods for generated results.
+By making the storage and retrieval of results part of the platform, with data effectively linked to the organisms of interest through a standardized ontology, we can mitigate the recomputing of analyses.
+Within Spfy, the output from all analyses is stored, and linked together in the context of a genome graph.
+This graph also stores metadata for each genome, facilitating inquiries ranging from population genomics to epidemiological investigations.
+Integrated data storage will be necessary as publicly available whole genome sequencing data for bacterial pathogens currently numbers in the tens of thousands, with hundreds of thousands set to be available within the next few years.
+\par
+
+The database is available at \url{https://lfz.corefacility.ca/superphy/spfy/}.
 
@@
 % 5. how we solved it
-In this study, we present an update to the SuperPhy platform, called Spfy.
+In this study, we present the Spfy platform, which was built with a graph database at its core.
 Spfy aims to solve the problem of real-time analysis and merging of results from different analysis methods.
@@
 \section{RESULTS}
-TODO: Update with all of Enterobase
-Spfy was tested with 59,5323 public \textit{E.
-coli} genomes, 5,353 genomes from GenBank and 54,181 genomes from Enterobase (\~596 GB), storing both the entire sequences and results for all included analysis modules.
-The resulting database had XYZ million nodes and XYZ million edges, with XYZ object properties, which worked out to XYZ TB of data stored.
+Spfy was tested with 4,622 public \textit{E. coli} assembled genomes from Enterobase, storing every sequence and results for all included analysis modules.
+Spfy provides real-time subtyping, and the results are immediately displayed to the user following their completion.
+Subtyping options include O-antigen, H-antigen, Shiga-toxin 1, Shiga-toxin 2, and Intimin typing.
+Reference-lab tests include virulence factor and anti-microbial resistance annotation.
+All genomes are analyzed within the pan-genome framework of \textit{E.
+coli}, and results from all analyses are automatically associated with the source genome.
+The resulting database had 1,333 nodes and 683,666 leaves, with 374,836,872 object properties. \par
+
+Spfy has been up since May 2017.
+The server accepts assembled \textit{E.
+coli} genomes with the \textit{.fasta} or \textit{.fna} extensions.
+Submissions are checked against a reference set of \textit{E.
+coli} gene sequences before running analyses.
+Outputs are displayed on the website in tables and can be downloaded as \textit{.csv} files.
+\par
 
@@
 % namely galaxy
-Scientific workflow technologies such as Galaxy \citep{goecks2010galaxy}, and bioinformatics pipelines IRIDA and BAP run analysis modules on WGS data.
+Existing scientific workflow technologies such as Galaxy \cite{goecks2010galaxy}, and pipelines such as the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial} and the Integrated Rapid Infectious Disease Analysis (IRIDA) platform \url{http://www.irida.ca/} help automate the use of WGS data for public-health surveillance.
 Galaxy aims to provide a reproducible, computation-based research environment which is accessible to individuals without programming knowledge.
@@
 
-In Spfy, we take a more opinionated approach to result storage and data integration.
-The benefits of real-time analysis pipelines as found in BAP, and the use of the GenEpiO ontology as in IRIDA, are included in Spfy.
+Like IRIDA and BAP, Spfy automates workflows for users, and like Galaxy, Spfy uses task queues to distribute selected analysis.
+File uploads begin through the ReactJS-based website, where user-defined analyses options are selected.
+To these concepts we add Docker containerization for task queue workers, allowing analysis software to safely run in parallel.
+% re: Matt "2. When comparing to IRIDA, BAP etc., can mention some differences with Superphy, e.g. the storage of interim result data that allows downstream integrated analysis"
+For result storage, existing workflow technologies use relational tables \cite{goecks2010galaxy}, or store resulting files to disk \cite{thomsen2016bacterial}.
+Because output from these programs is user-specific or transitory, results from identical comparisons are often recomputed.
+Additionally, output from different analyses are structured using distinct terminology and formats, which must be converted before they can be compared.
+Without a unified structure, these conversions quickly become impractical for broad usage.
+Graph-based storage of all results solves these problems.
+% as before
+To avoid proliferating ontologies, and to allow Spfy to integrate with existing ones, annotations from the GenEpiO \cite{griffiths2017context}, FALDO \cite{bolleman2016faldo}, and TypOn \cite{vaz2014typon} ontologies are used to describe biological data.
+The entire platform is packaged using Docker-Compose, and can be recreated with a simple command. \par
+
 On a per file basis, Spfy performs at a similar speed to BAP on predictive genomics tasks, though Spfy does not provide genome assembly services.

### 2018-01-16 paper-webserver.tex
EDIT: some things from proposal

@@
 
-Spfy, uses graph data structures to store and retrieve results for computational workflows, facilitating the management and querying of tens of thousands of whole-genome \textit{E.
+Current comparative computational workflows chain different analysis software, but lack storage and retrieval methods for generated results.
+Spfy uses graph data structures to store and retrieve results for computational workflows, facilitating the management and querying of tens of thousands of whole-genome \textit{E.
 coli} sequences, and efficient downstream processing.
-Current comparative computational workflows chain different analysis software, but lack storage and retrieval methods for generated results.
 By making the storage and retrieval of results part of the platform, with data effectively linked to the organisms of interest through a standardized ontology, we can mitigate the recomputing of analyses.
@@
 These graph objects, which consist of nodes and edges, must be annotated with common terminology.
-Spfy uses annotations from the GenEpiO \citep{griffiths2017context}, FALDO \citep{bolleman2016faldo}, and TypOn \citep{vaz2014typon} ontologies, as well as custom terminology to enable graph traversals and related data points to be inferred from the ontology (see Figure \ref{fig-ontology}).
+To avoid proliferating ontologies, and to allow Spfy to integrate with existing ones, annotations from the GenEpiO \cite{griffiths2017context}, FALDO \cite{bolleman2016faldo}, and TypOn \cite{vaz2014typon} ontologies are used to describe biological data.
 The use of semantic web technologies allows results to be linked to the genomes they were computed from, and for all data points to share a common data structure that can be queried together.
@@
 % as before
-To avoid proliferating ontologies, and to allow Spfy to integrate with existing ones, annotations from the GenEpiO \cite{griffiths2017context}, FALDO \cite{bolleman2016faldo}, and TypOn \cite{vaz2014typon} ontologies are used to describe biological data.
 The entire platform is packaged using Docker-Compose, and can be recreated with a simple command. \par

### 2018-01-16 paper-webserver.tex
STOP: going to use a vanilla tex format
FIX: latex builds now

@@
-\documentclass[a4,center,fleqn]{NAR}
+\documentclass{article}
 
-% Enter dates of publication
-\copyrightyear{2008}
-\pubdate{31 July 2009}
-\pubyear{2009}
-\jvolume{37}
-\jissue{12}
+\RequirePackage{hyperref}
+
+\usepackage[parfill]{parskip}
+\usepackage[letterpaper, portrait, margin=1in]{geometry}
+% Author
+\usepackage{authblk}
+% Images
+\usepackage{graphicx}
+% \usepackage{svg}
 
@@
 
-\title{Spfy: a graph-based genome sequence database for real-time prediction of \textit{E. coli} phenotypes}
+\title{Spfy: applying a graph-based sequence database for real-time prediction of \textit{E. coli} phenotypes}
 
-\author{%
-Corresponding Author\,$^{1,*}$,
-First Co-Author\,$^{2}$
-and Second Co-Author\,$^2$%
-\footnote{To whom correspondence should be addressed.
-Tel: +44 000 0000000; Fax: +44 000 0000000; Email: xxx@yyyy.ac.zz}}
+\author[1]{Kevin K Le\thanks{kevin.le@canada.ca}}
+\author[1]{Matthew D Whiteside\thanks{matthew.whiteside@canada.ca}}
+\author[1]{James Hopkins}
+\author[1]{Victor PJ Gannon}
+\author[1]{Chad R Laing}
+\affil[1]{National Microbiology Laboratory at Lethbridge, Public Health Agency of Canada, Twp Rd 9-1, Lethbridge, AB, T1J 3Z4, Canada}
 
-\address{%
-$^{1}$Affiliation of Corresponding Author
-and
-$^{2}$Affiliation of Both Co-Authors}
-% Affiliation must include:
-% Department name, institution name, full road and district address,
-% state, Zip or postal code, country
-
-\history{%
-Received January 1, 2009;
-Revised February 1, 2009;
-Accepted March 1, 2009}
+\renewcommand\Authands{ and }
 
@@
 Whole genome sequencing (WGS) can in theory provide the entire genetic content of an organism.
-This unparalleled resolution and sensitivity has recently transformed public-health surveillance and outbreak response \citep{ronholm2016navigating,lytsy2017time}.
-Additionally, the identification of novel disease mechanisms \citep{wang2014whole,yuen2015whole}, and rapid clinical diagnoses and reference lab tests based on the specific mechanism of disease are now possible.
-\citep{willig2015whole,dewey2014clinical}.
+This unparalleled resolution and sensitivity has recently transformed public-health surveillance and outbreak response \cite{ronholm2016navigating,lytsy2017time}.
+Additionally, the identification of novel disease mechanisms \cite{wang2014whole,yuen2015whole}, and rapid clinical diagnoses and reference lab tests based on the specific mechanism of disease are now possible.
+\cite{willig2015whole,dewey2014clinical}.
 
@@
 The rapid characterization and comparison of bacterial pathogens relies principally on the combination of outputs from multiple software programs that are targeted for specific applications.
-Examples include the Resistance Gene Identifier (RGI) \citep{mcarthur2013comprehensive} for antimicrobial resistance (AMR) gene prediction, VirulenceFinder \citep{kleinheinz2014applying} for virulence factor (VF) annotation, and Prokka for bacterial genome annotation with external tools \citep{doi:10.1093/bioinformatics/btu153}.
+Examples include the Resistance Gene Identifier (RGI) \cite{mcarthur2013comprehensive} for antimicrobial resistance (AMR) gene prediction, VirulenceFinder \cite{kleinheinz2014applying} for virulence factor (VF) annotation, and Prokka for bacterial genome annotation with external tools \cite{doi:10.1093/bioinformatics/btu153}.
 In particular, RGI and VirulenceFinder represent a series of \textit{in-silico} methods which have been developed to replicate the results of traditional wet-lab tests; this allows new WGS results to be viewed in the context of historical tests.
@@
 Comprehensive platforms that combine individual programs into a cohesive whole also exist.
-These include free platforms such as the Bacterium Analysis Pipeline (BAP) \citep{thomsen2016bacterial}, the Integrated Rapid Infectious Disease Analysis (IRIDA) project \url{http://www.irida.ca/}, and PATRIC \citep{wattam2013patric}.
-Commercial applications, such as Bionumerics, which is used by PulseNet international for the analyses of WGS data in outbreak situations also exist, and offer support as well as accredited, standardized tests \citep{swaminathan2001pulsenet}.
+These include free platforms such as the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial}, the Integrated Rapid Infectious Disease Analysis (IRIDA) project \url{http://www.irida.ca/}, and PATRIC \cite{wattam2013patric}.
+Commercial applications, such as Bionumerics, which is used by PulseNet international for the analyses of WGS data in outbreak situations also exist, and offer support as well as accredited, standardized tests \cite{swaminathan2001pulsenet}.
 
@@
 WGS data for bacterial pathogens of public health importance have recently accumulated in public databases in the tens of thousands, with hundreds of thousands set to be available within the next few years.
-For \textit{Escherichia coli}, there are over sixty thousand publicly available genomes in EnteroBase \url{https://enterobase.warwick.ac.uk/} and three million whole genomes in GenBank \citep{doi:10.1093/nar/gks1195}.
+For \textit{Escherichia coli}, there are over sixty thousand publicly available genomes in EnteroBase \url{https://enterobase.warwick.ac.uk/} and three million whole genomes in GenBank \cite{doi:10.1093/nar/gks1195}.
 
@@
 % 4. Our specific problem, previous work
-We have previously developed Superphy \citep{whiteside2016superphy}, an online predictive genomics platform targeting \textit{E.
+We have previously developed Superphy \cite{whiteside2016superphy}, an online predictive genomics platform targeting \textit{E.
 coli}.
@@
 Future work will focus on adding additional analyses modules, using machine learning and artificial neural networks to aid genotype to phenotype predictions, and supporting different species.
-While the integrated approach of storing and retrieving results provides enormous benefits, the developed analyses modules are self-contained and can easily be integrated into existing platforms such as Galaxy \citep{goecks2010galaxy}, and IRIDA.
+While the integrated approach of storing and retrieving results provides enormous benefits, the developed analyses modules are self-contained and can easily be integrated into existing platforms such as Galaxy \cite{goecks2010galaxy}, and IRIDA.
 The website is available at \url{https://lfz.corefacility.ca/superphy/spfy/}.
@@
 Prebuilt tools are available for a variety of reference laboratory tasks.
-Spfy performs: O-antigen typing, H-antigen typing, and VF gene determination using ECtyper \url{https://github.com/phac-nml/ecoli\_serotyping}, STX typing using Phylotyper \citep{whiteside2017phylotyper}, and AMR gene determination using the RGI program \citep{mcarthur2013comprehensive}.
-Spfy also performs bioinformatics analyses: pangenome generation using Panseq \citep{laing2010pan}, statistical significance testing using SciPy \citep{jones2014scipy}, and support vector machine (SVM)-backed AMR predictions using Scikit-learn \citep{pedregosa2011scikit}.
+Spfy performs: O-antigen typing, H-antigen typing, and VF gene determination using ECtyper \url{https://github.com/phac-nml/ecoli\_serotyping}, STX typing using Phylotyper \cite{whiteside2017phylotyper}, and AMR gene determination using the RGI program \cite{mcarthur2013comprehensive}.
+Spfy also performs bioinformatics analyses: pangenome generation using Panseq \cite{laing2010pan}, statistical significance testing using SciPy \cite{jones2014scipy}, and support vector machine (SVM)-backed AMR predictions using Scikit-learn \cite{pedregosa2011scikit}.
 For the larger population comparisons, we store and aggregate the results from individual genome analyses.
@@
 % para: the semantic web
-Semantic web technologies describe the relations between different data \citep{berners2001semantic}.
+Semantic web technologies describe the relations between different data \cite{berners2001semantic}.
 For biological data, individual data points such as genome, contiguous DNA sequence, or gene, can be linked together in a query-able graph structure.
-This allows novel data to be seamlessly incorporated into the existing graph, and has the use of graph databases for relational information has been proposed as a common standard for the open sharing of data \citep{horrocks2005semantic}.
+This allows novel data to be seamlessly incorporated into the existing graph, and has the use of graph databases for relational information has been proposed as a common standard for the open sharing of data \cite{horrocks2005semantic}.
 
@@
 \begin{center}
-\includegraphics{images/spfy_ontology-1.svg}
+\includegraphics{images/ontology}
 \end{center}
@@
 % it's not quite a full OS
-Docker \url{https://www.docker.com/} is a virtualization technology to simulate self-contained operating systems on the same host computer, without the overhead of full hardware virtualization \citep{felter2015updated}.
+Docker \url{https://www.docker.com/} is a virtualization technology to simulate self-contained operating systems on the same host computer, without the overhead of full hardware virtualization \cite{felter2015updated}.
 The Spfy platform depends on a series of webservers, databases, and task workers and uses Docker to compartmentalize these services which are then networked together using Docker-Compose \url{https://docs.docker.com/compose/}.
 (see Figure \ref{fig-docker})
-Docker integration ensures that software dependencies, which are typically manually installed \citep{doi:10.1093/bioinformatics/btu153,laing2010pan,inouye2014srst2,naccache2014cloud}, are instead handled automatically.
+Docker integration ensures that software dependencies, which are typically manually installed \cite{doi:10.1093/bioinformatics/btu153,laing2010pan,inouye2014srst2,naccache2014cloud}, are instead handled automatically.
 With compartmentalization of service runtimes, code failures do not propagate to other services.
@@
 \begin{center}
-\includegraphics{images/docker.svg}
+\includegraphics{images/docker}
 \end{center}
@@
 \section{DISCUSSION}
-% Gist, we address the the "Limitations" of \citep{de2015trends}.
+% Gist, we address the the "Limitations" of \cite{de2015trends}.
 % Namely:
@@
 % drawbacks - big data
-Many previous bioinformatics software programs have been developed \textit{ad hoc}, with individual researchers and laboratories developing software specific to their environment \citep{de2015trends}.
-Such tools were often script-based, with custom data formats, and only suitable for small collections of data \citep{de2015trends}.
-While this was acceptable for smaller analyses, bioinformatic pipelines utilizing WGS data are larger and involve linked dependencies, which require the application of systems engineering principles \citep{schatz2015biological}.
-Additionally, many subsets of biology now require the analyses of big-data, where the ability to perform computations in real-time, store data in flexible databases, and utilize a common application programming interface (API) linking resources are required \citep{swaminathan2016review}.
+Many previous bioinformatics software programs have been developed \textit{ad hoc}, with individual researchers and laboratories developing software specific to their environment \cite{de2015trends}.
+Such tools were often script-based, with custom data formats, and only suitable for small collections of data \cite{de2015trends}.
+While this was acceptable for smaller analyses, bioinformatic pipelines utilizing WGS data are larger and involve linked dependencies, which require the application of systems engineering principles \cite{schatz2015biological}.
+Additionally, many subsets of biology now require the analyses of big-data, where the ability to perform computations in real-time, store data in flexible databases, and utilize a common application programming interface (API) linking resources are required \cite{swaminathan2016review}.
 
@@
 The isolation and characterization of bacterial pathogens are critical for Public Health laboratories to rapidly respond to outbreaks, and to effectively monitor known and emerging pathogens through surveillance programs.
-Until recently, public-health agencies relied on laboratory tests such as serotyping, pulsed-field gel electrophoresis (PFGE) banding patterns, or Polymerase Chain Reaction (PCR)-based amplification to identify known VFs or AMR genes, along with other tests \citep{ronholm2016navigating}, to characterize bacterial isolates in outbreak and surveillance settings.
+Until recently, public-health agencies relied on laboratory tests such as serotyping, pulsed-field gel electrophoresis (PFGE) banding patterns, or Polymerase Chain Reaction (PCR)-based amplification to identify known VFs or AMR genes, along with other tests \cite{ronholm2016navigating}, to characterize bacterial isolates in outbreak and surveillance settings.
 
@@
 Current efforts are focused on predictive genomics, where the relevant phenotypic information can be determined through examination of the whole-genome sequence.
-WGS methods allow a single laboratory method - the sequencing step - to provide a variety of analysis options, such as AMR, VF, and Shiga-toxin results, and as such can be used to evaluate the spread of outbreaks with better resolution and context than traditional methods \citep{ronholm2016navigating}.
+WGS methods allow a single laboratory method - the sequencing step - to provide a variety of analysis options, such as AMR, VF, and Shiga-toxin results, and as such can be used to evaluate the spread of outbreaks with better resolution and context than traditional methods \cite{ronholm2016navigating}.
 
@@
 BAP takes a combined approach to genome analysis including different programs, such as for VF and AMR determination, by default into the pipeline.
-A record of the different files generated by an analysis are stored in a relational database, and the record is queried to determine which files are returned to the user \citep{thomsen2016bacterial}.
+A record of the different files generated by an analysis are stored in a relational database, and the record is queried to determine which files are returned to the user \cite{thomsen2016bacterial}.
 
@@
 To approach the challenge of integrating different bioinformatics programs, Spfy instead uses technologies prevalent in common web services not necessarily related to scientific workflows.
-While we prefer to package individual bioinformatics programs using language-agnostic package managers such as Conda \url{https://conda.io/}, we do agree that suites of bioinformatics software benefit from having all dependencies installed and deployable in Docker containers \citep{di2015impact}.
+While we prefer to package individual bioinformatics programs using language-agnostic package managers such as Conda \url{https://conda.io/}, we do agree that suites of bioinformatics software benefit from having all dependencies installed and deployable in Docker containers \cite{di2015impact}.
 These programs are connected to form a workflow using task queues; a design methodology used to offload long-running or non-instantaneous tasks over available computing infrastructure.
@@
 Spfy packages compute nodes, which manage a collection of task queue workers, as reproducible Docker containers which can be networked together to form a compute cluster.
-Docker containerization has a negligible impact on performance \citep{di2015impact}, and allows the platform to easily scale to demand.
+Docker containerization has a negligible impact on performance \cite{di2015impact}, and allows the platform to easily scale to demand.
 In contrast to Galaxy, Spfy is a targeted approach building a predictive analytics platform for \textit{E.
-coli} which leverages generic web technologies to favor a responsive, big-data design strategy tackling the challenge \citep{fricke2014bacterial} of integrating results from multiple genomes and examining shared connections.
+coli} which leverages generic web technologies to favor a responsive, big-data design strategy tackling the challenge \cite{fricke2014bacterial} of integrating results from multiple genomes and examining shared connections.
 By packaging the individual bioinformatics software as Conda packages, they can be easily ported to Galaxy.

### 2018-01-17 paper-webserver.tex
FIX: bibliography
FIX: image builds & no warnings
EDIT: emails

@@
 
-\bibliography{paper-webserver.bib}
-
 %\articlesubtype{This is the article type (optional)}
+% \bibliography{paper-webserver.bib}
 
@@
 
-\title{Spfy: applying a graph-based sequence database for real-time prediction of \textit{E. coli} phenotypes}
+\title{Spfy: applying a graph-based sequence database for real-time prediction of Escherichia coli phenotypes}
 
 \author[1]{Kevin K Le\thanks{kevin.le@canada.ca}}
-\author[1]{Matthew D Whiteside\thanks{matthew.whiteside@canada.ca}}
+\author[1]{Matthew D Whiteside}
 \author[1]{James Hopkins}
 \author[1]{Victor PJ Gannon}
-\author[1]{Chad R Laing}
+\author[1]{Chad R Laing\thanks{chad.laing@canada.ca}}
 \affil[1]{National Microbiology Laboratory at Lethbridge, Public Health Agency of Canada, Twp Rd 9-1, Lethbridge, AB, T1J 3Z4, Canada}
@@
 \begin{center}
-\includegraphics{images/ontology}
+\includegraphics[width=\textwidth]{images/ontology}
 \end{center}
@@
 \begin{center}
-\includegraphics{images/results.png}
+\includegraphics[width=\textwidth]{images/results.png}
 \end{center}
@@
 \begin{center}
-\includegraphics{images/tables.png}
+\includegraphics[width=\textwidth]{images/tables.png}
 \end{center}
@@
 \begin{center}
-\includegraphics{images/docker}
+\includegraphics[width=\textwidth]{images/docker}
 \end{center}
@@
 
+\textit{Conflict of interest}. None declared.
+
 \section{ACKNOWLEDGEMENTS}
 
-\subsubsection{Conflict of interest statement.} None declared.
 \newpage
 
-\begin{thebibliography}{4}
-
-% Format for article
-\bibitem{1}
-Author,A.B. and Author,C. (1992)
-Article title.
-\textit{Abbreviated Journal Name}, \textbf{5}, 300--330.
-
-% Format for book
-\bibitem{2}
-Author,D., Author,E.F. and Author,G. (1995)
-\textit{Book Title}.
-Publisher Name, Publisher Address.
-
-% Format for chapter in book
-\bibitem{3}
-Author,H. and Author,I. (2005)
-Chapter title.
-In
-Editor,A. and Editor,B. (eds),
-\textit{Book Title},
-Publisher Name, Publisher Address,
-pp.\ 60--80.
-
-% Another article
-\bibitem{4}
-Author,Y. and Author,Z. (2002)
-Article title.
-\textit{Abbreviated Journal Name}, \textbf{53}, 500--520.
-
-\end{thebibliography}
+\bibliographystyle{unsrt}
+\bibliography{paper-webserver}
 

### 2018-01-19 paper-webserver.tex
ADD: example table
ADD: trimmed table

@@
 \usepackage{graphicx}
+\usepackage{lscape}
+\usepackage{longtable}
 % \usepackage{svg}
@@
 \section{RESULTS}
-Spfy was tested with 4,622 public \textit{E. coli} assembled genomes from Enterobase, storing every sequence and results for all included analysis modules.
+Spfy was tested with 10,243 public \textit{E. coli} assembled genomes from Enterobase, storing every sequence and results for all included analysis modules.
 Spfy provides real-time subtyping, and the results are immediately displayed to the user following their completion.
@@
 coli}, and results from all analyses are automatically associated with the source genome.
-The resulting database had 1,333 nodes and 683,666 leaves, with 374,836,872 object properties. \par
+The resulting database had 17,820 nodes and 3,811,473 leaves, with 1,125,909,074 object properties. \par
+
+\small \input{tables/database}
 
@@
 
+\newpage
+
+\section{APPENDIX}
+
 \end{document}

### 2018-01-30 paper-webserver.tex
ADD: new figures
CHANGE: shorten the intro (& move some of its material to other sections

@@
 
-\title{Spfy: applying a graph-based sequence database for real-time prediction of Escherichia coli phenotypes}
+\title{Spfy: an integrated graph-based sequence database for real-time prediction of Escherichia coli phenotypes}
 
@@
 % 3. 100,000 genomes, (lookup number for Enterobase, GenBank), how do we run analysis on it
-WGS data for bacterial pathogens of public health importance have recently accumulated in public databases in the tens of thousands, with hundreds of thousands set to be available within the next few years.
+Traditionally, these platforms have been applied to samples numbering only a few dozens while WGS data for bacterial pathogens of public health importance have recently accumulated in public databases in the tens of thousands, with hundreds of thousands set to be available within the next few years.
 For \textit{Escherichia coli}, there are over sixty thousand publicly available genomes in EnteroBase \url{https://enterobase.warwick.ac.uk/} and three million whole genomes in GenBank \cite{doi:10.1093/nar/gks1195}.
 
-To begin to solve this "big-data" problem, platforms such as BAP and IRIDA provide hosted solutions that compute results in real-time, and distribute analyses across computing resources.
-While effective for self-contained workflows, many of the comparative analyses that are run are broadly useful, and therefore computed multiple times.
-An effective method to mitigate the recomputing of analyses is to make the storage and retrieval of results part of the platform, and effectively linked to the organisms of interest with a standardized ontology.
+Many of the comparative analyses that are run are broadly useful, and therefore computed multiple times.
+To begin to compute results on all available genomes, an effective method to mitigate the recomputing of analyses is to make the storage and retrieval of results part of the platform, and effectively linked to the organisms of interest with a standardized ontology.
 Such measures can help ensure the rapid response times required for public health applications, and allow results to be integrated and progressively updated as new data becomes available.
@@
 % 5. how we solved it
-In this study, we present the Spfy platform, which was built with a graph database at its core.
-Spfy aims to solve the problem of real-time analysis and merging of results from different analysis methods.
-By merging results and resolving connections between genome data, Spfy is able to identify relationships between all genomes sequenced in the past, present, and future.
-This graph-based result storage allows retrospective comparisons as more genomes are sequenced or populations change, and is flexible to accommodate new analysis methods as they are developed.
+In this study, we present the Spfy platform, which integrates a graph database with real-time analysis options.
+By integrating analysis options with data storage, we avoid recomputing analysis for identical runs.
+Graph-based result storage also allows retrospective comparisons as more genomes are sequenced or populations change, and is flexible to accommodate new analysis modules as they are developed.
+The website is available at \url{https://lfz.corefacility.ca/superphy/spfy/}.
 
 % {needs to be beefed up, in language a biologist / public health worker would care about}
-
-By supporting multiple \textit{in-silico} subtyping options, the platform functions similar to a reference laboratory, while adding support for big-data analyses.
-Subtyping options for E.coli are O-antigen, H-antigen, Shiga-toxin 1 (Stx1), Shiga-toxin 2 (Stx2), and Intimin typing, VF and AMR annotation, pan-genome generation, group comparisons via Fisher's, and machine-learning (ML) based predictions for OmniLog AMR assays.
-These tasks are divided into subtasks, and distributed across a built-in task queue.
-Results are also decomposed into graph structures and stored within a larger graph database.
-By integrating task distribution with graph storage, Spfy enables large-scale analyses, such as epidemiological associations between specific genotypes, biomarkers, host, source, and other metadata, and statistical significance testing of genome markers for user-defined groups, in real-time.
-Currently, the platform has been tested with XXX genome files and result storage for XX analyses modules.
-
-Future work will focus on adding additional analyses modules, using machine learning and artificial neural networks to aid genotype to phenotype predictions, and supporting different species.
-While the integrated approach of storing and retrieving results provides enormous benefits, the developed analyses modules are self-contained and can easily be integrated into existing platforms such as Galaxy \cite{goecks2010galaxy}, and IRIDA.
-The website is available at \url{https://lfz.corefacility.ca/superphy/spfy/}.
-Spfy's codebase is provided at \url{https://github.com/superphy/backend} and a developer guide is provided at \url{https://superphy.readthedocs.io/en/latest/}.
 
@@
 % para covering everything
+
+By supporting multiple \textit{in-silico} subtyping options, the platform functions similar to a reference laboratory, while adding support for population analyses.
+Subtyping options for E.coli are O-antigen, H-antigen, Shiga-toxin 1 (Stx1), Shiga-toxin 2 (Stx2), and Intimin typing, VF and AMR annotation, pan-genome generation, group comparisons via Fisher's, and machine-learning (ML) based predictions for OmniLog AMR assays.
+These tasks are divided into subtasks, and distributed across a built-in task queue.
+Results are also decomposed into graph structures and stored within a larger graph database.
+By integrating task distribution with graph storage, Spfy enables large-scale analyses, such as epidemiological associations between specific genotypes, biomarkers, host, source, and other metadata, and statistical significance testing of genome markers for user-defined groups, in real-time.
+Currently, the platform has been tested with XXX genome files and result storage for XX analyses modules.
 
@@
 
+Future work will focus on adding additional analyses modules, using machine learning and artificial neural networks to aid genotype to phenotype predictions, and supporting different species.
+While the integrated approach of storing and retrieving results provides enormous benefits, the developed analyses modules are self-contained and can easily be integrated into existing platforms such as Galaxy \cite{goecks2010galaxy}, and IRIDA.
+Spfy's codebase is provided at \url{https://github.com/superphy/backend} and a developer guide is provided at \url{https://superphy.readthedocs.io/en/latest/}.
+
 \textit{Conflict of interest}. None declared.

### 2018-01-30 paper-webserver.tex
EDIT: merged 2 functionality paragraphs

@@
 By supporting multiple \textit{in-silico} subtyping options, the platform functions similar to a reference laboratory, while adding support for population analyses.
-Subtyping options for E.coli are O-antigen, H-antigen, Shiga-toxin 1 (Stx1), Shiga-toxin 2 (Stx2), and Intimin typing, VF and AMR annotation, pan-genome generation, group comparisons via Fisher's, and machine-learning (ML) based predictions for OmniLog AMR assays.
+Subtyping options for E.coli are O-antigen, H-antigen, and VF gene determination using ECtyper \url{https://github.com/phac-nml/ecoli\_serotyping}, Shiga-toxin 1 (Stx1), Shiga-toxin 2 (Stx2), and Intimin typing using Phylotyper \cite{whiteside2017phylotyper}, and AMR annotation using the RGI program \cite{mcarthur2013comprehensive}.
+Spfy also performs bioinformatics analyses: pangenome generation using Panseq \cite{laing2010pan}, statistical significance testing using SciPy \cite{jones2014scipy}, and support vector machine (SVM)-backed AMR predictions for OmniLog AMR assays using Scikit-learn \cite{pedregosa2011scikit}.
+
 These tasks are divided into subtasks, and distributed across a built-in task queue.
-Results are also decomposed into graph structures and stored within a larger graph database.
+Results are converted to individual graphs and stored within a larger graph database.
 By integrating task distribution with graph storage, Spfy enables large-scale analyses, such as epidemiological associations between specific genotypes, biomarkers, host, source, and other metadata, and statistical significance testing of genome markers for user-defined groups, in real-time.
 Currently, the platform has been tested with XXX genome files and result storage for XX analyses modules.
-
-% Need a sentance introducing the tools otherwise this sounds just like the intro section
-Prebuilt tools are available for a variety of reference laboratory tasks.
-Spfy performs: O-antigen typing, H-antigen typing, and VF gene determination using ECtyper \url{https://github.com/phac-nml/ecoli\_serotyping}, STX typing using Phylotyper \cite{whiteside2017phylotyper}, and AMR gene determination using the RGI program \cite{mcarthur2013comprehensive}.
-Spfy also performs bioinformatics analyses: pangenome generation using Panseq \cite{laing2010pan}, statistical significance testing using SciPy \cite{jones2014scipy}, and support vector machine (SVM)-backed AMR predictions using Scikit-learn \cite{pedregosa2011scikit}.
-For the larger population comparisons, we store and aggregate the results from individual genome analyses.
 

### 2018-01-31 paper-webserver.tex
DEL: 3rd para in functionality section is somewhat redundant. Merged into 2nd

@@
 By supporting multiple \textit{in-silico} subtyping options, the platform functions similar to a reference laboratory, while adding support for population analyses.
-Subtyping options for E.coli are O-antigen, H-antigen, and VF gene determination using ECtyper \url{https://github.com/phac-nml/ecoli\_serotyping}, Shiga-toxin 1 (Stx1), Shiga-toxin 2 (Stx2), and Intimin typing using Phylotyper \cite{whiteside2017phylotyper}, and AMR annotation using the RGI program \cite{mcarthur2013comprehensive}.
-Spfy also performs bioinformatics analyses: pangenome generation using Panseq \cite{laing2010pan}, statistical significance testing using SciPy \cite{jones2014scipy}, and support vector machine (SVM)-backed AMR predictions for OmniLog AMR assays using Scikit-learn \cite{pedregosa2011scikit}.
+Subtyping options for \textit{E.coli} are O-antigen, H-antigen, and VF gene determination using ECtyper \url{https://github.com/phac-nml/ecoli\_serotyping}, Shiga-toxin 1 (Stx1), Shiga-toxin 2 (Stx2), and Intimin typing using Phylotyper \cite{whiteside2017phylotyper}, and AMR annotation using the RGI program \cite{mcarthur2013comprehensive}.
+Spfy also performs bioinformatics analyses: pangenome generation using Panseq \cite{laing2010pan} and support vector machine (SVM)-backed AMR predictions for OmniLog AMR assays using Scikit-learn \cite{pedregosa2011scikit}.
 
@@
 Results are converted to individual graphs and stored within a larger graph database.
-By integrating task distribution with graph storage, Spfy enables large-scale analyses, such as epidemiological associations between specific genotypes, biomarkers, host, source, and other metadata, and statistical significance testing of genome markers for user-defined groups, in real-time.
+By integrating task distribution with graph storage, Spfy enables large-scale analyses, such as epidemiological associations between specific genotypes, biomarkers, host, source, and other metadata, and statistical significance testing of genome markers for user-defined groups using SciPy \cite{jones2014scipy}.
+Any data type or relation in the graph is valid for analysis, such as the presence or absence of pan-genome regions against determined serotypes or other subtyping options.
+In addition, Spfy supports the submission of user-specified metadata, such as location or source information.
 Currently, the platform has been tested with XXX genome files and result storage for XX analyses modules.
@@
 % phylotyper also has its own paper
-
-% group comparisons
-Knowledge graphs semantically link data points using relations defined in an ontology.
-Spfy allows users to compare the results associated with particular genome types or metadata by storing all results in a graph structure.
-Comparisons can be made in real-time, as new genomes are submitted for analysis and the results stored in the existing graph.
-Any data type or relation in the graph is valid for analysis, such as the presence or absence of pan-genome regions against determined serotypes or other subtyping options.
-In addition, Spfy supports the submission of user-specified metadata, such as location or source information.
-Results are displayed on-screen and available for download.
 

### 2018-01-31 paper-webserver.tex
DEL/MERGE: redundancies in the Data Storage section

@@
 Semantic web technologies describe the relations between different data \cite{berners2001semantic}.
-For biological data, individual data points such as genome, contiguous DNA sequence, or gene, can be linked together in a query-able graph structure.
-This allows novel data to be seamlessly incorporated into the existing graph, and has the use of graph databases for relational information has been proposed as a common standard for the open sharing of data \cite{horrocks2005semantic}.
-
-% para: semantic web in spfy
-In Spfy, the results from all the analysis modules are converted into graph objects, which are used to update the main graph database.
-These graph objects, which consist of nodes and edges, must be annotated with common terminology.
-To avoid proliferating ontologies, and to allow Spfy to integrate with existing ones, annotations from the GenEpiO \cite{griffiths2017context}, FALDO \cite{bolleman2016faldo}, and TypOn \cite{vaz2014typon} ontologies are used to describe biological data.
-The use of semantic web technologies allows results to be linked to the genomes they were computed from, and for all data points to share a common data structure that can be queried together.
+For biological data, individual data points such as genome, contiguous DNA sequence, or gene, can be linked together in a searchable graph structure.
+This allows novel data to be incorporated into the existing graph, and has the use of graph databases for semantic information has been proposed as a open standard for sharing public information \cite{horrocks2005semantic}.
+Data points are annotated with terms from published ontologies.
+To avoid proliferating ontologies, and to allow Spfy to integrate with existing ones, annotations from the GenEpiO \cite{griffiths2017context}, FALDO \cite{bolleman2016faldo}, and TypOn \cite{vaz2014typon} ontologies are used for our annotations.
 
@@
 
-% note: below is a feature in development
-The permanent storage of results is as a one-time cost, and allows population-wide analyses of all stored genomes; result storage also enables Spfy to avoid recomputation when the same analysis is re-run.
-To identify analysis tasks, Spfy fingerprints requests by computing its SHA1 hash and tags the results in the graph database with this hash.
-Subsequent queries are then compared through fingerprinting, and results can be rebuilt from the graph by searching for matching fingerprints.
-
-For population analyses, Spfy searches the graph for all nodes annotated with the queried ontology term.
-Depending on the given analysis, this data is then vectorized into NumPy numerical arrays for use with the SciPy scientific computing library, or the Scikit-learn machine learning library.
+% Permanent storage + why a graph is preferable over tables
+The permanent storage of results is as a one-time cost to avoid recomputation when the same analysis is re-run.
+For population analyses, Spfy searches the graph for all data points annotated with the queried ontology term.
+This graph data is then converted into the required structure, usually numerical arrays, for the given analysis modules.
 

### 2018-01-31 paper-webserver.tex
EDIT: some changes to the web design paragraph

@@
 
+% TODO: add figure illustrating search.
+
 \subsection{Web design}
@@
 
-The front-end website is written using the React JavaScript library \url{https://facebook.github.io/react/} as a single-page application to allow efficient data-flow without reloading the website.
+The front-end website is written using the React JavaScript library \url{https://facebook.github.io/react/} as a single-page application resembling a traditional desktop application.
 To ensure a familiar user interface, we followed the Material Design specification \url{https://material.io/}, published by Google, surrounding a card-based design.
 (see Figure \ref{fig-results})
-Both the task selection and result displays follow this card-based design: while data storage is graph-based, the results of various analysis modules are presented to users in a familiar tabular structure and available for download as .csv spreadsheet files.
+Both the task selection and result displays follow the same design pattern: while data storage is graph-based, the results of various analysis modules are presented to users in a familiar tabular structure and available for download as .csv spreadsheet files.
 (see Figure \ref{fig-tables})

### 2018-02-01 paper-webserver.tex
EDIT: misc

@@
 TravisCI \url{travisci.io}, a continuous integration (CI) platform, is used to ensure that Spfy \url{https://github.com/superphy/backend} does not break with any changes to the codebase, and runs tests for functionality and backwards compatibility.
-The individual tests use PyTest \url{https://doc.pytest.org/}, and are run within TravisCI's virtual environment, and the current build status can be checked either on our Gtihub repository or at \url{https://travis-ci.org/superphy/backend}.
-CI is also used to automatically build Spfy's core Docker images, and upload them to Docker Hub \url{https://hub.docker.com/u/superphy/}.
+The individual tests use PyTest \url{https://doc.pytest.org/}, and are run within TravisCI's virtual environment.
+The current build status can be checked either on our Gtihub repository or at \url{https://travis-ci.org/superphy/backend}.
+The CI is also used to automatically build Spfy's core Docker images, and upload them to Docker Hub \url{https://hub.docker.com/u/superphy/}.
 
@@
 Spfy was tested with 10,243 public \textit{E. coli} assembled genomes from Enterobase, storing every sequence and results for all included analysis modules.
-Spfy provides real-time subtyping, and the results are immediately displayed to the user following their completion.
-Subtyping options include O-antigen, H-antigen, Shiga-toxin 1, Shiga-toxin 2, and Intimin typing.
-Reference-lab tests include virulence factor and anti-microbial resistance annotation.
-All genomes are analyzed within the pan-genome framework of \textit{E.
+All subtyping options: O-antigen, H-antigen, Shiga-toxin 1, Shiga-toxin 2, and Intimin typing, VF and AMR annotation were ran on the reference set.
+The genomes were also analyzed within the pan-genome framework of \textit{E.
 coli}, and results from all analyses are automatically associated with the source genome.
@@
 Such tools were often script-based, with custom data formats, and only suitable for small collections of data \cite{de2015trends}.
-While this was acceptable for smaller analyses, bioinformatic pipelines utilizing WGS data are larger and involve linked dependencies, which require the application of systems engineering principles \cite{schatz2015biological}.
+While this was acceptable for smaller analyses, bioinformatic pipelines utilizing WGS data are larger and involve linked dependencies, which would benefit from a more standardized design \cite{schatz2015biological}.
 Additionally, many subsets of biology now require the analyses of big-data, where the ability to perform computations in real-time, store data in flexible databases, and utilize a common application programming interface (API) linking resources are required \cite{swaminathan2016review}.

### 2018-02-01 paper-webserver.tex
ADD: few sentances on the sharable links

@@
 
+No account creation is required for using the platform.
+A sharable token is automically created for users upon opening the website, and is embeded into the website address.
+Users can share results with their colleagues by simply copying the url, and files can be submitted for analysis from different computers with results reflected on each others' screens.
+
 \subsection{Service Virtualization}
@@
 
-\subsection{Continuous integration and testing}
+\subsection{Continuous integration}
 % Keep this short

### 2018-02-02 paper-webserver.tex
DEBUG: figures are everywhere
CHANGE: rewrote the intro for the Discussion section to focus on data storage / speed, as opposed to the analysis pipelines

--- 2018-02-01 paper-webserver.tex
@@
 
-\begin{figure}[t]
+\begin{figure}[!hb]
 \begin{center}
@@
 \end{figure}
+
+No account creation is required to use the platform.
+A sharable token is automically created for users upon opening the website, and is embeded into the website address.
+Users can share results with their colleagues by copying the url, and files can be submitted for analysis from different computers with results reflected on each others' screens.
 
@@
 % analysis run-time / throughput with different levels of parallelization
+% particularly for statistical tests
+% should do a 1 genome = X number VFs, Y number of genomes for Y*X retrieval/analysis
 
@@
 
-One of the key goals in developing Spfy is to maintain instantaneity, as modern websites have accustomed users to immediate results.
-We attempt to use innovations in web development to bring a similar experience to Spfy as a predictive genomics platform for \textit{E. coli}.
-The use of Docker containerization, task queues, and a graph database, were all chosen to support instantaneity; modern analytics platforms must respond to user-requested analyses of big-data in the same efficiency as old analyses of single files.
+One of the key goals in developing Spfy was to accomadate a variety of result formats, from storing subtyping data to retrieving results as inputs for downstream analyses, such as statistical significance testing.
+We've shown how a graph database can accomdate a variety of programs used by bioinformaticians and is performant for data retrieval on the results of 10,000+ genomes.
+Spfy maintains instantaneity, as modern analytics platforms must respond to user-requested analyses of big-data in the same efficiency as old analyses of single files.
 

### 2018-02-02 paper-webserver.tex
CHANGE: focus the first para of the discussion on this data storage as well

@@
 % drawbacks - big data
-Many previous bioinformatics software programs have been developed \textit{ad hoc}, with individual researchers and laboratories developing software specific to their environment \cite{de2015trends}.
+Many bioinformatics software programs have been developed \textit{ad hoc}, with individual researchers and laboratories developing software specific to their environment \cite{de2015trends}.
 Such tools were often script-based, with custom data formats, and only suitable for small collections of data \cite{de2015trends}.
-While this was acceptable for smaller analyses, bioinformatic pipelines utilizing WGS data are larger and involve linked dependencies, which would benefit from a more standardized design \cite{schatz2015biological}.
-Additionally, many subsets of biology now require the analyses of big-data, where the ability to perform computations in real-time, store data in flexible databases, and utilize a common application programming interface (API) linking resources are required \cite{swaminathan2016review}.
+Recent efforts \cite{goecks2010galaxy,thomsen2016bacterial} have focused on providing a common web interface for these programs, while still returning the same result files.
+However, many subsets of biology now require the analyses of big-data, where inputs are taken from a variety of analysis programs, and involve large-scale data warehousing \cite{schatz2015biological}.
+The ability to perform computations in real-time, store data in flexible databases, and utilize a common application programming interface (API) linking resources are required for these analyses \cite{swaminathan2016review}.
 

### 2018-02-02 paper-webserver.tex
START: cleanup of the comparison section (to BAP/IRIDA) to focus on the data storage

@@
 % namely galaxy
-Existing scientific workflow technologies such as Galaxy \cite{goecks2010galaxy}, and pipelines such as the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial} and the Integrated Rapid Infectious Disease Analysis (IRIDA) platform \url{http://www.irida.ca/} help automate the use of WGS data for public-health surveillance.
+Existing scientific workflow technologies such as Galaxy \cite{goecks2010galaxy}, and pipelines such as the Integrated Rapid Infectious Disease Analysis (IRIDA) platform \url{http://www.irida.ca/} and the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial} help automate the use of WGS data for public-health surveillance.
 Galaxy aims to provide a reproducible, computation-based research environment which is accessible to individuals without programming knowledge.
@@
 IRIDA is build on top of the Galaxy framework, and adds prebuilt pipelines specific to bioinformatics uses, as well as sequence and result storage.
-IRIDA takes a project-based approach, with sequences stored per project, and results stored linearly per sequence.
-IRIDA adds Controls for collaborating on projects, and uses common terms found in the GenEpiO ontology to describe results.
-
-The Bacterial Analysis Platform (BAP), developed out of the Technical University of Denmark, provides an integrated analysis pipeline for bacterial WGS data as a web service.
-BAP takes a combined approach to genome analysis including different programs, such as for VF and AMR determination, by default into the pipeline.
-A record of the different files generated by an analysis are stored in a relational database, and the record is queried to determine which files are returned to the user \cite{thomsen2016bacterial}.
+IRIDA takes a project-based approach, with sequences stored per project, and results stored linearly per whole-genome sequence.
+Simiarily, BAP provides an integrated analysis pipeline for bacterial WGS data as a web service.
+BAP intergrats different programs, such as for VF and AMR determination, into one pipeline.
+For result storage, these workflow technologies either use relational tables \cite{goecks2010galaxy}, or store resulting files to disk \cite{thomsen2016bacterial}.
 
 Like IRIDA and BAP, Spfy automates workflows for users, and like Galaxy, Spfy uses task queues to distribute selected analysis.
-File uploads begin through the ReactJS-based website, where user-defined analyses options are selected.
-To these concepts we add Docker containerization for task queue workers, allowing analysis software to safely run in parallel.
 % re: Matt "2. When comparing to IRIDA, BAP etc., can mention some differences with Superphy, e.g. the storage of interim result data that allows downstream integrated analysis"
-For result storage, existing workflow technologies use relational tables \cite{goecks2010galaxy}, or store resulting files to disk \cite{thomsen2016bacterial}.
 Because output from these programs is user-specific or transitory, results from identical comparisons are often recomputed.
@@
 Graph-based storage of all results solves these problems.
-% as before
-The entire platform is packaged using Docker-Compose, and can be recreated with a simple command. \par
 

### 2018-02-08 paper-webserver.tex
EDIT: changes to the discussion (IRIDA/BAP) to focus on data storage

@@
 
+% standard web tech means you can deploy to different cloud computing services
+One of the key benefits of using more common-place technologies is the compatibility with other infrastructure resources.
+Docker containerization is widely supported by cloud computing services: Amazon Web Services (AWS) \url{https://aws.amazon.com/docker/}, Google Cloud Platform (GCloud) \url{https://cloud.google.com/container-engine/}, and Microsoft Azure \url{https://azure.microsoft.com/en-us/services/container-service/}, and self-hosted cloud computing technologies such as OpenStack \url{https://wiki.openstack.org/wiki/Docker} all support Docker.
+% we may want to sign up for a free trial to give an example of this
+Spfy packages compute nodes, which manage a collection of task queue workers, as reproducible Docker containers which can be networked together to form a compute cluster.
+Docker containerization has a negligible impact on performance \cite{di2015impact}, and allows the platform to easily scale to demand.
+
 \subsection{Continuous integration}
@@
 Existing scientific workflow technologies such as Galaxy \cite{goecks2010galaxy}, and pipelines such as the Integrated Rapid Infectious Disease Analysis (IRIDA) platform \url{http://www.irida.ca/} and the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial} help automate the use of WGS data for public-health surveillance.
-Galaxy aims to provide a reproducible, computation-based research environment which is accessible to individuals without programming knowledge.
-Galaxy tackles the problem of linking different analysis software together, by defining interdependencies using a custom schema.
-A visual workflow editor is also provided for ease of use.
-IRIDA is build on top of the Galaxy framework, and adds prebuilt pipelines specific to bioinformatics uses, as well as sequence and result storage.
+Reproducibility is important to science, and Galaxy aims to provide a reproducible, computation-based research environment which is accessible to individuals without programming knowledge.
+Galaxy defines a formal schema for linking different analysis software together, so analyses can be replicated end-to-end.
+Building on top of the Galaxy framework, IRIDA adds prebuilt pipelines specific to bioinformatics uses, as well as sequence and result storage.
 IRIDA takes a project-based approach, with sequences stored per project, and results stored linearly per whole-genome sequence.
 Simiarily, BAP provides an integrated analysis pipeline for bacterial WGS data as a web service.
-BAP intergrats different programs, such as for VF and AMR determination, into one pipeline.
 For result storage, these workflow technologies either use relational tables \cite{goecks2010galaxy}, or store resulting files to disk \cite{thomsen2016bacterial}.
@@
 % re: Matt "2. When comparing to IRIDA, BAP etc., can mention some differences with Superphy, e.g. the storage of interim result data that allows downstream integrated analysis"
+On a per file basis, Spfy performs at a similar speed to BAP on predictive genomics tasks, though Spfy does not provide genome assembly services.
+Spfy also processes XXXX files over XX tasks in XXXX time by using a novel approach of distributing computations over a task queue and multiple Docker compartmentalized containers.
+We also focus on the data warehousing problem in bioinformatics: storing results as in IRIDA and BAP, but in a way where we can integrate results for further analyses.
+Like BAP, Spfy intergrates different programs, such as for VF and AMR determination, into one pipeline. 
 Because output from these programs is user-specific or transitory, results from identical comparisons are often recomputed.
@@
 
-On a per file basis, Spfy performs at a similar speed to BAP on predictive genomics tasks, though Spfy does not provide genome assembly services.
-Spfy also processes XXXX files over XX tasks in XXXX time by using a novel approach of distributing computations over a task queue and multiple Docker compartmentalized containers.
-In place of a relational database for storing the location of result files, Spfy parses result data into graph objects and integrates results into a persistent graph database.
+In place of a relational database for storing the location of result files, Spfy parses and maps results to formal ontologies, with the end result stored in a persistent graph database.
 This enables Spfy to create large datasets where analysis results are linked to genomes, and to perform population-wide analyses.
-In all, Spfy provides similar functionality with an expanded focus on integrated result storage and big-data analyses.
-
-To approach the challenge of integrating different bioinformatics programs, Spfy instead uses technologies prevalent in common web services not necessarily related to scientific workflows.
-While we prefer to package individual bioinformatics programs using language-agnostic package managers such as Conda \url{https://conda.io/}, we do agree that suites of bioinformatics software benefit from having all dependencies installed and deployable in Docker containers \cite{di2015impact}.
-These programs are connected to form a workflow using task queues; a design methodology used to offload long-running or non-instantaneous tasks over available computing infrastructure.
-Unlike platforms such as Galaxy which favor a drag-and-drop design for building workflows, Spfy favors a tightly coupled approach explicitly linking together different bioinformatics programs within the task queue implementation.
-In particular, Spfy uses asynchronous tasks queues which are tripped in response to user requests, thereby allowing immediate responsivity to users on the website, and processes generating human-readable results can be separated from processes handling long-term result storage.
-
-% standard web tech means you can deploy to different cloud computing services
-One of the key benefits of using more common-place technologies is the compatibility with other infrastructure resources.
-Docker containerization is widely supported by cloud computing services: Amazon Web Services (AWS) \url{https://aws.amazon.com/docker/}, Google Cloud Platform (GCloud) \url{https://cloud.google.com/container-engine/}, and Microsoft Azure \url{https://azure.microsoft.com/en-us/services/container-service/}, and self-hosted cloud computing technologies such as OpenStack \url{https://wiki.openstack.org/wiki/Docker} all support Docker.
-% we may want to sign up for a free trial to give an example of this
-Spfy packages compute nodes, which manage a collection of task queue workers, as reproducible Docker containers which can be networked together to form a compute cluster.
-Docker containerization has a negligible impact on performance \cite{di2015impact}, and allows the platform to easily scale to demand.
-In contrast to Galaxy, Spfy is a targeted approach building a predictive analytics platform for \textit{E.
-coli} which leverages generic web technologies to favor a responsive, big-data design strategy tackling the challenge \cite{fricke2014bacterial} of integrating results from multiple genomes and examining shared connections.
-By packaging the individual bioinformatics software as Conda packages, they can be easily ported to Galaxy.
+In all, Spfy provides similar functionality to IRIDA and BAP with an expanded focus on integrated result storage and big-data analyses.
 
@@
 
-Future work will focus on adding additional analyses modules, using machine learning and artificial neural networks to aid genotype to phenotype predictions, and supporting different species.
+Future work will focus on adding additional analyses modules to aid genotype to phenotype predictions, and supporting different species.
+We're currently working on an approach using support vector machiens to predict \textit{E.coli} phenotypes \url{https://github.com/superphy/kmer} which will be integerated into Spfy.
 While the integrated approach of storing and retrieving results provides enormous benefits, the developed analyses modules are self-contained and can easily be integrated into existing platforms such as Galaxy \cite{goecks2010galaxy}, and IRIDA.

### 2018-02-08 paper-webserver.tex

@@
 No account creation is required to use the platform.
-A sharable token is automically created for users upon opening the website, and is embeded into the website address.
-Users can share results with their colleagues by copying the url, and files can be submitted for analysis from different computers with results reflected on each others' screens.
+A sharable token is automatically created for users upon opening the website, and is embedded into the website address.
+Users can share results with their colleagues by copying the URL, and files can be submitted for analysis from different computers with results reflected on each others' screens.
 
@@
 The individual tests use PyTest \url{https://doc.pytest.org/}, and are run within TravisCI's virtual environment.
-The current build status can be checked either on our Gtihub repository or at \url{https://travis-ci.org/superphy/backend}.
+The current build status can be checked either on our GitHub repository or at \url{https://travis-ci.org/superphy/backend}.
 The CI is also used to automatically build Spfy's core Docker images, and upload them to Docker Hub \url{https://hub.docker.com/u/superphy/}.
@@
 
-One of the key goals in developing Spfy was to accomadate a variety of result formats, from storing subtyping data to retrieving results as inputs for downstream analyses, such as statistical significance testing.
-We've shown how a graph database can accomdate a variety of programs used by bioinformaticians and is performant for data retrieval on the results of 10,000+ genomes.
+One of the key goals in developing Spfy was to accommodate a variety of result formats, from storing subtyping data to retrieving results as inputs for downstream analyses, such as statistical significance testing.
+We've shown how a graph database can accommodate a variety of programs used by bioinformaticians and is performant for data retrieval on the results of 10,000+ genomes.
 Spfy maintains instantaneity, as modern analytics platforms must respond to user-requested analyses of big-data in the same efficiency as old analyses of single files.
@@
 IRIDA takes a project-based approach, with sequences stored per project, and results stored linearly per whole-genome sequence.
-Simiarily, BAP provides an integrated analysis pipeline for bacterial WGS data as a web service.
+Similarly, BAP provides an integrated analysis pipeline for bacterial WGS data as a web service.
 For result storage, these workflow technologies either use relational tables \cite{goecks2010galaxy}, or store resulting files to disk \cite{thomsen2016bacterial}.
@@
 We also focus on the data warehousing problem in bioinformatics: storing results as in IRIDA and BAP, but in a way where we can integrate results for further analyses.
-Like BAP, Spfy intergrates different programs, such as for VF and AMR determination, into one pipeline. 
+Like BAP, Spfy integrates different programs, such as for VF and AMR determination, into one pipeline. 
 Because output from these programs is user-specific or transitory, results from identical comparisons are often recomputed.
@@
 Future work will focus on adding additional analyses modules to aid genotype to phenotype predictions, and supporting different species.
-We're currently working on an approach using support vector machiens to predict \textit{E.coli} phenotypes \url{https://github.com/superphy/kmer} which will be integerated into Spfy.
+We're currently working on an approach using support vector machines to predict \textit{E.coli} phenotypes \url{https://github.com/superphy/kmer} which will be integrated into Spfy.
 While the integrated approach of storing and retrieving results provides enormous benefits, the developed analyses modules are self-contained and can easily be integrated into existing platforms such as Galaxy \cite{goecks2010galaxy}, and IRIDA.

### 2018-02-12 paper-webserver.tex
EDIT: some general edits, but mostly changes to the Comparison to BAP/IRIDA

@@
 Current comparative computational workflows chain different analysis software, but lack storage and retrieval methods for generated results.
-Spfy uses graph data structures to store and retrieve results for computational workflows, facilitating the management and querying of tens of thousands of whole-genome \textit{E.
-coli} sequences, and efficient downstream processing.
+Spfy uses graph data structures to store and retrieve results for computational workflows, facilitating the management and querying of tens of thousands of whole-genome \textit{Escherichia coli} sequences, and efficient downstream processing.
 By making the storage and retrieval of results part of the platform, with data effectively linked to the organisms of interest through a standardized ontology, we can mitigate the recomputing of analyses.
@@
 
-The database is available at \url{https://lfz.corefacility.ca/superphy/spfy/}.
+Database URL: \url{https://lfz.corefacility.ca/superphy/spfy/}.
 
@@
 Comprehensive platforms that combine individual programs into a cohesive whole also exist.
-These include free platforms such as the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial}, the Integrated Rapid Infectious Disease Analysis (IRIDA) project \url{http://www.irida.ca/}, and PATRIC \cite{wattam2013patric}.
-Commercial applications, such as Bionumerics, which is used by PulseNet international for the analyses of WGS data in outbreak situations also exist, and offer support as well as accredited, standardized tests \cite{swaminathan2001pulsenet}.
+These include free platforms such as the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial}, the Integrated Rapid Infectious Disease Analysis (IRIDA) project \url{http://www.irida.ca/}, and the Pathosystems Resource Integration Center (PATRIC) \cite{wattam2013patric}.
+Commercial applications, such as Bionumerics, which is used by PulseNet International for the analyses of WGS data in outbreak situations also exist, and offer support as well as accredited, standardized tests \cite{swaminathan2001pulsenet}.
 
 % 3. 100,000 genomes, (lookup number for Enterobase, GenBank), how do we run analysis on it
-Traditionally, these platforms have been applied to samples numbering only a few dozens while WGS data for bacterial pathogens of public health importance have recently accumulated in public databases in the tens of thousands, with hundreds of thousands set to be available within the next few years.
+Traditionally, these platforms have been applied to samples numbering only a few dozen while WGS data for bacterial pathogens of public health importance have recently accumulated in public databases in the tens of thousands, with hundreds of thousands set to be available within the next few years.
 For \textit{Escherichia coli}, there are over sixty thousand publicly available genomes in EnteroBase \url{https://enterobase.warwick.ac.uk/} and three million whole genomes in GenBank \cite{doi:10.1093/nar/gks1195}.
@@
 Results are converted to individual graphs and stored within a larger graph database.
-By integrating task distribution with graph storage, Spfy enables large-scale analyses, such as epidemiological associations between specific genotypes, biomarkers, host, source, and other metadata, and statistical significance testing of genome markers for user-defined groups using SciPy \cite{jones2014scipy}.
+By integrating task distribution with graph storage, Spfy enables large-scale analyses, such as epidemiological associations between specific genotypes, biomarkers, host, source, and other metadata, and statistical significance testing of genome markers for user-defined groups.
 Any data type or relation in the graph is valid for analysis, such as the presence or absence of pan-genome regions against determined serotypes or other subtyping options.
 In addition, Spfy supports the submission of user-specified metadata, such as location or source information.
-Currently, the platform has been tested with XXX genome files and result storage for XX analyses modules.
 
@@
 % para: intro to the spfy stack
-Spfy's server-side code is developed in Python and the front-end website is developed using the React JavaScript library. For the addition of new data to the database, the following steps are taken:
+Spfy's server-side code, graph generation, and analysis modules are developed in Python and the front-end website is developed using the React JavaScript library.
+For the addition of new data to the database, the following steps are taken:
 
@@
 For biological data, individual data points such as genome, contiguous DNA sequence, or gene, can be linked together in a searchable graph structure.
-This allows novel data to be incorporated into the existing graph, and has the use of graph databases for semantic information has been proposed as a open standard for sharing public information \cite{horrocks2005semantic}.
+This allows novel data to be incorporated into the existing graph, and the use of graph databases for semantic information has been proposed as a open standard for sharing public information \cite{horrocks2005semantic}.
 Data points are annotated with terms from published ontologies.
@@
 % TODO: add figure illustrating search.
+In a graph database, search begins...
+Comparison to SQL...
+NoSQL...
 
@@
 A sharable token is automatically created for users upon opening the website, and is embedded into the website address.
-Users can share results with their colleagues by copying the URL, and files can be submitted for analysis from different computers with results reflected on each others' screens.
+Users can share results by copying the URL, and files submitted from different computers using the same token will have the results visible to anyone with the same URL.
 
@@
 On a per file basis, Spfy performs at a similar speed to BAP on predictive genomics tasks, though Spfy does not provide genome assembly services.
-Spfy also processes XXXX files over XX tasks in XXXX time by using a novel approach of distributing computations over a task queue and multiple Docker compartmentalized containers.
-We also focus on the data warehousing problem in bioinformatics: storing results as in IRIDA and BAP, but in a way where we can integrate results for further analyses.
+Spfy processes XXXX files over XX tasks in XXXX time by using a novel approach of distributing computations over a task queue and multiple Docker compartmentalized containers.
+
+We also focus on the data warehousing problem in bioinformatics: storing results as in IRIDA and BAP, but using a graph database to integrate results for further analyses.
 Like BAP, Spfy integrates different programs, such as for VF and AMR determination, into one pipeline.
@@
 Without a unified structure, these conversions quickly become impractical for broad usage.
-Graph-based storage of all results solves these problems.
-
-In place of a relational database for storing the location of result files, Spfy parses and maps results to formal ontologies, with the end result stored in a persistent graph database.
-This enables Spfy to create large datasets where analysis results are linked to genomes, and to perform population-wide analyses.
-In all, Spfy provides similar functionality to IRIDA and BAP with an expanded focus on integrated result storage and big-data analyses.
+Graph-based storage of all results solves these problems and allows Spfy to perform population-wide analyses, regardless of the individual analysis software used to generate the datasets.
 

### 2018-02-21 paper-webserver.tex
START: edits

@@
 A sharable token is automatically created for users upon opening the website, and is embedded into the website address.
-Users can share results by copying the URL, and files submitted from different computers using the same token will have the results visible to anyone with the same URL.
+Users can share results by copying the URL, and files submitted from different computers using the same token will be visible to anyone with the same URL.
 
@@
 We're currently working on an approach using support vector machines to predict \textit{E.coli} phenotypes \url{https://github.com/superphy/kmer} which will be integrated into Spfy.
-While the integrated approach of storing and retrieving results provides enormous benefits, the developed analyses modules are self-contained and can easily be integrated into existing platforms such as Galaxy \cite{goecks2010galaxy}, and IRIDA.
-Spfy's codebase is provided at \url{https://github.com/superphy/backend} and a developer guide is provided at \url{https://superphy.readthedocs.io/en/latest/}.
+While the integrated approach of storing and retrieving results provides enormous benefits, the developed analyses modules are self-contained and can easily be integrated into existing platforms such as Galaxy, and IRIDA.
+Spfy's code hosted at \url{https://github.com/superphy/backend}, and is available for free under the open-source Apache 2.0 license; a developer guide is provided at \url{https://superphy.readthedocs.io/en/latest/}.
 

### 2018-02-21 paper-webserver.tex

--- 2018-02-12 paper-webserver.tex
@@
 Current comparative computational workflows chain different analysis software, but lack storage and retrieval methods for generated results.
-Spfy uses graph data structures to store and retrieve results for computational workflows, facilitating the management and querying of tens of thousands of whole-genome \textit{Escherichia coli} sequences, and efficient downstream processing.
-By making the storage and retrieval of results part of the platform, with data effectively linked to the organisms of interest through a standardized ontology, we can mitigate the recomputing of analyses.
-Within Spfy, the output from all analyses is stored, and linked together in the context of a genome graph.
-This graph also stores metadata for each genome, facilitating inquiries ranging from population genomics to epidemiological investigations.
+Spfy uses a graph database to store and retrieve results for computational workflows, facilitating the management and querying of tens of thousands of whole-genome \textit{Escherichia coli} sequences, and efficient downstream processing.
+By making the storage and retrieval of results part of the platform, with data effectively linked to the organisms of interest through a standardized ontology, we can facilitate inquiries ranging from population genomics to epidemiological investigations, and mitigate the recomputing of analyses.
+A graph approach is flexible to accommodate new analysis software as they are developed.
 Integrated data storage will be necessary as publicly available whole genome sequencing data for bacterial pathogens currently numbers in the tens of thousands, with hundreds of thousands set to be available within the next few years.
@@
 Graph-based result storage also allows retrospective comparisons as more genomes are sequenced or populations change, and is flexible to accommodate new analysis modules as they are developed.
-The website is available at \url{https://lfz.corefacility.ca/superphy/spfy/}.
+The database is available at \url{https://lfz.corefacility.ca/superphy/spfy/}.
 
@@
 By supporting multiple \textit{in-silico} subtyping options, the platform functions similar to a reference laboratory, while adding support for population analyses.
-Subtyping options for \textit{E.coli} are O-antigen, H-antigen, and VF gene determination using ECtyper \url{https://github.com/phac-nml/ecoli\_serotyping}, Shiga-toxin 1 (Stx1), Shiga-toxin 2 (Stx2), and Intimin typing using Phylotyper \cite{whiteside2017phylotyper}, and AMR annotation using the RGI program \cite{mcarthur2013comprehensive}.
+Subtyping options for \textit{E.coli} are O-antigen, H-antigen, and VF gene determination using ECTyper \url{https://github.com/phac-nml/ecoli\_serotyping}, Shiga-toxin 1 (Stx1), Shiga-toxin 2 (Stx2), and Intimin typing using Phylotyper \cite{whiteside2017phylotyper}, and AMR annotation using the RGI program \cite{mcarthur2013comprehensive}.
 Spfy also performs bioinformatics analyses: pangenome generation using Panseq \cite{laing2010pan} and support vector machine (SVM)-backed AMR predictions for OmniLog AMR assays using Scikit-learn \cite{pedregosa2011scikit}.
@@
 % para: intro to the spfy stack
-Spfy's server-side code, graph generation, and analysis modules are developed in Python and the front-end website is developed using the React JavaScript library.
+Spfy's server-side code, graph generation, and analysis modules are developed in Python and the front-end website is developed using the React JavaScript library \url{https://facebook.github.io/react/}.
 For the addition of new data to the database, the following steps are taken:
 
-i) The upload begins through the ReactJS-based website, where user-defined analyses options are selected.
+i) The upload begins through the website, where user-defined analyses options are selected.
 The results of these chosen analyses are immediately reported to the user following their completion, while the remaining analyses are subsequently completed and stored in the database without interaction from the user.
@@
 Data points are annotated with terms from published ontologies.
-To avoid proliferating ontologies, and to allow Spfy to integrate with existing ones, annotations from the GenEpiO \cite{griffiths2017context}, FALDO \cite{bolleman2016faldo}, and TypOn \cite{vaz2014typon} ontologies are used for our annotations.
+To avoid proliferating ontologies, and to allow Spfy to integrate with existing ones, we use annotations from the GenEpiO \cite{griffiths2017context}, FALDO \cite{bolleman2016faldo}, and TypOn \cite{vaz2014typon}.
 
@@
 \end{center}
-\caption{Caption for figure within column.}
+\caption{An example of how data is stored in Spfy.
+Brackets highlight the source of different data points and the software it was generated from.
+These parts are added in as the analyis modules complete, at varying times, and the overall connections are inferred by the database.}
 \label{fig-ontology}
@@
 For population analyses, Spfy searches the graph for all data points annotated with the queried ontology term.
-This graph data is then converted into the required structure, usually numerical arrays, for the given analysis modules.
-
-% TODO: add figure illustrating search.
-In a graph database, search begins...
-Comparison to SQL...
-NoSQL...
+This graph data is then converted into the required structure, usually numerical arrays, as required for the given analysis modules.
+In a graph database, search begins at any node or attribute.
+This is in contrast to a SQL database which requires a predefined schema, or a NoSQL database which treats data as documents with varying structure.
+For example, the addition of a new analysis module would typically require a new table definition in a SQL database, or the addition of a new document type in a NoSQL database.
+A graph database instead adds new nodes or attributes which can be connected to existing data, so explicit joins or data conversions are not required.
+Spfy also adds the graph results in parts, and allows the database to infer their connections.
 
@@
 
-The front-end website is written using the React JavaScript library \url{https://facebook.github.io/react/} as a single-page application resembling a traditional desktop application.
+The front-end website is written as a single-page application.
 To ensure a familiar user interface, we followed the Material Design specification \url{https://material.io/}, published by Google, surrounding a card-based design.
@@
 No account creation is required to use the platform.
-A sharable token is automatically created for users upon opening the website, and is embedded into the website address.
-Users can share results by copying the URL, and files submitted from different computers using the same token will have the results visible to anyone with the same URL.
+A sharable token is automatically created for users upon entering the website, and is embedded into the website address.
+Users can share results by copying their URL, and files submitted from different computers using the same token will be visible to anyone one the same link.
 
@@
 Docker integration ensures that software dependencies, which are typically manually installed \cite{doi:10.1093/bioinformatics/btu153,laing2010pan,inouye2014srst2,naccache2014cloud}, are instead handled automatically.
-With compartmentalization of service runtimes, code failures do not propagate to other services.
 
@@
 One of the key benefits of using more common-place technologies is the compatibility with other infrastructure resources.
-Docker containerization is widely supported by cloud computing services: Amazon Web Services (AWS) \url{https://aws.amazon.com/docker/}, Google Cloud Platform (GCloud) \url{https://cloud.google.com/container-engine/}, and Microsoft Azure \url{https://azure.microsoft.com/en-us/services/container-service/}, and self-hosted cloud computing technologies such as OpenStack \url{https://wiki.openstack.org/wiki/Docker} all support Docker.
+Docker containers are widely supported by cloud computing services: Amazon Web Services (AWS) \url{https://aws.amazon.com/docker/}, Google Cloud Platform (GCloud) \url{https://cloud.google.com/container-engine/}, and Microsoft Azure \url{https://azure.microsoft.com/en-us/services/container-service/}, and self-hosted cloud computing technologies such as OpenStack \url{https://wiki.openstack.org/wiki/Docker} all support Docker.
 % we may want to sign up for a free trial to give an example of this
-Spfy packages compute nodes, which manage a collection of task queue workers, as reproducible Docker containers which can be networked together to form a compute cluster.
-Docker containerization has a negligible impact on performance \cite{di2015impact}, and allows the platform to easily scale to demand.
+Spfy packages compute nodes as reproducible Docker containers, and allows the platform to easily scale to demand.
 
@@
 
-TravisCI \url{travisci.io}, a continuous integration (CI) platform, is used to ensure that Spfy \url{https://github.com/superphy/backend} does not break with any changes to the codebase, and runs tests for functionality and backwards compatibility.
-The individual tests use PyTest \url{https://doc.pytest.org/}, and are run within TravisCI's virtual environment.
-The current build status can be checked either on our GitHub repository or at \url{https://travis-ci.org/superphy/backend}.
-The CI is also used to automatically build Spfy's core Docker images, and upload them to Docker Hub \url{https://hub.docker.com/u/superphy/}.
+Our tests for functionality and backwards compatibility run on TravisCI \url{https://travisci.io}, a continuous integration (CI) platform.
+The individual tests use PyTest \url{https://doc.pytest.org/}, and the current build status can be checked either on our GitHub repository or at \url{https://travis-ci.org/superphy/backend}.
+TravisCI also builds Spfy's core Docker images, and uploads them to Docker Hub \url{https://hub.docker.com/u/superphy/}.
 
 \section{RESULTS}
-Spfy was tested with 10,243 public \textit{E. coli} assembled genomes from Enterobase, storing every sequence and results for all included analysis modules.
-All subtyping options: O-antigen, H-antigen, Shiga-toxin 1, Shiga-toxin 2, and Intimin typing, VF and AMR annotation were ran on the reference set.
+Spfy was tested with 10,243 public \textit{E. coli} assembled genomes from Enterobase, storing every sequence and the results for all included analysis modules.
+All subtyping options: O-antigen, H-antigen, Shiga-toxin 1, Shiga-toxin 2, and Intimin typing, VF and AMR annotation were ran on the test set.
 The genomes were also analyzed within the pan-genome framework of \textit{E.
-coli}, and results from all analyses are automatically associated with the source genome.
-The resulting database had 17,820 nodes and 3,811,473 leaves, with 1,125,909,074 object properties. \par
-
-\small \input{tables/database}
-
+coli}.
+The resulting database had 17,820 nodes and 3,811,473 leaves, with 1,125,909,074 object properties.
 Spfy has been up since May 2017.
@@
 coli} gene sequences before running analyses.
-Outputs are displayed on the website in tables and can be downloaded as \textit{.csv} files.
 \par
+
+\small \input{tables/database}
 
@@
 % namely galaxy
-Existing scientific workflow technologies such as Galaxy \cite{goecks2010galaxy}, and pipelines such as the Integrated Rapid Infectious Disease Analysis (IRIDA) platform \url{http://www.irida.ca/} and the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial} help automate the use of WGS data for public-health surveillance.
-Reproducibility is important to science, and Galaxy aims to provide a reproducible, computation-based research environment which is accessible to individuals without programming knowledge.
-Galaxy defines a formal schema for linking different analysis software together, so analyses can be replicated end-to-end.
+Existing scientific workflow technologies such as Galaxy \cite{goecks2010galaxy}, and pipelines such as the Integrated Rapid Infectious Disease Analysis (IRIDA) platform \url{http://www.irida.ca/} and the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial} help automate the use of WGS data.
+Reproducibility is important, and Galaxy aims to provide a reproducible, computation-based research environment which is accessible to individuals without programming knowledge.
+Galaxy defines a formal schema for linking different analysis software together, so the entire analysis pipeline can be replicated.
 Building on top of the Galaxy framework, IRIDA adds prebuilt pipelines specific to bioinformatics uses, as well as sequence and result storage.
@@
 On a per file basis, Spfy performs at a similar speed to BAP on predictive genomics tasks, though Spfy does not provide genome assembly services.
-Spfy processes XXXX files over XX tasks in XXXX time by using a novel approach of distributing computations over a task queue and multiple Docker compartmentalized containers.
+Spfy processes XXXX files over XX tasks in XXXX time, distributing computations over a task queue and multiple Docker compartmentalized containers.
+We also focus on the data warehousing problem in bioinformatics: storing results as in IRIDA and BAP, but using a graph database to integrate results for further analyses.
+Like BAP, Spfy integrates different programs, such as for VF and AMR determination, into one pipeline. \par
 
-We also focus on the data warehousing problem in bioinformatics: storing results as in IRIDA and BAP, but using a graph database to integrate results for further analyses.
-Like BAP, Spfy integrates different programs, such as for VF and AMR determination, into one pipeline.
 Because output from these programs is user-specific or transitory, results from identical comparisons are often recomputed.
@@
 We're currently working on an approach using support vector machines to predict \textit{E.coli} phenotypes \url{https://github.com/superphy/kmer} which will be integrated into Spfy.
-While the integrated approach of storing and retrieving results provides enormous benefits, the developed analyses modules are self-contained and can easily be integrated into existing platforms such as Galaxy \cite{goecks2010galaxy}, and IRIDA.
-Spfy's codebase is provided at \url{https://github.com/superphy/backend} and a developer guide is provided at \url{https://superphy.readthedocs.io/en/latest/}.
+While the integrated approach of storing and retrieving results provides enormous benefits, the developed analyses modules are self-contained and be transfered to existing platforms such as Galaxy, and IRIDA.
+Spfy's code is hosted at \url{https://github.com/superphy/backend}, and is available for free under the open-source Apache 2.0 license.
+A developer guide is provided at \url{https://superphy.readthedocs.io/en/latest/}.
 

### 2018-02-22 paper-webserver.tex
Abstract edits

@@
 
-\title{Spfy: an integrated graph-based sequence database for real-time prediction of Escherichia coli phenotypes}
+\title{Spfy: an integrated graph-based sequence database for real-time prediction of Escherichia coli phenotypes using the SuperPhy platform}
 
@@
 
-Current comparative computational workflows chain different analysis software, but lack storage and retrieval methods for generated results.
-Spfy uses a graph database to store and retrieve results for computational workflows, facilitating the management and querying of tens of thousands of whole-genome \textit{Escherichia coli} sequences, and efficient downstream processing.
-By making the storage and retrieval of results part of the platform, with data effectively linked to the organisms of interest through a standardized ontology, we can facilitate inquiries ranging from population genomics to epidemiological investigations, and mitigate the recomputing of analyses.
-A graph approach is flexible to accommodate new analysis software as they are developed.
-Integrated data storage will be necessary as publicly available whole genome sequencing data for bacterial pathogens currently numbers in the tens of thousands, with hundreds of thousands set to be available within the next few years.
+Current workflows in comparative computational genomics rely on chaining different analysis software together, but lack storage and retrieval methods for the generated results.
+Reference laboratories are currently moving to whole-genome sequencing based analyses, and require rapid analyses coupled to efficient storage and retrieval of data.
+To solve this problem, we have created Spfy, which uses a graph database to store and retrieve results from computational workflows.
+This facilitates rapid analyses, as well as the efficient management and downstream processing of tens of thousands of whole-genome sequences.
+Though generally applicable to bacterial genome sequences, Spfy currently contains X \textit{Escherichia coli} genomes, for which \textit{in-silico} serotype and Shiga-toxin subtype, as well as the presence of known virulence factors and antimicrobial resistance determinants have been computed.
+Spfy links the results and metadata to the genome sequences through a standardized ontology, which facilitates hypothesis testing in fields ranging from population genomics to epidemiology, while mitigating the recomputing of analyses.
+The graph approach is flexible, and can accommodate new analysis software modules as they are developed, and easily link new results to those already stored.
+Integrated data storage and analyses are currently necessary as the number of publicly available whole genome sequences is currently in the hundreds of thousands, with millions likely to be available within the next few years.
 \par

### 2018-02-22 paper-webserver.tex
Intro edits.

@@
 The rapid characterization and comparison of bacterial pathogens relies principally on the combination of outputs from multiple software programs that are targeted for specific applications.
-Examples include the Resistance Gene Identifier (RGI) \cite{mcarthur2013comprehensive} for antimicrobial resistance (AMR) gene prediction, VirulenceFinder \cite{kleinheinz2014applying} for virulence factor (VF) annotation, and Prokka for bacterial genome annotation with external tools \cite{doi:10.1093/bioinformatics/btu153}.
-In particular, RGI and VirulenceFinder represent a series of \textit{in-silico} methods which have been developed to replicate the results of traditional wet-lab tests; this allows new WGS results to be viewed in the context of historical tests.
+Examples include the identification of known antimircrobial (AMR) resistance genes, through software such as the Resistance Gene Identifier (RGI) \cite{mcarthur2013comprehensive}, {ResFinder}, {ARG-ANNOT}, and {ARIBA}; the identification of known virulence genes through software such as VirulenceFinder \cite{kleinheinz2014applying}, {SRST2}, and {an additional reference}; the annotation of bacterial genomes, through software such as Prokka \cite{doi:10.1093/bioinformatics/btu153}, {another}, and {another}.
+\todo{Add a section on serotyping and subtyping a la Phylotyper}.
+These methods represent \textit{in-silico} analogues of traditional wet-lab tests, which allows new WGS results to be viewed in the context of historical tests, and greatly expedites the analyses of newly sequenced genomes.
 
@@
 % 3. 100,000 genomes, (lookup number for Enterobase, GenBank), how do we run analysis on it
-Traditionally, these platforms have been applied to samples numbering only a few dozen while WGS data for bacterial pathogens of public health importance have recently accumulated in public databases in the tens of thousands, with hundreds of thousands set to be available within the next few years.
-For \textit{Escherichia coli}, there are over sixty thousand publicly available genomes in EnteroBase \url{https://enterobase.warwick.ac.uk/} and three million whole genomes in GenBank \cite{doi:10.1093/nar/gks1195}.
+Traditionally, these platforms have been applied to samples numbering only a few dozen \cite{} while WGS of bacterial pathogens have recently accumulated in public databases in the hundreds of thousands, with millions set to be available within the next few years.
+For \textit{Escherichia coli} alone, there are over sixty thousand publicly available genomes in EnteroBase \url{https://enterobase.warwick.ac.uk/} and three million sequenced genomes in GenBank \cite{doi:10.1093/nar/gks1195}.
 
-Many of the comparative analyses that are run are broadly useful, and therefore computed multiple times.
-To begin to compute results on all available genomes, an effective method to mitigate the recomputing of analyses is to make the storage and retrieval of results part of the platform, and effectively linked to the organisms of interest with a standardized ontology.
+Many of the comparative analyses that are currently used in the analyses of bacterial genomes are broadly useful, and therefore computed multiple times for the same genomes.
+An effective method to mitigate the recomputing of analyses, is to make the storage and retrieval of results part of the analyses platform, and effectively linked to the genomes of interest through a standardized ontology.
 Such measures can help ensure the rapid response times required for public health applications, and allow results to be integrated and progressively updated as new data becomes available.
@@
 % 5. how we solved it
-In this study, we present the Spfy platform, which integrates a graph database with real-time analysis options.
-By integrating analysis options with data storage, we avoid recomputing analysis for identical runs.
-Graph-based result storage also allows retrospective comparisons as more genomes are sequenced or populations change, and is flexible to accommodate new analysis modules as they are developed.
+In this study, we present the Spfy update to the SuperPhy platform, which integrates a graph database with real-time analyses; this integration avoids recomputing identical analyses.
+Graph-based result storage also allows retrospective comparisons as more genomes are sequenced or populations change, and is flexible, accommodating new analysis modules as they are developed.
 The database is available at \url{https://lfz.corefacility.ca/superphy/spfy/}.

### 2018-02-22 paper-webserver.tex
Implementation edits.

@@
 
-By supporting multiple \textit{in-silico} subtyping options, the platform functions similar to a reference laboratory, while adding support for population analyses.
-Subtyping options for \textit{E.coli} are O-antigen, H-antigen, and VF gene determination using ECTyper \url{https://github.com/phac-nml/ecoli\_serotyping}, Shiga-toxin 1 (Stx1), Shiga-toxin 2 (Stx2), and Intimin typing using Phylotyper \cite{whiteside2017phylotyper}, and AMR annotation using the RGI program \cite{mcarthur2013comprehensive}.
-Spfy also performs bioinformatics analyses: pangenome generation using Panseq \cite{laing2010pan} and support vector machine (SVM)-backed AMR predictions for OmniLog AMR assays using Scikit-learn \cite{pedregosa2011scikit}.
+Spfy provides rapid in-silco versions of common reference laboratory tests for the analyses of \textit{E.
+coli}.
+It supports the following \textit{in-silico} subtyping options: serotyping, through both O- and H-antigen identification {ectyper ref}; Shiga-toxin 1 (Stx1), Shiga-toxin 2 (Stx2), and Intimin typing using Phylotyper \cite{whiteside2017phylotyper}, VF gene determination using ECTyper \url{https://github.com/phac-nml/ecoli\_serotyping}, and AMR annotation using the RGI program \cite{mcarthur2013comprehensive}.
 
-These tasks are divided into subtasks, and distributed across a built-in task queue.
-Results are converted to individual graphs and stored within a larger graph database.
-By integrating task distribution with graph storage, Spfy enables large-scale analyses, such as epidemiological associations between specific genotypes, biomarkers, host, source, and other metadata, and statistical significance testing of genome markers for user-defined groups.
-Any data type or relation in the graph is valid for analysis, such as the presence or absence of pan-genome regions against determined serotypes or other subtyping options.
-In addition, Spfy supports the submission of user-specified metadata, such as location or source information.
+Spfy also performs pangenome analyses using Panseq \cite{laing2010pan}, and provides machine learning modules for biomarker discovery among groups using Scikit-learn \cite{pedregosa2011scikit}.
+
+Spfy handles all of the analyses tasks by dividing them into subtasks, which are subsequently distributed across a built-in task queue.
+Results are converted into individual graphs and stored within a larger graph database according to standard ontologies {list}, where metadata including genotypes, biomarkers, host, source, and statistical significance testing of genome markers for user-defined groups are stored.
+
+By integrating task distribution with graph storage, Spfy enables large-scale analyses, such as epidemiological association studies.
+Any data type or relation in the graph is a valid option for analysis.
+This means that genomes can be compared on the basis of the presence or absence of pan-genome regions, serotype, subtyping data, or provided metadata such as location or host-source.
 
@@
 % para: intro to the spfy stack
-Spfy's server-side code, graph generation, and analysis modules are developed in Python and the front-end website is developed using the React JavaScript library \url{https://facebook.github.io/react/}.
-For the addition of new data to the database, the following steps are taken:
+The server-side code for Spfy, graph generation, and analysis modules, are developed in Python, with the front-end website developed using the React JavaScript library \url{https://facebook.github.io/react/}.
+When new data is added to the database, the following steps are taken:
 
 i) The upload begins through the website, where user-defined analyses options are selected.
-The results of these chosen analyses are immediately reported to the user following their completion, while the remaining analyses are subsequently completed and stored in the database without interaction from the user.
-The public web service accepts up to 200 MB of genome files (50 genomes uncompressed, or 120 genomes compressed) at a time, though an unlimited amount of data can be submitted locally.
+The results of these analyses are immediately reported to the user following their completion, while all other non-selected analyses are subsequently completed in the background and stored in the database without interaction from the user.
+The public web service accepts uploads of up to 200 MB (approximately 50 \textit {E.
+coli} genomes uncompressed, or 120 genomes compressed) at a time, though an unlimited amount of data can be submitted to a local instance.
 
@@
 % para: the semantic web
-Semantic web technologies describe the relations between different data \cite{berners2001semantic}.
-For biological data, individual data points such as genome, contiguous DNA sequence, or gene, can be linked together in a searchable graph structure.
-This allows novel data to be incorporated into the existing graph, and the use of graph databases for semantic information has been proposed as a open standard for sharing public information \cite{horrocks2005semantic}.
-Data points are annotated with terms from published ontologies.
-To avoid proliferating ontologies, and to allow Spfy to integrate with existing ones, we use annotations from the GenEpiO \cite{griffiths2017context}, FALDO \cite{bolleman2016faldo}, and TypOn \cite{vaz2014typon}.
-
+Semantic web technologies describe the relationships between data \cite{berners2001semantic}, and graph databases for the storage of this information has been proposed as a open standard for sharing public information \cite{horrocks2005semantic}.
+For biological data, individual data points can be a genome, contiguous DNA sequence, or gene, and these are linked together in a searchable graph structure using existing ontologies, including annotations from GenEpiO \cite{griffiths2017context}, FALDO \cite{bolleman2016faldo}, and TypOn \cite{vaz2014typon}.
+This system is flexible and allows novel data to be incorporated into the existing graph.
 \begin{figure}[!hb]
@@
 The permanent storage of results is as a one-time cost to avoid recomputation when the same analysis is re-run.
-For population analyses, Spfy searches the graph for all data points annotated with the queried ontology term.
-This graph data is then converted into the required structure, usually numerical arrays, as required for the given analysis modules.
-In a graph database, search begins at any node or attribute.
+For analyses, Spfy searches the graph for all data points annotated with the queried ontology term.
+This graph data is then converted into the required structure, usually numerical arrays, or as required for the given analysis module.
+In a graph database, a search can begin at any node or attribute.
 This is in contrast to a SQL database which requires a predefined schema, or a NoSQL database which treats data as documents with varying structure.
 For example, the addition of a new analysis module would typically require a new table definition in a SQL database, or the addition of a new document type in a NoSQL database.
-A graph database instead adds new nodes or attributes which can be connected to existing data, so explicit joins or data conversions are not required.
-Spfy also adds the graph results in parts, and allows the database to infer their connections.
+With a graph database, new nodes or attributes are added and then connected to existing data, removing the need for explicit joins or data conversions.
+Additionally, data can be added to Spfy in parts, and the database will infer the correct connections between the data.
 
@@
 A sharable token is automatically created for users upon entering the website, and is embedded into the website address.
-Users can share results by copying their URL, and files submitted from different computers using the same token will be visible to anyone one the same link.
+Users can share results by copying their URL, and files submitted from different computers using the same token will be visible to anyone with the same link.
 
@@
 Docker \url{https://www.docker.com/} is a virtualization technology to simulate self-contained operating systems on the same host computer, without the overhead of full hardware virtualization \cite{felter2015updated}.
-The Spfy platform depends on a series of webservers, databases, and task workers and uses Docker to compartmentalize these services which are then networked together using Docker-Compose \url{https://docs.docker.com/compose/}.
+The Spfy platform depends on a series of webservers, databases, and task workers, and uses Docker to compartmentalize these services, which are then networked together using Docker-Compose \url{https://docs.docker.com/compose/}.
 (see Figure \ref{fig-docker})
@@
 % standard web tech means you can deploy to different cloud computing services
-One of the key benefits of using more common-place technologies is the compatibility with other infrastructure resources.
-Docker containers are widely supported by cloud computing services: Amazon Web Services (AWS) \url{https://aws.amazon.com/docker/}, Google Cloud Platform (GCloud) \url{https://cloud.google.com/container-engine/}, and Microsoft Azure \url{https://azure.microsoft.com/en-us/services/container-service/}, and self-hosted cloud computing technologies such as OpenStack \url{https://wiki.openstack.org/wiki/Docker} all support Docker.
+One of the key benefits of using common-place technologies is the compatibility with other infrastructure resources.
+Docker containers are widely supported by cloud computing services: Amazon Web Services (AWS) \url{https://aws.amazon.com/docker/}, Google Cloud Platform (GCloud) \url{https://cloud.google.com/container-engine/}, and Microsoft Azure \url{https://azure.microsoft.com/en-us/services/container-service/}, and self-hosted cloud computing technologies such as OpenStack \url{https://wiki.openstack.org/wiki/Docker}.
 % we may want to sign up for a free trial to give an example of this
@@
 The individual tests use PyTest \url{https://doc.pytest.org/}, and the current build status can be checked either on our GitHub repository or at \url{https://travis-ci.org/superphy/backend}.
-TravisCI also builds Spfy's core Docker images, and uploads them to Docker Hub \url{https://hub.docker.com/u/superphy/}.
+TravisCI also builds the core Docker images for Spfy, and uploads them to Docker Hub \url{https://hub.docker.com/u/superphy/}.
 

### 2018-02-22 paper-webserver.tex

@@
 \section{RESULTS}
-Spfy was tested with 10,243 public \textit{E. coli} assembled genomes from Enterobase, storing every sequence and the results for all included analysis modules.
-All subtyping options: O-antigen, H-antigen, Shiga-toxin 1, Shiga-toxin 2, and Intimin typing, VF and AMR annotation were ran on the test set.
-The genomes were also analyzed within the pan-genome framework of \textit{E.
+Spfy was tested with 10,243 public \textit{E.
+coli} assembled genomes from Enterobase, storing every sequence and the results for all included analysis modules.
+This included: serotypeing (O-antigen, H-antigen), toxin sub-typing (Shiga-toxin 1, Shiga-toxin 2, and Intimin), the identification of VF and AMR determinants, and determination of pan-genome content \textit{E.
 coli}.
@@
 coli} genomes with the \textit{.fasta} or \textit{.fna} extensions.
-Submissions are checked against a reference set of \textit{E.
-coli} gene sequences before running analyses.
+Submissions are subjected to quality control, ensuring the submitted genomes are \textit{E.
+coli} sequences before subsequent analyses are run.
 \par
@@
 However, many subsets of biology now require the analyses of big-data, where inputs are taken from a variety of analysis programs, and involve large-scale data warehousing \cite{schatz2015biological}.
-The ability to perform computations in real-time, store data in flexible databases, and utilize a common application programming interface (API) linking resources are required for these analyses \cite{swaminathan2016review}.
+The ability to perform computations in real-time, store data in flexible databases, and utilize a common application programming interface (API) linking resources are now required for types of analyses that need to be performed \cite{swaminathan2016review}.
 
-One of the key goals in developing Spfy was to accommodate a variety of result formats, from storing subtyping data to retrieving results as inputs for downstream analyses, such as statistical significance testing.
-We've shown how a graph database can accommodate a variety of programs used by bioinformaticians and is performant for data retrieval on the results of 10,000+ genomes.
-Spfy maintains instantaneity, as modern analytics platforms must respond to user-requested analyses of big-data in the same efficiency as old analyses of single files.
+One of the key goals in developing Spfy was to accommodate and store a variety of result formats, and then to make the data from these results retrievable and usable as inputs for downstream analyses, such as predictive biomarker discovery.
+We have shown how a graph database can accommodate the results given by a variety of bioinformatics programs, and how Spfy is performant for data retrieval on the results for multiple analyses among over 10,000 genomes, providing results from big-data comparisons in the same efficiency as old analyses on single files.
 
@@
 The isolation and characterization of bacterial pathogens are critical for Public Health laboratories to rapidly respond to outbreaks, and to effectively monitor known and emerging pathogens through surveillance programs.
-Until recently, public-health agencies relied on laboratory tests such as serotyping, pulsed-field gel electrophoresis (PFGE) banding patterns, or Polymerase Chain Reaction (PCR)-based amplification to identify known VFs or AMR genes, along with other tests \cite{ronholm2016navigating}, to characterize bacterial isolates in outbreak and surveillance settings.
+Until recently, public-health agencies relied on laboratory tests such as serotyping, pulsed-field gel electrophoresis (PFGE), PCR-based amplification of known VFs, and disc-diffusion assays to identify AMR, for the characterization of bacterial isolates in outbreak, surveillance, and reference laboratory settings \cite{ronholm2016navigating}.
 
-AMR testing, VF testing, and Shiga-toxin testing are of particular importance to surveillance efforts due to their role in a pathogen's lethality.
-Current efforts are focused on predictive genomics, where the relevant phenotypic information can be determined through examination of the whole-genome sequence.
-WGS methods allow a single laboratory method - the sequencing step - to provide a variety of analysis options, such as AMR, VF, and Shiga-toxin results, and as such can be used to evaluate the spread of outbreaks with better resolution and context than traditional methods \cite{ronholm2016navigating}.
-
-% application: results similar to a wet-lab
-Spfy uses WGS results.
-After initial sequencing of new isolates, Spfy can be used in place of a traditional reference laboratory, to determine the O-type and H-type, Stx type, and all known VFs and AMR genes in real-time.
-These results can be shared with other agencies and researchers over the internet.
-Furthermore, using Spfy's database of pre-processed genomes, Spfy can determine all strains a sample may be related to which is useful for monitoring the evolution of pathogens over time.
-A computational approach saves the time and cost associated with performing multiple tests per sample, and allows population comparisons to keep pace the current generation WGS data.
+Current efforts are focused on predictive genomics, where the relevant phenotypic information can be determined through examination of the whole-genome sequence without need for the traditional laboratory test.
+Spfy provides rapid and easy predictive genomic analyses of \textit{E.
+coli} genomes.
 
@@
 % namely galaxy
-Existing scientific workflow technologies such as Galaxy \cite{goecks2010galaxy}, and pipelines such as the Integrated Rapid Infectious Disease Analysis (IRIDA) platform \url{http://www.irida.ca/} and the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial} help automate the use of WGS data.
-Reproducibility is important, and Galaxy aims to provide a reproducible, computation-based research environment which is accessible to individuals without programming knowledge.
+The automated analyses of WGS is currently facilitated by existing scientific workflow technologies such as Galaxy \cite{goecks2010galaxy}, the Integrated Rapid Infectious Disease Analysis (IRIDA) platform \url{http://www.irida.ca/}, and the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial}.
+Galaxy aims to provide a reproducible, computation-based research environment which is accessible to individuals without programming knowledge.
 Galaxy defines a formal schema for linking different analysis software together, so the entire analysis pipeline can be replicated.
-Building on top of the Galaxy framework, IRIDA adds prebuilt pipelines specific to bioinformatics uses, as well as sequence and result storage.
-IRIDA takes a project-based approach, with sequences stored per project, and results stored linearly per whole-genome sequence.
+IRIDA is built on top of Galaxy and adds prebuilt pipelines for specific analyses, as well as sequence and result storage.
+IRIDA takes a project-based approach, with sequences stored per project, and results stored per project.
 Similarly, BAP provides an integrated analysis pipeline for bacterial WGS data as a web service.
-For result storage, these workflow technologies either use relational tables \cite{goecks2010galaxy}, or store resulting files to disk \cite{thomsen2016bacterial}.
+These workflow technologies store their results using relational tables \cite{goecks2010galaxy}, or as files on disk \cite{thomsen2016bacterial}.
 
-Like IRIDA and BAP, Spfy automates workflows for users, and like Galaxy, Spfy uses task queues to distribute selected analysis.
+Spfy is similar to these technologies in that it automates workflows for users, and like Galaxy, uses task queues to distribute selected analysis.
 % re: Matt "2. When comparing to IRIDA, BAP etc., can mention some differences with Superphy, e.g. the storage of interim result data that allows downstream integrated analysis"
@@
 Spfy processes XXXX files over XX tasks in XXXX time, distributing computations over a task queue and multiple Docker compartmentalized containers.
-We also focus on the data warehousing problem in bioinformatics: storing results as in IRIDA and BAP, but using a graph database to integrate results for further analyses.
-Like BAP, Spfy integrates different programs, such as for VF and AMR determination, into one pipeline. \par
-
-Because output from these programs is user-specific or transitory, results from identical comparisons are often recomputed.
-Additionally, output from different analyses are structured using distinct terminology and formats, which must be converted before they can be compared.
-Without a unified structure, these conversions quickly become impractical for broad usage.
-Graph-based storage of all results solves these problems and allows Spfy to perform population-wide analyses, regardless of the individual analysis software used to generate the datasets.
+However, unlike these workflow managers, Spfy is designed to help solve the needless recomputation of analyses by storing results in a graph database for future use.
+This allows Spfy to perform population-wide analyses, regardless of the individual analysis software used to generate the results for an individual genome.
 
@@
 
-Future work will focus on adding additional analyses modules to aid genotype to phenotype predictions, and supporting different species.
-We're currently working on an approach using support vector machines to predict \textit{E.coli} phenotypes \url{https://github.com/superphy/kmer} which will be integrated into Spfy.
+Future work will focus on adding additional analyses modules to aid genotype to phenotype predictions using machine learning modules, and supporting bacterial species such as \textit{Salmonella}, and \textit{Campylobater}.
 While the integrated approach of storing and retrieving results provides enormous benefits, the developed analyses modules are self-contained and be transfered to existing platforms such as Galaxy, and IRIDA.
-Spfy's code is hosted at \url{https://github.com/superphy/backend}, and is available for free under the open-source Apache 2.0 license.
+The source code for Spfy is hosted at \url{https://github.com/superphy/backend}, and is available for free under the open-source Apache 2.0 license.
 A developer guide is provided at \url{https://superphy.readthedocs.io/en/latest/}.

### 2018-02-22 paper-webserver.tex
ADD: matt's edits to cl-feb22 branch

@@
 
+\usepackage{xcolor}
+\newcommand\mwcomment[1]{\textcolor{red}{#1}}
+
 \begin{document}
 
-\title{Spfy: an integrated graph-based sequence database for real-time prediction of Escherichia coli phenotypes using the SuperPhy platform}
+\mwcomment{I would suggest we avoid the comparison to sequence databases or genome sequence platforms (e.g.
+assembly, gene annotation etc).
+We really don't have the capacity to replace EnteroBase, Genbank, IRIDA, etc.
+I see Spfy as a repository of E.coli phenotypes important for public health.
+It allows users to rapidly predict phenotypes and retrieve the phenotypes of historical strains to: track trends, observe context amoung evolutionary related strains, identify biomarkers, etc.
+In other words, Spfy provides WGS-based phenotypes + big picture comparative genomics.
+The big picture is only possible by 1.
+recomputing everything everytime, or 2.
+storing the results as genomes are analysed.
+With size of WGS data, option 2 is only viable option}
+\mwcomment{Title may need work. Should it mention comparative genomics? Can't be too long thou}
+\title{Spfy: an integrated graph database and analysis platform for real-time prediction of Escherichia coli phenotypes}
 
@@
 
+Public health laboratories are currently moving to whole-genome sequencing based analyses, and require rapid prediction of relevent pathogen phenotypes impacting health.
 Current workflows in comparative computational genomics rely on chaining different analysis software together, but lack storage and retrieval methods for the generated results.
-Reference laboratories are currently moving to whole-genome sequencing based analyses, and require rapid analyses coupled to efficient storage and retrieval of data.
+\mwcomment{Reference laboratories do not care about data retrieval, only real-time results.
+Researchers care more about context and thus would use a historical record of phenotypes for all predicted strains}
 To solve this problem, we have created Spfy, which uses a graph database to store and retrieve results from computational workflows.
-This facilitates rapid analyses, as well as the efficient management and downstream processing of tens of thousands of whole-genome sequences.
+The newly developed Spfy platform facilitates rapid phenotype identification, as well as the efficient storage and downstream comparative anlaysis of tens of thousands of genome sequences.
 Though generally applicable to bacterial genome sequences, Spfy currently contains X \textit{Escherichia coli} genomes, for which \textit{in-silico} serotype and Shiga-toxin subtype, as well as the presence of known virulence factors and antimicrobial resistance determinants have been computed.
@@
 Integrated data storage and analyses are currently necessary as the number of publicly available whole genome sequences is currently in the hundreds of thousands, with millions likely to be available within the next few years.
+\mwcomment{should we mention how many ecoli genomes are available? It might prompt reviewers to ask why we don't have them in spfy}
 \par
@@
 The rapid characterization and comparison of bacterial pathogens relies principally on the combination of outputs from multiple software programs that are targeted for specific applications.
-Examples include the identification of known antimircrobial (AMR) resistance genes, through software such as the Resistance Gene Identifier (RGI) \cite{mcarthur2013comprehensive}, {ResFinder}, {ARG-ANNOT}, and {ARIBA}; the identification of known virulence genes through software such as VirulenceFinder \cite{kleinheinz2014applying}, {SRST2}, and {an additional reference}; the annotation of bacterial genomes, through software such as Prokka \cite{doi:10.1093/bioinformatics/btu153}, {another}, and {another}.
+Examples include the identification of known antimircrobial (AMR) resistance genes, through software such as the Resistance Gene Identifier (RGI) \cite{mcarthur2013comprehensive}, {ResFinder}, {ARG-ANNOT}, and {ARIBA}; or the identification of known virulence genes through software such as VirulenceFinder \cite{kleinheinz2014applying}, {SRST2}, and {an additional reference}.
 \todo{Add a section on serotyping and subtyping a la Phylotyper}.
@@
 Comprehensive platforms that combine individual programs into a cohesive whole also exist.
-These include free platforms such as the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial}, the Integrated Rapid Infectious Disease Analysis (IRIDA) project \url{http://www.irida.ca/}, and the Pathosystems Resource Integration Center (PATRIC) \cite{wattam2013patric}.
+These include free platforms such as the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial}, and the Pathosystems Resource Integration Center (PATRIC) \cite{wattam2013patric}.
 Commercial applications, such as Bionumerics, which is used by PulseNet International for the analyses of WGS data in outbreak situations also exist, and offer support as well as accredited, standardized tests \cite{swaminathan2001pulsenet}.
+These platforms are designed to be applied to individual projects \cite{}.
 
 % 3. 100,000 genomes, (lookup number for Enterobase, GenBank), how do we run analysis on it
-Traditionally, these platforms have been applied to samples numbering only a few dozen \cite{} while WGS of bacterial pathogens have recently accumulated in public databases in the hundreds of thousands, with millions set to be available within the next few years.
+WGS of bacterial pathogens have recently accumulated in public databases in the hundreds of thousands, with millions set to be available within the next few years.
 For \textit{Escherichia coli} alone, there are over sixty thousand publicly available genomes in EnteroBase \url{https://enterobase.warwick.ac.uk/} and three million sequenced genomes in GenBank \cite{doi:10.1093/nar/gks1195}.
-
 Many of the comparative analyses that are currently used in the analyses of bacterial genomes are broadly useful, and therefore computed multiple times for the same genomes.
@@
 Such measures can help ensure the rapid response times required for public health applications, and allow results to be integrated and progressively updated as new data becomes available.
+\mwcomment{should we mention how many ecoli genomes are available? It might prompt reviewers to ask why we don't have them in spfy}
 
@@
 
+\mwcomment{I think this discussion sets the perfect tone for the big data applications provided by spfy.
+My only suggestion is to give examples of the types analysis that are mentioned in the review.
+One idea, instead of dropping "Impact on Public Heath Efforts" entirely, you can convert it into a use case example that highlights the benefit spfy provides.}
+
 \subsection{Impact on Public Health Efforts}
@@
 % namely galaxy
-The automated analyses of WGS is currently facilitated by existing scientific workflow technologies such as Galaxy \cite{goecks2010galaxy}, the Integrated Rapid Infectious Disease Analysis (IRIDA) platform \url{http://www.irida.ca/}, and the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial}.
-Galaxy aims to provide a reproducible, computation-based research environment which is accessible to individuals without programming knowledge.
-Galaxy defines a formal schema for linking different analysis software together, so the entire analysis pipeline can be replicated.
-IRIDA is built on top of Galaxy and adds prebuilt pipelines for specific analyses, as well as sequence and result storage.
-IRIDA takes a project-based approach, with sequences stored per project, and results stored per project.
-Similarly, BAP provides an integrated analysis pipeline for bacterial WGS data as a web service.
-These workflow technologies store their results using relational tables \cite{goecks2010galaxy}, or as files on disk \cite{thomsen2016bacterial}.
+The automated analyses of WGS is currently facilitated by existing scientific workflow technologies such as Galaxy \cite{goecks2010galaxy}.
+Galaxy aims to provide a reproducible, computational inferface which is accessible to individuals without programming knowledge.
+Galaxy defines a formal schema for linking different analysis software together, so the entire analysis pipeline can be replicated and also extended as new analysis tools are developed.
+The Galaxy worklow focus is on running an individual analysis pipeline.
+In does not include functionality to store and collate analysis results for large-scale comparative analysis.
+
+The Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial} provides an integrated analysis pipeline for bacterial WGS data as a web service.
+It provides an individual per-genome report of \mwcomment{list the features BAP provides}.
+\cite{thomsen2016bacterial}.
 
@@
 Spfy processes XXXX files over XX tasks in XXXX time, distributing computations over a task queue and multiple Docker compartmentalized containers.
-However, unlike these workflow managers, Spfy is designed to help solve the needless recomputation of analyses by storing results in a graph database for future use.
+However, unlike these workflow managers, Spfy is designed to help solve the needless recomputation of analyses by storing results in a graph database for downstream comparative analysis.
 This allows Spfy to perform population-wide analyses, regardless of the individual analysis software used to generate the results for an individual genome.
+
+\mwcomment{Maybe should include a comparison to PATRIC.
+Patric does offer some comparative genomic analysis, however they are not health focused (pathways, gene presence/absence etc), and you cannnot add your own data to the repository}
 
@@
 Future work will focus on adding additional analyses modules to aid genotype to phenotype predictions using machine learning modules, and supporting bacterial species such as \textit{Salmonella}, and \textit{Campylobater}.
-While the integrated approach of storing and retrieving results provides enormous benefits, the developed analyses modules are self-contained and be transfered to existing platforms such as Galaxy, and IRIDA.
+While the integrated approach of storing and retrieving results provides enormous benefits, the developed analyses modules are self-contained and be transfered to existing platforms such as Galaxy.
 The source code for Spfy is hosted at \url{https://github.com/superphy/backend}, and is available for free under the open-source Apache 2.0 license.

### 2018-02-28 paper-webserver.tex
ADD: a la phylo

@@
 The rapid characterization and comparison of bacterial pathogens relies principally on the combination of outputs from multiple software programs that are targeted for specific applications.
-Examples include the identification of known antimircrobial (AMR) resistance genes, through software such as the Resistance Gene Identifier (RGI) \cite{mcarthur2013comprehensive}, {ResFinder}, {ARG-ANNOT}, and {ARIBA}; or the identification of known virulence genes through software such as VirulenceFinder \cite{kleinheinz2014applying}, {SRST2}, and {an additional reference}.
-\todo{Add a section on serotyping and subtyping a la Phylotyper}.
+Examples include the identification of known antimircrobial resistance (AMR) genes, through software such as the Resistance Gene Identifier (RGI) \cite{mcarthur2013comprehensive}, \cite{kleinheinz2014applying}, \cite{gupta2014arg}, and ARIBA \cite{hunt2017ariba}; or the identification of known virulence factor genes (VF) through software such as VirulenceFinder \cite{kleinheinz2014applying}, SRST2 \cite{inouye2014srst2}, and GeneSippr \cite{lambert2015genesippr}.
+For clinical diagnoses and comparisons, individual species can be first divided into subtypes with complementing AMR and VF results.
+Software methods for subtyping rely on intraspecies genes or genomic regions, and are targeted through software such as Phylotyper \cite{whiteside2017phylotyper}, SerotypeFinder \cite{joensen2015rapid}, the EcOH dataset applied through SRST2 \cite{ingle2016silico}, and V-Typer \cite{carrillo2016comparative}.
 These methods represent \textit{in-silico} analogues of traditional wet-lab tests, which allows new WGS results to be viewed in the context of historical tests, and greatly expedites the analyses of newly sequenced genomes.

### 2018-02-28 paper-webserver.tex
Added captions to the figures

@@
 \mwcomment{Title may need work. Should it mention comparative genomics? Can't be too long thou}
-\title{Spfy: an integrated graph database and analysis platform for real-time prediction of Escherichia coli phenotypes}
+\title{Spfy: an integrated graph database for real-time prediction of Escherichia coli phenotypes and downstream comparative genomics}
 
@@
 Spfy handles all of the analyses tasks by dividing them into subtasks, which are subsequently distributed across a built-in task queue.
-Results are converted into individual graphs and stored within a larger graph database according to standard ontologies {list}, where metadata including genotypes, biomarkers, host, source, and statistical significance testing of genome markers for user-defined groups are stored.
+Results are converted into individual graphs and stored within a larger graph database according to the standard ontologies GenEpiO \cite{griffiths2017context}, FALDO \cite{bolleman2016faldo}, and TypOn \cite{vaz2014typon}, where metadata including genotypes, biomarkers, host, source, and statistical significance testing of genome markers for user-defined groups are stored.
 
@@
 Semantic web technologies describe the relationships between data \cite{berners2001semantic}, and graph databases for the storage of this information has been proposed as a open standard for sharing public information \cite{horrocks2005semantic}.
-For biological data, individual data points can be a genome, contiguous DNA sequence, or gene, and these are linked together in a searchable graph structure using existing ontologies, including annotations from GenEpiO \cite{griffiths2017context}, FALDO \cite{bolleman2016faldo}, and TypOn \cite{vaz2014typon}.
+For biological data, individual data points can be a genome, contiguous DNA sequence, or gene, and these are linked together in a searchable graph structure using existing ontologies, including annotations from given ontologies.
 This system is flexible and allows novel data to be incorporated into the existing graph.
@@
 \end{center}
-\caption{An example of how data is stored in Spfy.
+\caption{Structure of Spfy's graph database.
 Brackets highlight the source of different data points and the software it was generated from.
-These parts are added in as the analyis modules complete, at varying times, and the overall connections are inferred by the database.}
+These parts are added in as the analyis modules complete, at varying times, and the overall connections are inferred by the database.
+Non-bracketed sections are sourced from the uploaded genome files or user-supplied metadata.}
 \label{fig-ontology}
@@
 \end{center}
-\caption{Caption for figure within column.}
+\caption{First entry into the results interface for submitted tasks. Cards represent individual tasks and the entire collection can be viewed by sharing the URL with the embedded token.}
 \label{fig-results}
@@
 \end{center}
-\caption{Caption for figure within column.}
+\caption{An expanded result for a subtyping task.
+While data storage in Spfy is graph-based, a familiar tabular structure is presented to users.
+The result can also be shared using the URL with the embedded token.}
 \label{fig-tables}
@@
 \end{center}
-\caption{Caption for figure within column.}
+\caption{The various Docker containers used in Spfy.
+Arrows represent the connections between different containers, and the entire platform can be recreated with a single command using Docker-Compose.}
 \label{fig-docker}
@@
 The Galaxy worklow focus is on running an individual analysis pipeline.
-In does not include functionality to store and collate analysis results for large-scale comparative analysis.
+It does not include functionality to store and collate analysis results for large-scale comparative analysis.
 

### 2018-02-28 paper-webserver.tex
Added how we address the pts in the review paper.

@@
 However, many subsets of biology now require the analyses of big-data, where inputs are taken from a variety of analysis programs, and involve large-scale data warehousing \cite{schatz2015biological}.
-The ability to perform computations in real-time, store data in flexible databases, and utilize a common application programming interface (API) linking resources are now required for types of analyses that need to be performed \cite{swaminathan2016review}.
+The ability to integrate data from different source technologies, merge submissions from other labs, and distribute computations over fault-tolerant systems are now required for types of analyses that need to be performed \cite{schatz2015biological}.
 
@@
 
-\subsection{Impact on Public Health Efforts}
+\subsection{Impact on Public Heath Efforts}
 
@@
 Until recently, public-health agencies relied on laboratory tests such as serotyping, pulsed-field gel electrophoresis (PFGE), PCR-based amplification of known VFs, and disc-diffusion assays to identify AMR, for the characterization of bacterial isolates in outbreak, surveillance, and reference laboratory settings \cite{ronholm2016navigating}.
+Current efforts are focused on predictive genomics, where the relevant phenotypic information can be determined through examination of the whole-genome sequence without need for the traditional laboratory test.
 
-Current efforts are focused on predictive genomics, where the relevant phenotypic information can be determined through examination of the whole-genome sequence without need for the traditional laboratory test.
 Spfy provides rapid and easy predictive genomic analyses of \textit{E.
-coli} genomes.
+coli} genomes while also addressing the problem of large scale comparisons.
+With the larger datasets involved in population genomics, it is no longer viable for individual researchers to download data to perform comparisons.
+Instead, efforts have focused on storing biological data online and enabling analyses on that data \cite{schatz2015biological}.
+By using a graph database, Spfy can integrate results from  different technologies including software predicted phenotypes as well as laboratory results and user-submitted metadata.
+In addition, datasets can be built and submitted from multiple labs for joint analysis.
 
@@
 The automated analyses of WGS is currently facilitated by existing scientific workflow technologies such as Galaxy \cite{goecks2010galaxy}.
-Galaxy aims to provide a reproducible, computational inferface which is accessible to individuals without programming knowledge.
+Galaxy aims to provide a reproducible, computational interface which is accessible to individuals without programming knowledge.
 Galaxy defines a formal schema for linking different analysis software together, so the entire analysis pipeline can be replicated and also extended as new analysis tools are developed.
@@
 The Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial} provides an integrated analysis pipeline for bacterial WGS data as a web service.
-It provides an individual per-genome report of \mwcomment{list the features BAP provides}.
-\cite{thomsen2016bacterial}.
+It provides an individual per-genome report of the determined species, multilocus sequence type, VF and AMR genes \cite{thomsen2016bacterial}.
 
-Spfy is similar to these technologies in that it automates workflows for users, and like Galaxy, uses task queues to distribute selected analysis.
+Spfy is similar to these technologies in that it automates workflows for users, and as in Galaxy, uses task queues to distribute selected analyses.
 % re: Matt "2. When comparing to IRIDA, BAP etc., can mention some differences with Superphy, e.g. the storage of interim result data that allows downstream integrated analysis"

### 2018-03-01 paper-webserver.tex
ADD: touching on patric

@@
 \mwcomment{Title may need work. Should it mention comparative genomics? Can't be too long thou}
-\title{Spfy: an integrated graph database for real-time prediction of Escherichia coli phenotypes and downstream comparative genomics}
+\title{Spfy: an integrated graph database for real-time prediction of Escherichia coli phenotypes and downstream comparative analyses}
 
@@
 Comprehensive platforms that combine individual programs into a cohesive whole also exist.
-These include free platforms such as the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial}, and the Pathosystems Resource Integration Center (PATRIC) \cite{wattam2013patric}.
+These include free platforms such as the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial}, and the Pathosystems Resource Integration Center (PATRIC) \cite{wattam2016improvements}.
 Commercial applications, such as Bionumerics, which is used by PulseNet International for the analyses of WGS data in outbreak situations also exist, and offer support as well as accredited, standardized tests \cite{swaminathan2001pulsenet}.
@@
 
+PATRIC \cite{wattam2016improvements} adds support for comparing user-submitted genomes and pre-analyzed public data, enabling comparisons of gene annotations for user-submitted genomes against reference genomes.
+PATRIC indexes a NoSQL document store to comparison similar document types.
+Unlike PATRIC which is targeted towards comparative genomics, Spfy's graph database enables broader generic comparisons against any population groupings.
+For example, Spfy can compare all genomes with a particular subtype for gene annotations, source metadata for gene annotations, or any two (or more) node or attribute data.
+
+\small \input{tables/comparison}
+
 \mwcomment{Maybe should include a comparison to PATRIC.

### 2018-03-01 tables/comparison.tex
ADD: touching on patric

% Please add the following required packages to your document preamble:
% \usepackage{graphicx}
\begin{table}[!hb]
\centering
\caption{Comparison of various bioinformatic pipelines and their underlying database. Functionally, Spfy integrates different analysis modules as in BAP while also merging large datasets as in PATRIC.}
\label{Table 2.}
\resizebox{\textwidth}{!}{%
\begin{tabular}{lllll}
         & Spfy                & Galaxy              & BAP                 & PATRIC              \\
Database & Blazegraph          & PostgreSQL          & MySQL + File System & MongoDB + Shock     \\
Type     & Graph               & SQL                 & SQL + File System   & NoSQL               \\
Focus    & Integrated Analyses & Workflow Technology & Batch Analysis      & Integrated Analyses
\end{tabular}%
}
\end{table}

### 2018-03-01 paper-webserver.tex
EDIT: how Spfys graph struct enables broader analyses

@@
 
-PATRIC \cite{wattam2016improvements} adds support for comparing user-submitted genomes and pre-analyzed public data, enabling comparisons of gene annotations for user-submitted genomes against reference genomes.
-PATRIC indexes a NoSQL document store to comparison similar document types.
-Unlike PATRIC which is targeted towards comparative genomics, Spfy's graph database enables broader generic comparisons against any population groupings.
-For example, Spfy can compare all genomes with a particular subtype for gene annotations, source metadata for gene annotations, or any two (or more) node or attribute data.
+\small \input{tables/comparison}
 
-\small \input{tables/comparison}
+PATRIC \cite{wattam2016improvements} adds support for comparing up to nine user-submitted genomes against a reference genome, targeting various gene annotations.
+PATRIC indexes a NoSQL document store to compare similar document types and is beneficial for comparative genomics.
+Unlike PARTIC, Spfy is targeted towards statistical comparisons of any population grouping, and Spfy's graph database has no limit on the number of genomes grouped for comparison.
+For example, Spfy can compare all genomes with a particular subtype for gene annotations, source metadata targets for gene annotations, or any two (or more) node or attribute types as shown in \ref{fig-groupings}.
+
+\begin{figure}[!hb]
+\begin{center}
+\includegraphics[width=\textwidth]{images/groupings}
+\end{center}
+\caption{Genomes can be grouped by any node or attribute type for comparisons, and Spfy has no upper bound for the number of genomes in a group.
+Different types can also be joined into a group through logical connectives AND, OR, Negation, and any data type is also a valid target.
+This approach can be used to compare any data regardless of source software.}
+\label{fig-groupings}
+\end{figure}
 

### 2018-03-01 paper-webserver.tex
caption

@@
 \caption{The various Docker containers used in Spfy.
-Arrows represent the connections between different containers, and the entire platform can be recreated with a single command using Docker-Compose.}
+Arrows represent the connections between different containers, and the entire platform can be recreated with a single command using its Docker-Compose definition.}
 \label{fig-docker}

### 2018-03-01 paper-webserver.tex
EDIT: readthrough

@@
 
-Public health laboratories are currently moving to whole-genome sequencing based analyses, and require rapid prediction of relevent pathogen phenotypes impacting health.
+Public health laboratories are currently moving to whole-genome sequencing based analyses, and require rapid prediction of relevant pathogen phenotypes impacting health.
 Current workflows in comparative computational genomics rely on chaining different analysis software together, but lack storage and retrieval methods for the generated results.
@@
 To solve this problem, we have created Spfy, which uses a graph database to store and retrieve results from computational workflows.
-The newly developed Spfy platform facilitates rapid phenotype identification, as well as the efficient storage and downstream comparative anlaysis of tens of thousands of genome sequences.
-Though generally applicable to bacterial genome sequences, Spfy currently contains X \textit{Escherichia coli} genomes, for which \textit{in-silico} serotype and Shiga-toxin subtype, as well as the presence of known virulence factors and antimicrobial resistance determinants have been computed.
+The newly developed Spfy platform facilitates rapid phenotype identification, as well as the efficient storage and downstream comparative analysis of tens of thousands of genome sequences.
+Though generally applicable to bacterial genome sequences, Spfy currently contains 10,243 \textit{Escherichia coli} genomes, for which \textit{in-silico} serotype and Shiga-toxin subtype, as well as the presence of known virulence factors and antimicrobial resistance determinants have been computed.
 Spfy links the results and metadata to the genome sequences through a standardized ontology, which facilitates hypothesis testing in fields ranging from population genomics to epidemiology, while mitigating the recomputing of analyses.
 The graph approach is flexible, and can accommodate new analysis software modules as they are developed, and easily link new results to those already stored.
-Integrated data storage and analyses are currently necessary as the number of publicly available whole genome sequences is currently in the hundreds of thousands, with millions likely to be available within the next few years.
-\mwcomment{should we mention how many ecoli genomes are available? It might prompt reviewers to ask why we don't have them in spfy}
+Integrated data storage and analyses are currently necessary to match the rapid accumulation of whole genome sequences in public databases.
 \par
@@
 The rapid characterization and comparison of bacterial pathogens relies principally on the combination of outputs from multiple software programs that are targeted for specific applications.
-Examples include the identification of known antimircrobial resistance (AMR) genes, through software such as the Resistance Gene Identifier (RGI) \cite{mcarthur2013comprehensive}, \cite{kleinheinz2014applying}, \cite{gupta2014arg}, and ARIBA \cite{hunt2017ariba}; or the identification of known virulence factor genes (VF) through software such as VirulenceFinder \cite{kleinheinz2014applying}, SRST2 \cite{inouye2014srst2}, and GeneSippr \cite{lambert2015genesippr}.
+Examples include the identification of known antimicrobial resistance (AMR) genes, through software such as the Resistance Gene Identifier (RGI) \cite{mcarthur2013comprehensive}, \cite{kleinheinz2014applying}, \cite{gupta2014arg}, and ARIBA \cite{hunt2017ariba}; or the identification of known virulence factor genes (VF) through software such as VirulenceFinder \cite{kleinheinz2014applying}, SRST2 \cite{inouye2014srst2}, and GeneSippr \cite{lambert2015genesippr}.
 For clinical diagnoses and comparisons, individual species can be first divided into subtypes with complementing AMR and VF results.
-Software methods for subtyping rely on intraspecies genes or genomic regions, and are targeted through software such as Phylotyper \cite{whiteside2017phylotyper}, SerotypeFinder \cite{joensen2015rapid}, the EcOH dataset applied through SRST2 \cite{ingle2016silico}, and V-Typer \cite{carrillo2016comparative}.
+Software methods for subtyping rely on pre-selected intraspecies genes or genomic regions, and are targeted through software such as Phylotyper \cite{whiteside2017phylotyper}, SerotypeFinder \cite{joensen2015rapid}, the EcOH dataset applied through SRST2 \cite{ingle2016silico}, and V-Typer \cite{carrillo2016comparative}.
 These methods represent \textit{in-silico} analogues of traditional wet-lab tests, which allows new WGS results to be viewed in the context of historical tests, and greatly expedites the analyses of newly sequenced genomes.
@@
 Commercial applications, such as Bionumerics, which is used by PulseNet International for the analyses of WGS data in outbreak situations also exist, and offer support as well as accredited, standardized tests \cite{swaminathan2001pulsenet}.
-These platforms are designed to be applied to individual projects \cite{}.
+These platforms are designed to be applied to individual projects.
 
-% 3. 100,000 genomes, (lookup number for Enterobase, GenBank), how do we run analysis on it
-WGS of bacterial pathogens have recently accumulated in public databases in the hundreds of thousands, with millions set to be available within the next few years.
-For \textit{Escherichia coli} alone, there are over sixty thousand publicly available genomes in EnteroBase \url{https://enterobase.warwick.ac.uk/} and three million sequenced genomes in GenBank \cite{doi:10.1093/nar/gks1195}.
-Many of the comparative analyses that are currently used in the analyses of bacterial genomes are broadly useful, and therefore computed multiple times for the same genomes.
+Many of the comparative analyses that are currently used in the study of bacterial genomes are broadly useful, and therefore computed multiple times for the same genomes.
 An effective method to mitigate the recomputing of analyses, is to make the storage and retrieval of results part of the analyses platform, and effectively linked to the genomes of interest through a standardized ontology.
 Such measures can help ensure the rapid response times required for public health applications, and allow results to be integrated and progressively updated as new data becomes available.
-\mwcomment{should we mention how many ecoli genomes are available? It might prompt reviewers to ask why we don't have them in spfy}
 
@@
 % 5. how we solved it
-In this study, we present the Spfy update to the SuperPhy platform, which integrates a graph database with real-time analyses; this integration avoids recomputing identical analyses.
-Graph-based result storage also allows retrospective comparisons as more genomes are sequenced or populations change, and is flexible, accommodating new analysis modules as they are developed.
+In this article, we present a new platform merging Superphy's pre-computed results with a novel data storage and processing architecture, which we call Spfy; Spfy integrates a graph database with real-time analyses to avoid recomputing identical runs.
+Graph-based result storage also allows retrospective comparisons across stored results as more genomes are sequenced or populations change, and is flexible, accommodating new analysis modules as they are developed.
 The database is available at \url{https://lfz.corefacility.ca/superphy/spfy/}.
@@
 coli}.
-It supports the following \textit{in-silico} subtyping options: serotyping, through both O- and H-antigen identification {ectyper ref}; Shiga-toxin 1 (Stx1), Shiga-toxin 2 (Stx2), and Intimin typing using Phylotyper \cite{whiteside2017phylotyper}, VF gene determination using ECTyper \url{https://github.com/phac-nml/ecoli\_serotyping}, and AMR annotation using the RGI program \cite{mcarthur2013comprehensive}.
-
-Spfy also performs pangenome analyses using Panseq \cite{laing2010pan}, and provides machine learning modules for biomarker discovery among groups using Scikit-learn \cite{pedregosa2011scikit}.
+It supports the following \textit{in-silico} subtyping options: serotyping, through both O- and H-antigen identification using ECTyper \url{https://github.com/phac-nml/ecoli_serotyping} as well as VF gene determination; Shiga-toxin 1 (Stx1), Shiga-toxin 2 (Stx2), and Intimin typing using Phylotyper \cite{whiteside2017phylotyper}, and AMR annotation using the RGI program \cite{mcarthur2013comprehensive}.
+Spfy also performs pangenome analyses using Panseq \cite{laing2010pan}, with the entire pangenome stored in the database and associated with source genomes.
 
@@
 This means that genomes can be compared on the basis of the presence or absence of pan-genome regions, serotype, subtyping data, or provided metadata such as location or host-source.
+All results are displayed to users in real-time, usually within 2-3 minutes.
 
@@
 
-i) The upload begins through the website, where user-defined analyses options are selected.
+i) The upload begins through the website, where user-defined analysis options are selected.
 The results of these analyses are immediately reported to the user following their completion, while all other non-selected analyses are subsequently completed in the background and stored in the database without interaction from the user.
@@
 
-iv) Python functions parse the results and permanently store them in Blazegraph \url{https://www.blazegraph.com/}, the graph database used for Superphy.
+iv) Python functions parse the results and permanently store them in Blazegraph \url{https://www.blazegraph.com/}, the graph database used for Spfy.
 
@@
 Semantic web technologies describe the relationships between data \cite{berners2001semantic}, and graph databases for the storage of this information has been proposed as a open standard for sharing public information \cite{horrocks2005semantic}.
-For biological data, individual data points can be a genome, contiguous DNA sequence, or gene, and these are linked together in a searchable graph structure using existing ontologies, including annotations from given ontologies.
+For biological data, individual data points can be a genome, contiguous DNA sequence, or gene, and these are linked together in a searchable graph structure using existing ontologies.
 This system is flexible and allows novel data to be incorporated into the existing graph.
@@
 Brackets highlight the source of different data points and the software it was generated from.
-These parts are added in as the analyis modules complete, at varying times, and the overall connections are inferred by the database.
+These parts are added in as the analysis modules complete, at varying times, and the overall connections are inferred by the database.
 Non-bracketed sections are sourced from the uploaded genome files or user-supplied metadata.}
@@
 coli} assembled genomes from Enterobase, storing every sequence and the results for all included analysis modules.
-This included: serotypeing (O-antigen, H-antigen), toxin sub-typing (Shiga-toxin 1, Shiga-toxin 2, and Intimin), the identification of VF and AMR determinants, and determination of pan-genome content \textit{E.
+This included: serotyping (O-antigen, H-antigen), toxin sub-typing (Shiga-toxin 1, Shiga-toxin 2, and Intimin), the identification of VF and AMR determinants, and determination of the pan-genome content of \textit{E.
 coli}.
-The resulting database had 17,820 nodes and 3,811,473 leaves, with 1,125,909,074 object properties.
+The resulting database has 17,820 nodes and 3,811,473 leaves, with 1,125,909,074 object properties.
 Spfy has been up since May 2017.
@@
 However, many subsets of biology now require the analyses of big-data, where inputs are taken from a variety of analysis programs, and involve large-scale data warehousing \cite{schatz2015biological}.
-The ability to integrate data from different source technologies, merge submissions from other labs, and distribute computations over fault-tolerant systems are now required for types of analyses that need to be performed \cite{schatz2015biological}.
+The ability to integrate data from different source technologies, merge submissions from other labs, and distribute computations over fault-tolerant systems are now required for the types of analyses that need to be performed \cite{schatz2015biological}.
 
 One of the key goals in developing Spfy was to accommodate and store a variety of result formats, and then to make the data from these results retrievable and usable as inputs for downstream analyses, such as predictive biomarker discovery.
-We have shown how a graph database can accommodate the results given by a variety of bioinformatics programs, and how Spfy is performant for data retrieval on the results for multiple analyses among over 10,000 genomes, providing results from big-data comparisons in the same efficiency as old analyses on single files.
+We have shown how a graph database can accommodate the results from a variety of bioinformatics programs, and how Spfy is performant for data retrieval on the results for multiple analyses among over 10,000 genomes, providing results from big-data comparisons in the same efficiency as old analyses on single files.
 
-\mwcomment{I think this discussion sets the perfect tone for the big data applications provided by spfy.
+% \mwcomment{I think this discussion sets the perfect tone for the big data applications provided by spfy.
 My only suggestion is to give examples of the types analysis that are mentioned in the review.
@@
 Until recently, public-health agencies relied on laboratory tests such as serotyping, pulsed-field gel electrophoresis (PFGE), PCR-based amplification of known VFs, and disc-diffusion assays to identify AMR, for the characterization of bacterial isolates in outbreak, surveillance, and reference laboratory settings \cite{ronholm2016navigating}.
-Current efforts are focused on predictive genomics, where the relevant phenotypic information can be determined through examination of the whole-genome sequence without need for the traditional laboratory test.
+Current efforts are focused on predictive genomics, where the relevant phenotypic information can be determined through examination of the whole-genome sequence without need for the traditional laboratory tests.
 
@@
 By using a graph database, Spfy can integrate results from  different technologies including software predicted phenotypes as well as laboratory results and user-submitted metadata.
-In addition, datasets can be built and submitted from multiple labs for joint analysis.
+In addition, datasets can be built and submitted from multiple labs for joint analyses.
 
@@
 Galaxy aims to provide a reproducible, computational interface which is accessible to individuals without programming knowledge.
-Galaxy defines a formal schema for linking different analysis software together, so the entire analysis pipeline can be replicated and also extended as new analysis tools are developed.
-The Galaxy worklow focus is on running an individual analysis pipeline.
-It does not include functionality to store and collate analysis results for large-scale comparative analysis.
+Galaxy defines a formal schema for linking different analysis software together, so the entire pipeline can be replicated and also extended as new tools are developed.
+The Galaxy workflow focuses on running an individual analysis pipeline.
+It does not include functionality to store and collate analysis results for large-scale comparative studies.
 
@@
 % re: Matt "2. When comparing to IRIDA, BAP etc., can mention some differences with Superphy, e.g. the storage of interim result data that allows downstream integrated analysis"
-On a per file basis, Spfy performs at a similar speed to BAP on predictive genomics tasks, though Spfy does not provide genome assembly services.
-Spfy processes XXXX files over XX tasks in XXXX time, distributing computations over a task queue and multiple Docker compartmentalized containers.
-However, unlike these workflow managers, Spfy is designed to help solve the needless recomputation of analyses by storing results in a graph database for downstream comparative analysis.
-This allows Spfy to perform population-wide analyses, regardless of the individual analysis software used to generate the results for an individual genome.
+On a per file basis, Spfy performs at a similar speed to BAP on predictive genomics tasks, usually completing within 2-3 minutes, though Spfy does not provide genome assembly services.
+However, unlike these workflow managers, Spfy is designed to help solve the needless recomputation of analyses by storing results in a graph database for downstream comparative studies.
+This allows Spfy to perform population-wide analyses, regardless of the individual software used to generate the results for a given genome.
 
@@
 
-PATRIC \cite{wattam2016improvements} adds support for comparing up to nine user-submitted genomes against a reference genome, targeting various gene annotations.
-PATRIC indexes a NoSQL document store to compare similar document types and is beneficial for comparative genomics.
-Unlike PARTIC, Spfy is targeted towards statistical comparisons of any population grouping, and Spfy's graph database has no limit on the number of genomes grouped for comparison.
+PATRIC \cite{wattam2016improvements} and Spfy share the same goal of integrated analyses.
+PATRIC adds support for comparing up to nine user-submitted genomes against a reference genome, targeting various gene annotations.
+The platform indexes a NoSQL document store to compare similar document types and is beneficial for comparative genomics.
+Unlike PARTIC, Spfy is targeted towards statistical comparisons of any permutation of a population group based on the chosen data types, and Spfy's graph database has no limit on the number of genomes grouped for comparison.
 For example, Spfy can compare all genomes with a particular subtype for gene annotations, source metadata targets for gene annotations, or any two (or more) node or attribute types as shown in \ref{fig-groupings}.
@@
 
-\mwcomment{Maybe should include a comparison to PATRIC.
+% \mwcomment{Maybe should include a comparison to PATRIC.
 Patric does offer some comparative genomic analysis, however they are not health focused (pathways, gene presence/absence etc), and you cannnot add your own data to the repository}
@@
 
-Future work will focus on adding additional analyses modules to aid genotype to phenotype predictions using machine learning modules, and supporting bacterial species such as \textit{Salmonella}, and \textit{Campylobater}.
+Future work will focus on adding additional analysis modules to aid genotype to phenotype predictions using machine learning modules, and supporting bacterial species such as \textit{Salmonella}, and \textit{Campylobater}.
 While the integrated approach of storing and retrieving results provides enormous benefits, the developed analyses modules are self-contained and be transfered to existing platforms such as Galaxy.

### 2018-03-02 paper-webserver.tex
FIX: a couple spelling mistakes (e.g. silco) REMOVE: mwcomments

@@
 
-\mwcomment{I would suggest we avoid the comparison to sequence databases or genome sequence platforms (e.g.
-assembly, gene annotation etc).
-We really don't have the capacity to replace EnteroBase, Genbank, IRIDA, etc.
-I see Spfy as a repository of E.coli phenotypes important for public health.
-It allows users to rapidly predict phenotypes and retrieve the phenotypes of historical strains to: track trends, observe context amoung evolutionary related strains, identify biomarkers, etc.
-In other words, Spfy provides WGS-based phenotypes + big picture comparative genomics.
-The big picture is only possible by 1.
-recomputing everything everytime, or 2.
-storing the results as genomes are analysed.
-With size of WGS data, option 2 is only viable option}
-\mwcomment{Title may need work. Should it mention comparative genomics? Can't be too long thou}
 \title{Spfy: an integrated graph database for real-time prediction of Escherichia coli phenotypes and downstream comparative analyses}
@@
 Current workflows in comparative computational genomics rely on chaining different analysis software together, but lack storage and retrieval methods for the generated results.
-\mwcomment{Reference laboratories do not care about data retrieval, only real-time results.
-Researchers care more about context and thus would use a historical record of phenotypes for all predicted strains}
 To solve this problem, we have created Spfy, which uses a graph database to store and retrieve results from computational workflows.
@@
 
-Spfy provides rapid in-silco versions of common reference laboratory tests for the analyses of \textit{E.
+Spfy provides rapid in-silico versions of common reference laboratory tests for the analyses of \textit{E.
 coli}.
@@
 
-% \mwcomment{I think this discussion sets the perfect tone for the big data applications provided by spfy.
-My only suggestion is to give examples of the types analysis that are mentioned in the review.
-One idea, instead of dropping "Impact on Public Heath Efforts" entirely, you can convert it into a use case example that highlights the benefit spfy provides.}
-
 \subsection{Impact on Public Heath Efforts}
@@
 
-\subsection{Comparison with other bioinformatic pipeline technologies}
+\subsection{Comparison with other bioinformatics pipeline technologies}
 % trying to cut down on details that would be better suited for the "Implemntation" section.
@@
 
-% \mwcomment{Maybe should include a comparison to PATRIC.
-Patric does offer some comparative genomic analysis, however they are not health focused (pathways, gene presence/absence etc), and you cannnot add your own data to the repository}
-
 \section{CONCLUSIONS}
 
-Future work will focus on adding additional analysis modules to aid genotype to phenotype predictions using machine learning modules, and supporting bacterial species such as \textit{Salmonella}, and \textit{Campylobater}.
-While the integrated approach of storing and retrieving results provides enormous benefits, the developed analyses modules are self-contained and be transfered to existing platforms such as Galaxy.
+Future work will focus on adding additional analysis modules to aid genotype to phenotype predictions using machine learning modules, and supporting bacterial species such as \textit{Salmonella}, and \textit{Campylobacter}.
+While the integrated approach of storing and retrieving results provides enormous benefits, the developed analyses modules are self-contained and be transferred to existing platforms such as Galaxy.
 The source code for Spfy is hosted at \url{https://github.com/superphy/backend}, and is available for free under the open-source Apache 2.0 license.

### 2018-03-07 paper-webserver.tex
Edits and suggestions.

@@
 
-Public health laboratories are currently moving to whole-genome sequencing based analyses, and require rapid prediction of relevant pathogen phenotypes impacting health.
-Current workflows in comparative computational genomics rely on chaining different analysis software together, but lack storage and retrieval methods for the generated results.
-To solve this problem, we have created Spfy, which uses a graph database to store and retrieve results from computational workflows.
-The newly developed Spfy platform facilitates rapid phenotype identification, as well as the efficient storage and downstream comparative analysis of tens of thousands of genome sequences.
+Public health laboratories are currently moving to whole-genome sequence (WGS) based analyses, and require the rapid prediction of standard reference laboratory methods based solely on genomic data.
+Currently, these predictive genomics tasks rely on workflows that chain together multiple programs for the requisite analyses.
+While useful, these systems do not store the analyses in a genome-centric way, meaning the same analyses are often re-computed for the same genomes.
+
+To solve this problem, we created Spfy, which uses a graph database to store and retrieve results from computational workflows, and link data to individual genomes using standardized ontologies.
+
+The Spfy platform facilitates rapid phenotype identification, as well as the efficient storage and downstream comparative analysis of tens of thousands of genome sequences.
 Though generally applicable to bacterial genome sequences, Spfy currently contains 10,243 \textit{Escherichia coli} genomes, for which \textit{in-silico} serotype and Shiga-toxin subtype, as well as the presence of known virulence factors and antimicrobial resistance determinants have been computed.
-Spfy links the results and metadata to the genome sequences through a standardized ontology, which facilitates hypothesis testing in fields ranging from population genomics to epidemiology, while mitigating the recomputing of analyses.
-The graph approach is flexible, and can accommodate new analysis software modules as they are developed, and easily link new results to those already stored.
-Integrated data storage and analyses are currently necessary to match the rapid accumulation of whole genome sequences in public databases.
+Additionally, the presence / absence of then entire \textit{E.
+coli} pan-genome was computed and linked to each genome.
+
+Owing to its vast database of pre-computed results, and the ability to easily incorporate user data, Spfy facilitates hypothesis testing in fields ranging from population genomics to epidemiology, while mitigating the re-computation of analyses.
+
+The graph approach of Spfy is flexible, and can accommodate new analysis software modules as they are developed, easily linking new results to those already stored.
+Spfy provides a database and analyses approach for \textit{E.
+coli} that is able to match the rapid accumulation of WGS in public databases.
 \par
@@
 % 1. brief: WGS is standard
-Whole genome sequencing (WGS) can in theory provide the entire genetic content of an organism.
+Whole genome sequencing (WGS) can provide the entire genetic content of an organism.
 This unparalleled resolution and sensitivity has recently transformed public-health surveillance and outbreak response \cite{ronholm2016navigating,lytsy2017time}.
-Additionally, the identification of novel disease mechanisms \cite{wang2014whole,yuen2015whole}, and rapid clinical diagnoses and reference lab tests based on the specific mechanism of disease are now possible.
+Additionally, the identification of novel disease mechanisms \cite{wang2014whole,yuen2015whole}, and rapid clinical diagnoses and reference lab tests are now possible.
 \cite{willig2015whole,dewey2014clinical}.
@@
 % 2. Big Problem: but tools are for individual analysis
-The rapid characterization and comparison of bacterial pathogens relies principally on the combination of outputs from multiple software programs that are targeted for specific applications.
+The rapid characterization based on WGS relies on the outputs from multiple software programs that are targeted for specific applications.
 Examples include the identification of known antimicrobial resistance (AMR) genes, through software such as the Resistance Gene Identifier (RGI) \cite{mcarthur2013comprehensive}, \cite{kleinheinz2014applying}, \cite{gupta2014arg}, and ARIBA \cite{hunt2017ariba}; or the identification of known virulence factor genes (VF) through software such as VirulenceFinder \cite{kleinheinz2014applying}, SRST2 \cite{inouye2014srst2}, and GeneSippr \cite{lambert2015genesippr}.
@@
 
-Comprehensive platforms that combine individual programs into a cohesive whole also exist.
+Comprehensive platforms that combine individual analyses programs into a cohesive whole also exist.
 These include free platforms such as the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial}, and the Pathosystems Resource Integration Center (PATRIC) \cite{wattam2016improvements}.
@@
 
-Many of the comparative analyses that are currently used in the study of bacterial genomes are broadly useful, and therefore computed multiple times for the same genomes.
-An effective method to mitigate the recomputing of analyses, is to make the storage and retrieval of results part of the analyses platform, and effectively linked to the genomes of interest through a standardized ontology.
-Such measures can help ensure the rapid response times required for public health applications, and allow results to be integrated and progressively updated as new data becomes available.
+Many of the analyses that are currently used in the characterization and study of bacterial genomes are broadly useful, and therefore the same analyses are often computed multiple times for the same genome.
+An effective method to mitigate this re-computation, is to make the storage and retrieval of results part of the analyses platform, and effectively linked to the genomes of interest through a standardized ontology.
+Such measures help ensure the rapid response times required for public health applications, and allow results to be integrated and progressively updated as new data becomes available.
 
@@
 Superphy integrates pre-computed results with domain-specific knowledge to provide real-time exploration of publicly available genomes.
-While this tool has been useful for the thousands of pre-computed genomes in its database, the current pace of genome sequencing requires real-time predictive genomic analyses of tens-, and soon hundreds-of-thousands of genomes, and the long term storage and referencing of these results, something that the original SuperPhy platform was incapable of.
+While this tool has been useful for the thousands of pre-computed genomes in its database, the current pace of genome sequencing requires real-time predictive genomic analyses of tens-, and soon hundreds-of-thousands of genomes, and the long term storage and referencing of these results, which the original SuperPhy platform was incapable of.
 
@@
 % 5. how we solved it
-In this article, we present a new platform merging Superphy's pre-computed results with a novel data storage and processing architecture, which we call Spfy; Spfy integrates a graph database with real-time analyses to avoid recomputing identical runs.
+Here, we present a new platform merging Superphy's pre-computed results with a novel data storage and processing architecture, which we call Spfy; Spfy integrates a graph database with real-time analyses to avoid recomputing identical results.
 Graph-based result storage also allows retrospective comparisons across stored results as more genomes are sequenced or populations change, and is flexible, accommodating new analysis modules as they are developed.
@@
 \section{FUNCTIONALITY}
+% IMPORTANT: All Tables and Figues must be explicitly mentioned in the text.
+
 % ONLY FOCUS ON THE ANALYSIS MODULES
@@
 
-Spfy provides rapid in-silico versions of common reference laboratory tests for the analyses of \textit{E.
+Spfy provides rapid \textit{in-silico} versions of common reference laboratory tests for the analyses of \textit{E.
 coli}.
-It supports the following \textit{in-silico} subtyping options: serotyping, through both O- and H-antigen identification using ECTyper \url{https://github.com/phac-nml/ecoli_serotyping} as well as VF gene determination; Shiga-toxin 1 (Stx1), Shiga-toxin 2 (Stx2), and Intimin typing using Phylotyper \cite{whiteside2017phylotyper}, and AMR annotation using the RGI program \cite{mcarthur2013comprehensive}.
-Spfy also performs pangenome analyses using Panseq \cite{laing2010pan}, with the entire pangenome stored in the database and associated with source genomes.
+It supports the following \textit{in-silico} subtyping options: serotyping, through both O- and H-antigen identification using ECTyper \url{https://github.com/phac-nml/ecoli_serotyping} as well as VF gene determination; Shiga-toxin 1 (Stx1), Shiga-toxin 2 (Stx2), and Intimin typing using Phylotyper \cite{whiteside2017phylotyper}; and AMR annotation using the resistance gene identifier from the comprehensive antibiotic resistance database \cite{mcarthur2013comprehensive}.
+Spfy also performs pan-genome analyses using Panseq \cite{laing2010pan}, with the entire pan-genome stored in the database and associated with source genomes.
 
 Spfy handles all of the analyses tasks by dividing them into subtasks, which are subsequently distributed across a built-in task queue.
-Results are converted into individual graphs and stored within a larger graph database according to the standard ontologies GenEpiO \cite{griffiths2017context}, FALDO \cite{bolleman2016faldo}, and TypOn \cite{vaz2014typon}, where metadata including genotypes, biomarkers, host, source, and statistical significance testing of genome markers for user-defined groups are stored.
+Results are converted into individual graphs and stored within a larger graph database according to the previously created ontologies GenEpiO \cite{griffiths2017context}, FALDO \cite{bolleman2016faldo}, and TypOn \cite{vaz2014typon}, where metadata including genotypes, biomarkers, host, source are stored.
 
 By integrating task distribution with graph storage, Spfy enables large-scale analyses, such as epidemiological association studies.
-Any data type or relation in the graph is a valid option for analysis.
+Any data-type or relation in the graph is a valid option for analysis.
 This means that genomes can be compared on the basis of the presence or absence of pan-genome regions, serotype, subtyping data, or provided metadata such as location or host-source.
@@
 
-i) The upload begins through the website, where user-defined analysis options are selected.
-The results of these analyses are immediately reported to the user following their completion, while all other non-selected analyses are subsequently completed in the background and stored in the database without interaction from the user.
+i) The upload begins through the website, where user-defined analyses options are selected.
+The results of these analyses are immediately reported to the user, while all other non-selected analyses are subsequently completed in the background and stored in the database without interaction from the user.
 The public web service accepts uploads of up to 200 MB (approximately 50 \textit {E.
@@
 
-ii) User-selected analyses are enqueued into the Redis Queue \url{http://python-rq.org/} task queue.
+ii) User-selected analyses are enqueued with the Redis Queue \url{http://python-rq.org/} task queue.
 Redis Queue consists of a Redis Database \url{https://redis.io/} and task queue workers which run as Python processes.
@@
 % para: the semantic web
-Semantic web technologies describe the relationships between data \cite{berners2001semantic}, and graph databases for the storage of this information has been proposed as a open standard for sharing public information \cite{horrocks2005semantic}.
-For biological data, individual data points can be a genome, contiguous DNA sequence, or gene, and these are linked together in a searchable graph structure using existing ontologies.
+Semantic web technologies describe the relationships between data, and have been proposed as an open standard for sharing public information \cite{berners2001semantic}, while graph databases are a flexible means of storing this this information \cite{horrocks2005semantic}.
+Biological data can be a genome, contiguous DNA sequence, or gene, and these are linked together in a searchable graph structure using existing ontologies.
 This system is flexible and allows novel data to be incorporated into the existing graph.
+
+% The text in this figure is too small to read and seems blurry. It should be enlarged.
 \begin{figure}[!hb]
@@
 \end{center}
-\caption{Structure of Spfy's graph database.
+\caption{Structure of the Spfy graph database.
 Brackets highlight the source of different data points and the software it was generated from.
-These parts are added in as the analysis modules complete, at varying times, and the overall connections are inferred by the database.
+Data are added in as the analysis modules complete, at varying times, and the overall connections are inferred by the database.
 Non-bracketed sections are sourced from the uploaded genome files or user-supplied metadata.}
@@
 % Permanent storage + why a graph is preferable over tables
-The permanent storage of results is as a one-time cost to avoid recomputation when the same analysis is re-run.
-For analyses, Spfy searches the graph for all data points annotated with the queried ontology term.
-This graph data is then converted into the required structure, usually numerical arrays, or as required for the given analysis module.
+The permanent storage of results is as a one-time cost, which avoids re-computation when the same analysis is re-run.
+During analyses, Spfy searches the graph for all data points annotated with the queried ontology term.
+This graph data is then converted into the required structure, usually numerical arrays, for the given analysis module.
+
 In a graph database, a search can begin at any node or attribute.
-This is in contrast to a SQL database which requires a predefined schema, or a NoSQL database which treats data as documents with varying structure.
+This is in contrast to a SQL database, which requires a predefined schema, or a NoSQL database which treats data as documents with varying structure.
 For example, the addition of a new analysis module would typically require a new table definition in a SQL database, or the addition of a new document type in a NoSQL database.
@@
 The front-end website is written as a single-page application.
-To ensure a familiar user interface, we followed the Material Design specification \url{https://material.io/}, published by Google, surrounding a card-based design.
-(see Figure \ref{fig-results})
-Both the task selection and result displays follow the same design pattern: while data storage is graph-based, the results of various analysis modules are presented to users in a familiar tabular structure and available for download as .csv spreadsheet files.
-(see Figure \ref{fig-tables})
+To ensure a familiar user interface, we followed the Material Design specification \url{https://material.io/}, published by Google, built around a card-based design (Figure \ref{fig-results}).
+
+Both the task selection and result displays follow the same design pattern, where data storage is graph-based, but the results of analysis modules are presented to users in a familiar tabular structure and available for download as .csv spreadsheet files, as shown in Figure \ref{fig-tables}.
 
@@
 \end{center}
-\caption{First entry into the results interface for submitted tasks. Cards represent individual tasks and the entire collection can be viewed by sharing the URL with the embedded token.}
+\caption{The results interface for submitted tasks. Cards represent individual tasks and the entire collection can be accessed by sharing the URL with the embedded token.}
 \label{fig-results}
@@
 \end{center}
-\caption{An expanded result for a subtyping task.
+\caption{Detailed results for the virulence factor subtyping task.
 While data storage in Spfy is graph-based, a familiar tabular structure is presented to users.
@@
 \includegraphics[width=\textwidth]{images/docker}
-\end{center}
-\caption{The various Docker containers used in Spfy.
-Arrows represent the connections between different containers, and the entire platform can be recreated with a single command using its Docker-Compose definition.}
+\end{centerbut }
+\caption{The Docker containers used in Spfy.
+Arrows represent the connections between different containers, and the entire platform can be recreated with a single command using its Docker-Compose definition.
 \label{fig-docker}
@@
 
-Our tests for functionality and backwards compatibility run on TravisCI \url{https://travisci.io}, a continuous integration (CI) platform.
-The individual tests use PyTest \url{https://doc.pytest.org/}, and the current build status can be checked either on our GitHub repository or at \url{https://travis-ci.org/superphy/backend}.
+Tests for functionality and backwards compatibility are run on TravisCI \url{https://travisci.io}, a continuous integration (CI) platform.
+The individual tests use PyTest \url{https://doc.pytest.org/}, and the current build status can be checked on GitHub at \url{https://travis-ci.org/superphy/backend}.
 TravisCI also builds the core Docker images for Spfy, and uploads them to Docker Hub \url{https://hub.docker.com/u/superphy/}.
@@
 However, many subsets of biology now require the analyses of big-data, where inputs are taken from a variety of analysis programs, and involve large-scale data warehousing \cite{schatz2015biological}.
-The ability to integrate data from different source technologies, merge submissions from other labs, and distribute computations over fault-tolerant systems are now required for the types of analyses that need to be performed \cite{schatz2015biological}.
+The ability to integrate data from different source technologies, merge submissions from other labs, and distribute computations over fault-tolerant systems are now required \cite{schatz2015biological}.
 
 One of the key goals in developing Spfy was to accommodate and store a variety of result formats, and then to make the data from these results retrievable and usable as inputs for downstream analyses, such as predictive biomarker discovery.
+
+% This has not been shown in this paper; it has only been stated. We should probably have a large-scale analysis for an example, with accompanying times etc.
 We have shown how a graph database can accommodate the results from a variety of bioinformatics programs, and how Spfy is performant for data retrieval on the results for multiple analyses among over 10,000 genomes, providing results from big-data comparisons in the same efficiency as old analyses on single files.
@@
 The isolation and characterization of bacterial pathogens are critical for Public Health laboratories to rapidly respond to outbreaks, and to effectively monitor known and emerging pathogens through surveillance programs.
-Until recently, public-health agencies relied on laboratory tests such as serotyping, pulsed-field gel electrophoresis (PFGE), PCR-based amplification of known VFs, and disc-diffusion assays to identify AMR, for the characterization of bacterial isolates in outbreak, surveillance, and reference laboratory settings \cite{ronholm2016navigating}.
+Until recently, public-health agencies relied on laboratory tests such as serotyping, pulsed-field gel electrophoresis (PFGE), PCR-based amplification of known VFs, and disc-diffusion assays, for the characterization of bacterial isolates in outbreak, surveillance, and reference laboratory settings \cite{ronholm2016navigating}.
 Current efforts are focused on predictive genomics, where the relevant phenotypic information can be determined through examination of the whole-genome sequence without need for the traditional laboratory tests.
@@
 With the larger datasets involved in population genomics, it is no longer viable for individual researchers to download data to perform comparisons.
-Instead, efforts have focused on storing biological data online and enabling analyses on that data \cite{schatz2015biological}.
-By using a graph database, Spfy can integrate results from  different technologies including software predicted phenotypes as well as laboratory results and user-submitted metadata.
+Instead, efforts have focused on storing biological data online and enabling analyses of those data \cite{schatz2015biological}.
+By using a graph database, Spfy integrates results from  different technologies, as well as laboratory results and user-submitted metadata.
 In addition, datasets can be built and submitted from multiple labs for joint analyses.
@@
 % re: Matt "2. When comparing to IRIDA, BAP etc., can mention some differences with Superphy, e.g. the storage of interim result data that allows downstream integrated analysis"
+
+% If we mention speed comparisons, we need to have actual data supporting them
 On a per file basis, Spfy performs at a similar speed to BAP on predictive genomics tasks, usually completing within 2-3 minutes, though Spfy does not provide genome assembly services.
-However, unlike these workflow managers, Spfy is designed to help solve the needless recomputation of analyses by storing results in a graph database for downstream comparative studies.
+However, unlike these workflow managers, Spfy is designed to help solve the needless re-computation of analyses by storing results in a graph database for downstream comparative studies.
 This allows Spfy to perform population-wide analyses, regardless of the individual software used to generate the results for a given genome.
@@
 PATRIC \cite{wattam2016improvements} and Spfy share the same goal of integrated analyses.
-PATRIC adds support for comparing up to nine user-submitted genomes against a reference genome, targeting various gene annotations.
-The platform indexes a NoSQL document store to compare similar document types and is beneficial for comparative genomics.
-Unlike PARTIC, Spfy is targeted towards statistical comparisons of any permutation of a population group based on the chosen data types, and Spfy's graph database has no limit on the number of genomes grouped for comparison.
-For example, Spfy can compare all genomes with a particular subtype for gene annotations, source metadata targets for gene annotations, or any two (or more) node or attribute types as shown in \ref{fig-groupings}.
+PATRIC has support for comparing up to nine user-submitted genomes against a reference genome, based on gene annotations; the platform indexes a NoSQL document store to compare similar document types.
+Unlike PARTIC, Spfy provides the ability to perform statistical comparisons of any permutation of a population group based on the chosen data types, and Spfy's graph database has no limit on the number of genomes grouped for comparison.
+Spfy can compare all genomes with a particular subtype for gene annotations, source metadata targets for gene annotations, or any two (or more) node or attribute types as shown in Figure \ref{fig-groupings}.
 
@@
 \end{center}
-\caption{Genomes can be grouped by any node or attribute type for comparisons, and Spfy has no upper bound for the number of genomes in a group.
+\caption{
+% This needs to be specific to the image we are showing eg. the O157 vs. O26 comparison. We mention the general comparison in the text, and this can be offered as a specific example of that.
+Genomes can be grouped by any node or attribute type for comparisons, and Spfy has no upper bound for the number of genomes in a group.
 Different types can also be joined into a group through logical connectives AND, OR, Negation, and any data type is also a valid target.
@@
 
-Future work will focus on adding additional analysis modules to aid genotype to phenotype predictions using machine learning modules, and supporting bacterial species such as \textit{Salmonella}, and \textit{Campylobacter}.
-While the integrated approach of storing and retrieving results provides enormous benefits, the developed analyses modules are self-contained and be transferred to existing platforms such as Galaxy.
+The integrated approach taken in the creation of Spfy, where the storage and retrieval of results provides enormous benefits for the large-scale analyes of _E.
+coli_.
+The developed analyses modules are also self-contained and be transferred to existing platforms such as Galaxy.
+Future work will focus on adding machine learning modules to improve genotype / phenotype predictions, and supporting bacterial species such as \textit{Salmonella}, and \textit{Campylobacter}.
 The source code for Spfy is hosted at \url{https://github.com/superphy/backend}, and is available for free under the open-source Apache 2.0 license.

### 2018-03-24 scripts/timing.py
ADD: basic outline script for timings

@@ -1,0 +1,70 @@
+import os
+import time
+import random
+import numpy as np
+
+from bokeh.plotting import figure, show, output_file
+
+GENOME_POOL = os.getenv(
+    'GENOME_POOL',
+    'data/'
+)
+
+def _plot(x):
+    pass
+
+def _spfy(list_genomes):
+    pass
+
+def _bap(list_genomes):
+    pass
+
+def _time(func, list_genomes):
+    '''func provided should block until complete.
+    '''
+    start_time = time.time()
+    func(list_genomes)
+    stop_time = time.time()
+    elapsed_time = stop_time - start_time
+    return elapsed_time
+
+def _seed_genomes(size):
+    st = set()
+    pick = ''
+    for i in range(size):
+        while pick in st:
+            pick = random.choice(os.listdir(GENOME_POOL))
+        st.add(pick)
+    return list(st)
+
+class Result:
+    def __init__(self):
+        self.x_values = []
+        self.y_values = []
+
+    def update(self, x, y):
+        self.x_values.append(x)
+        self.y_values.append(y)
+
+    def as_np(self):
+        return np.array([self.x_values, self.y_values])
+
+def _timing(func, seeds):
+    r = Result()
+    for list_genomes in seeds:
+        # x-axis is the number of files.
+        x = len(list_genomes)
+        # y-axis is the execution time.
+        y = _time(func, list_genomes)
+        r.update(x,y)
+    return r.as_np()
+
+def main(start=1, stop=22, step=5):
+    # Create groups of seed genomes.
+    seeds = [_seed_genomes(i) for i in range(start,stop,step)]
+    # Run timings
+    r_spfy = _timing(_spfy, seeds)
+    r_bap = _timing(_bap, seeds)
+
+if __name__ == '__main__':
+    main()

### 2018-03-25 scripts/timing.py
STOP: still need to do timings, but addresses cl comments for merge

@@ -16,6 +16,12 @@ def _plot(x):
 def _spfy(list_genomes):
     pass
 
+def _spfy_gc():
+    # analysis run-time / throughput with different levels of parallelization
+    # particularly for statistical tests
+    # should do a 1 genome = X number VFs, Y number of genomes for Y*X retrieval/analysis
+    pass
+
 def _bap(list_genomes):
     pass
 

### 2018-03-25 paper-webserver.tex

@@
 All results are displayed to users in real-time, usually within 2-3 minutes.
+For example, Spfy can determine if there's a statistically significant difference, using Fisher's exact test, in the presence of all known AMR genes in collected O157 \textit{E.
+coli} genomes vs.
+O26 as shown in Figure \ref{fig-groupings}.
+Different types can also be joined into a group through logical connectives AND, OR, Negation, and any data type is also a valid target.
+This approach can be used to compare any data regardless of source software.
 
-% para covering ectyper & RGI -- don't really need this
-% RGI has its own paper and was in the previous SuperPhy paper
-% ectyper will get its own paper
-% phylotyper also has its own paper
+\begin{figure}[!hb]
+\begin{center}
+\includegraphics[width=\textwidth]{images/o157vo26_amrgenes}
+\end{center}
+\caption{
+% This needs to be specific to the image we are showing eg. the O157 vs. O26 comparison. We mention the general comparison in the text, and this can be offered as a specific example of that.
+Genomes with the predicted subtype O157 are compared against the subtype O26.
+In this example, there are 903 O157 genomes and 291 O26 genomes which are compared for the presence of 129 AMR genes.
+Fisher's Exact Test is used for the comparison.}
+\label{fig-groupings}
+\end{figure}
 
@@
 \includegraphics[width=\textwidth]{images/docker}
-\end{centerbut }
+\end{center}
 \caption{The Docker containers used in Spfy.
-Arrows represent the connections between different containers, and the entire platform can be recreated with a single command using its Docker-Compose definition.
+Arrows represent the connections between different containers, and the entire platform can be recreated with a single command using its Docker-Compose definition.}
 \label{fig-docker}
@@
 \par
-
-\small \input{tables/database}
 
@@
 Unlike PARTIC, Spfy provides the ability to perform statistical comparisons of any permutation of a population group based on the chosen data types, and Spfy's graph database has no limit on the number of genomes grouped for comparison.
-Spfy can compare all genomes with a particular subtype for gene annotations, source metadata targets for gene annotations, or any two (or more) node or attribute types as shown in Figure \ref{fig-groupings}.
-
-\begin{figure}[!hb]
-\begin{center}
-\includegraphics[width=\textwidth]{images/groupings}
-\end{center}
-\caption{
-% This needs to be specific to the image we are showing eg. the O157 vs. O26 comparison. We mention the general comparison in the text, and this can be offered as a specific example of that.
-Genomes can be grouped by any node or attribute type for comparisons, and Spfy has no upper bound for the number of genomes in a group.
-Different types can also be joined into a group through logical connectives AND, OR, Negation, and any data type is also a valid target.
-This approach can be used to compare any data regardless of source software.}
-\label{fig-groupings}
-\end{figure}
 
@@
 
-The integrated approach taken in the creation of Spfy, where the storage and retrieval of results provides enormous benefits for the large-scale analyes of _E.
-coli_.
+The integrated approach taken in the creation of Spfy, where the storage and retrieval of results provides enormous benefits for the large-scale analyes of \textit{E.
+coli}.
 The developed analyses modules are also self-contained and be transferred to existing platforms such as Galaxy.

### 2018-03-26 scripts/timing.py
ADD: docker call for BAP

@@ -1,6 +1,7 @@
 import os
 import time
 import random
+import subprocess
 import numpy as np
 
 from bokeh.plotting import figure, show, output_file
@@ -23,7 +24,9 @@ def _spfy_gc():
     pass
 
 def _bap(list_genomes):
-    pass
+    # BAP throws an error without KmerFinder.
+    r = subprocess.check_call("""docker run -ti --rm -w /workdir -v $(pwd):/workdir    cgetools BAP --dbdir /usr/src/cgepipeline/test/databases  --services KmerFinder,ResFinder,VirulenceFinder  --fa /usr/src/cgepipeline/test/test.fa""", shell=True)
+    return True
 
 def _time(func, list_genomes):
     '''func provided should block until complete.

### 2018-03-26 scripts/timing.py
ADD: func to time and size gc queries

@@ -2,8 +2,10 @@ import os
 import time
 import random
 import subprocess
+import requests
 import numpy as np
 
+from time import sleep
 from bokeh.plotting import figure, show, output_file
 
 GENOME_POOL = os.getenv(
@@ -11,16 +13,79 @@ GENOME_POOL = os.getenv(
     'data/'
 )
 
+ROOT = os.getenv(
+    'SPFY_API',
+    'https://localhost:9000/'
+)
+API = ROOT + 'api/v0/'
+
 def _plot(x):
     pass
 
 def _spfy(list_genomes):
     pass
 
+def _run_gc(rl_a, attr_a, rl_b, attr_b, target):
+    """
+    POSTs the comparison and retrieves its runtime.
+    """
+    groups = [
+        [
+            {
+                'negated': 'false',
+                'relation': rl_a,
+                'attribute': attr_a,
+                'logical': 'null'
+            }
+        ],
+        [
+            {
+                'negated': 'false',
+                'relation': rl_b,
+                'attribute': attr_b,
+                'logical': 'null'
+            }
+        ]
+    ]
+    data = {
+        'groups': groups,
+        'target': target
+    }
+    # POST and get the jobid from Spfy.
+    jobid = requests.post(API + 'newgroupcomparison')
+
+    # Loop until complete.
+    while requests.get(API + 'results/' + jobid) == 'pending':
+        # The length we sleep doesn't matter, as timing is retrieved directly
+        # from RQ.
+        sleep(4)
+    # Grab the result.
+    r = requests.get(API + 'results/' + jobid)
+    # Tell me how many rows (ie. how many found targets) there were.
+    size_targets = len(r['index'])
+    # Find the number of genoems for a given attribute.
+    row = r['data'][0]
+    size_attr_a = row[3] + row[4]
+    size_attr_b = row[5] + row[6]
+
+    # Request the time it took to run, in seconds.
+    sec = requests.get(API + 'timings/' + jobid)
+    return (size_attr_a, size_attr_b, size_targets, sec)
+
 def _spfy_gc():
-    # analysis run-time / throughput with different levels of parallelization
-    # particularly for statistical tests
-    # should do a 1 genome = X number VFs, Y number of genomes for Y*X retrieval/analysis
+    """
+    analysis run-time / throughput with different levels of parallelization
+    particularly for statistical tests
+    should do a 1 genome = X number VFs, Y number of genomes for Y*X retrieval/analysis
[…]

### 2018-03-26 scripts/timing.py
ADD: func to time across all o/htypes and build np.array

@@ -19,16 +19,39 @@ ROOT = os.getenv(
 )
 API = ROOT + 'api/v0/'
 
+class Result:
+    def __init__(self):
+        self.x_values = []
+        self.y_values = []
+
+    def update(self, x, y):
+        self.x_values.append(x)
+        self.y_values.append(y)
+
+    def as_np(self):
+        return np.array([self.x_values, self.y_values])
+
 def _plot(x):
     pass
 
 def _spfy(list_genomes):
     pass
 
-def _run_gc(rl_a, attr_a, rl_b, attr_b, target):
+def _find_rl(attr):
+    if attr[0] == 'O':
+        return 'http://purl.obolibrary.org/obo/GENEPIO_0001076'
+    elif attr[0] == 'H':
+        return 'http://purl.obolibrary.org/obo/GENEPIO_0001077'
+    else:
+        raise Exception('_find_rl() couldnt find relation for {0}'.format(attr))
+
+def _run_gc(attr_a, attr_b, target):
     """
     POSTs the comparison and retrieves its runtime.
     """
+    rl_a = _find_rl(attr_a)
+    rl_b = _find_rl(attr_b)
+
     groups = [
         [
             {
@@ -72,7 +95,7 @@ def _run_gc(rl_a, attr_a, rl_b, attr_b, target):
     sec = requests.get(API + 'timings/' + jobid)
     return (size_attr_a, size_attr_b, size_targets, sec)
 
-def _spfy_gc():
+def _attr_gc():
     """
     analysis run-time / throughput with different levels of parallelization
     particularly for statistical tests
@@ -86,7 +109,24 @@ def _spfy_gc():
     o_types = requests.get(API + 'get_attribute_values/type/http://purl.obolibrary.org/obo/GENEPIO_0001076')
     o_types = o_types.values()
 
-    pass
+    l = h_types + o_types
+    return l
+
+def _time_gc():
+    attributes = _attr_gc()
+    r = Result()
+    for target in ('https://www.github.com/superphy#AntimicrobialResistanceGene', 'https://www.github.com/superphy#VirulenceFactor'):
+        p = 0
+        q = 1
+        while p <= len(attributes) - 2:
+            while q <= len(attribuetes) - 1:
+                st = _run_gc(attributes[p], attributes[q], target)
+                attr_x_targets = (st(0) + st(1)) * st(3)
+                r.update(attr_x_targets, st(4))
+                q += 1
+            p += 1
+            q = p + 1
+    return r.as_np()
 
[…]

### 2018-03-26 scripts/timing.py
ADD: lets be safe and dump the raws as well

@@ -4,8 +4,10 @@ import random
 import subprocess
 import requests
 import numpy as np
+import cPickle as pickle
 
 from time import sleep
+from datetime import datetime
 from bokeh.plotting import figure, show, output_file
 
 GENOME_POOL = os.getenv(
@@ -115,6 +117,10 @@ def _attr_gc():
 def _time_gc():
     attributes = _attr_gc()
     r = Result()
+    raws = []
+    now = datetime.now()
+    now = now.strftime("%Y-%m-%d-%H-%M-%S")
+    # Loop.
     for target in ('https://www.github.com/superphy#AntimicrobialResistanceGene', 'https://www.github.com/superphy#VirulenceFactor'):
         p = 0
         q = 1
@@ -122,11 +128,17 @@ def _time_gc():
             while q <= len(attribuetes) - 1:
                 st = _run_gc(attributes[p], attributes[q], target)
                 attr_x_targets = (st(0) + st(1)) * st(3)
+                # Update our main Result instance with the values.
                 r.update(attr_x_targets, st(4))
+                # Also saves the raws for writing out.
+                raws.append(st)
                 q += 1
             p += 1
             q = p + 1
-    return r.as_np()
+    array = r.as_np()
+    pickle.dump(raws, open(now + '_raws.p', "wb" ))
+    pickle.dump(array, open(now + '_array.p', "wb" ))
+    return array
 
 def _bap(list_genomes):
     # BAP throws an error without KmerFinder.

### 2018-03-26 scripts/timing.py
FIX: tupple calls in timing

@@ -125,11 +125,11 @@ def _time_gc():
         p = 0
         q = 1
         while p <= len(attributes) - 2:
-            while q <= len(attribuetes) - 1:
+            while q <= len(attributes) - 1:
                 st = _run_gc(attributes[p], attributes[q], target)
-                attr_x_targets = (st(0) + st(1)) * st(3)
+                attr_x_targets = (st[0] + st[1]) * st[3]
                 # Update our main Result instance with the values.
-                r.update(attr_x_targets, st(4))
+                r.update(attr_x_targets, st[4])
                 # Also saves the raws for writing out.
                 raws.append(st)
                 q += 1

### 2018-03-27 scripts/timing.py
EDIT: a bit more data into the raws

@@ -131,7 +131,7 @@ def _time_gc():
                 # Update our main Result instance with the values.
                 r.update(attr_x_targets, st[4])
                 # Also saves the raws for writing out.
-                raws.append(st)
+                raws.append(st + (attributes[p], attributes[q], target))
                 q += 1
             p += 1
             q = p + 1

### 2018-03-27 scripts/timing.py
FIX: port

@@ -17,7 +17,7 @@ GENOME_POOL = os.getenv(
 
 ROOT = os.getenv(
     'SPFY_API',
-    'https://localhost:9000/'
+    'https://localhost:8000/'
 )
 API = ROOT + 'api/v0/'
 

### 2018-03-27 scripts/timing.py
FIX: right http

@@ -17,7 +17,7 @@ GENOME_POOL = os.getenv(
 
 ROOT = os.getenv(
     'SPFY_API',
-    'https://localhost:8000/'
+    'http://localhost:8000/'
 )
 API = ROOT + 'api/v0/'
 

### 2018-03-27 scripts/timing.py
FIX: response types

@@ -77,15 +77,15 @@ def _run_gc(attr_a, attr_b, target):
         'target': target
     }
     # POST and get the jobid from Spfy.
-    jobid = requests.post(API + 'newgroupcomparison')
+    jobid = requests.post(API + 'newgroupcomparison').text
 
     # Loop until complete.
-    while requests.get(API + 'results/' + jobid) == 'pending':
+    while requests.get(API + 'results/' + jobid).text == 'pending':
         # The length we sleep doesn't matter, as timing is retrieved directly
         # from RQ.
         sleep(4)
     # Grab the result.
-    r = requests.get(API + 'results/' + jobid)
+    r = requests.get(API + 'results/' + jobid).json()
     # Tell me how many rows (ie. how many found targets) there were.
     size_targets = len(r['index'])
     # Find the number of genoems for a given attribute.
@@ -94,7 +94,7 @@ def _run_gc(attr_a, attr_b, target):
     size_attr_b = row[5] + row[6]
 
     # Request the time it took to run, in seconds.
-    sec = requests.get(API + 'timings/' + jobid)
+    sec = requests.get(API + 'timings/' + jobid).text
     return (size_attr_a, size_attr_b, size_targets, sec)
 
 def _attr_gc():

### 2018-03-27 scripts/timing.py
FIX: response types

@@ -104,11 +104,11 @@ def _attr_gc():
     should do a 1 genome = X number VFs, Y number of genomes for Y*X retrieval/analysis
     """
     # Get all H-Types in the database.
-    h_types = requests.get(API + 'get_attribute_values/type/http://purl.obolibrary.org/obo/GENEPIO_0001077')
+    h_types = requests.get(API + 'get_attribute_values/type/http://purl.obolibrary.org/obo/GENEPIO_0001077').json()
     # This is a weird json format due to the Human-Readable conversion, convert it to a simple list.
     h_types = h_types.values()
     # Get all O-Types in the database.
-    o_types = requests.get(API + 'get_attribute_values/type/http://purl.obolibrary.org/obo/GENEPIO_0001076')
+    o_types = requests.get(API + 'get_attribute_values/type/http://purl.obolibrary.org/obo/GENEPIO_0001076').json()
     o_types = o_types.values()
 
     l = h_types + o_types

### 2018-03-27 scripts/timing.py
DEBUG: json conv

@@ -85,7 +85,12 @@ def _run_gc(attr_a, attr_b, target):
         # from RQ.
         sleep(4)
     # Grab the result.
-    r = requests.get(API + 'results/' + jobid).json()
+    r = requests.get(API + 'results/' + jobid)
+    try:
+        r = r.json()
+    except:
+        print(r)
+        raise Exception('failed')
     # Tell me how many rows (ie. how many found targets) there were.
     size_targets = len(r['index'])
     # Find the number of genoems for a given attribute.

### 2018-03-27 scripts/timing.py
DEBUG: 404?

@@ -78,7 +78,7 @@ def _run_gc(attr_a, attr_b, target):
     }
     # POST and get the jobid from Spfy.
     jobid = requests.post(API + 'newgroupcomparison').text
-
+    print("jobid", jobid)
     # Loop until complete.
     while requests.get(API + 'results/' + jobid).text == 'pending':
         # The length we sleep doesn't matter, as timing is retrieved directly

### 2018-03-27 scripts/timing.py
FIX: didnt add payload

@@ -77,7 +77,7 @@ def _run_gc(attr_a, attr_b, target):
         'target': target
     }
     # POST and get the jobid from Spfy.
-    jobid = requests.post(API + 'newgroupcomparison').text
+    jobid = requests.post(API + 'newgroupcomparison', data=data).text
     print("jobid", jobid)
     # Loop until complete.
     while requests.get(API + 'results/' + jobid).text == 'pending':

### 2018-03-27 scripts/timing.py
FIX: json encode the data instead of forms

@@ -77,7 +77,7 @@ def _run_gc(attr_a, attr_b, target):
         'target': target
     }
     # POST and get the jobid from Spfy.
-    jobid = requests.post(API + 'newgroupcomparison', data=data).text
+    jobid = requests.post(API + 'newgroupcomparison', json=data).text
     print("jobid", jobid)
     # Loop until complete.
     while requests.get(API + 'results/' + jobid).text == 'pending':

### 2018-03-27 scripts/timing.py
DEBUG: string?

@@ -88,6 +88,7 @@ def _run_gc(attr_a, attr_b, target):
     r = requests.get(API + 'results/' + jobid)
     try:
         r = r.json()
+        print ("r", r)
     except:
         print(r)
         raise Exception('failed')

### 2018-03-27 scripts/timing.py
DEBUG: unicode

@@ -80,9 +80,10 @@ def _run_gc(attr_a, attr_b, target):
     jobid = requests.post(API + 'newgroupcomparison', json=data).text
     print("jobid", jobid)
     # Loop until complete.
-    while requests.get(API + 'results/' + jobid).text == 'pending':
+    while requests.get(API + 'results/' + jobid).text == unicode('pending'):
         # The length we sleep doesn't matter, as timing is retrieved directly
         # from RQ.
+        print "sleeping"
         sleep(4)
     # Grab the result.
     r = requests.get(API + 'results/' + jobid)

### 2018-03-27 scripts/timing.py
FIX: response

@@ -80,19 +80,13 @@ def _run_gc(attr_a, attr_b, target):
     jobid = requests.post(API + 'newgroupcomparison', json=data).text
     print("jobid", jobid)
     # Loop until complete.
-    while requests.get(API + 'results/' + jobid).text == unicode('pending'):
+    while requests.get(API + 'results/' + jobid).json() == unicode('pending'):
         # The length we sleep doesn't matter, as timing is retrieved directly
         # from RQ.
         print "sleeping"
         sleep(4)
     # Grab the result.
-    r = requests.get(API + 'results/' + jobid)
-    try:
-        r = r.json()
-        print ("r", r)
-    except:
-        print(r)
-        raise Exception('failed')
+    r = requests.get(API + 'results/' + jobid).json()
     # Tell me how many rows (ie. how many found targets) there were.
     size_targets = len(r['index'])
     # Find the number of genoems for a given attribute.

### 2018-03-27 scripts/timing.py
DEBUG: prints

@@ -87,6 +87,7 @@ def _run_gc(attr_a, attr_b, target):
         sleep(4)
     # Grab the result.
     r = requests.get(API + 'results/' + jobid).json()
+    print(r)
     # Tell me how many rows (ie. how many found targets) there were.
     size_targets = len(r['index'])
     # Find the number of genoems for a given attribute.

### 2018-03-27 scripts/timing.py
DEBUG: getting blank response

@@ -76,6 +76,7 @@ def _run_gc(attr_a, attr_b, target):
         'groups': groups,
         'target': target
     }
+    print(data)
     # POST and get the jobid from Spfy.
     jobid = requests.post(API + 'newgroupcomparison', json=data).text
     print("jobid", jobid)

### 2018-03-27 scripts/timing.py
FIX: workaround for empty datasets

@@ -92,9 +92,14 @@ def _run_gc(attr_a, attr_b, target):
     # Tell me how many rows (ie. how many found targets) there were.
     size_targets = len(r['index'])
     # Find the number of genoems for a given attribute.
-    row = r['data'][0]
-    size_attr_a = row[3] + row[4]
-    size_attr_b = row[5] + row[6]
+    if r['data']:
+        row = r['data'][0]
+        size_attr_a = row[3] + row[4]
+        size_attr_b = row[5] + row[6]
+    else:
+        # data is empty. May happen with a small db.
+        size_attr_a = 0
+        size_attr_b = 0
 
     # Request the time it took to run, in seconds.
     sec = requests.get(API + 'timings/' + jobid).text
@@ -124,7 +129,9 @@ def _time_gc():
     now = datetime.now()
     now = now.strftime("%Y-%m-%d-%H-%M-%S")
     # Loop.
-    for target in ('https://www.github.com/superphy#AntimicrobialResistanceGene', 'https://www.github.com/superphy#VirulenceFactor'):
+    targets = ('https://www.github.com/superphy#AntimicrobialResistanceGene', 'https://www.github.com/superphy#VirulenceFactor')
+    targets = ('https://www.github.com/superphy#AntimicrobialResistanceGene', 'https://www.github.com/superphy#VirulenceFactor')
+    for target in targets:
         p = 0
         q = 1
         while p <= len(attributes) - 2:

### 2018-03-27 scripts/timing.py
FIX: workaround for empty datasets

@@ -130,7 +130,6 @@ def _time_gc():
     now = now.strftime("%Y-%m-%d-%H-%M-%S")
     # Loop.
     targets = ('https://www.github.com/superphy#AntimicrobialResistanceGene', 'https://www.github.com/superphy#VirulenceFactor')
-    targets = ('https://www.github.com/superphy#AntimicrobialResistanceGene', 'https://www.github.com/superphy#VirulenceFactor')
     for target in targets:
         p = 0
         q = 1

### 2018-03-27 scripts/timing.py
FIX: indexes

@@ -136,9 +136,9 @@ def _time_gc():
         while p <= len(attributes) - 2:
             while q <= len(attributes) - 1:
                 st = _run_gc(attributes[p], attributes[q], target)
-                attr_x_targets = (st[0] + st[1]) * st[3]
+                attr_x_targets = (st[0] + st[1]) * st[2]
                 # Update our main Result instance with the values.
-                r.update(attr_x_targets, st[4])
+                r.update(attr_x_targets, st[3])
                 # Also saves the raws for writing out.
                 raws.append(st + (attributes[p], attributes[q], target))
                 q += 1

### 2018-03-27 scripts/timing.py
FIX: was seeing as negated

@@ -57,18 +57,18 @@ def _run_gc(attr_a, attr_b, target):
     groups = [
         [
             {
-                'negated': 'false',
+                'negated': False,
                 'relation': rl_a,
                 'attribute': attr_a,
-                'logical': 'null'
+                'logical': None
             }
         ],
         [
             {
-                'negated': 'false',
+                'negated': False,
                 'relation': rl_b,
                 'attribute': attr_b,
-                'logical': 'null'
+                'logical': None
             }
         ]
     ]
@@ -120,6 +120,7 @@ def _attr_gc():
     o_types = o_types.values()
 
     l = h_types + o_types
+    print(l)
     return l
 
 def _time_gc():

### 2018-03-27 scripts/timing.py
FIX: check for unsuitable attributes

@@ -120,6 +120,8 @@ def _attr_gc():
     o_types = o_types.values()
 
     l = h_types + o_types
+    # Rm unsuitable attributes.
+    l = [s for s in l if s[0] in ('O', 'H')]
     print(l)
     return l
 

### 2018-03-27 scripts/timing.py
CHANGE: larger try/except block since were going to run for awhile

@@ -80,29 +80,34 @@ def _run_gc(attr_a, attr_b, target):
     # POST and get the jobid from Spfy.
     jobid = requests.post(API + 'newgroupcomparison', json=data).text
     print("jobid", jobid)
-    # Loop until complete.
-    while requests.get(API + 'results/' + jobid).json() == unicode('pending'):
-        # The length we sleep doesn't matter, as timing is retrieved directly
-        # from RQ.
-        print "sleeping"
-        sleep(4)
-    # Grab the result.
-    r = requests.get(API + 'results/' + jobid).json()
-    print(r)
-    # Tell me how many rows (ie. how many found targets) there were.
-    size_targets = len(r['index'])
-    # Find the number of genoems for a given attribute.
-    if r['data']:
+    # Sleep at least 1 second.
+    sleep(1)
+    try:
+        # Loop until complete.
+        while requests.get(API + 'results/' + jobid).json() == unicode('pending'):
+            # The length we sleep doesn't matter, as timing is retrieved directly
+            # from RQ.
+            print "sleeping"
+            sleep(4)
+        # Grab the result.
+        r = requests.get(API + 'results/' + jobid).json()
+        print(r)
+        # Tell me how many rows (ie. how many found targets) there were.
+        size_targets = len(r['index'])
+        # Find the number of genoems for a given attribute.
         row = r['data'][0]
         size_attr_a = row[3] + row[4]
         size_attr_b = row[5] + row[6]
-    else:
-        # data is empty. May happen with a small db.
+        # Request the time it took to run, in seconds.
+        sec = requests.get(API + 'timings/' + jobid).text
+    except:
+        # Either:
+        # 1) data is empty. This can be due to a small db.
+        # 2) some other error. Likely something timed out.
+        size_targets = 0
         size_attr_a = 0
         size_attr_b = 0
-
-    # Request the time it took to run, in seconds.
-    sec = requests.get(API + 'timings/' + jobid).text
+        sec = 0
     return (size_attr_a, size_attr_b, size_targets, sec)
 
 def _attr_gc():
@@ -125,7 +130,7 @@ def _attr_gc():
     print(l)
     return l
 
-def _time_gc():
+def time_gc():
     attributes = _attr_gc()
     r = Result()
     raws = []

### 2018-03-27 scripts/timing.py
CHANGE: print tuple

@@ -108,7 +108,9 @@ def _run_gc(attr_a, attr_b, target):
         size_attr_a = 0
         size_attr_b = 0
         sec = 0
-    return (size_attr_a, size_attr_b, size_targets, sec)
+    ret = (size_attr_a, size_attr_b, size_targets, sec)
+    print(ret)
+    return ret
 
 def _attr_gc():
     """

### 2018-04-01 scripts/timing.py
DEBUG: bypass Job.timing failures for now

@@ -248,7 +248,7 @@ def _timing(func, seeds):
         r.update(d)
     return r
 
-def main(spfy=True, bap=True, start=1, stop=22, step=5):
+def main(spfy=True, bap=False, start=1, stop=22, step=5):
     # Create groups of seed genomes.
     seeds = [_seed_genomes(i) for i in range(start,stop,step)]
     # Run timings

### 2018-04-02 paper-webserver.tex
EDIT: note that bap uses https://en.wikipedia.org/wiki/TORQUE
ADD: some specs on bap timings

@@
 
-Spfy is similar to these technologies in that it automates workflows for users, and as in Galaxy, uses task queues to distribute selected analyses.
+Spfy is similar to these technologies in that it automates workflows for users and uses task queues to distribute selected analyses.
 % re: Matt "2. When comparing to IRIDA, BAP etc., can mention some differences with Superphy, e.g. the storage of interim result data that allows downstream integrated analysis"
@@
 % If we mention speed comparisons, we need to have actual data supporting them
-On a per file basis, Spfy performs at a similar speed to BAP on predictive genomics tasks, usually completing within 2-3 minutes, though Spfy does not provide genome assembly services.
+On a per file basis, Spfy performs at a similar speed to BAP on predictive genomics tasks, though Spfy does not provide genome assembly services.
+After accounting for assembly services, BAP reported \cite{thomsen2016bacterial} an average runtime of 8-9 minutes per genome over 476 runs.
+ResFinder, which is similar to the RGI program used by Spfy, took an average of 3-4 minutes.
 However, unlike these workflow managers, Spfy is designed to help solve the needless re-computation of analyses by storing results in a graph database for downstream comparative studies.

### 2018-04-02 scripts/timefishers.py
CHANGE: split fishers and spfy timings into separate scripts

@@ -1,15 +1,10 @@
 import os
-import time
-import random
-import subprocess
 import requests
-import zipfile
 import numpy as np
 import cPickle as pickle
 
 from time import sleep
 from datetime import datetime
-from bokeh.plotting import figure, show, output_file
 
 GENOME_POOL = os.getenv(
     'GENOME_POOL',
@@ -24,24 +19,12 @@ API = ROOT + 'api/v0/'
 
 # Utility functions.
 
-def _plot(x):
-    pass
-
 def _now():
     now = datetime.now()
     now = now.strftime("%Y-%m-%d-%H-%M-%S-%f")
     return now
 
-def _seed_genomes(size):
-    st = set()
-    pick = random.choice(os.listdir(GENOME_POOL))
-    for i in range(size):
-        while pick in st:
-            pick = random.choice(os.listdir(GENOME_POOL))
-        st.add(pick)
-    return list(st)
-
-# Result classes.
+# Result class.
 
 class Result:
     def __init__(self):
@@ -55,65 +38,7 @@ class Result:
     def as_np(self):
         return np.array([self.x_values, self.y_values])
 
-class BarResult:
-    def __init__(self, list_sizes):
-        self.list_sizes = list_sizes
-        self.subtasks = None
-        self.data = {
-            'size_increments': size_increments
-        }
-
-    def update(self, timing):
-        '''Variable "timing" should be of shape {sample_size(str): list_subtask_times(list)}
-        '''
-        assert isinstance(timing, dict)
-        self.data.update(timing)
-
 # Calling functions.
-
-def _run_spfy(list_genomes):
-    '''POSTs to Spfy's API.
-    '''
-    # Zip the files.
-    name = _now() + '.zip'
-    with ZipFile(name, 'w') as z:
-        for f in list_genomes:
-            z.write(f)
-    # File payload.
-    files = {'file': open(name), 'rb'}
-    # Options payload.
-    data = {
-        'options.pi': 90,
-        'options.amr': True,
-        'options.serotype': True,
-        'options.vf': True,
-        'options.stx1': True,
-        'options.stx2': True,
-        'options.eae': True,
-        'options.groupresults': True,
-        'options.bulk': False,
-        'options.pan': True
-    }
-
-    # POST.
-    r = requests.post(API + 'upload', data=data, files=files)
-    pipeline_id = r.text
-    print("pipeline_id", pipeline_id)
[…]

### 2018-04-02 scripts/timespfy.py

@@ -1,0 +1,138 @@
+import os
+import time
+import random
+import subprocess
+import requests
+import zipfile
+import numpy as np
+import cPickle as pickle
+
+from time import sleep
+from datetime import datetime
+
+GENOME_POOL = os.getenv(
+    'GENOME_POOL',
+    'data/'
+)
+
+ROOT = os.getenv(
+    'SPFY_API',
+    'http://localhost:8000/'
+)
+API = ROOT + 'api/v0/'
+
+# Utility functions.
+
+def _now():
+    now = datetime.now()
+    now = now.strftime("%Y-%m-%d-%H-%M-%S-%f")
+    return now
+
+def _seed_genomes(size):
+    st = set()
+    pick = random.choice(os.listdir(GENOME_POOL))
+    for i in range(size):
+        while pick in st:
+            pick = random.choice(os.listdir(GENOME_POOL))
+        st.add(pick)
+    return list(st)
+
+# Result classes.
+
+class Result:
+    def __init__(self):
+        self.x_values = []
+        self.y_values = []
+
+    def update(self, x, y):
+        self.x_values.append(x)
+        self.y_values.append(y)
+
+    def as_np(self):
+        return np.array([self.x_values, self.y_values])
+
+class BarResult:
+    def __init__(self, list_sizes):
+        self.list_sizes = list_sizes
+        self.subtasks = None
+        self.data = {
+            'size_increments': size_increments
+        }
+
+    def update(self, timing):
+        '''Variable "timing" should be of shape {sample_size(str): list_subtask_times(list)}
+        '''
+        assert isinstance(timing, dict)
+        self.data.update(timing)
+
+# Calling functions.
+
+def _run_spfy(list_genomes):
+    '''POSTs to Spfy's API.
+    '''
+    # Zip the files.
+    name = _now() + '.zip'
+    with ZipFile(name, 'w') as z:
+        for f in list_genomes:
+            z.write(f)
+    # File payload.
+    files = {'file': open(name), 'rb'}
+    # Options payload.
+    data = {
+        'options.pi': 90,
+        'options.amr': True,
+        'options.serotype': True,
+        'options.vf': True,
+        'options.stx1': True,
+        'options.stx2': True,
+        'options.eae': True,
+        'options.groupresults': True,
+        'options.bulk': False,
+        'options.pan': True
+    }
+
+    # POST.
+    r = requests.post(API + 'upload', data=data, files=files)
[…]

### 2018-04-02 scripts/timespfy.py
FIX: basic fixes to the spfy timing script

@@ -3,12 +3,12 @@ import time
 import random
 import subprocess
 import requests
-import zipfile
 import numpy as np
 import cPickle as pickle
 
 from time import sleep
 from datetime import datetime
+from zipfile import ZipFile
 
 GENOME_POOL = os.getenv(
     'GENOME_POOL',
@@ -56,7 +56,7 @@ class BarResult:
         self.list_sizes = list_sizes
         self.subtasks = None
         self.data = {
-            'size_increments': size_increments
+            'size_increments': list_sizes
         }
 
     def update(self, timing):
@@ -70,13 +70,18 @@ class BarResult:
 def _run_spfy(list_genomes):
     '''POSTs to Spfy's API.
     '''
-    # Zip the files.
-    name = _now() + '.zip'
-    with ZipFile(name, 'w') as z:
-        for f in list_genomes:
-            z.write(f)
-    # File payload.
-    files = {'file': open(name), 'rb'}
+    # Zip files if more than 1 genome.
+    if len(list_genomes)>1:
+        # Zip the files.
+        name = _now() + '.zip'
+        with ZipFile(name, 'w') as z:
+            for f in list_genomes:
+                z.write(f)
+        # File payload.
+        files = {'file': open(name, 'rb')}
+    else:
+        files = {'file': open(list_genomes[0], 'rb')}
+
     # Options payload.
     data = {
         'options.pi': 90,
@@ -99,13 +104,13 @@ def _run_spfy(list_genomes):
     sleep(4)
     try:
         # Loop until complete.
-        while requests.get(API + 'results/' + jobid).json() == unicode('pending'):
+        while requests.get(API + 'results/' + pipeline_id).json() == unicode('pending'):
             # The length we sleep doesn't matter, as timing is retrieved directly
             # from RQ.
             print "sleeping"
             sleep(4)
         # Request to timings for various sub-jobs.
-        timings = requests.get(API + 'timings/' + jobid).json()
+        timings = requests.get(API + 'timings/' + pipeline_id).json()
         return timings
     except:
         # Case connection broke.

### 2018-04-03 scripts/timespfy.py
ADD: singlerun() func

@@ -129,9 +129,11 @@ def _timing(func, seeds):
         r.update(d)
     return r
 
-def main(spfy=True, bap=False, start=1, stop=22, step=5):
+def main(spfy=True, bap=False, rn=None):
+    if not rn:
+        rn = range(1,22,5)
     # Create groups of seed genomes.
-    seeds = [_seed_genomes(i) for i in range(start,stop,step)]
+    seeds = [_seed_genomes(i) for i in rn]
     # Run timings
     if spfy:
         r_spfy = _timing(_run_spfy, seeds)
@@ -139,5 +141,22 @@ def main(spfy=True, bap=False, start=1, stop=22, step=5):
         r_bap = _timing(_bap, seeds)
     return r_spfy, r_bap
 
+def singlerun(n=100):
+    '''Times Spfy 1-by-1 for up to n genomes. Used to find average/module.
+    '''
+    l = []
+    now = _now()
+    seeds = _seed_genomes(n)
+    for index, genome in enumerate(seeds):
+        # Run Spfy with a single genome.
+        print('{0}/{1} Running Spfy with file: {2}'.format(index,len(seeds),genome))
+        r = _run_spfy(genome)
+        l.append(r)
+    # Pickle file.
+    p = '{0}_singlerun_{1}.p'.format(now,n)
+    pickle.dump(l, open(p, "wb" ))
+    return l
+
 if __name__ == '__main__':
-    main()
+    rn = range(1,22,5)
+    main(rn=rn)

### 2018-04-03 scripts/timespfy.py
FIX: account for single files in _run_spfy()

@@ -71,16 +71,19 @@ def _run_spfy(list_genomes):
     '''POSTs to Spfy's API.
     '''
     # Zip files if more than 1 genome.
-    if len(list_genomes)>1:
-        # Zip the files.
-        name = _now() + '.zip'
-        with ZipFile(name, 'w') as z:
-            for f in list_genomes:
-                z.write(f)
-        # File payload.
-        files = {'file': open(name, 'rb')}
+    if isinstance(list_genomes, list):
+        if len(list_genomes)>1:
+            # Zip the files.
+            name = _now() + '.zip'
+            with ZipFile(name, 'w') as z:
+                for f in list_genomes:
+                    z.write(f)
+            # File payload.
+            files = {'file': open(name, 'rb')}
+        else:
+            files = {'file': open(list_genomes[0], 'rb')}
     else:
-        files = {'file': open(list_genomes[0], 'rb')}
+        files = {'file': open(list_genomes, 'rb')}
 
     # Options payload.
     data = {

### 2018-04-03 scripts/timespfy.py
FIX: abs paths

@@ -33,7 +33,7 @@ def _seed_genomes(size):
     pick = random.choice(os.listdir(GENOME_POOL))
     for i in range(size):
         while pick in st:
-            pick = random.choice(os.listdir(GENOME_POOL))
+            pick = os.path.join(GENOME_POOL, random.choice(os.listdir(GENOME_POOL)))
         st.add(pick)
     return list(st)
 

### 2018-04-03 scripts/timespfy.py
FIX: id retreiveal

@@ -29,11 +29,14 @@ def _now():
     return now
 
 def _seed_genomes(size):
+    def _pick():
+        f = os.path.join(GENOME_POOL, random.choice(os.listdir(GENOME_POOL)))
+        return f
     st = set()
-    pick = random.choice(os.listdir(GENOME_POOL))
+    pick = _pick()
     for i in range(size):
         while pick in st:
-            pick = os.path.join(GENOME_POOL, random.choice(os.listdir(GENOME_POOL)))
+            pick = _pick()
         st.add(pick)
     return list(st)
 
@@ -101,8 +104,8 @@ def _run_spfy(list_genomes):
 
     # POST.
     r = requests.post(API + 'upload', data=data, files=files)
-    pipeline_id = r.text
-    print("pipeline_id", pipeline_id)
+    pipeline_id = r.json.keys()[0]
+    print("pipeline_id: ".format(pipeline_id))
     # Sleep at least 4 second.
     sleep(4)
     try:
@@ -114,6 +117,7 @@ def _run_spfy(list_genomes):
             sleep(4)
         # Request to timings for various sub-jobs.
         timings = requests.get(API + 'timings/' + pipeline_id).json()
+        print("timings: {0}".format(timings))
         return timings
     except:
         # Case connection broke.

### 2018-04-03 scripts/timespfy.py
FIX: id retreiveal

@@ -104,7 +104,7 @@ def _run_spfy(list_genomes):
 
     # POST.
     r = requests.post(API + 'upload', data=data, files=files)
-    pipeline_id = r.json.keys()[0]
+    pipeline_id = r.json().keys()[0]
     print("pipeline_id: ".format(pipeline_id))
     # Sleep at least 4 second.
     sleep(4)
@@ -156,7 +156,7 @@ def singlerun(n=100):
     seeds = _seed_genomes(n)
     for index, genome in enumerate(seeds):
         # Run Spfy with a single genome.
-        print('{0}/{1} Running Spfy with file: {2}'.format(index,len(seeds),genome))
+        print('{0}/{1} Running Spfy with file: {2}'.format(index+1,len(seeds),genome))
         r = _run_spfy(genome)
         l.append(r)
     # Pickle file.

### 2018-04-03 scripts/timespfy.py
FIX: id retreiveal

@@ -104,7 +104,11 @@ def _run_spfy(list_genomes):
 
     # POST.
     r = requests.post(API + 'upload', data=data, files=files)
-    pipeline_id = r.json().keys()[0]
+    r_json = r.json()
+    print("r_json: {0}".format(r_json))
+    r_keys = r_json.keys()
+    print("r_keys: {0}".format(r_keys))
+    pipeline_id = r_keys[0]
     print("pipeline_id: ".format(pipeline_id))
     # Sleep at least 4 second.
     sleep(4)

### 2018-04-03 scripts/timespfy.py
FIX: id retreiveal

@@ -105,11 +105,9 @@ def _run_spfy(list_genomes):
     # POST.
     r = requests.post(API + 'upload', data=data, files=files)
     r_json = r.json()
-    print("r_json: {0}".format(r_json))
     r_keys = r_json.keys()
-    print("r_keys: {0}".format(r_keys))
     pipeline_id = r_keys[0]
-    print("pipeline_id: ".format(pipeline_id))
+    print("pipeline_id: {0}".format(pipeline_id))
     # Sleep at least 4 second.
     sleep(4)
     try:

### 2018-04-03 scripts/timespfy.py
ADD: small sample of raw single runtimes

@@ -104,9 +104,7 @@ def _run_spfy(list_genomes):
 
     # POST.
     r = requests.post(API + 'upload', data=data, files=files)
-    r_json = r.json()
-    r_keys = r_json.keys()
-    pipeline_id = r_keys[0]
+    pipeline_id = r.json().keys()[0]
     print("pipeline_id: {0}".format(pipeline_id))
     # Sleep at least 4 second.
     sleep(4)

### 2018-04-03 scripts/timespfy.py
ADD: setup for figures

@@ -158,7 +158,7 @@ def singlerun(n=100):
         # Run Spfy with a single genome.
         print('{0}/{1} Running Spfy with file: {2}'.format(index+1,len(seeds),genome))
         r = _run_spfy(genome)
-        l.append(r)
+        l.append({genome:r})
     # Pickle file.
     p = '{0}_singlerun_{1}.p'.format(now,n)
     pickle.dump(l, open(p, "wb" ))

### 2018-04-03 scripts/timespfy.py
DEBUG: see why we cant find pipelineid from production

@@ -104,7 +104,10 @@ def _run_spfy(list_genomes):
 
     # POST.
     r = requests.post(API + 'upload', data=data, files=files)
-    pipeline_id = r.json().keys()[0]
+    try:
+        pipeline_id = r.json().keys()[0]
+    except:
+        raise Exception('Could not find pipeline_id from response {0}'.format(r.text))
     print("pipeline_id: {0}".format(pipeline_id))
     # Sleep at least 4 second.
     sleep(4)

### 2018-04-03 scripts/timespfy.py
ADD: batch script

@@ -55,18 +55,17 @@ class Result:
         return np.array([self.x_values, self.y_values])
 
 class BarResult:
-    def __init__(self, list_sizes):
-        self.list_sizes = list_sizes
+    def __init__(self, seeds, now):
+        self.list_sizes = [len(l) for l in seeds]
+        self.seeds = seeds
         self.subtasks = None
-        self.data = {
-            'size_increments': list_sizes
-        }
+        self.data = []
+        self.now = now
 
     def update(self, timing):
         '''Variable "timing" should be of shape {sample_size(str): list_subtask_times(list)}
         '''
-        assert isinstance(timing, dict)
-        self.data.update(timing)
+        self.data.append(timing)
 
 # Calling functions.
 
@@ -132,24 +131,17 @@ def _bap(list_genomes):
     return True
 
 def _timing(func, seeds):
-    list_sizes = [len(l) for l in seeds]
-    r = BarResult(list_sizes)
+    now = _now()
+    r = BarResult(seeds, now)
+    raws = []
     for list_genomes in seeds:
         d = func(list_genomes)
         r.update(d)
+        raws.append(d)
+    # Pickle file.
+    pickle.dump(raws, open('{0}_spfy_raws_{1}.p'.format(now,len(seeds)), "wb" ))
+    pickle.dump(r, open('{0}_spfy_class_{1}.p'.format(now,len(seeds)), "wb" ))
     return r
-
-def main(spfy=True, bap=False, rn=None):
-    if not rn:
-        rn = range(1,22,5)
-    # Create groups of seed genomes.
-    seeds = [_seed_genomes(i) for i in rn]
-    # Run timings
-    if spfy:
-        r_spfy = _timing(_run_spfy, seeds)
-    if bap:
-        r_bap = _timing(_bap, seeds)
-    return r_spfy, r_bap
 
 def singlerun(n=100):
     '''Times Spfy 1-by-1 for up to n genomes. Used to find average/module.
@@ -167,6 +159,17 @@ def singlerun(n=100):
     pickle.dump(l, open(p, "wb" ))
     return l
 
+def main(spfy=True, bap=False, n=101):
+    rn = range(0,n+1,10)
+    rn[0]=1
+    # Create groups of seed genomes.
+    seeds = [_seed_genomes(i) for i in rn]
+    # Run timings
+    if spfy:
+        r_spfy = _timing(_run_spfy, seeds)
+    if bap:
[…]

### 2018-04-04 scripts/timespfy.py
FIX: dont write path

@@ -79,7 +79,7 @@ def _run_spfy(list_genomes):
             name = _now() + '.zip'
             with ZipFile(name, 'w') as z:
                 for f in list_genomes:
-                    z.write(f)
+                    z.write(f, os.path.basename(f))
             # File payload.
             files = {'file': open(name, 'rb')}
         else:
@@ -135,6 +135,7 @@ def _timing(func, seeds):
     r = BarResult(seeds, now)
     raws = []
     for list_genomes in seeds:
+        print('{0}/{1} Spfy Batch with files: {2}'.format(len(list_genomes),len(seeds[-1]),list_genomes))
         d = func(list_genomes)
         r.update(d)
         raws.append(d)

### 2018-04-04 scripts/timespfy.py
FIX: try/except for not found

@@ -163,14 +163,17 @@ def singlerun(n=100):
 def main(spfy=True, bap=False, n=101):
     rn = range(0,n+1,10)
     rn[0]=1
+    r = {}
     # Create groups of seed genomes.
     seeds = [_seed_genomes(i) for i in rn]
     # Run timings
     if spfy:
         r_spfy = _timing(_run_spfy, seeds)
+        r.update({'r_spfy':r_spfy})
     if bap:
         r_bap = _timing(_bap, seeds)
-    return r_spfy, r_bap
+        r.update({'r_bap':r_bap})
+    return r
 
 if __name__ == '__main__':
     main()

### 2018-04-04 scripts/timespfy.py
EDIT: more info in print messages

@@ -69,7 +69,7 @@ class BarResult:
 
 # Calling functions.
 
-def _run_spfy(list_genomes):
+def _run_spfy(list_genomes, on=''):
     '''POSTs to Spfy's API.
     '''
     # Zip files if more than 1 genome.
@@ -98,7 +98,7 @@ def _run_spfy(list_genomes):
         'options.eae': True,
         'options.groupresults': True,
         'options.bulk': False,
-        'options.pan': True
+        'options.pan': False
     }
 
     # POST.
@@ -108,6 +108,7 @@ def _run_spfy(list_genomes):
     except:
         raise Exception('Could not find pipeline_id from response {0}'.format(r.text))
     print("pipeline_id: {0}".format(pipeline_id))
+    started = datetime.now()
     # Sleep at least 4 second.
     sleep(4)
     try:
@@ -115,7 +116,7 @@ def _run_spfy(list_genomes):
         while requests.get(API + 'results/' + pipeline_id).json() == unicode('pending'):
             # The length we sleep doesn't matter, as timing is retrieved directly
             # from RQ.
-            print "sleeping"
+            print "On {0}, sleeping. Elapsed: {1}".format(on, started-datetime.now())
             sleep(4)
         # Request to timings for various sub-jobs.
         timings = requests.get(API + 'timings/' + pipeline_id).json()
@@ -125,7 +126,7 @@ def _run_spfy(list_genomes):
         # Case connection broke.
         return {}
 
-def _bap(list_genomes):
+def _bap(list_genomes, on=''):
     # BAP throws an error without KmerFinder.
     r = subprocess.check_call("""docker run -ti --rm -w /workdir -v $(pwd):/workdir    cgetools BAP --dbdir /usr/src/cgepipeline/test/databases  --services KmerFinder,ResFinder,VirulenceFinder  --fa /usr/src/cgepipeline/test/test.fa""", shell=True)
     return True
@@ -135,8 +136,9 @@ def _timing(func, seeds):
     r = BarResult(seeds, now)
     raws = []
     for list_genomes in seeds:
-        print('{0}/{1} Spfy Batch with files: {2}'.format(len(list_genomes),len(seeds[-1]),list_genomes))
-        d = func(list_genomes)
+        on = '{0}/{1}'.format(len(list_genomes),len(seeds[-1]))
+        print('{0} Spfy Batch with files: {1}'.format(on,list_genomes))
+        d = func(list_genomes, on)
         r.update(d)
         raws.append(d)
     # Pickle file.

### 2018-04-04 scripts/timespfy.py
EDIT: more info in print messages

@@ -163,6 +163,7 @@ def singlerun(n=100):
     return l
 
 def main(spfy=True, bap=False, n=101):
+    print('Starting timing run for {0} files\nFrom: {1}\nAgainst API: {2}'.format(n,GENOME_POOL,API))
     rn = range(0,n+1,10)
     rn[0]=1
     r = {}

### 2018-04-04 scripts/timespfy.py
EDIT: more info in print messages

@@ -116,7 +116,7 @@ def _run_spfy(list_genomes, on=''):
         while requests.get(API + 'results/' + pipeline_id).json() == unicode('pending'):
             # The length we sleep doesn't matter, as timing is retrieved directly
             # from RQ.
-            print "On {0}, sleeping. Elapsed: {1}".format(on, started-datetime.now())
+            print "On {0}, sleeping. Elapsed: {1}".format(on, datetime.now()-started)
             sleep(4)
         # Request to timings for various sub-jobs.
         timings = requests.get(API + 'timings/' + pipeline_id).json()
@@ -137,7 +137,7 @@ def _timing(func, seeds):
     raws = []
     for list_genomes in seeds:
         on = '{0}/{1}'.format(len(list_genomes),len(seeds[-1]))
-        print('{0} Spfy Batch with files: {1}'.format(on,list_genomes))
+        print('\n{0} Spfy Batch with files: {1}\n'.format(on,list_genomes))
         d = func(list_genomes, on)
         r.update(d)
         raws.append(d)

### 2018-04-11 scripts/timespfy.py
EDIT: incr delay for checks

@@ -109,7 +109,9 @@ def _run_spfy(list_genomes, on='', resultcls=None, phylotyper=True):
         if resultcls:
             resultcls.pipelines.append(pipeline_id)
     except:
-        raise Exception('Could not find pipeline_id from response {0}'.format(r.text))
+        print('Could not find pipeline_id from response {0}'.format(r.text))
+        sleep(20)
+        return {}
     print("pipeline_id: {0}".format(pipeline_id))
     started = datetime.now()
     # Sleep at least 4 second.
@@ -120,7 +122,7 @@ def _run_spfy(list_genomes, on='', resultcls=None, phylotyper=True):
             # The length we sleep doesn't matter, as timing is retrieved directly
             # from RQ.
             print "On {0}, sleeping. Elapsed: {1}".format(on, datetime.now()-started)
-            sleep(4)
+            sleep(20)
         # Request to timings for various sub-jobs.
         r = requests.get(API + 'timings/' + pipeline_id)
         try:

### 2018-04-12 scripts/timespfy.py
print msg + some diags

@@ -164,9 +164,10 @@ def singlerun(n=100):
     now = _now()
     seeds = _seed_genomes(n)
     for index, genome in enumerate(seeds):
+        on = '{0}/{1}'.format(index+1,len(seeds)
         # Run Spfy with a single genome.
-        print('{0}/{1} Running Spfy with file: {2}'.format(index+1,len(seeds),genome))
-        r = _run_spfy(genome)
+        print('{0} Running Spfy with file: {1}'.format(on,genome))
+        r = _run_spfy(genome, on=on)
         l.append({genome:r})
     # Pickle file.
     p = '{0}_singlerun_{1}.p'.format(now,n)

### 2018-04-12 scripts/timespfy.py
print msg + some diags

@@ -157,7 +157,7 @@ def _timing(func, seeds, phylotyper=True):
     pickle.dump(r, open('{0}_spfy_class_{1}.p'.format(now,len(seeds)), "wb" ))
     return r
 
-def singlerun(n=100):
+def singlerun(n=100, phylotyper=True):
     '''Times Spfy 1-by-1 for up to n genomes. Used to find average/module.
     '''
     l = []
@@ -167,7 +167,7 @@ def singlerun(n=100):
         on = '{0}/{1}'.format(index+1,len(seeds)
         # Run Spfy with a single genome.
         print('{0} Running Spfy with file: {1}'.format(on,genome))
-        r = _run_spfy(genome, on=on)
+        r = _run_spfy(genome, on=on, phylotyper=phylotyper)
         l.append({genome:r})
     # Pickle file.
     p = '{0}_singlerun_{1}.p'.format(now,n)

### 2018-04-12 scripts/timespfy.py
print msg + some diags

@@ -164,7 +164,7 @@ def singlerun(n=100, phylotyper=True):
     now = _now()
     seeds = _seed_genomes(n)
     for index, genome in enumerate(seeds):
-        on = '{0}/{1}'.format(index+1,len(seeds)
+        on = '{0}/{1}'.format(index+1,len(seeds))
         # Run Spfy with a single genome.
         print('{0} Running Spfy with file: {1}'.format(on,genome))
         r = _run_spfy(genome, on=on, phylotyper=phylotyper)

### 2018-04-13 paper-webserver.tex
ADD: fishers images into tex

@@
 
+On comparison tasks, Spfy can retrieve and compare 1 million nodes or attributes in the graph database in aproximately 70 seconds \ref{fig:fishers_performance}.
+Performance scales linearly with the number of genomes involved in the comparison, or the number of target nodes retrieved per genome.
+
+\begin{figure}[!htb]
+\minipage{0.32\textwidth}
+  \includegraphics[width=\linewidth]{images/fishers_genomes.png}
+\endminipage\hfill
+\minipage{0.32\textwidth}
+  \includegraphics[width=\linewidth]{images/fishers_targets.png}
+\endminipage\hfill
+\minipage{0.32\textwidth}%
+  \includegraphics[width=\linewidth]{images/fishers_overall.png}
+\endminipage
+\caption{Runtimes of Fisher's Exact Test depending on the number of nodes/attributes involved in the comparison.
+Left: Runtimes as the number of genomes involved increases for a fixed (107) number of targets per genome.
+Center: Runtimes as the number of targets involved increased for a fixed (116) number of genomes per target.
+Right: Runtimes as the total number of targets retrieved increases; the total number of targets is calculated as follows: (Number Genomes Group A + Number Genomes Group B) x Number Targets per Genome.
+In all cases, we see a linear increase in runtime as the number of targets or the number of genomes involved increases.}\label{fig:fishers_performance}
+\end{figure}
+
 \section{DISCUSSION}
@@
 % 1. Custom formats.
-% 2. Use of legacy tech (SQL) means youu can't accomadate "Big-Data" goals.
+% 2. Use of legacy tech (SQL) means youu can't accomadate "Big-Data" goals.{}
 

### 2018-04-13 paper-webserver.tex
ADD: spfy performance images into te

@@
 
+On analysis tasks, Spfy runs all included modules in 130 seconds on average, per file \ref{fig:spfy_performance}.
+Spfy can also queue up batches of files for analysis, decreasing the average runtime to 54 seconds on average with a degree of parallelization.
+
+\begin{figure}[!htb]
+\minipage{0.5\textwidth}
+  \includegraphics[width=\linewidth]{images/spfy_indivs.png}
+\endminipage\hfill
+\minipage{0.5\textwidth}%
+  \includegraphics[width=\linewidth]{images/spfy_batches.png}
+\endminipage
+\caption{Runtimes of Spfy's analysis modules.
+Left: 100 genome files are ran one-by-one with different colors representing the runtimes of the analysis module; Blue indicates the total runtime and the line represents the average runtime per file.
+Right: batches of files are run of increasing size.
+As indicated by the line, the average runtime per genome in consistant when running more than one file.
+This is due to the degree of parallelization meeting the capacity of the host computer.}\label{fig:spfy_performance}
+\end{figure}
+
 \section{DISCUSSION}
@@
 % re: Matt "2. When comparing to IRIDA, BAP etc., can mention some differences with Superphy, e.g. the storage of interim result data that allows downstream integrated analysis"
-
 % If we mention speed comparisons, we need to have actual data supporting them
 On a per file basis, Spfy performs at a similar speed to BAP on predictive genomics tasks, though Spfy does not provide genome assembly services.
-After accounting for assembly services, BAP reported \cite{thomsen2016bacterial} an average runtime of 8-9 minutes per genome over 476 runs.
-ResFinder, which is similar to the RGI program used by Spfy, took an average of 3-4 minutes.
-However, unlike these workflow managers, Spfy is designed to help solve the needless re-computation of analyses by storing results in a graph database for downstream comparative studies.
+After accounting for assembly services, BAP reported \cite{thomsen2016bacterial} an average runtime of 8-9 minutes per genome over 476 runs which is in the same scale as Spfy's average of over 2 minutes.
+ResFinder, which is similar to the RGI program used by Spfy, took an average of 3-4 minutes whereas RGI took an average of 1 minute 30 seconds.
+
+However, unlike these workflow managers, Spfy is designed to help solve the needless re-computation of analyses by storing results in a graph database for downstream comparative studies \ref{db_comparison}.
 This allows Spfy to perform population-wide analyses, regardless of the individual software used to generate the results for a given genome.

### 2018-04-16 paper-webserver.tex
EDIT: fig positions
START: mongo + total runtimes

@@
 It supports the following \textit{in-silico} subtyping options: serotyping, through both O- and H-antigen identification using ECTyper \url{https://github.com/phac-nml/ecoli_serotyping} as well as VF gene determination; Shiga-toxin 1 (Stx1), Shiga-toxin 2 (Stx2), and Intimin typing using Phylotyper \cite{whiteside2017phylotyper}; and AMR annotation using the resistance gene identifier from the comprehensive antibiotic resistance database \cite{mcarthur2013comprehensive}.
+An example of the VF results is given in \ref{fig-tables}.
 Spfy also performs pan-genome analyses using Panseq \cite{laing2010pan}, with the entire pan-genome stored in the database and associated with source genomes.
+
+\begin{figure}[!htb]
+\begin{center}{}
+\includegraphics[width=\textwidth]{images/tables.png}
+\end{center}
+\caption{Detailed results for the virulence factor subtyping task.
+While data storage in Spfy is graph-based, a familiar tabular structure is presented to users.
+The result can also be shared using the URL with the embedded token.}
+\label{fig-tables}
+\end{figure}
 
@@
 
+Spfy primarily uses Blazegraph \url{https://github.com/blazegraph/database} for storage along with MongoDB to cache a hash table for duplicate checking.
+The cache allows Spfy to more efficiently check for duplicate files in Blazegraph, than would be possible through an exaustive search of the graph structure.
+
 \subsection{Web design}
@@
 
-Both the task selection and result displays follow the same design pattern, where data storage is graph-based, but the results of analysis modules are presented to users in a familiar tabular structure and available for download as .csv spreadsheet files, as shown in Figure \ref{fig-tables}.
-
-\begin{figure}[!hb]
+\begin{figure}[!htb]
 \begin{center}
@@
 
-\begin{figure}[!hb]
-\begin{center}
-\includegraphics[width=\textwidth]{images/tables.png}
-\end{center}
-\caption{Detailed results for the virulence factor subtyping task.
-While data storage in Spfy is graph-based, a familiar tabular structure is presented to users.
-The result can also be shared using the URL with the embedded token.}
-\label{fig-tables}
-\end{figure}
+Both the task selection and result displays follow the same design pattern, where data storage is graph-based, but the results of analysis modules are presented to users in a familiar tabular structure and available for download as .csv spreadsheet files, as shown in Figure \ref{fig-tables}.
 
@@
 
-On comparison tasks, Spfy can retrieve and compare 1 million nodes or attributes in the graph database in aproximately 70 seconds \ref{fig:fishers_performance}.
-Performance scales linearly with the number of genomes involved in the comparison, or the number of target nodes retrieved per genome.
+Spfy is running on a virtual machine (VM) with 8 vCPUs emulating single-core Intel® Xeon® E3-series processors, and with 32 GB of RAM.
+The VM is running Linux with CentOS.
+On comparison tasks, Spfy can retrieve and compare 1 million nodes/attributes in the graph database in aproximately 70 seconds.
+As shown in Figure \ref{fig:fishers_performance}, performance scales linearly with the number of genomes involved in the comparison, or the number of target nodes retrieved per genome; 1.5 million nodes/attributes can be compared in aproximately 90 seconds, and 2 million nodes/attributes in aproximately 110 seconds.
 
@@
 Spfy can also queue up batches of files for analysis, decreasing the average runtime to 54 seconds on average with a degree of parallelization.
+% TODO: add totals
 

### 2018-04-16 paper-webserver.tex
ADD: fig for totals

@@
 
-On analysis tasks, Spfy runs all included modules in 130 seconds on average, per file \ref{fig:spfy_performance}.
-Spfy can also queue up batches of files for analysis, decreasing the average runtime to 54 seconds on average with a degree of parallelization.
-% TODO: add totals
+On analysis tasks, Spfy runs all included modules in 130 seconds on average, per file \ref{fig:spfy_performance_indivs}.
+Spfy can also queue up batches of files for analysis, decreasing the average runtime to an average of 54 seconds per file due to parallelization of analysis runs \ref{fig:spfy_performance_batches}.
+In total, its took 45 minutes to analyze 50 files and 89 minutes to analyze 100 files.
+
+\begin{figure}[!hb]
+\begin{center}
+\includegraphics[width=\textwidth]{images/spfy_indivs.png}
+\end{center}
+\caption{Runtimes of Spfy's analysis modules.
+100 genome files were run one-by-one with the different colors representing the runtimes of the different analysis module; The blue line indicates the average actual runtime per file after accounting for parallelization.}\label{fig:spfy_performance_indivs}
+\end{figure}
 
@@
 \minipage{0.5\textwidth}
-  \includegraphics[width=\linewidth]{images/spfy_indivs.png}
+  \includegraphics[width=\linewidth]{images/spfy_batches_totals.png}
 \endminipage\hfill
@@
 \endminipage
-\caption{Runtimes of Spfy's analysis modules.
-Left: 100 genome files are ran one-by-one with different colors representing the runtimes of the analysis module; Blue indicates the total runtime and the line represents the average runtime per file.
-Right: batches of files are run of increasing size.
+\caption{Runtimes of Spfy's analysis modules for batches of files.
+Left: Total runtime for analyzing all files in the given batch.
+The blue line indicates the actual time to completion after accounting for parallelization; 50 files are analyzed in 45 minutes and 100 files in 89 minutes.
+Right: Average runtime per file in the batch.
 As indicated by the line, the average runtime per genome in consistant when running more than one file.
-This is due to the degree of parallelization meeting the capacity of the host computer.}\label{fig:spfy_performance}
+This is due to the degree of parallelization reaching the capacity of the host computer.}\label{fig:spfy_performance_batches}
 \end{figure}

### 2018-04-18 paper-webserver.tex
EDIT: text only, still need to edit figure captions

@@
 It supports the following \textit{in-silico} subtyping options: serotyping, through both O- and H-antigen identification using ECTyper \url{https://github.com/phac-nml/ecoli_serotyping} as well as VF gene determination; Shiga-toxin 1 (Stx1), Shiga-toxin 2 (Stx2), and Intimin typing using Phylotyper \cite{whiteside2017phylotyper}; and AMR annotation using the resistance gene identifier from the comprehensive antibiotic resistance database \cite{mcarthur2013comprehensive}.
-An example of the VF results is given in \ref{fig-tables}.
+An example of the VF results is given in Figure \ref{fig-tables}.
 Spfy also performs pan-genome analyses using Panseq \cite{laing2010pan}, with the entire pan-genome stored in the database and associated with source genomes.
@@
 O26 as shown in Figure \ref{fig-groupings}.
-Different types can also be joined into a group through logical connectives AND, OR, Negation, and any data type is also a valid target.
-This approach can be used to compare any data regardless of source software.
+Different types can also be joined into a group through logical connectives AND, OR, or Negation. This approach can be used to compare any data regardless of source software.
 
@@
 Spfy primarily uses Blazegraph \url{https://github.com/blazegraph/database} for storage along with MongoDB to cache a hash table for duplicate checking.
-The cache allows Spfy to more efficiently check for duplicate files in Blazegraph, than would be possible through an exaustive search of the graph structure.
+The cache allows Spfy to more efficiently check for duplicate files in Blazegraph than would be possible through a search of the graph structure.
 
@@
 Spfy is running on a virtual machine (VM) with 8 vCPUs emulating single-core Intel® Xeon® E3-series processors, and with 32 GB of RAM.
-The VM is running Linux with CentOS.
+The VM is running CentOS version 7.3, with Docker version 17.06.1-ce and Docker-Compose version 1.12.0.
 On comparison tasks, Spfy can retrieve and compare 1 million nodes/attributes in the graph database in aproximately 70 seconds.
@@
 
-On analysis tasks, Spfy runs all included modules in 130 seconds on average, per file \ref{fig:spfy_performance_indivs}.
-Spfy can also queue up batches of files for analysis, decreasing the average runtime to an average of 54 seconds per file due to parallelization of analysis runs \ref{fig:spfy_performance_batches}.
-In total, its took 45 minutes to analyze 50 files and 89 minutes to analyze 100 files.
+On analysis tasks, Spfy runs all included modules in an average of 130 seconds per genome, as shown in Figure \ref{fig:spfy_performance_indivs}.
+Spfy can also queue batches of genomes for analysis, decreasing the average runtime to an average of 54 seconds per file due to parallelization of analysis runs \ref{fig:spfy_performance_batches}.
+In total, the platform analyzed 50 genomes in 45 minutes, and 100 genomes in 89 minutes.
 
@@
 After accounting for assembly services, BAP reported \cite{thomsen2016bacterial} an average runtime of 8-9 minutes per genome over 476 runs which is in the same scale as Spfy's average of over 2 minutes.
-ResFinder, which is similar to the RGI program used by Spfy, took an average of 3-4 minutes whereas RGI took an average of 1 minute 30 seconds.
+Note that the two platforms are now directly comparable, due to the differences in analysis tasks invovled.
+In similar tasks such as AMR determination, the ResFinder program included in BAP took an average of 3-4 minutes \cite{thomsen2016bacterial} and is similar to the RGI program included in Spfy, which took an average of 1 minute 30 seconds.
 

### 2018-04-18 paper-webserver.tex
EDIT: figure captions

@@
 
+No account creation is required to use the platform.
+A sharable token is automatically created for users upon entering the website, and is embedded into the website address.
+Users can share results by copying their URL, and files submitted from different computers using the same token will be visible to anyone with the same link.
+
 \begin{figure}[!htb]
@@
 \end{center}
-\caption{Detailed results for the virulence factor subtyping task.
+\caption{Detailed results for the serotyping and virulence factor subtyping task.
 While data storage in Spfy is graph-based, a familiar tabular structure is presented to users.
-The result can also be shared using the URL with the embedded token.}
+The genome file, GCA_001911825.1_ASM191182v1_genome.fna, was analyzed with the determined serotype associated with the file, and virulence factors associated with the contigouous DNA sequences they were found on.
+The Start/Stop positions on the contig., are provided along with the Percent Identity (Cutoff) used in the analysis.}
 \label{fig-tables}
@@
 In this example, there are 903 O157 genomes and 291 O26 genomes which are compared for the presence of 129 AMR genes.
-Fisher's Exact Test is used for the comparison.}
+Fisher's Exact Test is used for the comparison.
+Of a group A or B, Spfy reports the  number of genomes in that group with the particular Target, and the P-Value and Odds Ratio from Fisher's test.}
 \label{fig-groupings}
@@
 The cache allows Spfy to more efficiently check for duplicate files in Blazegraph than would be possible through a search of the graph structure.
+MongoDB is also used to support the
 
@@
 To ensure a familiar user interface, we followed the Material Design specification \url{https://material.io/}, published by Google, built around a card-based design (Figure \ref{fig-results}).
+Both the task selection and result displays follow the same design pattern, where data storage is graph-based, but the results of analysis modules are presented to users in a familiar tabular structure and available for download as .csv spreadsheet files.
 
@@
 \end{center}
-\caption{The results interface for submitted tasks. Cards represent individual tasks and the entire collection can be accessed by sharing the URL with the embedded token.}
+\caption{The results interface for submitted tasks.
+Cards represent individual tasks submitted by the user, such as checking the Database status, subtyping of single genomes, population comparisons, or subtyping of multiple genomes.}
 \label{fig-results}
 \end{figure}
-
-Both the task selection and result displays follow the same design pattern, where data storage is graph-based, but the results of analysis modules are presented to users in a familiar tabular structure and available for download as .csv spreadsheet files, as shown in Figure \ref{fig-tables}.
-
-No account creation is required to use the platform.
-A sharable token is automatically created for users upon entering the website, and is embedded into the website address.
-Users can share results by copying their URL, and files submitted from different computers using the same token will be visible to anyone with the same link.
 
@@
 \caption{The Docker containers used in Spfy.
-Arrows represent the connections between different containers, and the entire platform can be recreated with a single command using its Docker-Compose definition.}
+Arrows represent the connections between different containers, and the entire platform can be recreated with a single command using its Docker-Compose definition.
+Users access the platform using the ReactJS based website, which makes requests to the Flask API.
+Any requested analysis task is distributed to the Redis Task Queue and data files are stored in a Docker volume.
+MongoDB stores a hash table for efficient duplicate checking of results in Blazegraph.}
 \label{fig-docker}

### 2018-04-19 paper-webserver.tex
final edits and suggestions

--- 2018-04-18 paper-webserver.tex
@@
 
-To solve this problem, we created Spfy, which uses a graph database to store and retrieve results from computational workflows, and link data to individual genomes using standardized ontologies.
+To solve this problem, we created Spfy, a platform that rapidly perform the common reference lab tests, uses a graph database to store and retrieve the results from the computational workflows, and links data to individual genomes using standardized ontologies.
 
@@
 Though generally applicable to bacterial genome sequences, Spfy currently contains 10,243 \textit{Escherichia coli} genomes, for which \textit{in-silico} serotype and Shiga-toxin subtype, as well as the presence of known virulence factors and antimicrobial resistance determinants have been computed.
-Additionally, the presence / absence of then entire \textit{E.
+Additionally, the presence / absence of the entire \textit{E.
 coli} pan-genome was computed and linked to each genome.
@@
 The rapid characterization based on WGS relies on the outputs from multiple software programs that are targeted for specific applications.
-Examples include the identification of known antimicrobial resistance (AMR) genes, through software such as the Resistance Gene Identifier (RGI) \cite{mcarthur2013comprehensive}, \cite{kleinheinz2014applying}, \cite{gupta2014arg}, and ARIBA \cite{hunt2017ariba}; or the identification of known virulence factor genes (VF) through software such as VirulenceFinder \cite{kleinheinz2014applying}, SRST2 \cite{inouye2014srst2}, and GeneSippr \cite{lambert2015genesippr}.
-For clinical diagnoses and comparisons, individual species can be first divided into subtypes with complementing AMR and VF results.
-Software methods for subtyping rely on pre-selected intraspecies genes or genomic regions, and are targeted through software such as Phylotyper \cite{whiteside2017phylotyper}, SerotypeFinder \cite{joensen2015rapid}, the EcOH dataset applied through SRST2 \cite{ingle2016silico}, and V-Typer \cite{carrillo2016comparative}.
-These methods represent \textit{in-silico} analogues of traditional wet-lab tests, which allows new WGS results to be viewed in the context of historical tests, and greatly expedites the analyses of newly sequenced genomes.
+Examples include the identification of known antimicrobial resistance (AMR) genes, through software such as the Resistance Gene Identifier (RGI) \cite{mcarthur2013comprehensive}, ResFinder \cite{kleinheinz2014applying}, ARG-ANNOT \cite{gupta2014arg}, and ARIBA \cite{hunt2017ariba}; or the identification of known virulence factor genes (VF) through software such as VirulenceFinder \cite{kleinheinz2014applying}, SRST2 \cite{inouye2014srst2}, and GeneSippr \cite{lambert2015genesippr}.
+
+For subtyping in clinical diagnoses and reference lab environments, software relies on pre-selected intraspecies genes or genomic regions, which are targeted through software such as Phylotyper \cite{whiteside2017phylotyper}, SerotypeFinder \cite{joensen2015rapid}, the EcOH dataset applied through SRST2 \cite{ingle2016silico}, and V-Typer \cite{carrillo2016comparative}.
+These methods represent \textit{in-silico} analogues of traditional wet-lab tests, which allow new whole-genome sequences to be viewed in the context of historical tests without the need for the time and labor of the traditional wet-lab tests.
 
@@
 
-Many of the analyses that are currently used in the characterization and study of bacterial genomes are broadly useful, and therefore the same analyses are often computed multiple times for the same genome.
+Many of the analyses used in the characterization and study of bacterial genomes are broadly useful, and therefore they are often computed multiple times for the same genome across different studies.
 An effective method to mitigate this re-computation, is to make the storage and retrieval of results part of the analyses platform, and effectively linked to the genomes of interest through a standardized ontology.
@@
 % 5. how we solved it
-Here, we present a new platform merging Superphy's pre-computed results with a novel data storage and processing architecture, which we call Spfy; Spfy integrates a graph database with real-time analyses to avoid recomputing identical results.
-Graph-based result storage also allows retrospective comparisons across stored results as more genomes are sequenced or populations change, and is flexible, accommodating new analysis modules as they are developed.
+Here, we present a new platform merging the pre-computed results of SuperPhy, with a novel data storage and processing architecture, which we call Spfy; Spfy integrates a graph database with real-time analyses to avoid recomputing identical results.
+Additionally, graph-based result storage allows retrospective comparisons across stored results as more genomes are sequenced or populations change.
+Spfy is flexible, accommodating new analysis modules as they are developed.
 The database is available at \url{https://lfz.corefacility.ca/superphy/spfy/}.
@@
 Spfy handles all of the analyses tasks by dividing them into subtasks, which are subsequently distributed across a built-in task queue.
-Results are converted into individual graphs and stored within a larger graph database according to the previously created ontologies GenEpiO \cite{griffiths2017context}, FALDO \cite{bolleman2016faldo}, and TypOn \cite{vaz2014typon}, where metadata including genotypes, biomarkers, host, source are stored.
+Results are converted into individual graphs and stored within a larger graph database according to the previously created ontologies GenEpiO \cite{griffiths2017context}, FALDO \cite{bolleman2016faldo}, and TypOn \cite{vaz2014typon}, which includes metadata for genotypes, location, biomarkers, host, and source, among others.
 
@@
 All results are displayed to users in real-time, usually within 2-3 minutes.
-For example, Spfy can determine if there's a statistically significant difference, using Fisher's exact test, in the presence of all known AMR genes in collected O157 \textit{E.
-coli} genomes vs.
-O26 as shown in Figure \ref{fig-groupings}.
+For example, Spfy can determine if a statistically significant difference exists, using Fisher's exact test, among any identified AMR genes, between \textit{E.
+coli} genomes of serotype O157, and genomes of serotype O26, as shown in Figure \ref{fig-groupings}.
 Different types can also be joined into a group through logical connectives AND, OR, or Negation. This approach can be used to compare any data regardless of source software.
@@
 % This needs to be specific to the image we are showing eg. the O157 vs. O26 comparison. We mention the general comparison in the text, and this can be offered as a specific example of that.
-Genomes with the predicted subtype O157 are compared against the subtype O26.
-In this example, there are 903 O157 genomes and 291 O26 genomes which are compared for the presence of 129 AMR genes.
-Fisher's Exact Test is used for the comparison.}
+\textit{E.
+coli} genomes of serotype O157 (903) compared against genome of serotype O26 (291), for statistically significant differences in the carriage of 129 AMR genes.
+Fisher's Exact Test is used by Spfy for these comparisons.}
 \label{fig-groupings}
@@
 % para: intro to the spfy stack
-The server-side code for Spfy, graph generation, and analysis modules, are developed in Python, with the front-end website developed using the React JavaScript library \url{https://facebook.github.io/react/}.
+The server-side code for Spfy, graph generation, and analysis modules, were developed in Python, with the front-end website developed using the React JavaScript library \url{https://facebook.github.io/react/}.
 When new data is added to the database, the following steps are taken:
@@
 % para: the semantic web
-Semantic web technologies describe the relationships between data, and have been proposed as an open standard for sharing public information \cite{berners2001semantic}, while graph databases are a flexible means of storing this this information \cite{horrocks2005semantic}.
+Semantic web technologies describe the relationships between data, and have been proposed as an open standard for sharing public information \cite{berners2001semantic}, while graph databases are a flexible means of storing this information \cite{horrocks2005semantic}.
 Biological data can be a genome, contiguous DNA sequence, or gene, and these are linked together in a searchable graph structure using existing ontologies.
@@
 
-% The text in this figure is too small to read and seems blurry. It should be enlarged.
+% This is still true. The text needs to be larger: --> The text in this figure is too small to read and seems blurry. It should be enlarged.
 \begin{figure}[!hb]
@@
 coli} assembled genomes from Enterobase, storing every sequence and the results for all included analysis modules.
-This included: serotyping (O-antigen, H-antigen), toxin sub-typing (Shiga-toxin 1, Shiga-toxin 2, and Intimin), the identification of VF and AMR determinants, and determination of the pan-genome content of \textit{E.
+These included: serotyping (O-antigen, H-antigen), toxin sub-typing (Shiga-toxin 1, Shiga-toxin 2, and Intimin), the identification of VF and AMR determinants, and determination of the pan-genome content of \textit{E.
 coli}.
@@
 The VM is running CentOS version 7.3, with Docker version 17.06.1-ce and Docker-Compose version 1.12.0.
-On comparison tasks, Spfy can retrieve and compare 1 million nodes/attributes in the graph database in aproximately 70 seconds.
-As shown in Figure \ref{fig:fishers_performance}, performance scales linearly with the number of genomes involved in the comparison, or the number of target nodes retrieved per genome; 1.5 million nodes/attributes can be compared in aproximately 90 seconds, and 2 million nodes/attributes in aproximately 110 seconds.
+On comparison tasks, Spfy can retrieve and compare 1 million nodes/attributes in the graph database in approximately 70 seconds.
+As shown in Figure \ref{fig:fishers_performance}, performance scales linearly with the number of genomes involved in the comparison, or the number of target nodes retrieved per genome; 1.5 million nodes/attributes can be compared in approximately 90 seconds, and 2 million nodes/attributes in approximately 110 seconds.
 
+% The X and Y axes labels need to be slightly larger.
+Designate the panes a), b), and c) and then reference them in the caption.
+The caption should contain the information that the titles on each pane currently do.
 \begin{figure}[!htb]
@@
 \caption{Runtimes of Fisher's Exact Test depending on the number of nodes/attributes involved in the comparison.
-Left: Runtimes as the number of genomes involved increases for a fixed (107) number of targets per genome.
-Center: Runtimes as the number of targets involved increased for a fixed (116) number of genomes per target.
-Right: Runtimes as the total number of targets retrieved increases; the total number of targets is calculated as follows: (Number Genomes Group A + Number Genomes Group B) x Number Targets per Genome.
-In all cases, we see a linear increase in runtime as the number of targets or the number of genomes involved increases.}\label{fig:fishers_performance}
+Left: Runtimes as the number of genomes increased for a fixed (107) number of targets per genome.
+Center: Runtimes as the number of targets increased for a fixed (116) number of genomes per target.
+Right: Runtimes as the total number of targets retrieved increased; the total number of targets was calculated as follows: (Number Genomes Group A + Number Genomes Group B) x Number Targets per Genome.
+In all cases, a linear increase in runtime was observered as the number of targets or genomes increased.}\label{fig:fishers_performance}
 \end{figure}
 
-On analysis tasks, Spfy runs all included modules in an average of 130 seconds per genome, as shown in Figure \ref{fig:spfy_performance_indivs}.
+On analysis tasks, Spfy runs all included analyses modules in an average of 130 seconds per genome, as shown in Figure \ref{fig:spfy_performance_indivs}.
 Spfy can also queue batches of genomes for analysis, decreasing the average runtime to an average of 54 seconds per file due to parallelization of analysis runs \ref{fig:spfy_performance_batches}.
@@
 \caption{Runtimes of Spfy's analysis modules.
-100 genome files were run one-by-one with the different colors representing the runtimes of the different analysis module; The blue line indicates the average actual runtime per file after accounting for parallelization.}\label{fig:spfy_performance_indivs}
+100 genomes were run individually, with each coloured dot representing the runtime of a separate analysis module.
+The blue line indicates the average actual runtime per file after accounting for parallelization.}\label{fig:spfy_performance_indivs}
 \end{figure}
 
+% Again, label panes as a) and b), then reference in the caption
 \begin{figure}[!htb]
@@
 Right: Average runtime per file in the batch.
-As indicated by the line, the average runtime per genome in consistant when running more than one file.
-This is due to the degree of parallelization reaching the capacity of the host computer.}\label{fig:spfy_performance_batches}
+As indicated by the line, the average runtime per genome in consistent when running more than one file due to the parallelization of analyses.}\label{fig:spfy_performance_batches}
 \end{figure}
@@
 % This has not been shown in this paper; it has only been stated. We should probably have a large-scale analysis for an example, with accompanying times etc.
-We have shown how a graph database can accommodate the results from a variety of bioinformatics programs, and how Spfy is performant for data retrieval on the results for multiple analyses among over 10,000 genomes, providing results from big-data comparisons in the same efficiency as old analyses on single files.
+We have shown how a graph database can accommodate the results from a variety of bioinformatics programs, and how Spfy is performant for data retrieval of the results from multiple analyses among over 10,000 genomes.
+Spfy provides results from big-data comparisons with the same efficiency as old analyses on single files.
 
@@
 The automated analyses of WGS is currently facilitated by existing scientific workflow technologies such as Galaxy \cite{goecks2010galaxy}.
-Galaxy aims to provide a reproducible, computational interface which is accessible to individuals without programming knowledge.
+Galaxy aims to provide a reproducible, computational interface that is accessible to individuals without programming knowledge.
 Galaxy defines a formal schema for linking different analysis software together, so the entire pipeline can be replicated and also extended as new tools are developed.
@@
 After accounting for assembly services, BAP reported \cite{thomsen2016bacterial} an average runtime of 8-9 minutes per genome over 476 runs which is in the same scale as Spfy's average of over 2 minutes.
-Note that the two platforms are now directly comparable, due to the differences in analysis tasks invovled.
+Note that the two platforms are not directly comparable, due to the differences in analysis tasks involved.
 In similar tasks such as AMR determination, the ResFinder program included in BAP took an average of 3-4 minutes \cite{thomsen2016bacterial} and is similar to the RGI program included in Spfy, which took an average of 1 minute 30 seconds.
@@
 PATRIC has support for comparing up to nine user-submitted genomes against a reference genome, based on gene annotations; the platform indexes a NoSQL document store to compare similar document types.
-Unlike PARTIC, Spfy provides the ability to perform statistical comparisons of any permutation of a population group based on the chosen data types, and Spfy's graph database has no limit on the number of genomes grouped for comparison.
+Unlike PARTIC, Spfy provides the ability to perform statistical comparisons of any permutation of a population group based on the chosen data types, and the graph database of Spfy has no limit on the number of genomes grouped for comparison.
 
@@
 
-The integrated approach taken in the creation of Spfy, where the storage and retrieval of results provides enormous benefits for the large-scale analyes of \textit{E.
+The integrated approach taken in the creation of Spfy, where the analyses, storage, and retrieval of results is combined, provides enormous benefits for the large-scale analyses of \textit{E.
 coli}.
-The developed analyses modules are also self-contained and be transferred to existing platforms such as Galaxy.
+The developed analyses modules are also self-contained and can be used in existing platforms such as Galaxy.
 Future work will focus on adding machine learning modules to improve genotype / phenotype predictions, and supporting bacterial species such as \textit{Salmonella}, and \textit{Campylobacter}.

### 2018-04-19 paper-webserver.tex
ADD: matt's edits. Typos and a few suggestions

--- 2018-04-18 paper-webserver.tex
@@
 Though generally applicable to bacterial genome sequences, Spfy currently contains 10,243 \textit{Escherichia coli} genomes, for which \textit{in-silico} serotype and Shiga-toxin subtype, as well as the presence of known virulence factors and antimicrobial resistance determinants have been computed.
-Additionally, the presence / absence of then entire \textit{E.
+Additionally, the presence / absence of the entire \textit{E.
 coli} pan-genome was computed and linked to each genome.
 
-Owing to its vast database of pre-computed results, and the ability to easily incorporate user data, Spfy facilitates hypothesis testing in fields ranging from population genomics to epidemiology, while mitigating the re-computation of analyses.
+Owing to its database of diverse pre-computed results, and the ability to easily incorporate user data, Spfy facilitates hypothesis testing in fields ranging from population genomics to epidemiology, while mitigating the re-computation of analyses.
 
@@
 Spfy provides a database and analyses approach for \textit{E.
-coli} that is able to match the rapid accumulation of WGS in public databases.
+coli} that is able to match the rapid accumulation of WGS data in public databases.
 \par
@@
 
+% MW: I would remove this section. Its not that critical to anyone that uses Spfy and would be an individual
+% choice for developers that obtain the code
+
 Tests for functionality and backwards compatibility are run on TravisCI \url{https://travisci.io}, a continuous integration (CI) platform.
@@
 In total, the platform analyzed 50 genomes in 45 minutes, and 100 genomes in 89 minutes.
+
+% MW I would move the next 2 figures to supplementary information to cut down on figures
+% Often there is a limit on the # of figs.
 
@@
 After accounting for assembly services, BAP reported \cite{thomsen2016bacterial} an average runtime of 8-9 minutes per genome over 476 runs which is in the same scale as Spfy's average of over 2 minutes.
-Note that the two platforms are now directly comparable, due to the differences in analysis tasks invovled.
+Note, however, that the two platforms are not directly comparable, due to the differences in analysis tasks involved.
 In similar tasks such as AMR determination, the ResFinder program included in BAP took an average of 3-4 minutes \cite{thomsen2016bacterial} and is similar to the RGI program included in Spfy, which took an average of 1 minute 30 seconds.
 
-However, unlike these workflow managers, Spfy is designed to help solve the needless re-computation of analyses by storing results in a graph database for downstream comparative studies \ref{db_comparison}.
-This allows Spfy to perform population-wide analyses, regardless of the individual software used to generate the results for a given genome.
+However, unlike these workflow managers, Spfy is designed to help solve the re-computation of analyses by storing results in a graph database for downstream comparative studies \ref{db_comparison}.
+This allows Spfy to, for example, perform population-wide analyses on varied data from multiple diverse individual software.
 
@@
 coli}.
-The developed analyses modules are also self-contained and be transferred to existing platforms such as Galaxy.
+The developed analyses modules are also self-contained and can be transferred to existing platforms such as Galaxy.
 Future work will focus on adding machine learning modules to improve genotype / phenotype predictions, and supporting bacterial species such as \textit{Salmonella}, and \textit{Campylobacter}.

### 2018-04-19 paper-webserver.tex

--- 2018-04-19 paper-webserver.tex
@@
 
-Owing to its vast database of pre-computed results, and the ability to easily incorporate user data, Spfy facilitates hypothesis testing in fields ranging from population genomics to epidemiology, while mitigating the re-computation of analyses.
+Owing to its database of diverse pre-computed results, and the ability to easily incorporate user data, Spfy facilitates hypothesis testing in fields ranging from population genomics to epidemiology, while mitigating the re-computation of analyses.
 
@@
 After accounting for assembly services, BAP reported \cite{thomsen2016bacterial} an average runtime of 8-9 minutes per genome over 476 runs which is in the same scale as Spfy's average of over 2 minutes.
-Note that the two platforms are not directly comparable, due to the differences in analysis tasks involved.
+Note, however, that the two platforms are not directly comparable, due to the differences in analysis tasks involved.
 In similar tasks such as AMR determination, the ResFinder program included in BAP took an average of 3-4 minutes \cite{thomsen2016bacterial} and is similar to the RGI program included in Spfy, which took an average of 1 minute 30 seconds.
 
-However, unlike these workflow managers, Spfy is designed to help solve the needless re-computation of analyses by storing results in a graph database for downstream comparative studies \ref{db_comparison}.
-This allows Spfy to perform population-wide analyses, regardless of the individual software used to generate the results for a given genome.
+However, unlike these workflow managers, Spfy is designed to help solve the re-computation of analyses by storing results in a graph database for downstream comparative studies \ref{db_comparison}.
+This allows Spfy to, for example, perform population-wide analyses on varied data from multiple diverse individual software.
 

### 2018-04-19 paper-webserver.tex
STOP: fix for latex compilation

--- 2018-04-19 paper-webserver.tex
@@
 Though generally applicable to bacterial genome sequences, Spfy currently contains 10,243 \textit{Escherichia coli} genomes, for which \textit{in-silico} serotype and Shiga-toxin subtype, as well as the presence of known virulence factors and antimicrobial resistance determinants have been computed.
-Additionally, the presence / absence of the entire \textit{E.
+Additionally, the presence / absence of then entire \textit{E.
 coli} pan-genome was computed and linked to each genome.
 
-Owing to its database of diverse pre-computed results, and the ability to easily incorporate user data, Spfy facilitates hypothesis testing in fields ranging from population genomics to epidemiology, while mitigating the re-computation of analyses.
+Owing to its vast database of pre-computed results, and the ability to easily incorporate user data, Spfy facilitates hypothesis testing in fields ranging from population genomics to epidemiology, while mitigating the re-computation of analyses.
 
@@
 Spfy provides a database and analyses approach for \textit{E.
-coli} that is able to match the rapid accumulation of WGS data in public databases.
+coli} that is able to match the rapid accumulation of WGS in public databases.
 \par
@@
 \end{center}
-\caption{Detailed results for the virulence factor subtyping task.
+\caption{Detailed results for the serotyping and virulence factor subtyping task.
 While data storage in Spfy is graph-based, a familiar tabular structure is presented to users.
-The result can also be shared using the URL with the embedded token.}
+The genome file, GCA\_001911825.1\_ASM191182v1\_genome.fna, was analyzed with the determined serotype associated with the file, and virulence factors associated with the contigouous DNA sequences they were found on.
+The Start/Stop positions on the contig., are provided along with the Percent Identity (Cutoff) used in the analysis.}
 \label{fig-tables}
@@
 % 2. how we've implemented it, integration with github
-
-% MW: I would remove this section. Its not that critical to anyone that uses Spfy and would be an individual
-% choice for developers that obtain the code
 
@@
 In total, the platform analyzed 50 genomes in 45 minutes, and 100 genomes in 89 minutes.
-
-% MW I would move the next 2 figures to supplementary information to cut down on figures
-% Often there is a limit on the # of figs.
 
@@
 After accounting for assembly services, BAP reported \cite{thomsen2016bacterial} an average runtime of 8-9 minutes per genome over 476 runs which is in the same scale as Spfy's average of over 2 minutes.
-Note, however, that the two platforms are not directly comparable, due to the differences in analysis tasks involved.
+Note that the two platforms are now directly comparable, due to the differences in analysis tasks invovled.
 In similar tasks such as AMR determination, the ResFinder program included in BAP took an average of 3-4 minutes \cite{thomsen2016bacterial} and is similar to the RGI program included in Spfy, which took an average of 1 minute 30 seconds.
 
-However, unlike these workflow managers, Spfy is designed to help solve the re-computation of analyses by storing results in a graph database for downstream comparative studies \ref{db_comparison}.
-This allows Spfy to, for example, perform population-wide analyses on varied data from multiple diverse individual software.
+However, unlike these workflow managers, Spfy is designed to help solve the needless re-computation of analyses by storing results in a graph database for downstream comparative studies \ref{db_comparison}.
+This allows Spfy to perform population-wide analyses, regardless of the individual software used to generate the results for a given genome.
 
@@
 coli}.
-The developed analyses modules are also self-contained and can be transferred to existing platforms such as Galaxy.
+The developed analyses modules are also self-contained and be transferred to existing platforms such as Galaxy.
 Future work will focus on adding machine learning modules to improve genotype / phenotype predictions, and supporting bacterial species such as \textit{Salmonella}, and \textit{Campylobacter}.

### 2018-04-19 paper-webserver.tex
CHANGE: made suggestions rec by Matt

@@
 \begin{center}
-\includegraphics[width=\textwidth]{images/ontology}
+\includegraphics[width=\textwidth]{images/ontology_r3}
 \end{center}
@@
 
-\subsection{Continuous integration}
-% Keep this short
-% para
-% 1. goals: why CI, testing is important
-% 2. how we've implemented it, integration with github
-
-% MW: I would remove this section. Its not that critical to anyone that uses Spfy and would be an individual
-% choice for developers that obtain the code
-
-Tests for functionality and backwards compatibility are run on TravisCI \url{https://travisci.io}, a continuous integration (CI) platform.
-The individual tests use PyTest \url{https://doc.pytest.org/}, and the current build status can be checked on GitHub at \url{https://travis-ci.org/superphy/backend}.
-TravisCI also builds the core Docker images for Spfy, and uploads them to Docker Hub \url{https://hub.docker.com/u/superphy/}.
-
 \section{RESULTS}
@@
 
-On analysis tasks, Spfy runs all included analyses modules in an average of 130 seconds per genome, as shown in Figure \ref{fig:spfy_performance_indivs}.
-Spfy can also queue batches of genomes for analysis, decreasing the average runtime to an average of 54 seconds per file due to parallelization of analysis runs \ref{fig:spfy_performance_batches}.
+On analysis tasks, Spfy runs all included analyses modules in an average of 130 seconds per genome.
+Spfy can also queue batches of genomes for analysis, decreasing the average runtime to an average of 54 seconds per file due to parallelization of analysis runs \ref{fig:spfy_performance}.
 In total, the platform analyzed 50 genomes in 45 minutes, and 100 genomes in 89 minutes.
 
-% MW I would move the next 2 figures to supplementary information to cut down on figures
-% Often there is a limit on the # of figs.
-
-\begin{figure}[!hb]
-\begin{center}
-\includegraphics[width=\textwidth]{images/spfy_indivs.png}
-\end{center}
-\caption{Runtimes of Spfy's analysis modules.
-100 genomes were run individually, with each coloured dot representing the runtime of a separate analysis module.
-The blue line indicates the average actual runtime per file after accounting for parallelization.}\label{fig:spfy_performance_indivs}
-\end{figure}
-
-% Again, label panes as a) and b), then reference in the caption
 \begin{figure}[!htb]
-\minipage{0.5\textwidth}
+\minipage{0.5\textwidth}%
   \includegraphics[width=\linewidth]{images/spfy_batches_totals.png}
-\endminipage\hfill
-\minipage{0.5\textwidth}%
-  \includegraphics[width=\linewidth]{images/spfy_batches.png}
 \endminipage
-\caption{Runtimes of Spfy's analysis modules for batches of files.
-Left: Total runtime for analyzing all files in the given batch.
-The blue line indicates the actual time to completion after accounting for parallelization; 50 files are analyzed in 45 minutes and 100 files in 89 minutes.
-Right: Average runtime per file in the batch.
-As indicated by the line, the average runtime per genome in consistent when running more than one file due to the parallelization of analyses.}\label{fig:spfy_performance_batches}
+\caption{Total runtimes of Spfy's analysis modules for batches of files.
+The blue line indicates the actual time to completion after accounting for parallelization; 50 files are analyzed in 45 minutes and 100 files in 89 minutes.}\label{fig:spfy_performance}
 \end{figure}
@@
 
+\section{SUPPLEMENTARY INFORMATION}
+
+% Again, label panes as a) and b), then reference in the caption
+\begin{figure}[!htb]
+\minipage{0.5\textwidth}
+  \includegraphics[width=\linewidth]{images/spfy_indivs.png}
+\endminipage\hfill
+\minipage{0.5\textwidth}%
+  \includegraphics[width=\linewidth]{images/spfy_batches.png}
+\endminipage
+\caption{Runtimes of Spfy's analysis modules.
+A: 100 genomes were run individually, with each coloured dot representing the runtime of a separate analysis module.
+The blue line indicates the average actual runtime per file after accounting for parallelization.
+B: Average runtime when multiple genomes are enqueued together.
+As indicated by the line, the average runtime per genome in consistent when running more than one file.}\label{fig:supp_spfy_performance}
+\end{figure}
+
 \end{document}

### 2018-04-19 paper-webserver.tex
STOP: off to DG

@@
 \author[1]{Matthew D Whiteside}
-\author[1]{James Hopkins}
+\author[1]{James E Hopkins}
 \author[1]{Victor PJ Gannon}
@@
 
-% This is still true. The text needs to be larger: --> The text in this figure is too small to read and seems blurry. It should be enlarged.
 \begin{figure}[!hb]
 \begin{center}
-\includegraphics[width=\textwidth]{images/ontology_r3}
+\includegraphics[width=\textwidth]{images/ontology_r6}
 \end{center}
@@
 Data are added in as the analysis modules complete, at varying times, and the overall connections are inferred by the database.
-Non-bracketed sections are sourced from the uploaded genome files or user-supplied metadata.}
+Non-bracketed sections are sourced from the uploaded genome files or user-supplied metadata.
+Figure was generated using \url{http://www.visualdataweb.de/webvowl/}.}
 \label{fig-ontology}
@@
 \caption{Runtimes of Fisher's Exact Test depending on the number of nodes/attributes involved in the comparison.
-Left: Runtimes as the number of genomes increased for a fixed (107) number of targets per genome.
-Center: Runtimes as the number of targets increased for a fixed (116) number of genomes per target.
-Right: Runtimes as the total number of targets retrieved increased; the total number of targets was calculated as follows: (Number Genomes Group A + Number Genomes Group B) x Number Targets per Genome.
+A: Runtimes as the number of genomes increased for a fixed (107) number of targets per genome.
+B: Runtimes as the number of targets increased for a fixed (116) number of genomes per target.
+C: Overall runtimes as the total number of targets retrieved increased; the total number of targets was calculated as follows: (Number Genomes Group A + Number Genomes Group B) x Number Targets per Genome.
 In all cases, a linear increase in runtime was observered as the number of targets or genomes increased.}\label{fig:fishers_performance}
@@
 \begin{figure}[!htb]
+\begin{center}
 \minipage{0.5\textwidth}%
@@
 The blue line indicates the actual time to completion after accounting for parallelization; 50 files are analyzed in 45 minutes and 100 files in 89 minutes.}\label{fig:spfy_performance}
+\end{center}
 \end{figure}

### 2018-04-19 paper-webserver.tex
EDIT: prefer the e+ scaling

@@
 \textit{E.
-coli} genomes of serotype O157 (903) compared against genome of serotype O26 (291), for statistically significant differences in the carriage of 129 AMR genes.
+coli} genomes of serotype O157 (903 genomes) compared against genome of serotype O26 (291 genomes), for statistically significant differences in the carriage of 129 AMR genes.
 Fisher's Exact Test is used by Spfy for these comparisons.}

### 2018-04-22 paper-webserver.tex

@@
 
-\title{Spfy: an integrated graph database for real-time prediction of Escherichia coli phenotypes and downstream comparative analyses}
+\title{Spfy: an integrated graph database for real-time prediction of bacterial phenotypes and downstream comparative analyses}
 
@@
 
-To solve this problem, we created Spfy, a platform that rapidly perform the common reference lab tests, uses a graph database to store and retrieve the results from the computational workflows, and links data to individual genomes using standardized ontologies.
+To solve this problem, we created Spfy, a platform that rapidly performs the common reference lab tests, uses a graph database to store and retrieve the results from the computational workflows, and links data to individual genomes using standardized ontologies.
 

### 2018-04-22 tables/comparison.tex

@@
 \centering
-\caption{Comparison of various bioinformatic pipelines and their underlying database.
-Functionally, Spfy integrates different analysis modules as in BAP while also merging large datasets as in PATRIC.}
-\label{Table 2.}
+\caption{Comparison of four bioinformatic pipelines and their underlying database. Functionally, Spfy integrates different analysis modules as in BAP while also merging large datasets as in PATRIC.}
+\label{db_comparison}
 \resizebox{\textwidth}{!}{%
 \begin{tabular}{lllll}
-         & Spfy                & Galaxy              & BAP                 & PATRIC              \\
+         & Spfy                & Galaxy              & BAP                 & PATRIC              \\ \hline
 Database & Blazegraph          & PostgreSQL          & MySQL + File System & MongoDB + Shock     \\

### 2018-06-15 paper-webserver.tex

@@
 
-\author[1]{Kevin K Le\thanks{kevin.le@canada.ca}}
+\author[1]{Kevin K Le\thanks{kle009@uottawa.ca}}
 \author[1]{Matthew D Whiteside}
@@
 The cache allows Spfy to more efficiently check for duplicate files in Blazegraph than would be possible through a search of the graph structure.
-MongoDB is also used to support the
+MongoDB is also used to support the synchronized user sessions feature of the website.
 
@@
 
-However, unlike these workflow managers, Spfy is designed to help solve the re-computation of analyses by storing results in a graph database for downstream comparative studies \ref{db_comparison}.
+However, unlike these workflow managers, Spfy is designed to help solve the re-computation of analyses by storing results in a graph database for downstream comparative studies (Table \ref{db_comparison}).
 This allows Spfy to, for example, perform population-wide analyses on varied data from multiple diverse individual software.
@@
 
-\section{ACKNOWLEDGEMENTS}
-
 \newpage
@@
 
-\newpage
-
-\section{APPENDIX}
-
-\section{SUPPLEMENTARY INFORMATION}
-
-% Again, label panes as a) and b), then reference in the caption
-\begin{figure}[!htb]
-\minipage{0.5\textwidth}
-  \includegraphics[width=\linewidth]{images/spfy_indivs.png}
-\endminipage\hfill
-\minipage{0.5\textwidth}%
-  \includegraphics[width=\linewidth]{images/spfy_batches.png}
-\endminipage
-\caption{Runtimes of Spfy's analysis modules.
-A: 100 genomes were run individually, with each coloured dot representing the runtime of a separate analysis module.
-The blue line indicates the average actual runtime per file after accounting for parallelization.
-B: Average runtime when multiple genomes are enqueued together.
-As indicated by the line, the average runtime per genome in consistent when running more than one file.}\label{fig:supp_spfy_performance}
-\end{figure}
-
 \end{document}
